### 2021-06-04 pass_through_paper/paper_draft_jan_7_up.tex
script updates

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage[sort&compress]{natbib}
\usepackage{amssymb, amsmath, amsthm}
\usepackage[breaklinks=true]{hyperref}
\hypersetup{
    colorlinks,
    citecolor= black,
    filecolor= black,
    linkcolor= black,
    urlcolor= black
}
\usepackage{setspace}
\usepackage{tikz}
\usetikzlibrary{patterns,arrows,decorations.pathreplacing}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage[normalem]{ulem}

\usepackage{siunitx}
\sisetup{group-separator={,},group-minimum-digits=3}

\usepackage[multiple]{footmisc}

\newcommand{\insitu}{\textit{in situ} }
%%%%%%%%%%

% FOR COMMENTING:
\usepackage{comment}
\usepackage{ifthen}
\usepackage{color}
\usepackage{tabularx}

\specialcomment{ALnote}{\begingroup\sffamily\colorbox{blue}{Andrew: } \color{blue}}{\endgroup}
\specialcomment{BBnote}{\begingroup\sffamily\colorbox{red}{Branko: }\color{red}}{\endgroup}

\setlength{\abovecaptionskip}{10pt}
\setlength{\belowcaptionskip}{10pt}

\addtolength{\parskip}{5pt}

%\usepackage[normalem]{ulem}

%%%% TITLE %%%%
\title{Carbon Price Pass-Through in Alberta's Electricity Market\thanks{We are grateful to Sam Harrison for excellent
research assistance. We thank seminar and conference participants at the 2019 AERE Summer Meetings for comments and suggestions. Financial support for this work was provided by the Future Energy Systems Canada First Research Excellence Fund grant.
}
}

%%%% DATE %%%%
\date{\today}

%%%% AUTHOR INFORMATION %%%%
% AUTHOR INFO
\author{Andrew Leach\\
 University of Alberta
 \and
Blake Shaffer\\University of Calgary and\\Stanford University
}
\begin{document}
\maketitle

%%%% ABSTRACT %%%%
\begin{abstract}
\noindent We evaluate .\\
\vspace{0.2in}
\noindent Keywords: climate change, social cost of carbon, oil sands\\
JEL classification: Q3, Q4, Q54
\end{abstract}

\thispagestyle{empty}
\newpage
\onehalfspacing

\section{Introduction}

We evaluate the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.  Between 2015 and 2018, changes affecting carbon prices, output-based allocations of emissions credits, and which facilities are covered by the policies.  These combinations imply changes both across and within facility types, and across and within portfolios held by major players in the market.  We show that even after adjusting for actual costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.

The Alberta wholesale power market provides an excellent laboratory for study because it is small, isolated, and the market is settled on a real-time, energy-only basis so there are limited confounding factors. Alberta's market is winter-peaking, with peak 2018 load of just under 12,000 MW.  It has only three small interconnections to British Columbia, Montana, and Saskatchewan.\footnote{The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan intertie has a path rating of 150 MW.} This means that Alberta's market will be more affected by Alberta's policies than would be the case if traded power had a more important market share. We do face a challenge in that Alberta was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies that led to record-low prices in the market. Alberta's generation mix is also somewhat unique, with a significant portion of internal load served by combined heat and power (or cogeneration) plants associated with industrial facilities, primarily located in the oil sands.

We compile data describing market offer behaviour and a wide range of relevant covariates.  We measure and describe facility level hourly offers of power into the market over nearly 10 full years.  We combine these data with plant characteristics compiled from regulatory data, emissions data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data include hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.  We combine these data with observed carbon policy costs at the facility level.

In order to isolate the impact of the changes in policies on offer behaviour, we develop a empirical strategy based on power portfolios.  For each hour of each day, we create synthetic power portfolios either by plant type or by which entity holds offer control on the units in question.  We are then able to examine, for example, how changes in policies affected the offer of power across all coal- or gas-fired plants in the province, or across all facilities owned by particular actors in the market.  We normalize these portfolios by percentile, so we are able to compare behaviour across portfolios of different sizes without loss of generality.

\section{Alberta's GHG Policy Changes}

Alberta has had carbon pricing in place since the Specified Gas Emitters Regulation (SGER) took effect on July 1, 2007. That regulation, the first industrial carbon price in North America, implemented a price of \$15/tonne, and allocated emissions credits to covered facilities at a rate equal to 88\% of the facility's historic (2003-2005) emissions intensity or, for new facilities, a rate equal to 88\% of their average year 3 emissions intensity. This system remained in place until June, 2015 when the newly-elected government of Premier Rachel Notley introduced changes to the existing regulation which increased the price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for emissions allocation to 85\% and 80\% respectively for 2016 and 2017.  Combined, these changes implied at increase in the average cost of carbon in each of the years 2016 and 2017.

In addition to carbon pricing on industrial emissions, the SGER also included an offset protocol which provided emissions credits for deemed emissions reductions due to certain activities.  In the case of combined heat and power plants, the deemed emissions reductions where equivalent to an allocation of 0.418t/MWh for net-to-grid electricity.  New renewable power facilities were also eligible for offset credits.

The government also adopted, in November of 2015, a more comprehensive change to GHG emissions policies. Two changes in this iteration of policies affected power markets.  Most importantly, the \textit{Carbon Competitiveness Incentive Regulation (CCIR)} replaced the SGER and these regulations levelized the output-based allocation of emissions credits across all power generators at 0.37t/MWh, the emissions-intensity of the best-in-class combined cycle natural gas generation facility.  This implied that coal producers saw a steep increase in their average costs of carbon, while impacts on gas power plants varied depending on the heat rates of the facility. Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease under the CCIR.  The second important change was that an economy-wide carbon price was introduced for facilities not covered under the CCIR - those without historic emissions in any previous year greater than 100,000 tonnes. These facilities did not, by default, receive output-based allocations of emissions credits to offset the cost of the carbon price, so their average costs of carbon could be much higher than their larger competitors.  An opt-in provision allowing these smaller firms to be covered under the CCIR was available, should facilities wish to undertake the more comprehensive emissions reporting.

COVERED FACILITIES TABLE

\section{Data}

The majority of our data for this paper comes from the Alberta Electricity System Operator (AESO). The AESO provides, with a 60 day lag, offers made by power plants into the pool on an hourly basis. Facilities offer their power in up to 7 increasing price blocks, with a price cap of \$1000/MWh. Blocks may be either flexible or not, and non-flexible blocks will only be dispatched when demand allows the entire block to be used for the hour. Plants are dispatched in merit order, with the lowest price blocks dispatched first. We have hourly merit order and dispath data from September 1, 2009 through February 28, 2020 where we cut off the data to avoid impacts of the COVID-19 pandemic.\footnote{For a discussion of Alberta market responses to the COVID-19 pandemic, see Leach et al. (2020)} A subset of the data, from 2015 onward, allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.

We also use three other AESO data sets to build our analysis sample. First, the AESO issues hourly price and load data including lagged load and price forecasts which we merge with merit order data. Next, we also included combined intertie capability rating data which allow us to account for hours of limited import or export capability. Finally, and most importantly, we used metered volumes data at the facility level to incorporate renewable generation into the merit order. The AESO currently treats non-dispatchable renewables (wind and solar power in our case) as negative load, but lists them in the merit order data as \$0 offers at full nameplate capacity. Prior to 2015, non-dispatchable renewables were not included at all in the merit order. We correct this omission using hourly metered net-to-grid volumes for each facility-hour pair in our data set, which we use to create \$0 offers which we merge into the merit order. This accurately reflects how the Alberta grid operator treats non-dispatchable renewables - generators do not bear any quantity risk and simply act as price takers.

We supplement these data with a variety of other information.  Most importantly, we add facility-level information including fuel source, ownership, nameplate capacity, heat rate, and estimated emissions-intensity based on heat rates and fuel sources.  ADD ON-LINE APPENDIX TABLE OF PLANT INFO. We also use compliance data from the Alberta Specified Gas Emitters Regulations and the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for facilities.  Since these data are both reported at the facility level, not by units or blocks, we apply these measurements in the same way in our data.  As such, we do not allow emissions intensity to vary with the intensity of use of a particular unit in a particular hour nor do we account for any ramping rates of units.

We add weather data from Environment Canada weather stations in Edmonton, Fort McMurray and Calgary, the three major demand centers in Alberta.  We take the average of two measurements in each area in order to maximize the number of hours we can cover with weather data.  With this approach, we lose only X hours without at least one observation in each of the three areas.

Finally, we merge daily natural gas prices for the Nova Inventory Transfer (NIT) hub.  We access daily spot prices from NRGStream.

All data save the natural gas prices which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK. For the publicly available code, we substitute Henry Hub gas prices as a placeholder, to which we apply the average discount observed between Henry Hub and NIT to adjust the values so that regression results may be more closely replicated.

WE MIGHT BE ABLE TO GET PERMISSION TO INCLUDE THE NGX DATA.

\section{Alberta's Power Market}

Alberta's wholesale market is a single price, energy-only market. There is no day-ahead energy market, but there is a separate ancillary services market which we do not model in this paper. Alberta's record internal load, reached on January 11, 2018, is 11,697 MW, so Alberta's is a relatively small power market.  Alberta has seen significant growth in average and peak loads, with X cumulative annual average load growth rate during our sample period, although the 2016 through 2020 period has seen much less growth in load. Alberta is also a winter-peaking system which differentiates it from most North American power grids. Finally,
consumption in Alberta is relatively stable due to the large industrial base.

\begin{figure}[!h]%%
	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_time.png}
\label{fig:AB load}
\vspace{-0.75cm}	\caption{Alberta internal load, average and range.  Source: Alberta Electric System Operator data.}
\end{figure}

Our sample period covers ten years with no major regulatory changes in the power market design. Alberta briefly considered an addition of a capacity market, but this was never implemented. However, market conditions have varied over the sample period. There are three distinct periods during our sample - relatively high load growth and tight market conditions from 2009 through 2013, followed by sharply constrained growth and high reserve margins after 2014 until 2018, followed by a return to tighter market conditions after the spring of 2018. There was also a major short-term market event with the wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load offline. Some of these changes market coincide with policy changes in our experiment - in particular, the changes to the Specified Gas Emitters regulations introduced in June of 2015 which took effect in January of 2016 and 2017 respectively coincide with a period of over-supply, while the introduction of the CCIR in 2018 occurred during a period of tightening reserve margins and lingering uncertainty over market structure. From 2018 through pre-COVID 2020, both GHG policy and the wholesale electricity market were tightening, and so we need to be careful to disentangle these impacts.

\begin{figure}[!h]%%
	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices_2000_2020.png}
\label{fig:prices}
\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours.  Source: Alberta Electric System Operator data.}
\end{figure}

The generation mix in Alberta's power market is dominated by fossil fuels although, as shown in Figure \ref{fig:gen_mix}, the dominant fossil fuel has change from coal to natural gas. Within natural gas generation, the mix of plant types is also relevant to our study. X\% are combined cycle plants, y\% are simple-cycle plants, and Z\% are combined heat and power facilities which tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated relay on them for process heat. As will be discussed below, any net-to-grid power from these facilities is primarily offered into the market at a \$0 offer and accepts market price.

\begin{figure}[!h]%%
	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel.png}
\label{fig:gen_mix}
\vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: Alberta Electric System Operator data.}
\end{figure}

\section{Estimation}

\subsection{Market Merit Order}

\subsection{Synthetic Plants}

\subsection{Synthetic Portfolios}

\section{Results and Discussion}

\section{Analysis and Results}

\section{Conclusion}

\bibliographystyle{chicago}
\bibliography{../env}

\end{document}

\section{Introduction}
We examine the offer behaviour of competitive electricity suppliers in response to carbon prices. We exploit a unique set of significant policy variations over time and across facilities to measure carbon pricing pass-through to electricity prices. Beginning in 2007, the province of Alberta, Canada, instituted carbon pricing which affected generators in its energy-only wholesale electricity market. Changes in provincial governments in 2015 and 2019 led to the emissions pricing scheme being changed 3 times in 4 years, with the first 3 changes all announced during the 2015 calendar year, and a subsequent change announced in 2018. These changes were each implemented in such a way as to have firm- and facility-type-specific impacts which generates variation that we can exploit to separately characterize the role of carbon pricing and output-based allocations (OBAs) on pricing behaviour.

Compare to Mar and other literature.

First, in June 2015, it was announced that the carbon price would increase to \$20 per tonne and that output-based allocations would decrease beginning in 2016. The price was also set to increase further, to \$30 per tonne, in 2017.  Later in 2015, following the recommendations of an expert panel, the government announced that it would alter the formula by which output-based allocations were provided to one which was consistent across all electricity generators beginning in 2018. This change was material, with coal plant allocations reduced by more than 0.5t/MWh in many cases, leading to changes in average costs of output of more than \$15/MWh for some facilities.  This combination of policy changes implies that the effective marginal cost of emissions and the average emissions cost per unit output both changed significantly, with changes having different impacts on generators within and across fuel types.  We show the effect of the policy changes on average revenue by fuel type in Figure 1.

\begin{figure}[!h]%
	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture_ctax.png}
\label{fig:masnadi_1}
\vspace{-0.75cm}	\caption{Impact of changes in carbon pricing on plant revenues by type}
\end{figure}

 Fossil-fuel-fired generators were charged a carbon price of \$15/tonne on emissions net of allocations of emissions credits. Emissions credits were allocated per unit output at rates based on facility-specific emissions-intensity histories. Renewable generators were also able to qualify for offset credits which provided an alternative revenue stream.

\noindent

\newpage \noindent We use data including hourly electricity prices and loads, plant availability, as well as offer functions at the plant level, historic emissions intensities, and emissions policy compliance records to assess the degree to which carbon prices have been passed through to wholesale electricity prices. We also assess the degree to which changes in the composition of the market including changes to offer control and firm concentration, as well as changes due to facility exit, have altered the pass-through of carbon prices.  Our work builds largely on the analysis of electricity price pass-through of \citet{fabra_reguant}, but also draws on work by \citet{hintermann17} on market power, \citet{hintermann15} and \citet{hintermann16} on impacts of carbon pricing in the EU-ETS, \citet{nazifi16} on pass-through in Australia's electricity market, and \citet{woo_et_al} on pass-through in California's market.

\noindent Our preliminary analysis suggests that there has been both a significant effect on emissions (see Figure \ref{fig:monthly_ghg}) and on the composition of the market (see Figure \ref{fig:monthly_gen}) with significant reduction in the generation from coal plants and significant increase in the hourly daily cycling of coal plants as the change in carbon prices push coal plants higher in the merit order relative to gas generation, a dynamic examined in \citet{cullen_mansur}.

\begin{figure}[p]%
	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../images/monthly_ghg_mwh.png}
\vspace{-0.75cm}	\caption{Impact of changes in carbon pricing on GHG emissions}\label{fig:monthly_ghg}
\end{figure}

\begin{figure}[p]%
	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../images/gen_ghg_price.png}
\vspace{-.5cm}	\caption{Impact of changes in carbon pricing on market share by fuel. Combined heat and power (COGEN), simple-cycle gas turbines (SCGT) and natural gas combined cycle (NGCC) facilities are each natural-gas-fired.}\label{fig:monthly_gen}
\end{figure}

%\begin{figure}[!h]%
	%\centering \vspace{-.5cm} \includegraphics[width=6.5in]{CJE/masnadi_fig_1.png} %\label{fig:masnadi_1}
%	\caption{Figure 1 from Masnadi et al. (2018)}
%	\vspace{-.5cm}
%\end{figure}
\newpage
\vspace{10cm}

### 2021-06-04 ELS_script.R

@@ -1,0 +1,378 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+#base templates
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+
+#FORECASTS AND OTHER AESO SCRIPTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+
+
+#Load MERITS
+update<-0
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+
+
+update_forecasts()
+
+load("forecast_data.Rdata")
+
+ #Load AS MERIT 
+load(file="merit_AS.RData")
+if(update!=0){
+  merit_AS_data<-rbind(merit_AS_data,update_AS_merit(merit_AS_data))
+  save(merit_AS_data, file="merit_AS.RData")  
+}
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+
+#open AESO workbook
+tightest_hours<-function(){
+  data_list<-list()
+  #read excel workbook
+  for(year in seq(2013,2017)){
+  hours<-read_excel("HistoricPerformance.xlsx", sheet = as.character(year),n_max=250)
+  hours<- hours %>% clean_names() %>% select(local_date,capacity_year,calendar_year,supply_cushion_mw,rank)
+  data_list[[year]]<-hours
+  }
+  hours<-data.frame(do.call(rbind,data_list))
+  hours
+}
+
+
+augment_data<-function(merit_sent){
+  
+  #testing below this line
+  #merit_sent<-merit_filter
+  #bring in plant data
[…]

### 2021-06-04 ELS_script_old.R

@@ -1,0 +1,374 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/Alberta Power Data")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/Alberta Power Data")
+print(getwd())
+
+#base templates
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+
+#FORECASTS AND OTHER AESO SCRIPTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+
+
+#Load MERITS
+update<-0
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+
+
+update_forecasts()
+
+load("forecast_data.Rdata")
+
+ #Load AS MERIT 
+load(file="merit_AS.RData")
+if(update!=0){
+  merit_AS_data<-rbind(merit_AS_data,update_AS_merit(merit_AS_data))
+  save(merit_AS_data, file="merit_AS.RData")  
+}
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+
+#open AESO workbook
+tightest_hours<-function(){
+  data_list<-list()
+  #read excel workbook
+  for(year in seq(2013,2017)){
+  hours<-read_excel("HistoricPerformance.xlsx", sheet = as.character(year),n_max=250)
+  hours<- hours %>% clean_names() %>% select(local_date,capacity_year,calendar_year,supply_cushion_mw,rank)
+  data_list[[year]]<-hours
+  }
+  hours<-data.frame(do.call(rbind,data_list))
+  hours
+}
+
+
+augment_data<-function(merit_sent){
+  
+  #testing below this line
+  #merit_sent<-merit_filter
+  #bring in plant data
[…]

### 2021-06-04 Steven_AS_Pricing.R

@@ -1,0 +1,219 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+
+#PC ***Change user before running***
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/sdegroot/Google Drive/alberta_power")
+print(getwd())
+
+#base templates
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+
+#FORECASTS AND OTHER AESO SCRIPTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+
+
+#Load MERITS
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+
+
+
+ #Load AS MERIT 
+load(file="merit_AS.RData")
+if(update!=0){
+  merit_AS_data<-rbind(merit_AS_data,update_AS_merit(merit_AS_data))
+  save(merit_AS_data, file="merit_AS.RData")  
+}
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+#open AESO workbook
+tightest_hours<-function(){
+  data_list<-list()
+  #read excel workbook
+  for(year in seq(2013,2017)){
+  hours<-read_excel("HistoricPerformance.xlsx", sheet = as.character(year),n_max=250)
+  hours<- hours %>% clean_names() %>% select(local_date,capacity_year,calendar_year,supply_cushion_mw,rank)
+  data_list[[year]]<-hours
+  }
+  hours<-data.frame(do.call(rbind,data_list))
+  hours
+}
+
+
+augment_data<-function(merit_sent){
+  
+  #testing below this line
+  #merit_sent<-merit_filter
+  #bring in plant data
[…]

### 2021-06-04 Steven_aeso_scrapes.R

@@ -1,0 +1,618 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/sdegroot/Google Drive/alberta_power")
+print(getwd())
+source("../andrew_base.R")
+
+library(stringi)
+
+ 
+#aeso site scraper
+
+ get_metered_volumes_report <- function(start_date, end_date) {
+   
+   start_date <- as.Date(start_date)
+   end_date <- as.Date(start_date)
+   
+   GET(
+     url = "http://ets.aeso.ca/ets_web/ip/Market/Reports/PublicSummaryAllReportServlet",
+     query = list(
+       beginDate = format(start_date, "%m%d%Y"),
+       endDate = format(end_date, "%m%d%Y"),
+       contentType = "csv"
+     )
+   ) -> res
+   
+   stop_for_status(res)
+   
+   # Neither the CSV nor HTML output is all that great but the CSV
+   # can be made to work with (IMO) less effort than the HTML. You may
+   # need to do some extra checks for data format (for either CSV or
+   # HTML), though, in "production" mode.
+   
+   # From what I saw in the output, you likely need to modify 
+   # this attempt at munging since the "hours" seem off, but you
+   # at least now have the data.
+   
+   test<- content(res, as="text") %>% 
+     stri_split_lines() %>% 
+     flatten_chr()
+   #headers<-paste(c(paste(test[8:9], collapse=",")), collapse="\n")
+   #headers<- gsub("\"-\",", "", headers)
+   data<-gsub("\"-\",", "",paste(c(paste(test[8:9], collapse=","), test[13:length(test)]), collapse="\n"))
+   clean_data<-read.csv(text=data,header = TRUE, stringsAsFactors=FALSE)
+   clean_data<-janitor::clean_names(clean_data)%>%   tbl_df()
+   clean_data$date<-start_date
+   
+   #clean_data$he<-hour(clean_data$date)
+   #clean_data$time<-clean_data$date
+   #clean_data$date<-as.Date(clean_data$time)
[…]

### 2021-06-04 aeso_data_cpp.R

@@ -1,0 +1,779 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+source("../andrew_base.R")
+
+library(grid)
+library(gridExtra)
+library(ggpubr)
+#library(gganimate)
+library(timeDate)
+library(broom)
+library(ggthemes)
+library(directlabels)
+
+library(directlabels)
+blakes_blue<-"#4477AA"
+
+
+graph_grey<-"grey20"
+graph_black<-"black"
+
+
+blake_theme<-function(){
+  theme_hc(20)+
+    theme(plot.subtitle = element_text(color="grey10",size=rel(.7)),
+          plot.title = element_text(face="bold"),
+          plot.caption = element_text(color="grey50",size=rel(.5)),
+          legend.title = element_text(color="grey10",size=rel(.5)),
+          legend.text = element_text(color="grey10",size=rel(.5)),
+          axis.title = element_text(size=rel(.8)),
+          axis.ticks = element_blank(),
+          panel.spacing = unit(2,"lines"),
+          legend.position = "none",
+          plot.margin = margin(t = .5, r = 1, b = .25, l = 1,unit= "cm"),
+          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))
+          
+    )
+}
+
+#load(file="metered_vols_data.Rdata")
+load(file="gen.Rdata") 
+
+load(file="forecast_data.Rdata") 
+forecast_data <- forecast_data %>% filter (he!="02*")
+
+
+
+
+#direct label graph with AESO data
+
+#align variable names
+df2 <- gen_data %>% 
+  select(time=Time,vol=gen,ID,AESO_Name,Plant_Type,Plant_Fuel) %>%
+  filter(Plant_Type %in% c("COAL","COGEN","NGCC","WIND","SCGT","TRADE","HYDRO","SOLAR","OTHER"))%>%
+  filter(year(time)>=2010) %>% left_join(forecast_data) %>% filter(!is.na(date))%>%
[…]

### 2021-06-04 aeso_data_proc.R

@@ -1,0 +1,1359 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+library(grid)
+library(gridExtra)
+library(ggpubr)
+library(gganimate)
+library(timeDate)
+
+
+#seasons
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+#FORECASTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+
+
+#METERED VOLUMES
+
+update_forecasts()
+#re-load
+load("forecast_data.Rdata")
+
+load(file="metered_vols_data.Rdata" ) 
+all_vols<-update_vols(all_vols)
+
+save(all_vols, file="metered_vols_data.Rdata" ) 
+
+
+
+
+renew_vols<-all_vols %>% filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,vol)%>%
+  rename("renew_gen"="vol")
+save(renew_vols, file="renew_vols.Rdata" ) 
+
+
+
+
+
+
+latest_vols<-all_vols %>% filter(year(date)>=2018)
+brooks_vols<-latest_vols%>% filter(asset_id=="BSC1")
+brooks_vols$time<-as.POSIXct(paste(brooks_vols$date," ",as.numeric(brooks_vols$hour),":00",sep=""),format="%Y-%m-%d %H:%M")
+
+
+#FORECASTS
+
+
+
+#forecast_data$time<-as.POSIXct(paste(forecast_data$date," ",as.numeric(forecast_data$he),":00",sep=""),format="%Y-%m-%d %H:%M")
+filename<-paste("forecast_data",".RData",sep = "")
+update_forecasts()
+load(filename)
+forecast_data<-assign_date_time_days(forecast_data)
+forecast_data<-assign_peaks(forecast_data)
+
+
+
+#calculate peak prices
+
+peak_data<-forecast_data %>% group_by(year,month,on_peak) %>% summarize(pool_price=mean(actual_posted_pool_price,na.rm = T),
[…]

### 2021-06-04 aeso_scrapes.R

@@ -36,7 +36,7 @@ library(stringi)
    #headers<- gsub("\"-\",", "", headers)
    data<-gsub("\"-\",", "",paste(c(paste(test[8:9], collapse=","), test[13:length(test)]), collapse="\n"))
    clean_data<-read.csv(text=data,header = TRUE, stringsAsFactors=FALSE)
-   clean_data<-janitor::clean_names(clean_data)%>%   tbl_df()
+   clean_data<-janitor::clean_names(clean_data)%>%   as_tibble()
    clean_data$date<-start_date
    
    #clean_data$he<-hour(clean_data$date)
@@ -59,7 +59,7 @@ library(stringi)
  }
  
 get_all_data<-function() {
-  years<-seq(2010,2019)
+  years<-seq(2020,2020)
   for(year_id in years){
    print(paste("Starting year ",year_id))
    days<-seq.Date(as.Date(paste(year_id,"-01-01",sep="")),as.Date(paste(year_id,"-12-31",sep="")),by="1 day")
@@ -81,44 +81,55 @@ get_all_data<-function() {
    #write.xlsx(data_store, file = filename, colNames = TRUE, borders = "columns") 
   }
 }
+#get_all_data()
+
+#rebuild the data
+build_all<-function(){
+years<-seq(2004,2020)
+data_list<-list()
+i<-1
+load("forecast_data.Rdata")
+for(year_id in years){
+  filename<-paste("measured_vols_",year_id,".RData",sep = "")
+  print(filename)
+  load(file= filename)
+  #Need to process time and dates to have date,he,hour,dest
+  data_list[[i]]<-process_data(data_store)
+  i<-i+1
+}
+all_vols<-data.frame(do.call(rbind,data_list))
+save(all_vols, file="metered_vols_data.Rdata" ) 
+#save(all_vols, file="new_metered_vols_data.Rdata" ) 
+}
 
 
 #if you want to add the latest metered volumes data to the existing data set
 update_vols <- function(data_sent) {
+  #data_sent<-tail(all_vols,10000)
   max_date<-max(data_sent$date)
   data_sent<-data_sent %>% filter(date<max_date)
-  days<-seq.Date(max_date,Sys.Date()-days(3),by="1 day")
+  days<-seq.Date(max_date,Sys.Date()-days(4),by="1 day")
+  #shifted to 4 days lag here to correct potential for early day errors
[…]

### 2021-06-04 aeso_update.R

@@ -1,0 +1,101 @@
+#library(plyr)
+library(dplyr)
+library(openxlsx)
+library(reshape2)
+library(lubridate)
+
+library(tidyr)
+library(ggplot2)
+library(httr)
+
+library(viridis)
+library(scales)
+
+library(stringi)
+library(janitor)
+
+
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive")
+print(getwd())
+
+
+set_png<-function(file_sent,w_sent=1400,h_sent=750,res_sent=130){
+  #MAC
+  if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+    png(file=file_sent, width = w_sent, height = h_sent,res=res_sent)
+  #PC
+  if(R.version$platform ==  "x86_64-w64-mingw32")
+    png(file=file_sent, width = w_sent, height = h_sent,res=res_sent,type='cairo')
+}
+
+set_pdf<-function(file_sent,w_sent=1400,h_sent=750,res_sent=130){
+  #MAC
+  if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+    pdf(file=file_sent, width = w_sent, height = h_sent,res=res_sent)
+  #PC
+  if(R.version$platform ==  "x86_64-w64-mingw32")
+    pdf(file=file_sent, width = w_sent, height = h_sent,res=res_sent)
+}
+
+
+
+## Make breaks from a starting date at a given hour, occuring by interval,
+## length.out is days
+make_breaks <- function(strt, hour, interval="day", length.out=31) {
+  strt <- as.POSIXlt(strt - 60*60*24)  # start back one day
+  strt <- ISOdatetime(strt$year+1900L, strt$mon+1L, strt$mday, hour=hour, min=0, sec=0, tz="UTC")
+  seq.POSIXt(strt, strt+(1+length.out)*60*60*24, by=interval)
+}
+
+
+if(!exists("get_metered_volumes_report", mode="function")) source("aeso_scrapes.R")
+
+
+#METERED VOLUMES
+
+update_vols()
+
+#all_vols<-data_frame()
+years<-seq(2004,2018)
+for(year_id in years){
+  filename<-paste("measured_vols_",year_id,".RData",sep = "")
+  load(file= filename) 
[…]

### 2021-06-04 albert_data.R

@@ -1,0 +1,249 @@
+
+
+start_time<-Sys.time()
+#update merit order data
+
+update<-1
+load("all_merit.RData")  
+if(update!=1){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  #remove the 02* hours
+  #  merit_data<-merit_data%>%filter(he!="02*")
+  save(merit_data, file="all_merit.RData")  
+  
+}
+
+
+
+  singles<-seq(1,9)
+  for(hour in singles){
+    merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+    #  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+  }
+
+    merit_aug<-merit_data%>% filter(year(date)>=2015) %>% #testing just with 2019 data
+      mutate(import_export=case_when(
+    is.na(import_export) ~ "",
+    TRUE                      ~  import_export
+  ))%>%
+  clean_merit_trade() 
+  
+    #checked here - all imports are zero.
+    
+  
+  merit_data<-NULL #drop merit data out of memory
+  
+  print(paste("Loaded Merit Data. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
+  
+    
+  #2 Fix renewable generation so that blocks reflect actual generation, not capacity
+  
+  #load volumes
+  
+  #load(file="metered_vols_data.Rdata" ) 
+  
+  #isolate renewable volumes from metered volumes
+  renew_vols<- all_vols %>% #filter(year(date)>=2016) %>%
+    filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,vol)%>%
+    rename("renew_gen"="vol")
+    save(renew_vols, file="renew_vols.RData")  
+  load("renew_vols.RData")
+  
+  
+  merit_aug<-merit_aug%>%
+    left_join(renew_vols,by=c("date","he","asset_id")) %>%
+    mutate(
+      to=ifelse(!is.na(renew_gen),renew_gen,to), #use the renewable gen if you can
+      size=ifelse(!is.na(renew_gen),renew_gen,size), #use the renewable gen if you can
+      available_mw=ifelse(!is.na(renew_gen),renew_gen,available_mw), #use the renewable gen if you can
[…]

### 2021-06-04 generate_func.R

@@ -1,0 +1,3 @@
+generate_func<-function(x,y){
+  return(stepfun(x, y, f = as.numeric(0), ties = "ordered",right = FALSE))
+}

### 2021-06-04 get_season.R

@@ -1,0 +1,27 @@
+getSeason <- function(DATES) {
+  WS <- as.Date("2012-12-15", format = "%Y-%m-%d") # Winter Solstice
+  SE <- as.Date("2012-3-15",  format = "%Y-%m-%d") # Spring Equinox
+  SS <- as.Date("2012-6-15",  format = "%Y-%m-%d") # Summer Solstice
+  FE <- as.Date("2012-9-15",  format = "%Y-%m-%d") # Fall Equinox
+  
+  # Convert dates from any year to 2012 dates
+  d <- as.Date(strftime(DATES, format="2012-%m-%d"))
+  
+  ifelse (d >= WS | d < SE, "Winter",
+          ifelse (d >= SE & d < SS, "Spring",
+                  ifelse (d >= SS & d < FE, "Summer", "Fall")))
+}
+
+getSeasonPOSIXct <- function(DATES) {
+  WS <- as.POSIXct("2012-12-15", format = "%Y-%m-%d") # Winter Solstice
+  SE <- as.POSIXct("2012-3-15",  format = "%Y-%m-%d") # Spring Equinox
+  SS <- as.POSIXct("2012-6-15",  format = "%Y-%m-%d") # Summer Solstice
+  FE <- as.POSIXct("2012-9-15",  format = "%Y-%m-%d") # Fall Equinox
+  
+  # Convert dates from any year to 2012 dates
+  d <- as.POSIXct(strftime(DATES, format="2012-%m-%d"))
+  
+  ifelse (d >= WS | d < SE, "Winter",
+          ifelse (d >= SE & d < SS, "Spring",
+                  ifelse (d >= SS & d < FE, "Summer", "Fall")))
+}

### 2021-06-04 ghgs_merit.R

@@ -1,0 +1,159 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive")
+print(getwd())
+
+#FORECASTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+#seasons
+if(!exists("ajl_hourly", mode="function")) source("andrew_base.R")
+
+
+#Load MERITS
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+#augment_data<-function(merit_sent){
+  
+  #testing below this line
+  merit_sent<-merit_data %>% filter(year(date)==2018)
+  #bring in plant data
+  plant_data <- read.xlsx(xlsxFile = "AB_Plant_Info.xlsx", sheet = "Plant_info", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
+  colnames(plant_data)<-gsub("\\.", " ", colnames(plant_data)) 
+  plant_info<-data.frame(t(plant_data[(1:12),-1]))
+  #fix names
+  plant_info<-setNames(plant_info,t(plant_data[(1:12),1]))
+  plant_info$Capacity <- as.numeric(as.character(plant_info$Capacity))
+  plant_info<-plant_info %>%
+    mutate(NRG_Stream = ifelse(grepl("AB - H R Milner Hr Avg MW",NRG_Stream),"AB Milner Hr Avg MW",NRG_Stream))%>%
+    mutate(NRG_Stream = ifelse(grepl("AB - NPC1 Denis St Pierre Hr Avg MW",NRG_Stream),"AB - NPC1 Denis St  Pierre Hr Avg MW",NRG_Stream))  
+  plant_info<-arrange(plant_info,NRG_Stream)
+  
+  #bring in ghg data
[…]

### 2021-06-04 hourly_supply_cushion.R

@@ -1,0 +1,130 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+library(dplyr)
+library(openxlsx)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+
+#Load temprature data  
+temp_data <- read.table('C:/Users/safton/Documents/R/Merit Scripts/Temp Data.txt', header = TRUE)
+  temp_data$hourly <- ifelse (temp_data$AM_PM == "PM", temp_data$Hour + 12, temp_data$Hour + 0)
+
+  
+#Load price data for hourly price and AECO-C NG prices  
+price_data <- read.table('C:/Users/safton/Documents/R/Merit Scripts/historic_prices.txt', header = TRUE)
+AECO_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/ATCOC.txt', header = TRUE)
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_source.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+#average temp data for Ft Mac, Cgy, Edm 
+hour_temp <- aggregate(temp_data$Temp, by=list(temp_data$Month, temp_data$Day, temp_data$Year, temp_data$hourly ), FUN=mean)
[…]

### 2021-06-04 ieso_data_proc.R

@@ -1,0 +1,389 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+library(grid)
+library(gridExtra)
+#library(ggpubr)
+#library(gganimate)
+library(timeDate)
+library(broom)
+
+#seasons
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+#seasons
+#if(!exists("getSeason", mode="function")) source("get_season.R")
+
+
+
+#METERED VOLUMES AND CAPABILITIES
+
+get_ieso_reports<- function(start_date, end_date) {
+  
+  
+#current data
+  date_list<-as.character(seq.Date(ymd("2019-05-01"),Sys.Date(),"1 month"))
+  ieso_scrapes<-list()
+  count<-1
+  for(date_start in date_list[1:2]){
+    link<-paste("http://reports.ieso.ca/public/GenOutputCapabilityMonth/PUB_GenOutputCapabilityMonth_",format(ymd(date_start),"%Y"),format(ymd(date_start),'%m'),".csv",sep="")
+    loc_file=paste("ieso_",date_start,".csv",sep="")
+    download.file(link,destfile=loc_file,mode="wb")
+    ieso_scrapes[[count]]<-read_csv(loc_file,skip=3,col_names = TRUE) %>% clean_names()
+    count<-count+1
+  }
+  ieso_new<-bind_rows(ieso_scrapes)%>%pivot_longer(cols=-c(delivery_date,generator,fuel_type,measurement),values_drop_na = T)%>%
+    pivot_wider(id_cols = -c(measurement, value), names_from = measurement, values_from = value) %>% clean_names() %>%
+    mutate(name=gsub("hour_","",name),name=as.numeric(name))%>% rename(date=delivery_date,hour=name) 
+  #older data
+  
+  year_seq<- c(2010,2018)
+  links<-paste("http://ieso.ca/-/media/Files/IESO/Power-Data/data-directory/GOC-",year_seq,".xlsx?la=en",sep="")
[…]

### 2021-06-04 interties.R

@@ -1,0 +1,140 @@
+
+#aeso intertie availability reports
+
+#http://itc.aeso.ca/itc/public/queryHistoricalIntertieReport.do
+
+#<form name="historicalIntertieReportForm" method="POST" action="/itc/public/queryHistoricalIntertieReport.do">
+#  <input type="hidden" name="availableEffectiveDate" value="943279200000 1999-11-22 07:00:00 MST (1999-11-22 14:00:00 GMT)">
+#  <input type="hidden" name="availableExpiryDate" value="1582354800000 2020-02-22 00:00:00 MST (2020-02-22 07:00:00 GMT)">
+#<input type="text" name="startDate" value="2000-01-01" id="startDate">
+
+
+library(httr)
+library(tidyverse)
+library(lubridate)
+start_date<-ymd("2000-01-01")
+end_date<-ymd("2000-01-31")
+query = list(
+  startDate = format(start_date, "%Y-%m-%d"),
+  endDate = format(end_date, "%Y-%m-%d"),
+  fileFormat = "csv",
+  availableEffectiveDate="943279200000+1999-11-22+07%3A00%3A00+MST+%281999-11-22+14%3A00%3A00+GMT%29",
+  availableExpiryDate="1582354800000+2020-02-22+00%3A00%3A00+MST+%282020-02-22+07%3A00%3A00+GMT%29"
+  )
+
+res <- POST("http://itc.aeso.ca/itc/public/historicalIntertieReport.do", body = query,
+           write_disk("doc2.txt",overwrite = T))
+
+
+#http://itc.aeso.ca/itc/public/historicalIntertieReport.do?fileFormat=CSV&startDate=2000-01-01&endDate=2000-01-31
+
+library(stringi)
+res <- POST("http://itc.aeso.ca/itc/public/queryHistoricalIntertieReport.do", body = query)
+stop_for_status(res)
+test<- content(res, as="text") %>% 
+  stri_split_lines() %>% 
+  flatten_chr()
+
+
+
+
+
+
+
+test<-paste(test[3:length(test)], collapse="\n")
+merit_data<-read.csv(text=test,header = TRUE, stringsAsFactors=FALSE)
+clean_data<-janitor::clean_names(merit_data) %>% 
+  tbl_df()
+#key_firms<-firms()
+#if(!is.null(key_firms))
+#{
+#  clean_data$key_firm<-grepl(paste(key_firms, collapse="|"), clean_data$offer_control)
[…]

### 2021-06-04 keras_merit.R

@@ -1,0 +1,37 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+library(grid)
+library(gridExtra)
+library(ggpubr)
+library(gganimate)
+library(timeDate)
+
+
+#seasons
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+#FORECASTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+library(keras)     # Neural Networks
+library(tidyverse) # Data cleaning / Visualization
+library(knitr)     # Table printing
+library(rmarkdown) # Misc. output utilities 
+library(ggridges)  # Visualization
+
+
+
+
+allObservations %>% 
+  mutate(recording_length = map_int(data,nrow)) %>% 
+  ggplot(aes(x = recording_length, y = activityName)) +
+  geom_density_ridges(alpha = 0.8)

### 2021-06-04 merit_ghgs.R

@@ -1,0 +1,336 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+
+#setwd("G:/My Drive/alberta_power")
+
+
+
+source("../andrew_base.R")
+source("aeso_scrapes.R")
+source("merit_scripts.R")
+
+
+  
+
+#analysing changes in the merit order placement by emissions
+
+ghg_merit<-function(){
+  #merit_sent<-head(merit_2018,10000)
+  start_time<-Sys.time()
+  
+  load("all_merit.RData")
+  #merit_data<-merit_data %>% filter(date==ymd("2009-09-16"),he=="18")
+  
+  singles<-seq(1,9)
+  for(hour in singles){
+    merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+    #  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+  }
+
+    merit_aug<-merit_data%>%
+      mutate(import_export=case_when(
+    is.na(import_export) ~ "",
+    TRUE                      ~  import_export
+  ))%>%
+  clean_merit_trade() 
+    
+  
+  
+  merit_data<-NULL #drop merit data out of memory
+  
+  print(paste("Loaded Merit Data. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
+  
+    
+  #2 Fix renewable generation so that blocks reflect actual generation, not capacity
+  
+  #isolate renewable volumes from metered volumes
+  #renew_gen<- all_vols %>% filter(year(date)>=2016) %>%
+  #  filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,vol)%>%
+  #  rename("renew_gen"="vol")
+  load("renew_vols.RData")
+  
+
+  #now, we need to take renewables out of merits, and then add in info from renewable_vols
+  
+  #checked and the effective times are odd
+  #dates<-merit_aug %>% group_by(date,he) %>% summarize()
+  
+  renew_vols <- renew_vols %>% filter(date>=min(merit_aug$date))%>%
+    mutate(
+    import_export="",
+    block_number=0,
+    price=0,
+    from=0,
+    to=renew_gen,
[…]

### 2021-06-04 merit_scripts.R

@@ -58,9 +58,13 @@ ngx_data_read<-function(){
   #testing stuff
   file_name<-"nit_gas_spot_old.csv"
   ngx_data_old <- read.csv(file_name,blank.lines.skip=T,stringsAsFactors=F,header=T)
+  file_name<-"nit_gas_15_19.csv"
+  ngx_data_15_19 <- read.csv(file_name,blank.lines.skip=T,stringsAsFactors=F,header=T)
   file_name<-"nit_gas_spot.csv"
   ngx_data <- read.csv(file_name,blank.lines.skip=T,stringsAsFactors=F,header=T)
-  ngx_data<-rbind(ngx_data_old,ngx_data)
+  ngx_data<-rbind(ngx_data_old,ngx_data_15_19,ngx_data)
+  
+  
   names(ngx_data)<-c("date_time","low","WAvg", "High","Open","Settle", "Volume","Inst_Date","Drop") 
   ngx_data<-ngx_data %>% select(date_time,Settle,Inst_Date)
   ngx_data$date<-as.Date(parse_date_time(ngx_data$date_time,c("ymd HM","dmy HMS")))

### 2021-06-04 merits_2020_11_09.R

@@ -1,0 +1,368 @@
+source("aeso_scrapes.R")
+source("merit_scripts.R")
+
+start_time<-Sys.time()
+#update merit order data
+
+update<-0
+load("all_merit.RData")  
+if(update==1){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  #remove the 02* hours
+  #  merit_data<-merit_data%>%filter(he!="02*")
+  save(merit_data, file="all_merit.RData")  
+}
+
+
+
+  singles<-seq(1,9)
+  for(hour in singles){
+    merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+    #  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+  }
+
+    merit_aug<-merit_data%>% #filter(year(date)>=2015) %>% #testing just with 2019 data
+      mutate(import_export=case_when(
+    is.na(import_export) ~ "",
+    TRUE                      ~  import_export
+  ),
+  effective_date_time=mdy_hm(effective_date_time,tz="America/Denver")
+  )%>%
+  clean_merit_trade() %>% ungroup() %>% select(-merit)# %>% filter(import_export!="E")
+  #remove merit variable since we're going to re-calculate
+  #remove exports from the merit - they're demand, not supply
+    
+    #checked here - all imports are zero.
+    
+  
+  #merit_data<-NULL #drop merit data out of memory
+  
+  print(paste("Loaded Merit Data. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
+  
+    
+  
+  
+  
+  
+  
+  #2 Fix renewable generation so that blocks reflect actual generation, not capacity
+  
+  #load volumes
+  
+  new_renew<-1
+  if(new_renew==1){
+    load(file="metered_vols_data.Rdata" ) 
+    #isolate renewable volumes from metered volumes
+    renew_vols<- all_vols %>% 
+    filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,dispatched_mw=vol,pool_participant_id,effective_date_time=time)%>%
+    #switch so that we have same data format as merit data
+    mutate(import_export="",
+          block_number=1,
+          price=0,
+          from=0,
+          to=dispatched_mw,
+          size=dispatched_mw,
+          available_mw=dispatched_mw,
[…]

### 2021-06-04 pctl_testing.R

@@ -1,0 +1,39 @@
+#merit_day<-filter(merit_data,date==ymd("2018-04-18"))
+#save(merit_day,file="merit_day.RData")
+#load(file="merit_day.RData")
+
+#testing data set creation code
+start_time<-Sys.time()
+iter<-100
+for (i in c(1:iter)) {
+  merit_aug<-merit_day%>% select(date,he,asset_id,from,to,size,price,offer_sum)%>% group_by(date,he,asset_id)%>% arrange(date,he,asset_id,price) %>%
+    mutate(bid_capacity=max(to))%>%
+    summarize( #summarizing alone takes .13 seconds
+      #offer_control=as.character(offer_sum)[1],
+      #price=price[1],
+      #adding this line takes about .2 seconds
+      from=list(from/bid_capacity*100),price=list(price),
+      #step_func=list(bid_func(from/bid_capacity*100,price)),
+      #total_mw=mean(bid_capacity
+    )%>% group_by(date,he,asset_id) %>%
+    
+    
+    
+    mutate(step_func=list(bid_func(from[[1]],price[[1]])))%>% select(-from,-price) %>%
+    #bids=map(step_func,~.x(c(10,25,40,55,70,85,100))))  %>%
+    ungroup()
+  merit_aug$bids<-merit_aug$step_func[[2]](c(10,25,40,55,70,85,100))
+  
+}
+paste("Average elapsed time is",time_length(interval(start_time, Sys.time())/iter, "seconds"),"seconds")
+
+merit_aug$step_func[[1]](2)
+
+start_time<-Sys.time()
+for (i in c(1:2500)) {
+  test_func=list(bid_func(seq(0,100,15),seq(0,100,150)))
+  #print(i)
+}
+#start_time<-Sys.time()
+paste("Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds")
+

### 2021-06-04 power_data.R

@@ -1,0 +1,1279 @@
+library(plyr)
+require(dplyr)
+require(tidyr)
+require(reshape2)
+library(openxlsx)
+library(ggplot2)
+library(lubridate)
+library(zoo)
+library(ggmap)
+library(ggjoy)
+library(viridis)
+library(RColorBrewer)
+
+
+#create long data file
+
+setwd("C:/Users/aleach/Google Drive")
+
+load("gen.RData") ## which is here *equivalent* to
+
+gen_data$AIL<-as.numeric(gen_data$AIL)
+gen_data$Price<-as.numeric(gen_data$Price)
+gen_data$Demand<-as.numeric(gen_data$Demand)
+
+
+
+
+#merge capacity information
+
+
+sub_samp<-subset(gen_data, Time > as.Date("2017-05-13"))
+sub_samp<-subset(sub_samp, Time < as.Date("2017-05-20"))
+sub_samp<-subset(sub_samp, sub_samp$`Plant_Type`=="COGEN")
+
+p <- ggplot(sub_samp, aes(Time,gen)) 
+p + geom_area(aes(fill= ID), position = 'stack') + labs(y = "Hourly Generation (MW)")+ labs(x = "Date")
+
+
+p <- ggplot(sub_samp, aes(Time,gen)) + geom_line(aes(colour= Plant_Type), position = 'stack') + labs(y = "Hourly Generation (MW)")+ labs(x = "Date")
+p + facet_grid(. ~ Plant_Fuel)
+
+
+sub_samp<-subset(gen_data, Time > as.Date("2017-05-1"))
+sub_samp<-subset(sub_samp, Time < as.Date("2017-05-31"))
+sub_samp<-subset(sub_samp, sub_samp$`Plant_Type`=="WIND")
+
+p <- ggplot(sub_samp, aes(Time,gen)) 
+p + geom_area(aes(fill= ID), position = 'stack') + labs(y = "Hourly Generation (MW)")+ labs(x = "Date")
+
+
+sub_samp<-subset(gen_data, Time > as.Date("2017-05-1"))
+sub_samp<-subset(sub_samp, Time < as.Date("2017-05-31"))
+sub_samp<-subset(sub_samp, sub_samp$`Plant_Type`=="COAL")
+
+p <- ggplot(sub_samp, aes(Time,gen)) 
+#p + geom_point(position = 'stack')
[…]

### 2021-06-04 power_data_gen.R

@@ -1,0 +1,485 @@
+library(plyr)
+require(dplyr)
+require(tidyr)
+library(ggplot2)
+library(httr)
+library(jsonlite)
+library(curlconverter)
+library(curl)
+library(lubridate)
+library(viridis)
+library(scales)
+library(openxlsx)
+
+
+#create long data file
+
+setwd("C:/Users/aleach/Google Drive")
+
+png<-0 #set to one to output to files, anything else to output to screen.
+
+#curl -i -X POST --data "grant_type=client_credentials&client_id=P8EEG3rvRxyqWd_beUBCJA&client_secret=ki6M5J_hTtiXuMAcHHPmtw" https://api.neur.io/v1/oauth2/token
+
+sensor_id <- "0x0000C47F510354AE"
+
+#
+#req <- make_req(straighten())[[1]]
+
+#status check, neurio API
+neurio_api <- GET("https://api.neur.io/v1/status")
+neurio_r <- fromJSON(rawToChar(neurio_api$content))
+
+
+#command line for token
+#curl -i -X POST -H "Content-Type:application/x-www-form-urlencoded" -d "grant_type=client_credentials&client_id=P8EEG3rvRxyqWd_beUBCJA&client_secret=HWnpmXfqSymeG8Iqjp46Dg" "https://api.neur.io/v1/oauth2/token"
+get_token<-function(){
+  client_id <- c("P8EEG3rvRxyqWd_beUBCJA")
+  client_secret <- c("HWnpmXfqSymeG8Iqjp46Dg")
+  neurio.data <- data.frame(client_id,client_secret)
+  
+  
+  # Use authorisation code to get (temporary) access token
+  req <- POST(url = "https://api.neur.io/v1/oauth2/token", encode = "form",
+              body = list(
+                client_id = neurio.data$client_id,
+                client_secret = neurio.data$client_secret,
+                redirect_uri = "https://api.neur.io/v1/oauth2/token",
+                grant_type = "client_credentials"))
+  raise <- content(req, as="text")
+  #convert to list using RJSONIO
+  fromJSON(raise) -> token_data
+  return(token_data$access_token)
+  }
+
+
+simon_token<-function(){
+  client_id <- c("woE3Ank7Ty-zKp_QoT-1Yw")
[…]

### 2021-06-04 power_data_merge.R

@@ -1,0 +1,288 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+source("../andrew_base.R")
+
+
+
+#routine to load plant info from Excel
+
+get_plant_info <-function() {
+  #load plant data
+  #plant_data <- read.xlsx(xlsxFile = "2014_2017_gen.xlsx", sheet = 4, startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE) 
+  plant_data <- read.xlsx(xlsxFile = "AB_Plant_Info_New.xlsx", sheet = "Plant_info", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
+  
+  colnames(plant_data)<-gsub("\\.", " ", colnames(plant_data)) 
+  
+  plant_info<-data.frame(t(plant_data[(1:10),-1]))
+  #fix names
+  plant_info<-setNames(plant_info,t(plant_data[(1:10),1]))
+  plant_info$Capacity <- as.numeric(as.character(plant_info$Capacity))
+  
+  #Some of the names had errors or inconsistencies so fixed here
+  #plant_info$NRG_Stream<-revalue(plant_info$NRG_Stream, c("AB - H R Milner Hr Avg MW"="AB Milner Hr Avg MW"))
+  #plant_info$NRG_Stream<-gsub("AB - H R Milner Hr Avg MW","AB - Milner Hr Avg MW",plant_info$NRG_Stream)
+  #plant_info$NRG_Stream<-revalue(plant_info$NRG_Stream, c("AB - NPC1 Denis St Pierre Hr Avg MW"="AB - NPC1 Denis St  Pierre Hr Avg MW"))
+  #plant_info$NRG_Stream<-gsub("AB - NPC1 Denis St Pierre Hr Avg MW","AB - NPC1 Denis St  Pierre Hr Avg MW",plant_info$NRG_Stream)
+  
+  plant_info<-arrange(plant_info,NRG_Stream)
+  #put plant info in the Global Environment
+  assign("plant_info", plant_info, envir = .GlobalEnv)
+}
+
+all_gen_data <- function() {
+  #get all generation data
[…]

### 2021-06-04 power_graphs.R

@@ -1,0 +1,3507 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+source("../andrew_base.R")
+
+
+
+oba_type<-function(plant_sent,year_sent){
+  grid_avg<-0.65
+  #nested case_when
+  oba_12<-
+    case_when(
+      plant_sent == "HYDRO" ~ grid_avg,
+      plant_sent == "COAL" ~ 1.07125*0.88,
+      plant_sent == "COGEN" ~ 0.418,
+      plant_sent == "NGCC" ~ 0.42*.88,
+      plant_sent == "SCGT" ~ 0.6*.88,
+      plant_sent == "SOLAR" ~ grid_avg,
+      plant_sent == "WIND" ~ grid_avg,
+      TRUE                      ~  0 
+    )
+  oba_20<-
+    case_when(
+      plant_sent == "HYDRO" ~ grid_avg,
+      plant_sent == "COAL" ~ 1.07125*0.8,
+      plant_sent == "COGEN" ~ 0.418,
+      plant_sent == "NGCC" ~ 0.42*.8,
+      plant_sent == "SCGT" ~ 0.6*.8,
+      plant_sent == "SOLAR" ~ grid_avg,
+      plant_sent == "WIND" ~ grid_avg,
+      TRUE                      ~  0 
+    )
+  oba_15<-
+    case_when(
+      plant_sent == "HYDRO" ~ grid_avg,
+      plant_sent == "COAL" ~ 1.07125*0.85,
+      plant_sent == "COGEN" ~ 0.418,
+      plant_sent == "NGCC" ~ 0.42*.85,
+      plant_sent == "SCGT" ~ 0.6*.85,
+      plant_sent == "SOLAR" ~ grid_avg,
+      plant_sent == "WIND" ~ grid_avg,
+      TRUE                      ~  0 
+    )
+  case_when(
+    year_sent <=2015     ~ oba_12,
+    year_sent ==2016     ~ oba_15,
+    year_sent ==2017    ~ oba_20,
+    TRUE                      ~  0.37 
+  )
+}
+
+
+ctax_year<-function(year_sent){
+  case_when(
[…]

### 2021-06-04 power_graphs.bak.R

@@ -1,0 +1,3163 @@
+library(tidyverse)
+library(lubridate)
+require(reshape2)
+library(openxlsx)
+library(ggplot2)
+library(zoo)
+library(ggmap)
+library(ggjoy)
+library(viridis)
+library(RColorBrewer)
+library(scales) 
+library(gganimate)
+library(rvest)
+library(httr)
+library(stringi)
+
+
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive")
+print(getwd())
+
+
+set_png<-function(file_sent,w_sent=1400,h_sent=750,res_sent=130){
+  #MAC
+  if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+    png(file=file_sent, width = w_sent, height = h_sent,res=res_sent)
+  #PC
+  if(R.version$platform ==  "x86_64-w64-mingw32")
+    png(file=file_sent, width = w_sent, height = h_sent,res=res_sent,type='cairo')
+}
+
+
+## Make breaks from a starting date at a given hour, occuring by interval,
+## length.out is days
+make_breaks <- function(strt, hour, interval="day", length.out=31) {
+  strt <- as.POSIXlt(strt - 60*60*24)  # start back one day
+  strt <- ISOdatetime(strt$year+1900L, strt$mon+1L, strt$mday, hour=hour, min=0, sec=0, tz="UTC")
+  seq.POSIXt(strt, strt+(1+length.out)*60*60*24, by=interval)
+}
+
+
+
+
+
+load("gen.RData") ## which is here *equivalent* to
+
+
+gen_data$gen[is.na(gen_data$gen)] <- 0
+
+gen_data$AIL<-as.numeric(gen_data$AIL)
+gen_data$Price<-as.numeric(gen_data$Price)
+gen_data$Demand<-as.numeric(gen_data$Demand)
+
+
+
+sub_samp<-subset(gen_data, Time > as.Date("2010-01-1"))
+sub_samp<-subset(sub_samp, Time < as.Date("2017-10-15"))
+sub_samp<-subset(sub_samp, sub_samp$`Plant_Fuel`=="WIND")
+sub_samp<-subset(sub_samp, sub_samp$Capacity>0)
+sub_samp<-na.omit(sub_samp)
+sub_samp$Year<-year(sub_samp$Time)
[…]

### 2021-06-04 power_merge.R

@@ -1,0 +1,106 @@
+library(plyr)
+require(dplyr)
+require(tidyr)
+require(reshape2)
+library(openxlsx)
+library(ggplot2)
+library(lubridate)
+library(zoo)
+library(ggmap)
+library(ggjoy)
+library(viridis)
+library(RColorBrewer)
+library(data.table)
+
+
+#create long data file
+
+setwd("C:/Users/aleach/Google Drive")
+
+#df <- read.xlsx(xlsxFile = "simple_test2.xlsx", sheet = 1, startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
+
+time_data <- read.xlsx(xlsxFile = "2014_2017_gen.xlsx", sheet = 4, startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
+cnames<-time_data[1,]
+cnames[1,1]<-"Time"
+time_data <- setNames(time_data,cnames)
+colnames(time_data)<-gsub("\\.", " ", colnames(time_data)) 
+
+m<-time_data[-(1:8),]
+cnames<-time_data[1,]
+cnames[1,1]<-"Time"
+m <- setNames(m,cnames)
+
+m2 <- melt(m,id=c("Time","Demand","Price","AIL"),variable_name="Plant_ID") 
+#m2 <- melt(m,id=c("Time"),variable_name="Plant_ID") 
+m2 <- setNames(m2,c("Time","Demand","Price","AIL","ID","gen"))
+m2$gen <- as.numeric(as.character(m2$gen))
+m2$Time <- as.numeric(as.character(m2$Time))
+m2[["Time"]] <- as.POSIXct(m2[["Time"]] * (60*60*24), origin="1899-12-30", tz="GMT")
+m2$Year <- year(m2$Time)
+
+m2$Price <- as.numeric(as.character(m2$Price))
+m2<-arrange(m2,Time,ID)
+
+m3<-data.frame(t(time_data[(1:8),-1]))
+#m3<-setNames(m3,c("ID","AESO_ID","Capacity_MW","Plant_Type","Plant_Fuel","NRG_Stream_Code"))
+m3<-setNames(m3,t(time_data[(1:8),1]))
+m3$Capacity <- as.numeric(as.character(m3$Capacity))
+
+
+
+new_data <- read.xlsx(xlsxFile = "hourly17_test.xlsx", sheet = 1, startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE,colNames = TRUE)
+#take out periods in column names
[…]

### 2021-06-04 regression_bids.R

@@ -1,0 +1,178 @@
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+#Maps location of zip file so that excel documents can be written
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+#Needed library, make sure it is loaded
+library(dplyr)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+# add fuel type to merit data, will need to remap if working on a different computer
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_source.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$price)
+      df2[,date_count]<-step_fun(df2$MW)
+      names(df2)[date_count]<-as.character(df_test$date[1])
+    }
+    else{
+      df2[,date_count]<-df_test$price
[…]

### 2021-06-04 ryan_fixed.R

@@ -1,0 +1,144 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+
+#Enter desired year and plant asset_id in quotes  (I took out the character years - that's slowing you down)
+input_year <- 2017
+input_plant <- "GN3"
+
+#Select data by plant
+plant <- merit_data %>% filter(asset_id == input_plant & he != "02*")  #use filter here - way faster
+weekdays<- c("Sunday","Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
+
+#Determine position of hours in week
+plant<- plant %>% mutate(day=day(date), weekday=factor(wday(date),labels=weekdays),he= as.numeric(as.character(he)))%>%
+                  mutate(month=month(date),year=year(date),weekhour=(wday(date)-1)*24+he)
+
+#again, using mutate to add columns as opposed to subsetting and merging
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    #print(paste(date_count,date_id))
+    df_test<-data_sent %>% filter(date==date_id)
[…]

### 2021-06-04 ryan_fixed_adder.R

@@ -1,0 +1,215 @@
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+library(dplyr)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+# add fuel type to merit data
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/Completed Plants.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+
+#a new function to process plant/hour data into means by MW
+data_proc_adder<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    
+    #print(paste(date_count,date_id))
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$adder)
+      df2[,date_count]<-step_fun(df2$MW)
+      names(df2)[date_count]<-as.character(df_test$date[1])
+    }
+    else{
+      df2[,date_count]<-df_test$adder
[…]

### 2021-06-04 ryan_fixed_all.R

@@ -1,0 +1,133 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+
+#load and update merits
+update<-0
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+
+#Enter desired year and plant asset_id in quotes  (I took out the character years - that's slowing you down)
+#input_year <- 2017
+input_plant <- "GN3"
+
+#Select data by plant
+plant <- merit_data %>% filter(asset_id == input_plant & he != "02*")  #use filter here - way faster
+weekdays<- c("Sunday","Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
+
+#Determine position of hours in week
+plant<- plant %>% mutate(day=day(date), weekday=factor(wday(date),labels=weekdays),he= as.numeric(as.character(he)))%>%
+                  mutate(month=month(date),year=year(date),weekhour=(wday(date)-1)*24+he)
+
+#again, using mutate to add columns as opposed to subsetting and merging
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    #print(paste(date_count,date_id))
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$price)
[…]

### 2021-06-04 ryan_fixed_bigloop.R

@@ -1,0 +1,323 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+library(dplyr)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+# add fuel type to merit data
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_source.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    
+    #print(paste(date_count,date_id))
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$price)
+      df2[,date_count]<-step_fun(df2$MW)
+      names(df2)[date_count]<-as.character(df_test$date[1])
+    }
+    else{
+      df2[,date_count]<-df_test$price
[…]

### 2021-06-04 ryan_fixed_bigloop_months.R

@@ -1,0 +1,432 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+#Maps location of zip file so that excel documents can be written
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+#Needed library, make sure it is loaded
+library(dplyr)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+# add fuel type to merit data, will need to remap if working on a different computer
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_source.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$price)
+      df2[,date_count]<-step_fun(df2$MW)
+      names(df2)[date_count]<-as.character(df_test$date[1])
+    }
+    else{
[…]

### 2021-06-04 ryan_fixed_bigloop_new_cutoffs.R

@@ -1,0 +1,323 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+library(dplyr)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+# add fuel type to merit data
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_source.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    
+    #print(paste(date_count,date_id))
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$price)
+      df2[,date_count]<-step_fun(df2$MW)
+      names(df2)[date_count]<-as.character(df_test$date[1])
+    }
+    else{
+      df2[,date_count]<-df_test$price
[…]

### 2021-06-04 ryan_fixed_correction_2016.R

@@ -1,0 +1,471 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+#Maps location of zip file so that excel documents can be written
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+#Needed library, make sure it is loaded
+library(dplyr)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+# add fuel type to merit data, will need to remap if working on a different computer
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_source.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$price)
+      df2[,date_count]<-step_fun(df2$MW)
+      names(df2)[date_count]<-as.character(df_test$date[1])
+    }
+    else{
+      df2[,date_count]<-df_test$price
[…]

### 2021-06-04 ryan_fixed_correction_tests.R

@@ -1,0 +1,434 @@
+
+#set your directories here
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/safton/Documents/R/Merit Scripts")
+print(getwd())
+
+#Maps location of zip file so that excel documents can be written
+Sys.setenv("R_ZIPCMD" = "C:/Users/safton/Documents/R/Rtools/bin/zip.exe")
+
+#makes sure these are in the same directory
+source("andrew_base.R")
+source("aeso_scrapes.R")
+if(!exists("generate_func", mode="function")) source("generate_func.R")
+
+#Needed library, make sure it is loaded
+library(dplyr)
+
+#load and update merits
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+# add fuel type to merit data, will need to remap if working on a different computer
+asset_fuel <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_source.txt', header = TRUE)
+fuel_price <- read.table('C:/Users/safton/Documents/R/Merit Scripts/fuel_price.txt', header = TRUE)
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+}
+
+#add plant type data into merit data
+merit_data <- merge(merit_data, asset_fuel)
+
+
+#a new function to process plant/hour data into means by MW
+data_proc<-function(data_sent,max_cap){    
+  #send it a maximum capacity
+  df2<-data.frame(seq(0,max_cap))
+  names(df2)[1]<-"MW"
+  data_sent<-arrange(data_sent,date,block_number)
+  #start at 2 since your filling the second column
+  date_count<-2  
+  #fill columns for bids by MW
+  for(date_id in unique(data_sent$date)) {
+    df_test<-data_sent %>% filter(date==date_id)
+    if(nrow(df_test)>1){
+      step_fun<-generate_func(df_test$from[-1],df_test$price)
+      df2[,date_count]<-step_fun(df2$MW)
+      names(df2)[date_count]<-as.character(df_test$date[1])
+    }
+    else{
[…]

### 2021-06-04 stuart_stuff.R

@@ -1,0 +1,708 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+source("../andrew_base.R")
+
+library(stringi)
+
+
+library(grid)
+library(gridExtra)
+library(ggpubr)
+#library(gganimate)
+library(timeDate)
+library(broom)
+library(ggthemes)
+library(directlabels)
+
+library(directlabels)
+blakes_blue<-"#4477AA"
+
+blake_theme<-function(){
+  theme_hc(20)+
+    theme(plot.subtitle = element_text(color="grey10",size=rel(.7)),
+          plot.title = element_text(face="bold"),
+          plot.caption = element_text(color="grey50",size=rel(.5)),
+          legend.title = element_text(color="grey10",size=rel(.5)),
+          legend.text = element_text(color="grey10",size=rel(.5)),
+          axis.title = element_text(size=rel(.8)),
+          axis.ticks = element_blank(),
+          panel.spacing = unit(2,"lines"),
+          legend.position = "none",
+          plot.margin = margin(t = .5, r = 1, b = .25, l = 1,unit= "cm"),
+          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))
+          
+    )
+}
+
+
+#rebuild the data
+build_all<-function(){
+  years<-seq(2015,2020)
+  data_list<-list()
+  i<-1
+  load("forecast_data.Rdata")
+  #year_id<-2019
+  for(year_id in years){
+    filename<-paste("measured_vols_",year_id,".RData",sep = "")
+    print(filename)
+    load(file= filename) 
+    #Need to process time and dates to have date,he,hour,dest
+    data_list[[i]]<-data_store%>% filter(asset_id=="TER1")
+    i<-i+1
+  }
+  ter1_vols<-data.frame(do.call(rbind,data_list))
+  save(ter1_vols, file="ter1.Rdata" ) 
+  #save(all_vols, file="new_metered_vols_data.Rdata" ) 
+}
[…]

### 2021-06-04 summary_stats.R

@@ -1,0 +1,256 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+
+source("aeso_scrapes.R")
+source("merit_scripts.R")
+
+
+
+
+#analysing changes in the merit order placement by emissions
+
+data_summary<-function(){
+  #merit_sent<-head(merit_2018,10000)
+  start_time<-Sys.time()
+  
+  load("all_merit.RData")
+
+  singles<-seq(1,9)
+  for(hour in singles){
+    merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+    #  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+  }
+
+    merit_aug<-merit_data%>% filter(year(date)>=2012) %>%  #why only after 2012? Just sampling? Don't recall
+      mutate(import_export=case_when(
+    is.na(import_export) ~ "",
+    TRUE                      ~  import_export
+  ))%>%
+  clean_merit_trade() 
+  
+  merit_data<-NULL #drop merit data out of memory
+  
+  print(paste("Loaded Merit Data. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
+  
+    
+  #2 Fix renewable generation so that blocks reflect actual generation, not capacity
+  
+  #isolate renewable volumes from metered volumes
+  #renew_gen<- all_vols %>% filter(year(date)>=2016) %>%
+  #  filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,vol)%>%
+  #  rename("renew_gen"="vol")
+  load("renew_vols.RData")
+  
+  
+  merit_aug<-merit_aug%>%
+    left_join(renew_vols,by=c("date","he","asset_id")) %>%
+    mutate(
+      to=ifelse(!is.na(renew_gen),renew_gen,to), #use the renewable gen if you can
+      size=ifelse(!is.na(renew_gen),renew_gen,size), #use the renewable gen if you can
+      available_mw=ifelse(!is.na(renew_gen),renew_gen,available_mw), #use the renewable gen if you can
[…]

### 2021-06-04 synth_plants.R

@@ -14,49 +14,134 @@ source("merit_scripts.R")
 #constructing the synthetic plants based on plant type
 
 synth_plants<-function(){
-  #merit_sent<-head(merit_2018,10000)
-  start_time<-Sys.time()
   
-  load("all_merit.RData")
+  start_time<-Sys.time()
+  #update merit order data
+  
+  update<-0
+  load("all_merit.RData")  
+  if(update==1){
+    merit_data<-rbind(merit_data,update_merit(merit_data))
+    #remove the 02* hours
+    #  merit_data<-merit_data%>%filter(he!="02*")
+    save(merit_data, file="all_merit.RData")  
+  }
+  
 
+  
   singles<-seq(1,9)
   for(hour in singles){
     merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
     #  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
   }
 
-    merit_aug<-merit_data%>% filter(year(date)>=2012) %>%
-      mutate(import_export=case_when(
-    is.na(import_export) ~ "",
-    TRUE                      ~  import_export
-  ))%>%
-  clean_merit_trade() 
+  merit_aug<-merit_data%>% #filter(year(date)>=2015) %>% #testing just with 2019 data
+    mutate(import_export=case_when(
+      is.na(import_export) ~ "",
+      TRUE                      ~  import_export
+    ),
+    effective_date_time=mdy_hm(effective_date_time,tz="America/Denver")
+    )%>%
+    clean_merit_trade() %>% ungroup() %>% select(-merit) %>% filter(import_export!="E")
+  #remove merit variable since we're going to re-calculate
+  #remove exports from the merit - they're demand, not supply
   
-  merit_data<-NULL #drop merit data out of memory
+  #checked here - all imports are zero.
+  
+  
+  #merit_data<-NULL #drop merit data out of memory
   
   print(paste("Loaded Merit Data. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
   
-    
+  
+  
+  
+  
+  
+  
   #2 Fix renewable generation so that blocks reflect actual generation, not capacity
   
-  #isolate renewable volumes from metered volumes
-  #renew_gen<- all_vols %>% filter(year(date)>=2016) %>%
[…]

### 2021-06-04 tightest_hours.R

@@ -1,0 +1,282 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+#base templates
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+
+#FORECASTS AND OTHER AESO SCRIPTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+
+
+#Load MERITS
+update<-1
+load("all_merit.RData")  
+if(update!=0){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  save(merit_data, file="all_merit.RData")  
+}
+
+
+
+update_forecasts()
+
+load("forecast_data.Rdata")
+
+ #Load AS MERIT 
+load(file="merit_AS.RData")
+if(update!=0){
+  merit_AS_data<-rbind(merit_AS_data,update_AS_merit(merit_AS_data))
+  save(merit_AS_data, file="merit_AS.RData")  
+}
+
+#check to make sure singles are consistent
+singles<-seq(1,9)
+for(hour in singles){
+  merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
+  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
+}
+
+
+
+#open AESO workbook
+tightest_hours<-function(){
+  data_list<-list()
+  #read excel workbook
+  for(year in seq(2013,2017)){
+  hours<-read_excel("HistoricPerformance.xlsx", sheet = as.character(year),n_max=250)
+  hours<- hours %>% clean_names() %>% select(local_date,capacity_year,calendar_year,supply_cushion_mw,rank)
+  data_list[[year]]<-hours
+  }
+  hours<-data.frame(do.call(rbind,data_list))
+  hours
+}
+
+
+
+
+#open AESO workbook
+supply_cushion<-function(){
+    data_read<-read_excel("AESO-Supply-Cushion.xlsx",sheet="All Hours (5 Years)")
+  data_read
+}
+
+
+
+
+
+
+augment_data<-function(merit_sent){
+  
+  #testing below this line
[…]

### 2021-06-05 paper/cv_apr_2021.tex
paper files

\documentclass{article}[12pt]
\usepackage{tabularx}
\usepackage{amsfonts,amsmath}
\usepackage{verbatim} % enables you to "comment" out sections.

\emergencystretch=20pt
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% to indent
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\ind}{\hspace*{15pt}}
\newcommand{\negind}{\hspace*{-15pt}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \education{ name of school }{ dates attended }{ degrees & description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\education}[4]{\negind \begin{bf} #1 \end{bf} \hfill  #2

\hfill \parbox[b]{.93\textwidth}{ #3 }\\[-.95ex]

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \experience{ position }{ institution }{ employment period }{ description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\experience}[4]{\negind \begin{bf} #1, \end{bf} #2 \hfill  #3

\hfill \parbox[b]{.93\textwidth}{ #4 }\hfill\\[-.25ex]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \presentation{ title }{ date }{ type & conference & location }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\presentation}[3]{\negind ``#1" \hfill  #2
\smallskip
\hfill \parbox[b]{.93\textwidth}{ #3 }
\smallskip }

\newcommand{\conference}[3]{\negind #1 \hfill  #2

\smallskip

\hfill \parbox[b]{.93\textwidth}{ #3 }

\smallskip }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\studentthesis}[3]{\negind  #1 \hfill  #2
\\
\hfill \parbox[b]{.93\textwidth}{ #3 }

\smallskip }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\student}[2]{ \begin{it} #1 \end{it} \hfill  #2

\smallskip

\hfill }

\newcommand{\publication}[6]{\parbox[b]{.93\textwidth}{\negind #1. (#2) #3, \textit{#4} #5, pages #6.}
\hfill \\[-.65ex]}

\newcommand{\book}[6]{\parbox[b]{.93\textwidth}{\negind #1. (#2) ``#3," \textit{#4} #5, #6.}
\hfill \\[-.65ex]}

\newcommand{\forthcoming}[5]{\parbox[b]{.93\textwidth}{\negind #1. (#2) ``#3," \textit{#4}}
\\[-.65ex] \hfill }

\newcommand{\inprogress}[3]{\parbox[b]{.93\textwidth}{\negind #1. ``#2" #3  }
\\[-.65ex] \hfill}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The purpose of this command is twofold:
%%   (1) Its really just a list, with no frills.
%%   (2) However, this list will justify (line up)
%%       with the formatting of the above commands.
%% Hence the name.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\justlist}[1]{\negind #1 \smallskip}

\newcommand{\courses}[3]{\negind  #1  \hfill  #2

\smallskip

\hfill \parbox[b]{.93\textwidth}{ #3 }

\smallskip }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \award{ title }{ date }{ reason }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\award}[3]{\negind #1 \hfill  #2

\smallskip

\hfill \parbox[b]{.93\textwidth}{  #3 }

\medskip }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% to resize the paper -- don't change
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addtolength{\hoffset}{-2.65cm} \addtolength{\textwidth}{5cm}
\addtolength{\voffset}{-1cm} \addtolength{\textheight}{3cm}

\pagestyle{myheadings}

\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}
\newcommand{\setlinespacing}[1]%
           {\setlength{\baselineskip}{#1 \defbaselineskip}}
\newcommand{\doublespacing}{\setlength{\baselineskip}%
                           {2.0 \defbaselineskip}}
\newcommand{\singlespacing}{\setlength{\baselineskip}{\defbaselineskip}}
\newcommand{\onehalfspacing}{\setlength{\baselineskip}%
                           {.95 \defbaselineskip}}

\begin{document}
\thispagestyle{empty}
 \onehalfspacing
%\setcounter{page}{30}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%   PERSONAL INFORMATION   %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%   (This is where you change things.)   %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
\textbf{{\large ANDREW J. LEACH, Ph.D.}}\\\smallskip \textbf{{\large Curriculum
vitae}}

\end{center}

\bigskip

{\large\scshape\hrule\vspace{0.8\baselineskip}}

\noindent  Alberta School of Business\hfill andrew.leach@ualberta.ca
\smallskip

\noindent University of Alberta \hfill http://business.ualberta.ca/aleach/
\smallskip

\noindent 3-20D Business Building \hfill Tel: 780-492-8489
\smallskip

\noindent Edmonton, Alberta T6G 2R6 (Canada)\hfill Fax: 780-492-3325
\medskip

{\large\scshape\hrule\vspace{0.8\baselineskip}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                                                           %%
\medskip
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%   Current Position %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large AFFILIATIONS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \experience{ position }{ institution }{ employment period }{ description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\experience{Alberta School of Business}{University of Alberta}{2011--}{Associate Professor, with tenure} %\vspace{-2\baselineskip}

\experience{Smart Prosperity}{Research Network}{2018--}{Member}

\experience{CABREE}{Centre for Applied Business Research in Energy and the Environment}{2014--}{Academic Director} %\vspace{-2\baselineskip}

\experience{CIRANO}{Center for Interuniversity Research and Analysis on Organizations}{2005--}{Researcher} %\vspace{-2\baselineskip}

%\experience{CIRPE\'{E}}{Centre interuniversitaire sur le risque, les politiques \'{e}conomique et l'emploi}{March, 2005-present}{Researcher} %\vspace{-2\baselineskip}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

                                                           %%
\bigskip                                                            %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%   Current Position %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large PAST AFFILIATIONS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\bigskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \experience{ position }{ institution }{ employment period }{ description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\experience{Government of Alberta}{Department of Environment and Climate Change}{2015}{Chair, Climate Leadership Panel} %\vspace{-2\baselineskip}

\experience{Alberta School of Business}{University of Alberta}{2013--2015}{Enbridge Professor of Energy Policy} %\vspace{-2\baselineskip}

\experience{Environment Canada}{Strategic Policy Branch}{2012--2013}{Visiting Scholar} %\vspace{-2\baselineskip}

\experience{Sustainable Prosperity}{Research Network}{2008--2017}{Member}
%\vspace{-2\baselineskip}

\experience{Alberta School of Business}{University of Alberta}{2006-- 2011}{Assistant Professor} %\vspace{-2\baselineskip}

\experience{HEC Montr\'{e}al}{Institute of Applied Economics
}{2003--2007}{Assistant Professor} %\vspace{-2\baselineskip}

\experience{Queen's University}{Economics Department}{2002 -- 2003} {Teaching
Fellow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

                                                           %%
\bigskip                                                            %%

%% This says the header should be my name (and the page number)    %%%
\markright{ANDREW J. LEACH} \vspace{1cm}                                 %%%

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%   EDUCATION   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large EDUCATION}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\bigskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \education{ name of school }{ dates attended }{ degrees & description }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\education{University of Alberta}{2021} {Master of Laws}

\education{Queen's University}{2004} {Ph.D.(Economics)}

 \education{University of Guelph}{1999} {Master of Arts (Economics)}

 \education{University
of Guelph}{1998} {Bachelor of Science (Environmental Science)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% I like to start a new page around here, but you can move this. %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%   PROJECTS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\noindent {\bf \large RESEARCH PROJECTS} $_{\begin{tabular}{l}      %%
%\hline \hspace{351pt} \\                                            %%
%\end{tabular}}$                                                     %%
                                                                    %%
%\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \presentation{ title }{ date }{ additional information }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\presentation{Single-Variable Linear Equations with Integer
%Coefficients}{September 20xx -- Spring 20xx}
%{ Senior Thesis, Department of Mathematics, Undergraduate Institution. \\
%Advisor: Professor Matt Amatician }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\medskip                                                            %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%   PUBLICATIONS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent \textbf{{\large PUBLICATIONS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\inprogress{Nigel Bankes, Andrew Leach and Martin Olszynski (forthcoming)}{Case Comment: \textit{References re Greenhouse Gas Pollution Pricing Act}, 2021 SCC 11}{\textit{Energy Regulation Quarterly}}

\publication{Leach, Andrew}{2021}{Environmental Policy is Economic Policy:
Climate Change Policy and the General Trade and Commerce Power}{\textit{Ottawa Law Review}}{52(2)}{97-156}

\publication{Leach, Andrew and Nic Rivers and Blake Shaffer}{2020}{Canadian Electricity Markets during the COVID-19 Pandemic: An Initial Assessment}{\textit{Canadian Public Policy}}{46(S2)}{S145-S159}

\publication{Olszynski, Martin, Nigel Bankes, and Andrew Leach}{2020}{Case Comment: \textit{Reference re Greenhouse Gas Pollution Pricing Act}, 2020 ABCA 74}{\textit{Journal of Environmental Law and Practice}}{33(2)}{159-184}

\publication{Leach, Andrew and Eric M. Adams}{2020}{Seeing Double: Peace, Order, and Good Government, and the Impact of Federal Greenhouse Emissions Legislation on Provincial Jurisdiction}{\textit{Constitutional Forum}}{29 (1)}{1-20}

\publication{Bo\v skovi\'{c}, Branko and Andrew Leach}{2020}{Leave it in the ground? Oil sands development under carbon pricing}{\textit{Canadian Journal of Economics}}{53(2)}{526-562}

\publication{Heyes, Anthony, Andrew Leach and Charles F. Mason}{2018}{The Economics of the Canadian Oil Sands}{\textit{Review of Environmental Economics and Policy}}{12 (2)}{242-263}

\publication{Leach, Andrew}{2016}{Canada in a Low-Carbon World: Impacts on new and existing resources}{\textit{Canadian Public Policy}}{42 (S1)}{S18-S23}

\publication{Leach, Andrew}{2015}{Canadian Hydrocarbon Resources in an Era of Manufactured Energy”, Chapter in Roland Paris and Taylor Owen (eds) The World Won't Wait: What Canada needs to rethink its international policies}{University of Toronto Press}{2015}{240}

\publication{Leach, Andrew and Branko Boskovic}{2014}{Carbon cost will not stop oil-sands work}{\textit{Nature} Letters}{511}{page 534-534}

\publication{Leach, Andrew}{2012}{Alberta's Specified Gas Emitters
Regulation}{\textit{Canadian Tax Journal}}{60(4)}{881-898}

\publication{Chakravorty, Ujjayant, Andrew Leach, and Michel
Moreaux}{2012}{Cycles in nonrenewable resource prices with pollution and
learning-by-doing}{\textit{Journal of Economic Dynamics and
Control}}{36(10)}{1448-1461}

\publication{Laurent-Lucchetti, Jeremy and Andrew Leach}{2011}{Generational
welfare and induced innovation under climate change
policy}{\textit{Scandinavian Journal of Economics}}{113(4)}{904-936}

\publication{Leach, Andrew, Charles F. Mason and Klaas van't
Veld}{2011}{Co-optimization of oil production and carbon
sequestration}{\textit{Resource and Energy Economics}}{33}{893-912}

\publication{Leach, Andrew, Joseph Doucet and Trevor Nickel}{2011}{The
effectiveness of Alberta biofuels policy}{\textit{Energy Policy}}{39(7)}
{4007-4015}

%\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}
%TENURE
%\noindent\makebox[\linewidth]{\rule{\textwidth}{1pt}}

\publication{Chakravorty, Ujjayant, Andrew Leach, and Michel
Moreaux}{2011}{Would Hotelling Kill the Electric Car?}{\textit{Journal of
Environmental Economics and Management}}{61(3)}{281-296}

\book{Chakravorty, Ujjayant, Shelby Gerking and Andrew Leach}{2010}{State Tax
Policy and Oil Production: The Role of the Severance Tax and Credits for
Drilling Expenses}{US Energy Tax Policy,}{Metcalf, Gilbert (ed), Cambridge
University Press}{414 pages.}

\publication{Leach, Andrew}{2009}{The Welfare Implications of Climate Change
Policy}{Journal of Environmental Economics and Management}{57(2)}{151-165.}

\publication{Bahn, Olivier and Andrew Leach}{2008}{The Secondary Benefits of
Climate Change Mitigation: An overlapping generations approach}{Computational
Management Science}{5(3)}{233-257}

\publication{Leach, Andrew}{2007}{The Climate Change Learning Curve}{Journal of
Economic Dynamics and Control}{33}{1728-1752}

\publication{Clark, C. Robert and Andrew Leach}{2007}{The Potential for
Electricity Market Restructuring in Quebec}{Canadian Public Policy}
{33(1)}{1-20}

\publication{Leach, Andrew}{2006}{Regulation by Reputation}{Computers and
Operations Research}{33}{409-425}

\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%   PROJECTS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large CURRENT PROJECTS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\inprogress{Andrew Leach}{“Economic Policy is Environmental Policy”}{Accepted for publication in the \textit{Ottawa Law Review}}

\inprogress{Andrew Leach}{The No More Pipelines Act? Bill C-69’s impact on oil sands pipelines}{Submitted, \textit{Alberta Law Review}}

\inprogress{Andrew Leach}{The Constitutional Limits on Federal GHG Emissions Policies}{In preparation for submission to the \textit{Canadian Journal of Economics}}

\inprogress{Fenner L Stewart, Andrew Leach and Jennifer Winter}{A Federal Backstop to Provincial Carbon Pricing: Decarbonization in an Age of Uncooperative Federalism}{Manuscript}

\inprogress{Andrew Leach and Blake Shaffer}{Carbon Price Pass-Through in Alberta's Electricity Market}{In preparation for presentations}

\inprogress{Branko Boskovic, Andrew Leach and Charles F. Mason}{Nonrenewable resources, reserves, and carbon pricing}{In preparation for submission}

\inprogress{Andrew Leach, and Alain-Desir\'{e} Nimubona}{Technology search and environmental regulation.}{In preparation for re-submission}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%   PROJECTS   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent \textbf{{\large SELECTED PROFESSIONAL AND ADVISORY ACTIVITY}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\smallskip %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\inprogress{Leach, Andrew and Angela Adams, Stephanie Cairns, Linda Coady and Gordon Lambert}{Climate Leadership Panel, Report to the Minister}{Prepared for Alberta Environment and Climate Change, 2015.}

\inprogress{Leach, Andrew}{Royalty Review Scoping Document}{Prepared for Alberta Energy, 2015.}

\inprogress{Leach, Andrew}{Emerging oilsands technology and prospects for GHG
	emissions reductions}{Prepared for Environment Canada, 2012.}

\inprogress{Leach, Andrew}{An evaluation of `A Climate Change Plan for the
Purposes of the Kyoto Protocol Implementation Act (Bill C-288)'}{prepared for
the National Roundtable on the Environment and the Economy annually from
2007-2011.}

\inprogress{Leach, Andrew}{GHG Emissions Policies for Alberta: Current
strengths and weaknesses, key future threats, and a proposed policy solution
for the Province of Alberta}{Prepared for Alberta Environment, July, 2009.}

\inprogress{Leach, Andrew and Dave Sawyer}{An evaluation of the modeling and
information transmittal process used for the development of ``Alberta's 2008
Climate Change Strategy"'}{Prepared for the Office of the Auditor General of
Alberta, July, 2008.}

\inprogress{Leach, Andrew and C. Robert Clark}{The determinants of demand for
natural gas: a review of the literature}{Prepared for Gaz M\'{e}tropolitain,
July, 2006.}

\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Research and Teaching Grants %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{{\large MAJOR GRANTS}}

\medskip
{\large\scshape\hrule\vspace{0.8\baselineskip}} %%
                                                                    %%
\medskip                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% usage: \justlist{ first line }{ paragraph below }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\justlist{\begin{bf}Canada First Excellence Research Fund\end{bf}
	Future of Energy Systems Institute, University of Alberta, \$75,000,000 over 7 years, 2016.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Insight Development Grant 430-2014-0275, \$58,630 over 2 years, 2014.}

\justlist{\begin{bf}Sustainable Prosperity\end{bf} Project grant with Branko
Boskovic, \$8000, 2014.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Insight Development Grant 430-2012-0079, \$45,612 over 2 years, 2012.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Collaborator, Alberta Climate Dialogue (ABCD), Community-University Research
Alliances,\$1 million over 6 years, 2010.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Standard Research Grant, Grant Number 410-2009-1044, \$18600 over 3 years,
2009.}

\justlist{\begin{bf}Social Sciences and Humanities Research Council\end{bf}
Sustainable Prosperity Network Grant, \$1.8 million over 7 years, 2008.}

\justlist{\begin{bf}Fonds qu\'{e}becois de la recherche sur la soci\'{e}t\'{e}
et la culture (FQRSC)\end{bf} Team Grant, \$332 200 over 4 years, with G.
Dionne, R. Gagn\'{e}, P. Lanoie, P. Fran\c{c}ois, and B. Dostie, 2005.}

%During the course of the year, I have given a number of academic presentations %and invited lectures as follows:
%
%University of Ottawa, Economics of Environmental Policy Class Presentation, %July 23
%University of Ottawa, Economics of Environmental Policy Workshop, July 25
%Presentation to Ivey School of Business Students, UWO, October 3
%AIMCO Investor Conference, October 16
%Better Policy for a Better World: Why good ideas aren't enough, Max Bell School %of Public Policy, McGill, November 14
%Federal Cabinet Retreat 2020, Winnipeg, January 20
%Rural Economy and Environmental Sociology (REES) department seminar, February 7
%Canadian Bar Association, Edmonton Environmental Law Chapter, February 13
%Alberta Corporate Growth Summit, March 11
%Harry Ainlay High School, March 13
%UBC Political Science class, April 1
%Canadian Bar Association, Edmonton Constitutional Law Chapter, May 7
%Alberta Electrical System Operator (AESO) University, June 5.
%Economic Society of Northern Alberta, June 16

%Writing

%Alberta Court of Appeal Opines That Federal Carbon Pricing Legislation %Unconstitutional (with Martin Olszynski and Nigel Bankes), ABlawg, March 17, %2020

%Clearing the Air on Teck Frontier (Extended ABlawg Edition) (with Martin %Olszynski), ABlawg, February 13, 2020

%Clearing the air on Teck Frontier: Here are the expected benefits — and harms — %of the oilsands project (with Martin Olszynski), CBC, Feb 12, 2020

%Alberta's TIER regulations good on electricity, not so good on oilsands
%CBC, October 30, 2019

%Election 2019: A national reckoning on climate change
%CBC, October 16, 2019

%How The Four Federal Parties' Climate Plans Stack Up (with Katharine Hayhoe), %Chatelaine, October 3, 2019

%Liberals' climate balancing act: The world is watching, CBC September 25, 2019

%NDP climate plan tries (and fails) to carve out middle ground, CBC,September %13, %2019

%What's really holding back the oilsands? It's not the bill of goods you're being %sold, CBC, September 4, 2019

%Conservative climate plan is cloaked in mystery, choked with irony, CBC,
%August 12, 2019

%Greens' climate plan adds up to Mission: Improbable, CBC, August 12, 2019

\end{document}

### 2021-06-05 paper/paper_jan_7.tex
paper files

--- 2021-06-04 pass_through_paper/paper_draft_jan_7_up.tex
@@
 We evaluate the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.
-Between 2015 and 2018, changes affecting carbon prices, output-based allocations of emissions credits, and which facilities are covered by the policies.
-These combinations imply changes both across and within facility types, and across and within portfolios held by major players in the market.
-We show that even after adjusting for actual costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.
+Between 2015 and 2019, legislative changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes.
+Changes in legislation created varying effects both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
+We exploit this variance in treatment to provide unique evidence of offer response to carbon pricing in power markets.
+We show that even after adjusting for costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.
 
-The Alberta wholesale power market provides an excellent laboratory for study because it is small, isolated, and the market is settled on a real-time, energy-only basis so there are limited confounding factors.
-Alberta's market is winter-peaking, with peak 2018 load of just under 12,000 MW.
-It has only three small interconnections to British Columbia, Montana, and Saskatchewan.\footnote{The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan intertie has a path rating of 150 MW.} This means that Alberta's market will be more affected by Alberta's policies than would be the case if traded power had a more important market share.
-We do face a challenge in that Alberta was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies that led to record-low prices in the market.
-Alberta's generation mix is also somewhat unique, with a significant portion of internal load served by combined heat and power (or cogeneration) plants associated with industrial facilities, primarily located in the oil sands.
+The Alberta wholesale power market provides an excellent laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis so there are limited confounding factors.
+Alberta's market is winter-peaking, with peak 2020 load of just under 12,000 MW.\footnote{The maximium Alberta internal load per AESO data was 11698MW, which is also the peak value ever recorded in Alberta.}  Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan.
+The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.
+The relative isolation of Alberta's market will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
+The downside to this is that we do face a challenge in that Alberta was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies, and this economic downturn led to record-low prices in the market.
+Alberta's generation mix is also somewhat unique, with a significant portion of internal load served by combined heat and power (or cogeneration) plants associated with industrial facilities, primarily located in the oil sands region in Northern Alberta.
 
 We compile data describing market offer behaviour and a wide range of relevant covariates.
-We measure and describe facility level hourly offers of power into the market over nearly 10 full years.
+We measure and describe facility-level hourly offers of power into the market over nearly 10 full years.
 We combine these data with plant characteristics compiled from regulatory data, emissions data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data include hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.
-We combine these data with observed carbon policy costs at the facility level.
+Finally, we include observed carbon policy costs at the facility level.
 
-In order to isolate the impact of the changes in policies on offer behaviour, we develop a empirical strategy based on power portfolios.
+In order to isolate the impact of the changes in policies on offer behaviour, we develop a unique empirical strategy based on synthetic power plants and synthetic power portfolios.
 For each hour of each day, we create synthetic power portfolios either by plant type or by which entity holds offer control on the units in question.
@@
 
+Results preview...
+
 \section{Alberta's GHG Policy Changes}
@@
 Alberta has had carbon pricing in place since the Specified Gas Emitters Regulation (SGER) took effect on July 1, 2007.
-That regulation, the first industrial carbon price in North America, implemented a price of \$15/tonne, and allocated emissions credits to covered facilities at a rate equal to 88\% of the facility's historic (2003-2005) emissions intensity or, for new facilities, a rate equal to 88\% of their average year 3 emissions intensity.
-This system remained in place until June, 2015 when the newly-elected government of Premier Rachel Notley introduced changes to the existing regulation which increased the price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for emissions allocation to 85\% and 80\% respectively for 2016 and 2017.
-Combined, these changes implied at increase in the average cost of carbon in each of the years 2016 and 2017.
-
+That regulation, the first industrial carbon price in North America, implemented a price of \$15/tonne, and allocated emissions credits to covered facilities at a rate equal to 88\% of an individual facility's historic (2003-2005) emissions intensity.
+For new facilities built after 2005, they were allocated emissions credits per unit output at a rate equal to 88\% of their average year 3 emissions intensity.
 In addition to carbon pricing on industrial emissions, the SGER also included an offset protocol which provided emissions credits for deemed emissions reductions due to certain activities.
@@
 
-The government also adopted, in November of 2015, a more comprehensive change to GHG emissions policies.
+This system remained in place until June, 2015 when the newly-elected government of Premier Rachel Notley introduced changes to the existing regulation which increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for emissions allocation to 85\% and 80\% respectively for 2016 and 2017.
+Combined, these changes implied at increase in the average cost of carbon in each of the years 2016 and 2017.
+The government subsequently adopted, in November of 2015, a more comprehensive change to GHG emissions policies.
 Two changes in this iteration of policies affected power markets.
-Most importantly, the \textit{Carbon Competitiveness Incentive Regulation (CCIR)} replaced the SGER and these regulations levelized the output-based allocation of emissions credits across all power generators at 0.37t/MWh, the emissions-intensity of the best-in-class combined cycle natural gas generation facility.
+Most importantly, the \textit{Carbon Competitiveness Incentive Regulation (CCIR)} replaced the SGER and these regulations levelized the output-based allocation of emissions credits across all power generators at 0.37t/MWh, the emissions intensity of the best-in-class combined cycle natural gas generation facility in the province.
 This implied that coal producers saw a steep increase in their average costs of carbon, while impacts on gas power plants varied depending on the heat rates of the facility.
-Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease under the CCIR.
+Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease slightly under the CCIR.
 The second important change was that an economy-wide carbon price was introduced for facilities not covered under the CCIR - those without historic emissions in any previous year greater than 100,000 tonnes.
@@
 
-The majority of our data for this paper comes from the Alberta Electricity System Operator (AESO).
-The AESO provides, with a 60 day lag, offers made by power plants into the pool on an hourly basis.
+The core of our data for this paper comes from the Alberta Electricity System Operator (AESO).
+The AESO provides, with a 60-day lag, information on the offers made by power plants into the pool on an hourly basis.
 Facilities offer their power in up to 7 increasing price blocks, with a price cap of \$1000/MWh.
-Blocks may be either flexible or not, and non-flexible blocks will only be dispatched when demand allows the entire block to be used for the hour.
-Plants are dispatched in merit order, with the lowest price blocks dispatched first.
-We have hourly merit order and dispath data from September 1, 2009 through February 28, 2020 where we cut off the data to avoid impacts of the COVID-19 pandemic.\footnote{For a discussion of Alberta market responses to the COVID-19 pandemic, see Leach et al.
-(2020)} A subset of the data, from 2015 onward, allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.
+Plants are dispatched according the merit order of these offer blocks, from lowest to highest offered prices.\footnote{Blocks may be either flexible or not, and non-flexible blocks will only be dispatched when demand allows the entire block to be used for the hour.
+If what would otherwise be the marginal block is not a flexible block, the next-highest-priced offer block will be dispatched.} We have hourly merit order and dispatch data from September 1, 2009 through February 28, 2020 where we cut off the data to avoid impacts of the COVID-19 pandemic.\footnote{For a discussion of Alberta market responses to the COVID-19 pandemic, see Leach et al.
+(2020).
+Our raw data set extends to November, 2020.} The data also allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.
 
 We also use three other AESO data sets to build our analysis sample.
-First, the AESO issues hourly price and load data including lagged load and price forecasts which we merge with merit order data.
-Next, we also included combined intertie capability rating data which allow us to account for hours of limited import or export capability.
+First, the AESO issues hourly price and load data which we merge with merit order data.
+Next, we also combined intertie capability rating data which allow us to account for hours of limited import or export capability.
 Finally, and most importantly, we used metered volumes data at the facility level to incorporate renewable generation into the merit order.
-The AESO currently treats non-dispatchable renewables (wind and solar power in our case) as negative load, but lists them in the merit order data as \$0 offers at full nameplate capacity.
-Prior to 2015, non-dispatchable renewables were not included at all in the merit order.
-We correct this omission using hourly metered net-to-grid volumes for each facility-hour pair in our data set, which we use to create \$0 offers which we merge into the merit order.
-This accurately reflects how the Alberta grid operator treats non-dispatchable renewables - generators do not bear any quantity risk and simply act as price takers.
+The AESO treats non-dispatchable renewables (wind and solar power in our case) as negative load, but lists them in the merit order data as \$0 offers at full nameplate capacity.
+For each facility-hour pair in our data set, we merge metered generation into the merit order and treat renewable generators as having offered their metered volumes each hour.
 
@@
 ADD ON-LINE APPENDIX TABLE OF PLANT INFO.
+
 We also use compliance data from the Alberta Specified Gas Emitters Regulations and the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for facilities.
@@
 We add weather data from Environment Canada weather stations in Edmonton, Fort McMurray and Calgary, the three major demand centers in Alberta.
-We take the average of two measurements in each area in order to maximize the number of hours we can cover with weather data.
+We take the average of available measurements in each area in order to maximize the number of hours we can cover with weather data.
 With this approach, we lose only X hours without at least one observation in each of the three areas.
 
-Finally, we merge daily natural gas prices for the Nova Inventory Transfer (NIT) hub.  We access daily spot prices from NRGStream.
+Finally, we merge daily natural gas prices for the Nova Inventory Transfer hub.  We access daily spot prices from NRGStream.
 
@@
 
+Using our dataset, we can provide a statistical portrait of Alberta's power market.
 Alberta's wholesale market is a single price, energy-only market.
 There is no day-ahead energy market, but there is a separate ancillary services market which we do not model in this paper.
-Alberta's record internal load, reached on January 11, 2018, is 11,697 MW, so Alberta's is a relatively small power market.
-Alberta has seen significant growth in average and peak loads, with X cumulative annual average load growth rate during our sample period, although the 2016 through 2020 period has seen much less growth in load.
-Alberta is also a winter-peaking system which differentiates it from most North American power grids.
-Finally,
-consumption in Alberta is relatively stable due to the large industrial base.
-
-\begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_time.png}
-\label{fig:AB load}
-\vspace{-0.75cm}	\caption{Alberta internal load, average and range.  Source: Alberta Electric System Operator data.}
-\end{figure}
+Alberta's record internal load, reached on January 11, 2018, is 11,697 MW, so Alberta's is a relatively small power market compared to ERCOT, for example, with 75,000 MW peak loads.
+Alberta has seen significant growth in average and peak loads, with X cumulative annual average load growth rate during our sample period, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
 
@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices_2000_2020.png}
-\label{fig:prices}
+\label{fig:masnadi_1}
+\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours.  Source: Alberta Electric System Operator data.}
+\end{figure}
+
+\begin{figure}[!h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices_2000_2020.png}
+\label{fig:masnadi_1}
 \vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours.  Source: Alberta Electric System Operator data.}
@@
 
+Consumption in Alberta is relatively stable due to the large industrial base. For example, here's AB load compared to ERCOT...what does this mean?
+
+\begin{figure}[!h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_compare.png}
+\label{fig:load_pattern}
+\vspace{-0.75cm}	\caption{Monthly and hourly load patterns in Alberta compared to ERCOT and MISO.}
+\end{figure}
+
 \section{Estimation}
-
-\subsection{Market Merit Order}
-
-\subsection{Synthetic Plants}
-
-\subsection{Synthetic Portfolios}
 
@@
 \begin{figure}[!h]%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture_ctax.png}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../alberta_power/price_capture_ctax.png}
 \label{fig:masnadi_1}
@@
 \begin{figure}[p]%
-	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../images/monthly_ghg_mwh.png}
+	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../../alberta_power//monthly_ghg_mwh.png}
 \vspace{-0.75cm}	\caption{Impact of changes in carbon pricing on GHG emissions}\label{fig:monthly_ghg}
@@
 \begin{figure}[p]%
-	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../images/gen_ghg_price.png}
+	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../../alberta_power/gen_ghg_price.png}
 \vspace{-.5cm}	\caption{Impact of changes in carbon pricing on market share by fuel.
@@
 \vspace{10cm}
+
+\begin{figure}[!h]%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../alberta_power/peak_prices_2009_2019.png}
+\label{fig:masnadi_1}
+\vspace{-0.75cm}	\caption{Impact of changes in carbon pricing on plant revenues by type}
+\end{figure}
+
+\newpage

### 2021-06-05 paper/paper_mar_2.tex
paper files

--- 2021-06-05 paper/paper_jan_7.tex
@@
 We evaluate the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.
-Between 2015 and 2019, legislative changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes.
-Changes in legislation created varying effects both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
+Between 2015 and 2019, legislative changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes, creating variation in average and marginal carbon prices both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
 We exploit this variance in treatment to provide unique evidence of offer response to carbon pricing in power markets.
@@
 
-The Alberta wholesale power market provides an excellent laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis so there are limited confounding factors.
-Alberta's market is winter-peaking, with peak 2020 load of just under 12,000 MW.\footnote{The maximium Alberta internal load per AESO data was 11698MW, which is also the peak value ever recorded in Alberta.}  Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan.
-The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.
-The relative isolation of Alberta's market will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
+The Alberta wholesale power market provides an excellent laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis, so there are limited confounding factors beyond what is observeable in our data.
+Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.\footnote{The 2021 maximum Alberta internal load per AESO 2021 data was 11729MW at time of writing, which is also the peak value ever recorded in Alberta.}  Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
 The downside to this is that we do face a challenge in that Alberta was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies, and this economic downturn led to record-low prices in the market.
 Alberta's generation mix is also somewhat unique, with a significant portion of internal load served by combined heat and power (or cogeneration) plants associated with industrial facilities, primarily located in the oil sands region in Northern Alberta.
+Despite these anomalies, the transparency and isolation of Alberta's market provide a unique opportunity to study carbon pricing pass-through to prices and the induced changes in market participation.
 
@@
 In addition to carbon pricing on industrial emissions, the SGER also included an offset protocol which provided emissions credits for deemed emissions reductions due to certain activities.
-In the case of combined heat and power plants, the deemed emissions reductions where equivalent to an allocation of 0.418t/MWh for net-to-grid electricity.
-New renewable power facilities were also eligible for offset credits.
+In the case of combined heat and power plants, facilities received an allocation of credits which amounted to 0.418t/MWh for net-to-grid electricity.
+New renewable power facilities were also eligible for offset credits under the SGER, with a deemed emissions reduction rate of 0.65t/MWh.\footnote{Cite.
+Subsequently reduced to 0.59t/MWh (2015) and 0.53t/MWh (2019)}
 
-This system remained in place until June, 2015 when the newly-elected government of Premier Rachel Notley introduced changes to the existing regulation which increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for emissions allocation to 85\% and 80\% respectively for 2016 and 2017.
-Combined, these changes implied at increase in the average cost of carbon in each of the years 2016 and 2017.
+This system remained in place until June, 2015 when the government of Premier Rachel Notley introduced a series of changes to the existing regulation.
+The first set of changes increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for emissions allocation to 85\% and 80\% respectively for 2016 and 2017.
+Combined, these changes implied a material increase in the average cost of carbon in each of the years 2016 and 2017.
 The government subsequently adopted, in November of 2015, a more comprehensive change to GHG emissions policies.
@@
 This implied that coal producers saw a steep increase in their average costs of carbon, while impacts on gas power plants varied depending on the heat rates of the facility.
-Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease slightly under the CCIR.
+Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease under the CCIR.
 The second important change was that an economy-wide carbon price was introduced for facilities not covered under the CCIR - those without historic emissions in any previous year greater than 100,000 tonnes.
@@
 An opt-in provision allowing these smaller firms to be covered under the CCIR was available, should facilities wish to undertake the more comprehensive emissions reporting.
+While not directly material to our sample period, the government also announced an accelerated coal phase-out which would see all coal-fired generation shut-down, equipped with carbon capture and storage, or re-fired with natural gas by December 31, 2030.
 
@@
 The core of our data for this paper comes from the Alberta Electricity System Operator (AESO).
-The AESO provides, with a 60-day lag, information on the offers made by power plants into the pool on an hourly basis.
-Facilities offer their power in up to 7 increasing price blocks, with a price cap of \$1000/MWh.
+The AESO provides, with a 60-day lag, information on the offers made by power plant owners or controlling entities into the pool on an hourly basis.
+Facilities offer their power in up to 7 increasing-price blocks, with a price cap of \$1000/MWh and a floor of \$0/MWh.
 Plants are dispatched according the merit order of these offer blocks, from lowest to highest offered prices.\footnote{Blocks may be either flexible or not, and non-flexible blocks will only be dispatched when demand allows the entire block to be used for the hour.
@@
 (2020).
-Our raw data set extends to November, 2020.} The data also allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.
+Our raw data set extends to November, 2020.} The data also allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.\footnote{The offer control identifiers are only published after 2013, so we...}
 
@@
 First, the AESO issues hourly price and load data which we merge with merit order data.
-Next, we also combined intertie capability rating data which allow us to account for hours of limited import or export capability.
-Finally, and most importantly, we used metered volumes data at the facility level to incorporate renewable generation into the merit order.
+Next, we combine intertie capability rating data which allow us to account for hours of limited import or export capacity.
+Finally, and most importantly, we use metered volumes data at the facility level to incorporate renewable generation into the merit order.
 The AESO treats non-dispatchable renewables (wind and solar power in our case) as negative load, but lists them in the merit order data as \$0 offers at full nameplate capacity.
-For each facility-hour pair in our data set, we merge metered generation into the merit order and treat renewable generators as having offered their metered volumes each hour.
+For each facility-hour pair in our data set, we merge metered generation into the merit order and treat renewable generators as having offered their metered volumes each hour, still at a \$0MWh offer.
 
@@
 Since these data are both reported at the facility level, not by units or blocks, we apply these measurements in the same way in our data.
-As such, we do not allow emissions intensity to vary with the intensity of use of a particular unit in a particular hour nor do we account for any ramping rates of units.
+As such, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility in a particular hour nor do we account for any ramping rates of units.
 
@@
 There is no day-ahead energy market, but there is a separate ancillary services market which we do not model in this paper.
-Alberta's record internal load, reached on January 11, 2018, is 11,697 MW, so Alberta's is a relatively small power market compared to ERCOT, for example, with 75,000 MW peak loads.
-Alberta has seen significant growth in average and peak loads, with X cumulative annual average load growth rate during our sample period, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
+Alberta's record internal load, reached on February 9, 2021, is 11,729MW, so Alberta's is a relatively small power market compared to ERCOT, for example, with 58,597 MW peak 2021 load.\footnote{The AESO previous record, reached on January 11, 2018, was 11,697 MW, so peak loads have been quite stable in recent years after a period of rapid growth} Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_load}, with X cumulative annual average load growth rate during our sample period, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
+There was also a major short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, which explains the markedly-lower spring trough in that year shown in Figure \ref{fig:ab_load}.
+
+\begin{figure}[!h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_time.png}
+\label{fig:ab_load}
+\vspace{-0.75cm}	\caption{Alberta Internal Load. Source: Alberta Electric System Operator data.}
+\end{figure}
 
@@
 There are three distinct periods during our sample - relatively high load growth and tight market conditions from 2009 through 2013, followed by sharply constrained growth and high reserve margins after 2014 until 2018, followed by a return to tighter market conditions after the spring of 2018.
-There was also a major short-term market event with the wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load offline.
+These conditions are reflected in peak and off-peak prices shown in Figure \ref{fig:ab_prices}.
 Some of these changes market coincide with policy changes in our experiment - in particular, the changes to the Specified Gas Emitters regulations introduced in June of 2015 which took effect in January of 2016 and 2017 respectively coincide with a period of over-supply, while the introduction of the CCIR in 2018 occurred during a period of tightening reserve margins and lingering uncertainty over market structure.
@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices_2000_2020.png}
-\label{fig:masnadi_1}
+\label{fig:ab_prices}
 \vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours.  Source: Alberta Electric System Operator data.}
@@
 
-\begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices_2000_2020.png}
-\label{fig:masnadi_1}
-\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours.  Source: Alberta Electric System Operator data.}
-\end{figure}
+The generation mix in Alberta's power market is dominated by fossil fuels although, as shown in Figure \ref{fig:gen_mix}, the dominant fossil fuel has changed from coal to natural gas.
+Within natural gas generation, the mix of plant types is also relevant to our study.
+X\% are combined cycle plants, y\% are simple-cycle plants, and Z\% are combined heat and power facilities which tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.
+As will be discussed below, any net-to-grid power from these facilities is primarily offered into the market at a \$0 offer and accepts market price.
 
-The generation mix in Alberta's power market is dominated by fossil fuels although, as shown in Figure \ref{fig:gen_mix}, the dominant fossil fuel has change from coal to natural gas.
-Within natural gas generation, the mix of plant types is also relevant to our study.
-X\% are combined cycle plants, y\% are simple-cycle plants, and Z\% are combined heat and power facilities which tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated relay on them for process heat.
-As will be discussed below, any net-to-grid power from these facilities is primarily offered into the market at a \$0 offer and accepts market price.
+ADJUST FIGURE TO SHOW SAMPLE PERIOD
 
@@
 
-Consumption in Alberta is relatively stable due to the large industrial base. For example, here's AB load compared to ERCOT...what does this mean?
+Consumption in Alberta is relatively stable due to the large industrial base.
+For example, in Figure \ref{fig:load_pattern}, we compare Alberta's hourly load pattern to ERCOT and MISO equivalents.
+In the left-hand panel, we see that there is less variability through the day in Alberta as industrial load dominates cycles in residential and commercial load.
+Seasonal patterns are also both less pronounced and exhibit a winter peak, compare to summer-peaking ERCOT and MISO systems.
 
@@
 
-\section{Estimation}
+The nature of the wholesale market is such that different generators will derive different levels of average revenue from the sale of their power.
+Those that dispatch only in high-price times will have higher revenues on average than plants without dispatch control, or plants which run most hours of the day.
+Since prices have varied substantially over time, as shown above in Figure \ref{fig:ab_prices}, captured average prices at the plant level have varied as well.
+These are shown below in Figure \ref{fig:capture_prices}.
+We can see evidence here of how plants of different types offer their power into the market: hydro and combined- and simple-cycle gas turbines follow prices, and dispatch more MW in higher-priced hours, and imports also tend to occur when prices are higher.
+Lower prices tend to prevail when wind generation is highest, and so their capture prices are lower than other forms of generation, and exports tend to be scheduled when prices are expected to be low.
+Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewables since its generation pattern coincides with daily and seasonal peak prices.
+
+\begin{figure}[!h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture.png}
+\label{fig:capture_prices}
+\vspace{-0.75cm}	\caption{Annual average revenue by plant type}
+\end{figure}
+
+\section{The Impact of Carbon Pricing Policy Changes}
+
+Alberta's three carbon pricing regimes affect generating facilities in
+
+\section{Estimation Strategy}
+There are two factors which confound our ability to draw inference from the merit order data.
+The first is that plants of the same type (e.g.
+coal or combined-cycle natural gas plants) may be of different sizes (e.g.
+a 350MW plant vs a 200MW plant) while facing similar optimization constraints, and they may be jointly dispatched by individual operators.
+The second issue is that the strategy space for each facility is very large, given that they may decide both on the break-points for each of up to 7 blocks of power, and then decide on a price for the first block and an adder for each subsequent block offered into the market.
+For each hour, then, the optimization problem consists of 12 choice variables and we have data on these choices for over 18.5 million facility-hour-block pairs.
+We render our problem more parsimonious, and at the same time overcome the portfolio bidding issue, both how the overall merit order and how power from a particular type of plant has been offered into the market over time.
+
+This is best understood with a series of figures, and for this we will again use one coal plant (Capital Power's Genessee 3 facility) as well as Alberta's overall coal fleet to demonstrate the approach we take here.
 

### 2021-06-05 merit_build.R

@@ -209,7 +209,7 @@ load("data/forecast_data.RData")
   
   #add updated temperature data
   source("cdn_weather.R")
-  load("ab_power_temps.RData")
+  load("data/ab_power_temps.RData")
   temps_power<-update_weather_data(temps_power)
   save(temps_power,file="ab_power_temps.RData")  
   

### 2021-06-05 paper/paper_draft.tex
files

--- 2021-06-05 paper/paper_mar_2.tex
@@
 \begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_time.png}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{images/load_time.png}
 \label{fig:ab_load}
@@
 \begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices_2000_2020.png}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{images/peak_prices_2000_2020.png}
 \label{fig:ab_prices}
@@
 \begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel.png}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{images/gen_fuel.png}
 \label{fig:gen_mix}
@@
 \begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_compare.png}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{images/load_compare.png}
 \label{fig:load_pattern}
@@
 \begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture.png}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{images/price_capture.png}
 \label{fig:capture_prices}

### 2021-06-05 presentations/aere_deck.tex
files

\documentclass{beamer}

\usepackage{graphicx,hyperref,fes,url}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
%\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{multirow}
\usepackage{capt-of}
\usepackage{graphicx}
\usepackage{array}
\usepackage{tikz}
\usepackage{hyperref}

% The title of the presentation:
%  - first a short version which is visible at the bottom of each slide;
%  - second the full title shown on the title slide;
\title[FES style for Beamer]{
  UAlberta FES style for Beamer \LaTeX}

% Optional: a subtitle to be dispalyed on the title slide

% The author(s) of the presentation:
%  - again first a short version to be displayed at the bottom;
%  - next the full list of authors, which may include contact information;

\author[Author Short]{
  Pim Vullers MSc \and Martijn Sprengers \\\medskip
  {\small \url{p.vullers@cs.ru.nl} \and \url{m.sprengers@student.ru.nl}} \\
  {\small \url{http://www.cs.ru.nl/~pim/}}}

% The institute:
%  - to start the name of the university as displayed on the top of each slide
%    this can be adjusted such that you can also create a Dutch version
%  - next the institute information as displayed on the title slide
\institute[University of Alberta]{
  Future Energy Systems Symposium\\
  University of Alberta}

% Add a date and possibly the name of the event to the slides
%  - again first a short version to be shown at the bottom of each slide
%  - second the full date and event name for the title slide

\renewcommand{\(}{\begin{columns}}
\renewcommand{\)}{\end{columns}}
\newcommand{\<}[1]{\begin{column}{#1}}
\renewcommand{\>}{\end{column}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\degC}{$^o$C$\,$}
\newcommand{\co}{$\text{CO}_{2}\,$}
\newcommand{\tx}{$G_{2\times\text{CO}_2}$}
\newcommand{\Real}{\mathbb R}

% The main document

\begin{document}

\title{Carbon price pass through in electricity with known prices}
%\author{Branko Bo\v skovi\'{c} \and Andrew Leach \and Charles F.\ Mason}
%\institute{University of Alberta  \and University of Alberta  \and University of Wyoming}

%\author[shortname]{author1 \inst{1} \and author2 \inst{2}}
%\institute[shortinst]{\inst{1} affiliation for author1 \and %
 %                     \inst{2} affiliation for author2}

\author[Short Name (U ABC)]{%
  \texorpdfstring{%
    \begin{columns}
      \column{.1\linewidth}
      \column{.4\linewidth}
      \centering
      Andrew Leach \\University of Alberta
      \column{.4\linewidth}
      \centering
      Blake Shaffer \\University of Calgary
      \column{.1\linewidth}
      \end{columns}
 }
 {Leach and Shaffer (2019)}
}
\date{\small May, 2019}
\frame[plain]{
  \titlepage
}

\section{Introduction}
\begin{frame}{Carbon Pricing, Electricity Markets and Output-Based Allocations}
\vbox to \textheight{
  \begin{block}{}
    \begin{itemize}
    \item Impacts of carbon pricing has become a central policy question in Canada and globally
    \item Electricity markets, at least in certain jurisdictions are ideal environments in which to study responses to carbon prices
    \item Rich, hourly electricity market data allows us to observe firm-level decisions with respect to carbon price pass through
    \item Fabra and Reguant (2014) shows how firms respond to emissions credit prices and that lump-sum credit allocations were not distortionary in the short run
    \item Alberta policy changes alter both carbon prices and output-based credit allocations
    \item Brown et al. (2018) have looked at the dynamics of such a market in a simplified, simulation model
    \end{itemize}
  \end{block}
\vfill}
\end{frame}

%\begin{frame}{Oil sands reserves\hspace{3cm} \small Source: AER (2013)}
%\pgfputat{\pgfxy(0,.25)}{\pgfimage[width=4.25in]{AB_reserves.pdf}}
%\end{frame}

\begin{frame}{Alberta's Power Market is an Ideal Laboratory}
\vbox to \textheight{
  \begin{block}{Alberta's power market:}
    \begin{itemize}
    \item Alberta's power market is relatively small, isolated, and features a mix of generation technologies
    \item Our peak internal load for 2018 was 11,697 MW, while our lowest observed internal load was 	7,819 MW
    \item We have relatively minimal intertie capacity.  Alberta is connected to WECC (1325 MW export and 1500 MW import path-rating) via Montana and BC) and to Saskatchewan (150 MW path-rating) but significant hourly variability in capacity which helps us with identification
    \item Supply mix (2018) was 44.8\% coal, 21.1\% net-to-grid from combined heat and power plants, 18.5\% natural gas, 2.9\% hydro, 6.4\% wind, 	1.0\% other sources including solar and 5.2\% imports.
  \end{itemize}
  \end{block}
\vfill}
\end{frame}

\begin{frame}{Alberta's Policy Changes as Treatments}
\vbox to \textheight{
  \begin{block}{Alberta's GHG policy changes:}
    \begin{itemize}
    \item From 2007 to 2015, Alberta's \textit{Specified Gas Emitters Regulation} imposed a carbon price of \$15 per tonne with output-based allocations at a rate of 88\% of a facility's historic emissions intensity if the facility's annual emissions were more than 100,000 tonnes per year
    \item In 2015, these parameters were changed to \$20 per tonne with allocations equal to 85\% of historic emissions intensities in 2016 and \$30 per tonne with allocations of 80\% in 2017.
    \item 2007-2017 system also had generous emissions credits for combined heat and power systems
    \item In 2018, the system changed again, to one with output-based allocations fixed for all generators (0.37t/MWh) including for net generation from combined heat and power.
    \end{itemize}
  \end{block}
\vfill}
\end{frame}

\begin{frame}{Data Richness}
\vbox to \textheight{
  \begin{block}{We have data on:}
    \begin{itemize}
    \item Hourly merit order (power offers) by unit with data on who held offer control for that unit at that time
    \item Hourly forecast and actual loads and prices
    \item Hourly weather for Edmonton, Calgary, and Fort McMurray
    \item Hourly imports, exports, and intertie capabilities
    \item Hourly wind forecasts including 3 and 7 day advanced forecasts.
    \item Plant-level emissions intensities and historic allocation rates under \textit{SGER}
    \end{itemize}
  \end{block}
\vfill}
\end{frame}

\section{Results Preview}
\begin{frame}{Carbon pricing impacts on power markets}
\vbox to \textheight{
  \begin{block}{Our results show that:}
    \begin{itemize}
    \item ...
    \end{itemize}
  \end{block}
\vfill}
\end{frame}

\begin{frame}{Carbon Pricing Policy and Power Generation Sources}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.9\paperwidth]{../../alberta_power/gen_ghg_price.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Carbon Intensity of Alberta Power}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.9\paperwidth]{../../alberta_power/monthly_ghg_mwh.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Carbon pricing policy impacts on the merit order}
   %\tikz [remember picture,overlay]
   %# \node[yshift=-.5cm,xshift=0cm] at (current page.center)
    %    #{\includegraphics[width=.9\paperwidth]{tax_runs.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Within-group, plant-level impacts of policies}
   %\tikz [remember picture,overlay]
   %# \node[yshift=-.5cm,xshift=0cm] at (current page.center)
    %    #{\includegraphics[width=.9\paperwidth]{tax_runs.png}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Empirical Framework}
\begin{frame}{Empirical Framework}
    \begin{block}{}
    \begin{itemize}
    \item The goal of our empirical framework is to estimate the degree to which carbon prices pass through to merit order behaviour and thus to electricity prices
    \item We follow Fabra and Reguant (2014) which estimates:
    \begin{eqnarray*}\label{eq:production}
    \underset{\color{blue}\substack{\text{Cost}\\{\text{pass-through}}}} {p_{th}} &=&\underset{\color{blue}\substack{\text{Effect of GHG}\\{\text{price times rate}}}} {\rho\tau_te_{th}}
    +\underset{\color{blue}\substack{\text{Supply, demand and common}\\{\text{factor adjustments}}}} {\beta_0X_{th}+\beta_1X_{th}^D+\beta+2X_{th}^S}\\&+&
    \underset{\color{blue}\substack{\text{Fixed}\\{\text{effects}}}}{B_3I_{th}}
    + \underset{\color{blue}\substack{\text{Error}\\{\text{term}}}}{\epsilon_{th}}
    \end{eqnarray*}
    \item Alberta policy changes alter both carbon prices and output-based credit allocations over time, which changes the empirical specifications in our case slightly
    \item We have hourly forecast prices and loads (3 hour leads) which we can use as supply and demand indicators
    \item We also have both forecast and actual temperature and wind generation data over the sample time period
    \end{itemize}
  \end{block}
   \vfill
\end{frame}

\section{Empirical Framework}
\begin{frame}{Empirical Framework}
    \begin{block}{}
    \begin{itemize}
    \item Like Fabra and Reguant, we also have hourly bids by plant for up to 7 blocks - an individual supply function - which we can then use to estimate firm-level residual demand for the marginal producer
    \begin{eqnarray*}\label{eq:production}
    \underset{\color{blue}\substack{\text{Marginal bid of firm $i$}\\{\text{unit $j$, time $th$}}}} {b_{ijth}} &=&\underset{\color{blue}\substack{\text{Effect of GHG}\\{\text{price times rate}}}} {\gamma\tau_te_{j}}
    +\underset{\color{blue}\substack{\text{Marginal cost}\\{\text{estimates}}}} {\beta c_{jt}}\\&+&
    \underset{\color{blue}\substack{\text{Approximate}\\{\text{markup}}}}{\theta\hat{m}_{ijth}}
    + \underset{\color{blue}\substack{\text{Error}\\{\text{term}}}}{\epsilon_{ijth}}
    \end{eqnarray*}
\item in our case, emissions prices vary over time as do allocation rates, so we need to modify both structures
    \end{itemize}
  \end{block}
   \vfill
\end{frame}

\section{Results}

\begin{frame}{Results}
   %\tikz [remember picture,overlay]
   %# \node[yshift=-.5cm,xshift=0cm] at (current page.center)
    %    #{\includegraphics[width=.9\paperwidth]{tax_runs.png}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Discussion}

\begin{frame}{Discussion}
   %\tikz [remember picture,overlay]
   %# \node[yshift=-.5cm,xshift=0cm] at (current page.center)
    %    #{\includegraphics[width=.9\paperwidth]{tax_runs.png}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Other FES Theme 06 Research}
\begin{frame} Secanell group
\begin{block}{}
\footnotesize{\begin{itemize}
  \item Design, procurement and installation of low and high temperature electrolysis fabrication and testing facilities
  \item Fabrication of single-cell low temperature electrolyzer prototypes using iridium and iridium oxide catalysts
  \item Demonstrated, for the first time, the use of inkjet printing as a feasible technology to fabricate low temperature electrolyzer cells  (https://doi.org/10.1149/2.1101807jes)
   \item Development of a novel nano-pompon-like iridium catalyst superstructure for low temperature fuel cell electrolyzers (https://doi.org/10.1016/j.jcat.2019.01.018)
   \item Development of a novel technique for characterization of iridium catalyst degradation (https://doi.org/10.1016/j.jcat.2019.01.018)
 \item Development of novel micro-structures for high temperature solid oxide electrolyzers and fuel cells (https://doi.org/10.1016/j.electacta.2018.02.055)
  \end{itemize}}
The project currently involves 5 PhD students and 1 MSc, 5 faculty (1 in MECE, 3 in CHEME and 1 in Chem). We are also collaborating with Ionomr, a firm in BC (https://ionomr.com/).
  \end{block}
   \vfill
\end{frame}

\begin{frame}{Secanell Group}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
      {\includegraphics[height=3in]{../../secanell_1}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Other FES Theme 06 Research}
\begin{frame} Secanell group
\begin{block}{}
\footnotesize{\begin{itemize}
  \item Design, procurement and installation of low and high temperature electrolysis fabrication and testing facilities
  \item Fabrication of single-cell low temperature electrolyzer prototypes using iridium and iridium oxide catalysts
  \item Demonstrated, for the first time, the use of inkjet printing as a feasible technology to fabricate low temperature electrolyzer cells  (https://doi.org/10.1149/2.1101807jes)
   \item Development of a novel nano-pompon-like iridium catalyst superstructure for low temperature fuel cell electrolyzers (https://doi.org/10.1016/j.jcat.2019.01.018)
   \item Development of a novel technique for characterization of iridium catalyst degradation (https://doi.org/10.1016/j.jcat.2019.01.018)
 \item Development of novel micro-structures for high temperature solid oxide electrolyzers and fuel cells (https://doi.org/10.1016/j.electacta.2018.02.055)
  \end{itemize}}
The project currently involves 5 PhD students and 1 MSc, 5 faculty (1 in MECE, 3 in CHEME and 1 in Chem). We are also collaborating with Ionomr, a firm in BC (https://ionomr.com/).
  \end{block}
   \vfill
\end{frame}

\begin{frame}{Secanell Group}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
      {\includegraphics[height=3in]{../../secanell_2}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Other FES Theme 06 Research}
\begin{frame} Secanell group
\begin{block}{}
\footnotesize{\begin{itemize}
  \item Design, procurement and installation of low and high temperature electrolysis fabrication and testing facilities
  \item Fabrication of single-cell low temperature electrolyzer prototypes using iridium and iridium oxide catalysts
  \item Demonstrated, for the first time, the use of inkjet printing as a feasible technology to fabricate low temperature electrolyzer cells  (https://doi.org/10.1149/2.1101807jes)
   \item Development of a novel nano-pompon-like iridium catalyst superstructure for low temperature fuel cell electrolyzers (https://doi.org/10.1016/j.jcat.2019.01.018)
   \item Development of a novel technique for characterization of iridium catalyst degradation (https://doi.org/10.1016/j.jcat.2019.01.018)
 \item Development of novel micro-structures for high temperature solid oxide electrolyzers and fuel cells (https://doi.org/10.1016/j.electacta.2018.02.055)
  \end{itemize}}
The project currently involves 5 PhD students and 1 MSc, 5 faculty (1 in MECE, 3 in CHEME and 1 in Chem). We are also collaborating with Ionomr, a firm in BC (https://ionomr.com/).
  \end{block}
   \vfill
\end{frame}

\begin{frame}{Secanell Group}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
      {\includegraphics[width=4.5in]{../../secanell_3}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Other FES Theme 06 Research}
\begin{frame} Secanell group
\begin{block}{}
\footnotesize{\begin{itemize}
  \item Design, procurement and installation of low and high temperature electrolysis fabrication and testing facilities
  \item Fabrication of single-cell low temperature electrolyzer prototypes using iridium and iridium oxide catalysts
  \item Demonstrated, for the first time, the use of inkjet printing as a feasible technology to fabricate low temperature electrolyzer cells  (https://doi.org/10.1149/2.1101807jes)
   \item Development of a novel nano-pompon-like iridium catalyst superstructure for low temperature fuel cell electrolyzers (https://doi.org/10.1016/j.jcat.2019.01.018)
   \item Development of a novel technique for characterization of iridium catalyst degradation (https://doi.org/10.1016/j.jcat.2019.01.018)
 \item Development of novel micro-structures for high temperature solid oxide electrolyzers and fuel cells (https://doi.org/10.1016/j.electacta.2018.02.055)
  \end{itemize}}
The project currently involves 5 PhD students and 1 MSc, 5 faculty (1 in MECE, 3 in CHEME and 1 in Chem). We are also collaborating with Ionomr, a firm in BC (https://ionomr.com/).
  \end{block}
   \vfill
\end{frame}

\section{Other FES Theme 06 Research}
\begin{frame} Secanell group
\footnotesize{\begin{itemize}
  \item Design, procurement and installation of low and high temperature electrolysis fabrication and testing facilities
  \item Fabrication of single-cell low temperature electrolyzer prototypes using iridium and iridium oxide catalysts
  \item Demonstrated, for the first time, the use of inkjet printing as a feasible technology to fabricate low temperature electrolyzer cells  (https://doi.org/10.1149/2.1101807jes)
   \item Development of a novel nano-pompon-like iridium catalyst superstructure for low temperature fuel cell electrolyzers (https://doi.org/10.1016/j.jcat.2019.01.018)
   \item Development of a novel technique for characterization of iridium catalyst degradation (https://doi.org/10.1016/j.jcat.2019.01.018)
 \item Development of novel micro-structures for high temperature solid oxide electrolyzers and fuel cells (https://doi.org/10.1016/j.electacta.2018.02.055)
  \end{itemize}}
The project currently involves 5 PhD students and 1 MSc, 5 faculty (1 in MECE, 3 in CHEME and 1 in Chem). We are also collaborating with Ionomr, a firm in BC (https://ionomr.com/).
   \vfill
\end{frame}

\begin{frame}{Contact info}
\begin{center}
Andrew Leach\bigskip \\
School of Business, University of Alberta\bigskip \\
Email: \href{mailto:aleach@ualberta.ca}{andrew.leach@ualberta.ca}\bigskip \\
Twitter: \href{http://twitter.com/andrew_leach}{\url{@andrew_leach}}
\end{center}
\vfill
\end{frame}

\end{document}

### 2021-06-05 merit_build.R

@@ -400,7 +400,7 @@ load("data/forecast_data.RData")
   
 print(paste("Market Data Merged. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
 
-save<-0
+save<-1
 #save<-1
 
 if(save==1)

### 2021-06-07 caiso.R
fixing images

@@ -1,0 +1,75 @@
+
+#BLAKE CAISO CODE
+
+#set parameters
+queryname <- "SLD_REN_FCST"
+market_run_id <- "ACTUAL"
+startdatetime <- 20160601
+enddatetime <- 20160630
+
+#create URL
+apiURL<-paste("oasis.caiso.com/oasisapi/Singl.",queryname,"&market_run_id=",market_run_id,"&startdatetime=",startdatetime,"T07:00-0000","&enddatetime=",enddatetime,"T07:00-0000","&version=1",sep="")
+
+#download to temp file
+temp<-tempfile()
+download.file(apiURL,temp,method="curl")
+tempdata <- read.table(unzip(temp), sep = ",", header=TRUE)
+unlink(temp) #deletes 'temp' file
+
+
+
+
+
+
+
+
+
+caiso_mar <- read.xlsx(xlsxFile = "caiso_mar_11.xlsx", sheet = 1, startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
+df_caiso <- melt(caiso_mar,id=c("Hour"),measure=c("GEOTHERMAL" ,"BIOMASS","BIOGAS","SMALL.HYDRO","WIND.TOTAL","SOLAR.PV","SOLAR.THERMAL"),value.name = "generation") 
+df2_caiso <- melt(caiso_mar,id=c("Hour"),measure=c("RENEWABLES.TOTAL", "NUCLEAR","THERMAL","IMPORTS","HYDRO"),value.name = "generation") 
+
+png(file="caiso_day.png", width = 1400, height = 750)
+ggplot(df_caiso, aes(Hour,generation))+ 
+  geom_area(aes(fill=variable), position = 'stack')+ guides(fill=guide_legend(title=NULL))+
+  scale_fill_viridis("Generation Type",discrete = TRUE,option="viridis")+
+  scale_x_continuous(breaks=seq(1, 24, 2))+ # Ticks from 0-1200, every 200+
+  #scale_fill_brewer(palette="Spectral") +
+  theme_minimal()+theme(
+    legend.position = "right",
+    legend.margin=margin(c(0,0,0,0),unit="cm"),
+    legend.text = element_text(colour="black", size = 18, face = "bold"),
+    plot.caption = element_text(size = 20, face = "italic"),
+    plot.title = element_text(face = "bold"),
+    plot.subtitle = element_text(size = 20, face = "italic"),
+    panel.grid.minor = element_blank(),
+    text = element_text(size = 20,face = "bold"),
[…]

### 2021-06-07 nrgstream_gen.R

@@ -1,0 +1,266 @@
+#code to load and process NRGStream hourly gen data
+
+source("power_paper_base.R")
+#source("aeso_scrapes.R")
+
+
+#routine to load plant info from Excel
+
+get_plant_info <-function() {
+  #load plant data
+  plant_data <- read.xlsx(xlsxFile = "data/AB_Plant_Info_New.xlsx", sheet = "Plant_info", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
+  colnames(plant_data)<-gsub("\\.", " ", colnames(plant_data)) 
+  plant_info<-data.frame(t(plant_data[(1:10),-1]))
+  #fix names
+  plant_info<-setNames(plant_info,t(plant_data[(1:10),1]))
+  plant_info$Capacity <- as.numeric(as.character(plant_info$Capacity))
+  
+  #Some of the names had errors or inconsistencies so fixed here
+
+  plant_info<-arrange(plant_info,NRG_Stream)
+  #put plant info in the Global Environment
+  assign("plant_info", plant_info, envir = .GlobalEnv) #dump it to the global environment
+}
+
+all_gen_data <- function() {
+  #get all generation data
+  years<-c(2004:year(Sys.Date()))
+  count<-1
+  get_plant_info()
+  if(exists("stack"))
+    rm(stack)
+  for (y_data in years){
+  #testing y_data<-2004
+    file_name<-paste("nrgstream/hourly",sprintf('%02d', y_data %% 100),".csv",sep="")
+    print(paste("data file is ",file_name))
+    new_data <- read.csv(file_name,check.names = FALSE,stringsAsFactors = F)
+    new_data[,-1] <- sapply(new_data[,-1], as.numeric )
+    
+    #here, we need to carry forward to fill any missing data points
+    new_data[,-1] <- sapply(new_data[,-1], na.locf, na.rm = FALSE)
+    
+    #take out periods
+    colnames(new_data)<-gsub("\\.", " ", colnames(new_data)) 
+    #rename time
[…]

### 2021-06-07 nrgstream_graphs.R

@@ -1,0 +1,1790 @@
+source("power_paper_base.R")
+library(reshape2)
+
+oba_type<-function(plant_sent,year_sent){
+  grid_avg<-0.65
+  #nested case_when
+  oba_12<-
+    case_when(
+      plant_sent == "HYDRO" ~ grid_avg,
+      plant_sent == "COAL" ~ 1.07125*0.88,
+      plant_sent == "COGEN" ~ 0.418,
+      plant_sent == "NGCC" ~ 0.42*.88,
+      plant_sent == "SCGT" ~ 0.6*.88,
+      plant_sent == "SOLAR" ~ grid_avg,
+      plant_sent == "WIND" ~ grid_avg,
+      TRUE                      ~  0 
+    )
+  oba_20<-
+    case_when(
+      plant_sent == "HYDRO" ~ grid_avg,
+      plant_sent == "COAL" ~ 1.07125*0.8,
+      plant_sent == "COGEN" ~ 0.418,
+      plant_sent == "NGCC" ~ 0.42*.8,
+      plant_sent == "SCGT" ~ 0.6*.8,
+      plant_sent == "SOLAR" ~ grid_avg,
+      plant_sent == "WIND" ~ grid_avg,
+      TRUE                      ~  0 
+    )
+  oba_15<-
+    case_when(
+      plant_sent == "HYDRO" ~ grid_avg,
+      plant_sent == "COAL" ~ 1.07125*0.85,
+      plant_sent == "COGEN" ~ 0.418,
+      plant_sent == "NGCC" ~ 0.42*.85,
+      plant_sent == "SCGT" ~ 0.6*.85,
+      plant_sent == "SOLAR" ~ grid_avg,
+      plant_sent == "WIND" ~ grid_avg,
+      TRUE                      ~  0 
+    )
+  case_when(
+    year_sent <=2015     ~ oba_12,
+    year_sent ==2016     ~ oba_15,
+    year_sent ==2017    ~ oba_20,
+    TRUE                      ~  0.37 
+  )
+}
+
+
+ctax_year<-function(year_sent){
+  case_when(
+    year_sent <= 2015 ~ 15,
+    year_sent == 2016 ~ 20,
+    year_sent == 2017 ~ 30,
+    TRUE                      ~  30 
+  )
+}
+
+deemed_ei<-function(plant_sent,year_sent){
+  case_when(
+    plant_sent == "HYDRO" ~ 0,
[…]

### 2021-06-07 power_merge.R

@@ -15,7 +15,7 @@ library(data.table)
 
 #create long data file
 
-setwd("C:/Users/aleach/Google Drive")
+#setwd("C:/Users/aleach/Google Drive")
 
 #df <- read.xlsx(xlsxFile = "simple_test2.xlsx", sheet = 1, startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
 

### 2021-06-09 paper/paper_draft.tex
fixing images
working

--- 2021-06-05 paper/paper_mar_2.tex
@@
  \and
-Blake Shaffer\\University of Calgary and\\Stanford University
-}
+Blake Shaffer\\University of Calgary}
 \begin{document}
@@
 \begin{abstract}
-\noindent We evaluate .\\
-\vspace{0.2in}
-\noindent Keywords: climate change, social cost of carbon, oil sands\\
+\noindent We evaluate the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.
+Between 2015 and 2019, legislative changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes, creating variation in average and marginal carbon prices both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
+We exploit this variance in treatment to provide unique evidence of offer response to carbon pricing in power markets.
+We show that even after adjusting for costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.\\ \\
+Keywords: climate change, carbon pricing, electricity, coal\\ \\
 JEL classification: Q3, Q4, Q54
@@
 
-The Alberta wholesale power market provides an excellent laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis, so there are limited confounding factors beyond what is observeable in our data.
-Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.\footnote{The 2021 maximum Alberta internal load per AESO 2021 data was 11729MW at time of writing, which is also the peak value ever recorded in Alberta.}  Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
-The downside to this is that we do face a challenge in that Alberta was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies, and this economic downturn led to record-low prices in the market.
-Alberta's generation mix is also somewhat unique, with a significant portion of internal load served by combined heat and power (or cogeneration) plants associated with industrial facilities, primarily located in the oil sands region in Northern Alberta.
-Despite these anomalies, the transparency and isolation of Alberta's market provide a unique opportunity to study carbon pricing pass-through to prices and the induced changes in market participation.
+The Alberta wholesale power market provides an excellent laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis, so there are limited confounding factors beyond what is observable in our data.
+Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.\footnote{The 2021 maximum Alberta internal load per AESO 2021 data was 11,729 MW at time of writing, which is also the peak value ever recorded in Alberta.} Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
+The downside to this is that Alberta, a major oil-producing jurisdiction, was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies, and this economic downturn led to a period of low prices in the power market.
+Alberta's generation mix is also somewhat unique, with a significant portion of internal load served by combined heat and power (or co-generation) plants associated with industrial facilities, primarily located in the oil sands region in Northern Alberta.
+Despite these anomalies, the transparency and isolation of Alberta's market provide us with a unique opportunity to study carbon price pass-through to prices as well as the induced changes in market participation.
 
 We compile data describing market offer behaviour and a wide range of relevant covariates.
-We measure and describe facility-level hourly offers of power into the market over nearly 10 full years.
-We combine these data with plant characteristics compiled from regulatory data, emissions data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data include hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.
-Finally, we include observed carbon policy costs at the facility level.
+We measure and describe facility-level hourly offers of power into the market over slightly more than 10 years preceding the COVID-19 pandemic.
+We combine these data with detailed plant characteristics compiled from regulatory data, emissions compliance data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data including hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.
+Finally, we include both observed carbon policy costs and the value of output-based allocations at the facility level.
 
 In order to isolate the impact of the changes in policies on offer behaviour, we develop a unique empirical strategy based on synthetic power plants and synthetic power portfolios.
-For each hour of each day, we create synthetic power portfolios either by plant type or by which entity holds offer control on the units in question.
-We are then able to examine, for example, how changes in policies affected the offer of power across all coal- or gas-fired plants in the province, or across all facilities owned by particular actors in the market.
+For each hour of each day, we create synthetic power portfolios either by plant type .
+We are then able to examine, for example, how changes in policies affected the offer of power across all coal- or gas-fired plants in the province.
+
+or by which entity holds offer control on the units in question
+, or across all facilities owned by particular actors in the market.
 We normalize these portfolios by percentile, so we are able to compare behaviour across portfolios of different sizes without loss of generality.
@@
 
-Alberta has had carbon pricing in place since the Specified Gas Emitters Regulation (SGER) took effect on July 1, 2007.
+Alberta has had carbon pricing in place since the \emph{Specified Gas Emitters Regulation} (\emph{SGER}) took effect on July 1, 2007.
 That regulation, the first industrial carbon price in North America, implemented a price of \$15/tonne, and allocated emissions credits to covered facilities at a rate equal to 88\% of an individual facility's historic (2003-2005) emissions intensity.
 For new facilities built after 2005, they were allocated emissions credits per unit output at a rate equal to 88\% of their average year 3 emissions intensity.
-In addition to carbon pricing on industrial emissions, the SGER also included an offset protocol which provided emissions credits for deemed emissions reductions due to certain activities.
+In addition to carbon pricing on industrial emissions, the \emph{SGER} also included an offset protocol which provided emissions credits for deemed emissions reductions due to certain activities.
 In the case of combined heat and power plants, facilities received an allocation of credits which amounted to 0.418t/MWh for net-to-grid electricity.
@@
 This system remained in place until June, 2015 when the government of Premier Rachel Notley introduced a series of changes to the existing regulation.
-The first set of changes increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for emissions allocation to 85\% and 80\% respectively for 2016 and 2017.
+The first set of changes increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for the output-based allocation of emissions credits to 85\% and 80\% of historic facility level emissions intensity for 2016 and 2017 respectively.
 Combined, these changes implied a material increase in the average cost of carbon in each of the years 2016 and 2017.
@@
 Two changes in this iteration of policies affected power markets.
-Most importantly, the \textit{Carbon Competitiveness Incentive Regulation (CCIR)} replaced the SGER and these regulations levelized the output-based allocation of emissions credits across all power generators at 0.37t/MWh, the emissions intensity of the best-in-class combined cycle natural gas generation facility in the province.
+Most importantly, the \textit{Carbon Competitiveness Incentive Regulation (CCIR)} replaced the \emph{SGER} and these regulations levelized the output-based allocation of emissions credits across all power generators at 0.37t/MWh, the emissions intensity of the best-in-class combined cycle natural gas generation facility in the province.
 This implied that coal producers saw a steep increase in their average costs of carbon, while impacts on gas power plants varied depending on the heat rates of the facility.
-Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease under the CCIR.
-The second important change was that an economy-wide carbon price was introduced for facilities not covered under the CCIR - those without historic emissions in any previous year greater than 100,000 tonnes.
+Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease under the \emph{CCIR}.
+The second important change was that an economy-wide carbon price was introduced for facilities not covered under the \emph{CCIR} - those without historic emissions in any previous year greater than 100,000 tonnes.
 These facilities did not, by default, receive output-based allocations of emissions credits to offset the cost of the carbon price, so their average costs of carbon could be much higher than their larger competitors.
-An opt-in provision allowing these smaller firms to be covered under the CCIR was available, should facilities wish to undertake the more comprehensive emissions reporting.
+An opt-in provision allowing these smaller firms to be covered under the \emph{CCIR} was available, should facilities wish to undertake the more comprehensive emissions reporting.
 While not directly material to our sample period, the government also announced an accelerated coal phase-out which would see all coal-fired generation shut-down, equipped with carbon capture and storage, or re-fired with natural gas by December 31, 2030.
 
-COVERED FACILITIES TABLE
+The \emph{CCIR} remained in place until it was replaced by the \emph{
+Technology Innovation and Emissions Reduction Regulation} (\emph{TIER}) on January 1, 2020, outside of our sample period.
+However, it's important to not that while \emph{TIER} altered the design of output-based allocations in most sectors, it did not impose material changes in the electricity industry and so we should not expect any changes late in our sample period which abuts the implementation of \emph{TIER}.
 
@@
 
-The core of our data for this paper comes from the Alberta Electricity System Operator (AESO).
+The majority of data for this paper comes from the Alberta Electricity System Operator (AESO).
 The AESO provides, with a 60-day lag, information on the offers made by power plant owners or controlling entities into the pool on an hourly basis.
@@
 Plants are dispatched according the merit order of these offer blocks, from lowest to highest offered prices.\footnote{Blocks may be either flexible or not, and non-flexible blocks will only be dispatched when demand allows the entire block to be used for the hour.
-If what would otherwise be the marginal block is not a flexible block, the next-highest-priced offer block will be dispatched.} We have hourly merit order and dispatch data from September 1, 2009 through February 28, 2020 where we cut off the data to avoid impacts of the COVID-19 pandemic.\footnote{For a discussion of Alberta market responses to the COVID-19 pandemic, see Leach et al.
-(2020).
-Our raw data set extends to November, 2020.} The data also allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.\footnote{The offer control identifiers are only published after 2013, so we...}
+If what would otherwise be the marginal block is not a flexible block, the next-highest-priced offer block will be dispatched.} We have hourly merit order and dispatch data from September 1, 2009 through May 31, 2021, but we truncate the data at December 31, 2019 to avoid impacts of the COVID-19 pandemic.\footnote{For a discussion of Alberta market responses to the COVID-19 pandemic, see Leach et al.
+(2020).} The data also allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.\footnote{The offer control identifiers are only published after 2013, so our analysis on these data takes place over a smaller sample.}
 
 We also use three other AESO data sets to build our analysis sample.
-First, the AESO issues hourly price and load data which we merge with merit order data.
-Next, we combine intertie capability rating data which allow us to account for hours of limited import or export capacity.
+First, the AESO issues hourly price and load data, including 3-hour ahead forecast and actual prices and loads, which we merge with merit order data.
+Next, we include intertie capability rating data which allow us to account for hours of limited import or export capacity.
 Finally, and most importantly, we use metered volumes data at the facility level to incorporate renewable generation into the merit order.
 The AESO treats non-dispatchable renewables (wind and solar power in our case) as negative load, but lists them in the merit order data as \$0 offers at full nameplate capacity.
-For each facility-hour pair in our data set, we merge metered generation into the merit order and treat renewable generators as having offered their metered volumes each hour, still at a \$0MWh offer.
+For each facility-hour pair in our data set, we merge actual metered generation into the merit order and treat renewable generators as having offered exactly their metered volumes each hour, still at a \$0MWh offer.
 
@@
 We also use compliance data from the Alberta Specified Gas Emitters Regulations and the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for facilities.
-Since these data are both reported at the facility level, not by units or blocks, we apply these measurements in the same way in our data.
+Since these data are both reported at the facility level, not by generating unit, we apply these measurements in the same way in our data.
 As such, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility in a particular hour nor do we account for any ramping rates of units.
 
-We add weather data from Environment Canada weather stations in Edmonton, Fort McMurray and Calgary, the three major demand centers in Alberta.
+We add weather data from Environment Canada weather stations in Edmonton, Fort McMurray, Cold Lake, and Calgary, covering the three major demand centers in Alberta.
 We take the average of available measurements in each area in order to maximize the number of hours we can cover with weather data.
-With this approach, we lose only X hours without at least one observation in each of the three areas.
+With this approach, we lose fewer hours from our data set than if we allow unavailable weather data from one station to remove an hour from our sample.
+There is also no reason to believe that weather data availability is correlated with our variables of interest, and so the lost observations should have no leverage on the results.
 
-Finally, we merge daily natural gas prices for the Nova Inventory Transfer hub.  We access daily spot prices from NRGStream.
+Finally, we merge daily natural gas prices for the Nova Inventory Transfer hub.  We access daily spot prices via NRGStream, a commercial data aggregation service.
 
@@
 For the publicly available code, we substitute Henry Hub gas prices as a placeholder, to which we apply the average discount observed between Henry Hub and NIT to adjust the values so that regression results may be more closely replicated.
-
-WE MIGHT BE ABLE TO GET PERMISSION TO INCLUDE THE NGX DATA.
 
@@
 Alberta's wholesale market is a single price, energy-only market.
-There is no day-ahead energy market, but there is a separate ancillary services market which we do not model in this paper.
-Alberta's record internal load, reached on February 9, 2021, is 11,729MW, so Alberta's is a relatively small power market compared to ERCOT, for example, with 58,597 MW peak 2021 load.\footnote{The AESO previous record, reached on January 11, 2018, was 11,697 MW, so peak loads have been quite stable in recent years after a period of rapid growth} Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_load}, with X cumulative annual average load growth rate during our sample period, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
+There is no day-ahead energy market, but there is a separate ancillary services market which we do not address beyond a couple of tangential references in this paper.
+Alberta's record internal load, reached on February 9, 2021, is 11,729MW, so Alberta's is a relatively small power market compared to the similarly-structure Energy Reliability Council of Texas (ERCOT) market, for example, which had a 58,597 MW peak 2021 load.\footnote{The AESO previous record, reached on January 11, 2018, was 11,697 MW, so peak loads have been quite stable in recent years after a period of rapid growth.} Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_load}, with X cumulative annual average load growth rate during our sample period, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
 There was also a major short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, which explains the markedly-lower spring trough in that year shown in Figure \ref{fig:ab_load}.
+The 2020 COVID-19 demand shock is also evident in this graph, but lies outside the sample period for our analysis.
 
-\begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_time.png}
+\begin{figure}[t]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/loads_clean.png}
+\vspace{-0.75cm}	\caption{Alberta Internal Load. Source: Alberta Electric System Operator data.}
 \label{fig:ab_load}
-\vspace{-0.75cm}	\caption{Alberta Internal Load. Source: Alberta Electric System Operator data.}
 \end{figure}
@@
 However, market conditions have varied over the sample period.
-There are three distinct periods during our sample - relatively high load growth and tight market conditions from 2009 through 2013, followed by sharply constrained growth and high reserve margins after 2014 until 2018, followed by a return to tighter market conditions after the spring of 2018.
+There are three distinct periods during our sample - relatively high load growth and tight market conditions from 2009 through 2013, followed by sharply constrained growth and high reserve margins after 2014 through most of 2018, followed by a return to tighter market conditions after the spring of 2018.
 These conditions are reflected in peak and off-peak prices shown in Figure \ref{fig:ab_prices}.
-Some of these changes market coincide with policy changes in our experiment - in particular, the changes to the Specified Gas Emitters regulations introduced in June of 2015 which took effect in January of 2016 and 2017 respectively coincide with a period of over-supply, while the introduction of the CCIR in 2018 occurred during a period of tightening reserve margins and lingering uncertainty over market structure.
+Some of these market changes coincide with policy changes in our experiment - in particular, the changes to the \emph{SGER} introduced in June of 2015 which took effect in January of 2016 and 2017 respectively coincide with a period of over-supply, while the introduction of the \emph{CCIR} in 2018 occurred during a period of tightening reserve margins and lingering uncertainty over market structure.
 From 2018 through pre-COVID 2020, both GHG policy and the wholesale electricity market were tightening, and so we need to be careful to disentangle these impacts.
 
-\begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices_2000_2020.png}
+\begin{figure}[t]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices.png}
+\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours. Peak hours are 8am to 11pm other than on statutory holidays or Sundays. Source: Alberta Electric System Operator data.}
 \label{fig:ab_prices}
-\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours.  Source: Alberta Electric System Operator data.}
+\end{figure}
+
+\begin{figure}[t]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel.png}
+\vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: Alberta Electric System Operator data.}
+\label{fig:gen_mix}
 \end{figure}
@@
 Within natural gas generation, the mix of plant types is also relevant to our study.
-X\% are combined cycle plants, y\% are simple-cycle plants, and Z\% are combined heat and power facilities which tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.
-As will be discussed below, any net-to-grid power from these facilities is primarily offered into the market at a \$0 offer and accepts market price.
+X\% are combined cycle (NGCC) plants, y\% are simple-cycle (NGSC) plants, and Z\% are combined heat and power (COGEN) facilities.
+Combined heat and power plants tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.
+As will be discussed below, net-to-grid power from these facilities is primarily offered into the market at a \$0 offer and accepts market price, and so we see limited evidence of carbon price pass-through for most of the power offered by these facilities.
 
@@
 
-\begin{figure}[!h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel.png}
-\label{fig:gen_mix}
-\vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: Alberta Electric System Operator data.}
-\end{figure}
-
-Consumption in Alberta is relatively stable due to the large industrial base.
+Consumption in Alberta is relatively stable on daily and annual bases due to the large industrial base.
 For example, in Figure \ref{fig:load_pattern}, we compare Alberta's hourly load pattern to ERCOT and MISO equivalents.
@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/load_compare.png}
+\vspace{-0.75cm}	\caption{Monthly and hourly load patterns in Alberta compared to ERCOT and MISO.}
 \label{fig:load_pattern}
-\vspace{-0.75cm}	\caption{Monthly and hourly load patterns in Alberta compared to ERCOT and MISO.}
 \end{figure}
@@
 Those that dispatch only in high-price times will have higher revenues on average than plants without dispatch control, or plants which run most hours of the day.
-Since prices have varied substantially over time, as shown above in Figure \ref{fig:ab_prices}, captured average prices at the plant level have varied as well.
-These are shown below in Figure \ref{fig:capture_prices}.
-We can see evidence here of how plants of different types offer their power into the market: hydro and combined- and simple-cycle gas turbines follow prices, and dispatch more MW in higher-priced hours, and imports also tend to occur when prices are higher.
-Lower prices tend to prevail when wind generation is highest, and so their capture prices are lower than other forms of generation, and exports tend to be scheduled when prices are expected to be low.
-Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewables since its generation pattern coincides with daily and seasonal peak prices.
+Since prices have varied substantially over time, as shown above in Figure \ref{fig:ab_prices}, captured average prices at the plant level have varied as well, as shown in Figure \ref{fig:capture_prices}.
+We can see evidence here of how plants of different types offer their power into the market: hydro and combined- and simple-cycle gas turbines follow prices, and dispatch more capacity in higher-priced hours, just as imports tend to occur when prices are higher.
+Lower prices tend to prevail when wind generation is highest, and so captured prices for wind power facilities are lower than other forms of generation, and exports tend to be scheduled when prices are expected to be low.
+Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewable sources since its generation pattern coincides with daily and seasonal peak prices, although we would expect this to change as the \emph{duck curve} effects take hold over time.
 
@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture.png}
+\vspace{-0.75cm}	\caption{Annual average revenue by plant type}
 \label{fig:capture_prices}
-\vspace{-0.75cm}	\caption{Annual average revenue by plant type}
 \end{figure}

### 2021-06-09 aeso_data_proc.R

@@ -1,0 +1,517 @@
+source("power_paper_base.R")
+source("aeso_scrapes.R")
+
+
+
+#seasons
+if(!exists("ajl_hourly", mode="function")) source("power_paper_base.R")
+
+#FORECASTS
+if(!exists("get_forecast_report", mode="function")) source("aeso_scrapes.R")
+
+#seasons
+if(!exists("getSeason", mode="function")) source("get_season.R")
+
+
+
+#update prices and loads
+
+update<-0
+if(update==1)
+  update_forecasts()
+#re-load
+load("data/forecast_data.Rdata")
+forecast_data<-assign_date_time_days(forecast_data)
+forecast_data<-assign_peaks(forecast_data)
+
+
+#Update volumes
+
+load(file="data/metered_vols_data.Rdata" ) 
+if(update==1){
+  all_vols<-update_vols(all_vols)
+  save(all_vols, file="data/metered_vols_data.Rdata" ) 
+  }
+#Update renewable volumes as used in merit paper
+
+  renew_vols<- all_vols %>% 
+  filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,dispatched_mw=vol,pool_participant_id,effective_date_time=time)%>%
+  #switch so that we have same data format as merit data
+  mutate(import_export="",
+         block_number=1,
+         price=0,
+         from=0,
+         to=dispatched_mw,
+         size=dispatched_mw,
+         available_mw=dispatched_mw,
+         dispatched="Y",
+         flexible="N",
+         key_firm=FALSE,
+         renew_gen=dispatched_mw
+  )
+  if(update==1)
+    save(renew_vols, file="data/renew_vols.RData")  
+  
+
+
+
+#calculate peak prices
+
+peak_data<-forecast_data %>% group_by(year,month,on_peak) %>% summarize(pool_price=mean(actual_posted_pool_price,na.rm = T),
+                                                                        ail=mean(actual_ail,na.rm = T)) %>%
+  mutate(date=ymd(paste(year,month,1,sep="-"))) %>% ungroup()%>% 
+  filter(date<max(date))
+
+
+peaks<-
+  ggplot(peak_data) +
+  geom_line(aes(date,pool_price,colour=on_peak,linetype=on_peak),size=1.25)+
+  #scale_color_viridis("",option="D",discrete = TRUE)+
[…]

### 2021-06-09 aeso_scrapes.R

@@ -115,8 +115,7 @@ update_vols <- function(data_sent) {
       data_store<-rbind(data_store,clean_volume_data(xdf))
       #list_item<-list_item+1
     }
-  #testing
-  #  testing<-process_data(data_store)
+  #return the data
   rbind(data_sent,process_data(data_store))
 }
 

### 2021-06-09 merit_ghgs.R

@@ -1,336 +1,0 @@
-#Mac
-if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
-  setwd("/Users/aleach/Google Drive/alberta_power")
-#PC
-if(R.version$platform ==  "x86_64-w64-mingw32")
-  setwd("C:/Users/aleach/Google Drive/alberta_power")
-print(getwd())
-
-
-#setwd("G:/My Drive/alberta_power")
-
-
-
-source("../andrew_base.R")
-source("aeso_scrapes.R")
-source("merit_scripts.R")
-
-
-  
-
-#analysing changes in the merit order placement by emissions
-
-ghg_merit<-function(){
-  #merit_sent<-head(merit_2018,10000)
-  start_time<-Sys.time()
-  
-  load("all_merit.RData")
-  #merit_data<-merit_data %>% filter(date==ymd("2009-09-16"),he=="18")
-  
-  singles<-seq(1,9)
-  for(hour in singles){
-    merit_data$he[merit_data$he==hour]<-paste(0,hour,sep="")
-    #  merit_AS_data$he[merit_AS_data$he==hour]<-paste(0,hour,sep="")
-  }
-
-    merit_aug<-merit_data%>%
-      mutate(import_export=case_when(
-    is.na(import_export) ~ "",
-    TRUE                      ~  import_export
-  ))%>%
-  clean_merit_trade() 
-    
-  
-  
-  merit_data<-NULL #drop merit data out of memory
-  
-  print(paste("Loaded Merit Data. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
-  
-    
-  #2 Fix renewable generation so that blocks reflect actual generation, not capacity
-  
-  #isolate renewable volumes from metered volumes
-  #renew_gen<- all_vols %>% filter(year(date)>=2016) %>%
-  #  filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,vol)%>%
-  #  rename("renew_gen"="vol")
-  load("renew_vols.RData")
-  
-
-  #now, we need to take renewables out of merits, and then add in info from renewable_vols
-  
-  #checked and the effective times are odd
-  #dates<-merit_aug %>% group_by(date,he) %>% summarize()
-  
-  renew_vols <- renew_vols %>% filter(date>=min(merit_aug$date))%>%
-    mutate(
-    import_export="",
-    block_number=0,
-    price=0,
-    from=0,
-    to=renew_gen,
[…]

### 2021-06-09 power_paper_base.R

@@ -1,29 +1,64 @@
-#library(EIAdata)
-library(nleqslv)
-library(curl)
-library(janitor)
-library(viridis)
-library(scales)
-library(openxlsx)
-library(reshape2)
-library(zoo)
-library(RColorBrewer)
-library(scales) 
-library(pdfetch)
-library(tidyverse)
-library(XML)
 library(lubridate)
-library(httr)
-library(jsonlite)
 library(readxl)
-library(rvest)
-library(forcats)
-library(ggrepel)
-library(ggpubr)
+library(scales)
+library(grid)
 library(gridExtra)
+library(ggpubr)
+library(labeling)
 library(timeDate)
-library(cansim)
-library(stringi)
+library(cowplot)
+library(patchwork)
+
+library(gghighlight)
+library(viridis)
+
+library(tidyverse)
+
+
+paper_theme<-function(caption_align=1){
+  theme_minimal()+theme(
+    axis.title.x = element_text(size = 12,margin = margin(t = 2, b =2)),
+    axis.text.x = element_text(size = 12,margin = margin(t = 2, b = 2)),
+    axis.title.y = element_text(size = 12,margin = margin(r = 2, l = 2)),
+    axis.text.y = element_text(size = 12,margin = margin(r = 2, l = 2)),
+    legend.position = "bottom",
+    legend.text=element_text(size=12),
+    legend.margin=margin(c(0,0,-0.25,0),unit="cm"),
+    plot.caption = element_text(size = 9, color = "gray40",hjust=1),
+    plot.title = element_text(face = "bold"),
+    plot.subtitle = element_text(size = 10, color = "gray40"),
+    panel.grid.minor = element_blank(),
+    panel.grid.major.x = element_blank()
+  )
+}
+
+
+insertLayer <- function(P, after=0, ...) {
+  #  P     : Plot object
+  # after  : Position where to insert new layers, relative to existing layers
+  #  ...   : additional layers, separated by commas (,) instead of plus sign (+)
+  
+  if (after < 0)
+    after <- after + length(P$layers)
+  
+  if (!length(P$layers))
+    P$layers <- list(...)
+  else 
+    P$layers <- append(P$layers, list(...), after)
+  
+  return(P)
+}
+
+`-.gg` <- function(plot, layer) {
+  if (missing(layer)) {
[…]

### 2021-06-30 paper/paper_draft.tex

@@
 \newcommand{\insitu}{\textit{in situ} }
+
+\newcommand{\co}{\text{CO$_2$ }}
+\newcommand{\coe}{\text{CO$_2$e }}
+
 %%%%%%%%%%
@@
 Within natural gas generation, the mix of plant types is also relevant to our study.
-X\% are combined cycle (NGCC) plants, y\% are simple-cycle (NGSC) plants, and Z\% are combined heat and power (COGEN) facilities.
+By the end of our sample period, X\% of installed natural gas capacity comes from combined cycle (NGCC) plants, while y\% is from simple-cycle (NGSC) plants, and the remaining Z\% of capacity is from combined heat and power (COGEN) facilities.
 Combined heat and power plants tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.
@@
 
-Alberta's three carbon pricing regimes affect generating facilities in
+Alberta's three carbon pricing regimes affect generating facilities in two ways.
+First, each of three (\emph{SGER}, \emph{CCIR}, and \emph{TIER}) regimes imposes a price on emissions.
+Second, each regime allocates emissions credits based on output or deemed avoided emissions.
+The net effect of these policies creates variation within and across facility types over time.
+Consider, for example, Figure \ref{fig:sector_compliance_cost} which shows the mean value and range of annual compliance costs for larger generators in Alberta's fleet over the sample period.\footnote{Plants were only included in the sample for this figure if their annual emissions were sufficiently large ($\geq$100kt CO$_2$e per year) such that we have specific compliance cost data.
+For our estimate, we impute compliance costs for a wider set of facilities using emissions intensities from the compliance cost data.}
+
+\begin{figure}[!h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../SGER/sector_compliance_cost.png}
+\vspace{-0.75cm}	\caption{Annual average revenue by plant type}
+\label{fig:sector_compliance_cost}
+\end{figure}
+
+As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices, but are actually dominated by the changes in output-based allocations which occurred in 2018-2019 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe between 2016 and 2018.
 
@@
 \label{fig:masnadi_1}
-\vspace{-0.75cm}	\caption{Impact of changes in carbon pricing on plant revenues by type}
+\vspace{-0.75cm}	\caption{Peak and off-peak power prices, 2007-2019}
 \end{figure}

### 2021-07-08 paper/paper_draft.tex

@@
 We supplement these data with a variety of other information.
-Most importantly, we add facility-level information including fuel source, ownership, nameplate capacity, heat rate, and estimated emissions-intensity based on heat rates and fuel sources.
-ADD ON-LINE APPENDIX TABLE OF PLANT INFO.
+Most importantly, we add facility-level compliance data from Alberta's greenhouse gas emissions pricing policies, the SGER (2009-2017) and the CCIR (2018-2019).
+These data were provided for this analysis by the Government of Alberta, with some portions of the data also available publicly.
+We supplement these data with information from the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for smaller facilities.\footnote{The federal data are useful since the reporting threshold of 10kt \coe per annum is lower than the provincial threshold for coverage under the emissions pricing policies of 100kt\coe per annum.}  Since these data are both reported at the facility level, not by generating unit, we apply these measurements in the same way in our paper and so we cannot differentiate between multiple generating units on particular sites.\footnote{There are two exceptions to this.
+The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.}  Since we have only annual data, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units.
 
-We also use compliance data from the Alberta Specified Gas Emitters Regulations and the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for facilities.
-Since these data are both reported at the facility level, not by generating unit, we apply these measurements in the same way in our data.
-As such, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility in a particular hour nor do we account for any ramping rates of units.
-
-We add weather data from Environment Canada weather stations in Edmonton, Fort McMurray, Cold Lake, and Calgary, covering the three major demand centers in Alberta.
+We add weather data from Environment Canada weather stations in Edmonton, Fort McMurray, Cold Lake, and Calgary, covering the major demand centers in Alberta.
 We take the average of available measurements in each area in order to maximize the number of hours we can cover with weather data.
@@
 There is also no reason to believe that weather data availability is correlated with our variables of interest, and so the lost observations should have no leverage on the results.
+We convert temperature data to heating and cooling degree days for each region, which we then use in our model.
 
-Finally, we merge daily natural gas prices for the Nova Inventory Transfer hub.  We access daily spot prices via NRGStream, a commercial data aggregation service.
+Finally, we supplement our publicly-available data with several series from NRGStream, a commercial data aggregation service.
+Most importantly, this service provides daily natural gas prices for the Nova Inventory Transfer hub which are a better proxy for Alberta natural gas prices than publicly-available series for Henry Hub.
+NRGStream also scrapes the real-time data from the AESO, and thus provides slightly different data from the aggregate, publicly-available historic metered volumes provided by the grid operator.
+We use NRGStream real-time data to compile real-time intertie flows between Alberta and British Columbia, Saskatchewan, and Montana and to source real-time generation (as opposed to net-to-grid metered volumes) from all generating units in the province where needed.
 
-All data save the natural gas prices which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK.
+All data save the natural gas prices and NRGStream generation and trade date which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK.
 For the publicly available code, we substitute Henry Hub gas prices as a placeholder, to which we apply the average discount observed between Henry Hub and NIT to adjust the values so that regression results may be more closely replicated.
@@
 
-Using our dataset, we can provide a statistical portrait of Alberta's power market.
+A statistical portrait of Alberta's power market provides context for our study.
 Alberta's wholesale market is a single price, energy-only market.
 There is no day-ahead energy market, but there is a separate ancillary services market which we do not address beyond a couple of tangential references in this paper.
-Alberta's record internal load, reached on February 9, 2021, is 11,729MW, so Alberta's is a relatively small power market compared to the similarly-structure Energy Reliability Council of Texas (ERCOT) market, for example, which had a 58,597 MW peak 2021 load.\footnote{The AESO previous record, reached on January 11, 2018, was 11,697 MW, so peak loads have been quite stable in recent years after a period of rapid growth.} Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_load}, with X cumulative annual average load growth rate during our sample period, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
-There was also a major short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, which explains the markedly-lower spring trough in that year shown in Figure \ref{fig:ab_load}.
-The 2020 COVID-19 demand shock is also evident in this graph, but lies outside the sample period for our analysis.
+Alberta's record internal load, reached on February 9, 2021, is 11,729MW, so Alberta's is a relatively small power market compared to the similarly-structured Energy Reliability Council of Texas (ERCOT) market, for example, which had a 58,597 MW peak 2021 load.
+Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_load}, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
+There was also a major, short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, which explains the markedly-lower spring trough in that year's load profile as shown in Figure \ref{fig:ab_load}.
+The 2020 COVID-19 demand shock is also evident in Figure \ref{fig:ab_load}, but lies outside the sample period for our analysis.
 
@@
 These conditions are reflected in peak and off-peak prices shown in Figure \ref{fig:ab_prices}.
-Some of these market changes coincide with policy changes in our experiment - in particular, the changes to the \emph{SGER} introduced in June of 2015 which took effect in January of 2016 and 2017 respectively coincide with a period of over-supply, while the introduction of the \emph{CCIR} in 2018 occurred during a period of tightening reserve margins and lingering uncertainty over market structure.
+Some of these market changes coincide with policy changes of interest in our paper, in particular the changes to the \emph{SGER} introduced in June of 2015 which took effect in January of 2016 and 2017 respectively and are coincidental with a period of over-supply, and the introduction of the \emph{CCIR} in 2018 which occurred during a period of tightening reserve margins and lingering uncertainty over market structure.
 From 2018 through pre-COVID 2020, both GHG policy and the wholesale electricity market were tightening, and so we need to be careful to disentangle these impacts.
@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices.png}
-\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours. Peak hours are 8am to 11pm other than on statutory holidays or Sundays. Source: Alberta Electric System Operator data.}
+\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours. Peak hours are 8am to 11pm other than on statutory holidays or Sundays. Source: AESO data, authors' graph.}
 \label{fig:ab_prices}
@@
 \begin{figure}[t]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel.png}
-\vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: Alberta Electric System Operator data.}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_area_grey.png}
+\vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: AESO data provided by NRGStream, authors' graph.}
 \label{fig:gen_mix}
@@
 Within natural gas generation, the mix of plant types is also relevant to our study.
-By the end of our sample period, X\% of installed natural gas capacity comes from combined cycle (NGCC) plants, while y\% is from simple-cycle (NGSC) plants, and the remaining Z\% of capacity is from combined heat and power (COGEN) facilities.
-Combined heat and power plants tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.
-As will be discussed below, net-to-grid power from these facilities is primarily offered into the market at a \$0 offer and accepts market price, and so we see limited evidence of carbon price pass-through for most of the power offered by these facilities.
-
-ADJUST FIGURE TO SHOW SAMPLE PERIOD
+By the end of our sample period, a significant increase in generating capacity from combined cycle (NGCC) plants had been installed, and this is reflected in higher generation from those sources.
+The installed capacity of combined heat and power (COGEN) facilities also grew during the sample period, and these plants tend to operate with high capacity factors but also tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.
+As will be discussed below, net-to-grid power from COGEN facilities is primarily offered into the market at a \$0 offer and accepts market price, and so we see limited evidence of carbon price pass-through for most of the power offered by these facilities.
+On the other hand, peaking capacity in the market during our sample period was largely met through ramping of coal and combined-cycle plants, as well as simple cycle (SCGT) natural gas turbines.
+Each of these would be exposed to carbon pricing, and so we expect to see pass through affect offers of power from these facilities into the market.
 
@@
 
-The nature of the wholesale market is such that different generators will derive different levels of average revenue from the sale of their power.
+The nature of Alberta's wholesale market is such that different generators will derive different levels of average revenue from the sale of their power.
 Those that dispatch only in high-price times will have higher revenues on average than plants without dispatch control, or plants which run most hours of the day.
@@
 Lower prices tend to prevail when wind generation is highest, and so captured prices for wind power facilities are lower than other forms of generation, and exports tend to be scheduled when prices are expected to be low.
-Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewable sources since its generation pattern coincides with daily and seasonal peak prices, although we would expect this to change as the \emph{duck curve} effects take hold over time.
+Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewable sources since its generation pattern coincides with daily and seasonal peak prices, although we would expect this to change as \emph{duck curve} effects take hold over time.
 
-\begin{figure}[!h]%%
+\begin{figure}[t]%%
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture.png}
-\vspace{-0.75cm}	\caption{Annual average revenue by plant type}
+\vspace{-0.75cm}	\caption{Annual average revenue by plant type. source: AESO data provided by NRGStream, authors' graph.}
 \label{fig:capture_prices}
@@
 
-Alberta's three carbon pricing regimes affect generating facilities in two ways.
+Alberta's three carbon pricing regimes each affected generating facilities in two ways.
 First, each of three (\emph{SGER}, \emph{CCIR}, and \emph{TIER}) regimes imposes a price on emissions.
 Second, each regime allocates emissions credits based on output or deemed avoided emissions.
-The net effect of these policies creates variation within and across facility types over time.
+The net effect of these policies creates variation within and across facilities and facility types and over time.
 Consider, for example, Figure \ref{fig:sector_compliance_cost} which shows the mean value and range of annual compliance costs for larger generators in Alberta's fleet over the sample period.\footnote{Plants were only included in the sample for this figure if their annual emissions were sufficiently large ($\geq$100kt CO$_2$e per year) such that we have specific compliance cost data.
-For our estimate, we impute compliance costs for a wider set of facilities using emissions intensities from the compliance cost data.}
+For our estimation results, we impute compliance costs for a wider set of facilities using emissions intensities from the compliance cost data.}
 
-\begin{figure}[!h]%%
+\begin{figure}[t]%%
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../SGER/sector_compliance_cost.png}
-\vspace{-0.75cm}	\caption{Annual average revenue by plant type}
+\vspace{-0.75cm}	\caption{Upper: Annual average compliance cost (line) and range (shaded area) by plant type.
+Lower: Price charged on excess net annual emissions under Alberta's carbon pricing programs.
+Source: Alberta Environment compliance cost data}
 \label{fig:sector_compliance_cost}
@@
 As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices, but are actually dominated by the changes in output-based allocations which occurred in 2018-2019 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe between 2016 and 2018.
+The largest cost increases affect coal-fired power plants and some simple-cycle generators, while have limited effects on combined heat and power plants.
+Compliance costs for natural gas combined-cycle plants dropped between 2017 and 2019 as a result of a change in the formula by which output-based allocations were calculated.
 
@@
 coal or combined-cycle natural gas plants) may be of different sizes (e.g.
-a 350MW plant vs a 200MW plant) while facing similar optimization constraints, and they may be jointly dispatched by individual operators.
+a 350MW plant vs a 200MW plant) while facing similar optimization constraints, and they may be jointly dispatched by the same operator.
+In particular for coal plants, plant-level minimum-must-run constraints imply that facilities will always offer some portion of their generation into the market at a \$0/MWh price, effectively acting as price takers for that hour for that block of power.
+They will offer marginal blocks of power, for capacity above their minimum-must-run level, at higher prices.
+However, given that plants are of different sizes, and blocks are endogenously determined, it's challenging to identify a change in offers looking only at block offers.
 The second issue is that the strategy space for each facility is very large, given that they may decide both on the break-points for each of up to 7 blocks of power, and then decide on a price for the first block and an adder for each subsequent block offered into the market.
-For each hour, then, the optimization problem consists of 12 choice variables and we have data on these choices for over 18.5 million facility-hour-block pairs.
-We render our problem more parsimonious, and at the same time overcome the portfolio bidding issue, both how the overall merit order and how power from a particular type of plant has been offered into the market over time.
+For each hour, then, the optimization problem consists of 12 choice variables and we have data on these choices for almost 20 million facility-hour-block pairs.
+Finally, since multiple plants may be controlled by a single entity, the observed data are the product of a complex portfolio optimization problem and the offers of any individual plant may not reveal an overall portfolio-level attempt to pass through the costs of carbon prices to wholesale prices.
+We avoid these problems by constructing synthetic offer curves either by plant-type or by controlling entity, thus estimating the impact of policy changes on the supply behaviour of the aggregate fleet rather than on any individual plant.
+We can then use the synthetic merit order data we construct to examine how power has been offered into the system over time as climate policies have varied.
 
-This is best understood with a series of figures, and for this we will again use one coal plant (Capital Power's Genessee 3 facility) as well as Alberta's overall coal fleet to demonstrate the approach we take here.
+Our empirical strategy is best understood visually through a series of figures.
+Consider Figure \ref{fig:synth_merit}.
+In the left-hand panel, we show the merit order for a particular hour for all of the coal facilities in the province, with offers from multi-unit facilities combined for ease of visualization.
+We see facilities bidding some of their power at \$0, ensuring they are in the market, with marginal generation blocks offered at increasing marginal prices.
+In this particular hour, the price was such that most of the offered coal generation was dispatched, with only blocks offered above the \$802/MWh market clearing price not being dispatched.
+In the right hand panel, we show our synthetic merit order which summarizes all of the coal power offered into the market in a given hour using the 15 points shown in the Figure, which correspond to the value of the merit order at prescribed percentiles of the total offered power in that hour.\footnote{We store the 10th through 40th percentiles in 10\% intervals and the 50th percentile and above in 5\% intervals to better capture the curvature of the merit order.} Summarizing the data this way reduces our stored information by a factor of three, but more importantly it allows us to answer the relevant question in our study - by how much did the supply curve shift in response to changes in the value of carbon pricing costs and output-based allocation revenue - by looking at changes in the value of our sampling points across the portfolio of offered power.
+Or, visually, we want to ask how, on average, each of the points in the right-hand panel of Figure \ref{fig:synth_merit} change when carbon prices are increased or when output-based allocations of emissions credits change.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{../images/coal_merit.png}
+        %\caption{$dt=0.1$}
+        %\label{fig:prob1_6_2}
+    \end{minipage}%
+    \begin{minipage}{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{../images/coal_synth_merit}
+        %\caption{$dt =$}
+        %\label{fig:prob1_6_1}
+    \end{minipage}
+    \caption{Observed coal facility merit order (left panel) and synthetic coal merit order (right panel) for February 4, 2019 at 7pm.
+For each hour, we convert the empirical merit order to a step function, and then extract the y-intercept as at 14 unique percentiles of total offered power in that hour, using 5\% increments in the upper half of the merit order.}
+    \label{fig:synth_merit}
+\end{figure}
+
+There is a lot of day-to-day and hour-to-hour variability in how coal are offered into the market.
+Figure \ref{fig:synth_coal_2019}, which shows all of the 7pm merit orders (left panel), their synthetic equivalents, and the average synthetic merit order for the last year of our sample, provides a sense of the degree of variability in our sample.
+Similar variability exists in the offers of generation from other dispatchable sources.\footnote{See online graphical appendix, Figure A1, for a similar Figure for all dispatchable generating source types.} Figure \ref{fig:synth_coal_2019} shows the degree to which changes in total offered capacity would skew our results if we attempted to base our analysis on offered megawatts.
+Instead, ny using percentiles of offered capacity as our units of measure, we can more cleanly ask how different segments of the supply curve shift vertically in response to different events, while implicitly assuming that the total offered power available in the market in a given hour is independent of the carbon pricing policy on offer at the time.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{../images/all_coal_merit.png}
+        %\caption{$dt=0.1$}
+        %\label{fig:prob1_6_2}
+    \end{minipage}%
+    \begin{minipage}{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{../images/coal_synth_2019}
+        %\caption{$dt =$}
+        %\label{fig:prob1_6_1}
+    \end{minipage}
+    \caption{Observed coal facility merit order (left panel) and synthetic coal merit order (right panel) for all of the 7pm hours of 2019.}
+    \label{fig:synth_coal_2019}
+\end{figure}
+
+The next stage in our analysis is to transfer our information on facility compliance costs to the synthetic merit orders.
+Recall that, for the major emitters, we have specific facility-level information on their emissions-intensity their output-based allocations of emissions credits which determines their net emissions pricing liabilities.
+Since we are interested in the separate effects of both components of emissions policies, and since both vary at the facility level through our sample, we can separately identify the impacts of each on price pass-through behaviour.
+As such, for each of the synthetic merit orders we create for analysis purposes, we also capture and store both the marginal emissions charge and the marginal output-based allocation at each percentile support point.
+
+For our analysis sample, we create three other types of synthetic merit orders.
+First, for very general results, we characterize the entire merit order using the methodology outlined above.
+We also use the fact that, from 2013 through 2019, the AESO provides offer control by unit as part of their merit order releases.
+We use this information to construct synthetic supply curves for each of the major participants in the Alberta power market.
+This means that we have facility-level, generator-class-level and market-level synthetic merit orders to analyze.
+
+To these, synthetic merit orders, we add our hourly-level data on market status, weather, and fuel prices specified above.
+This then allows us to build regression equations for each percentile level of each synthetic merit order, asking the degree to which carbon price and output-based allocation changes impact the level of each step of the supply curves by facility, plant-type, by owner, or for the complete market.
 
@@
 
-\section{Analysis and Results}
+\subsection{Facility-level impacts}
+
+\subsection{Impacts By Generator Type}
+
+\subsection{Impacts By Generator Type}
+
+\subsection{Market-wide impacts}
 

### 2021-07-22 paper/paper_draft.tex
data redo complete

@@
 \newcommand{\insitu}{\textit{in situ} }
+\newcommand{\SGER}{\textit{SGER} }
+\newcommand{\CCIR}{\textit{CCIR} }
 
@@
 We supplement these data with a variety of other information.
-Most importantly, we add facility-level compliance data from Alberta's greenhouse gas emissions pricing policies, the SGER (2009-2017) and the CCIR (2018-2019).
+Most importantly, we add facility-level compliance data from Alberta's greenhouse gas emissions pricing policies, the \SGER (2009-2017) and the \CCIR (2018-2019).
 These data were provided for this analysis by the Government of Alberta, with some portions of the data also available publicly.
-We supplement these data with information from the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for smaller facilities.\footnote{The federal data are useful since the reporting threshold of 10kt \coe per annum is lower than the provincial threshold for coverage under the emissions pricing policies of 100kt\coe per annum.}  Since these data are both reported at the facility level, not by generating unit, we apply these measurements in the same way in our paper and so we cannot differentiate between multiple generating units on particular sites.\footnote{There are two exceptions to this.
-The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.}  Since we have only annual data, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units.
+We supplement these data with information from the Federal Greenhouse Gas Emissions Reporting Program to provide, where available, more precise emissions intensity data for smaller facilities.\footnote{The federal data are useful since the reporting threshold of 10kt \coe per annum is lower than the provincial threshold for coverage under the emissions pricing policies of 100kt\coe per annum.}
+
+There are three sources of incompleteness in our emissions data.
+First, we have no emissions data for smaller facilities, since provincial and federal reporting is not required below a minimum threshold.
+We have data on all facilities that fell under Alberta's industrial carbon pricing programs, but for smaller facilities which would have been subject only to the economy-wide carbon pricing program imposed in 2018, we have limited information.
+Federal reporting thresholds decreased to 10,000 tonnes per year which translates to an exemption from reporting for most natural gas generators with installed capacity of less than 20MW.
+Because reporting thresholds have changed over time, in some cases we have partial information on emissions intensities for smaller facilities, while in other cases we have no information at all.
+We use emissions-intensities from provincial reporting data as our first option, since the provincial reporting includes both verified production and emissions.
+Where we don't have provincial data, we use generation data aggregated by facility and federally-reported emissions to calculate an emissions intensity.
+Where we have partial information, we complete our dataset by filling first backward in time and then forward in time by plant by year such that, for example, a plant for which we only have 2019 data would see their 2019 emissions intensity used for every previous year.
+Where we have no information at all, we assume an emissions-intensity of 0.55 tonnes per MWh for simple cycle natural gas plants, the only type of plant for which we lack information of this type, based on facilities for which we do have information in our sample.
+The second source of incompleteness comes from the fact that data are reported at the facility level, not by generating unit.
+We apply the same emissions intensity to multiple generating units within a facility.\footnote{There are two exceptions to this.
+The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.} Since we have only annual data at the facility level, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units.
+Finally, combined heat and power units present a challenge since many units report their emissions for the entire production facility, not simply for the power plant.
+There are some stand-alone units, but these also vary in terms of the emissions-intensity of net-to-grid electricity depending on the design of the specific unit and what share of the produced heat is used to generate electricity vs process steam.
+The emissions reporting data tell us the total electricity and heat produced in these facilities, and the emissions attributed to them, but do not tell us what the emissions intensity of an incremental unit of electricity would be.
+The output-based allocations for cogeneration facilities are determined by the carbon pricing policies, so we know the value of these for each regulated entity.
+For the purposes of our analysis, we adopt an assumed emissions intensity of electricity of 0.418 tonnes per MWh for combined heat and power units.
+Biomass plants, as well as wind, solar and hydroelectric plants are assigned a deemed emissions intensity of zero.
+
+The compliance data also identifies two other attributes of importance to our analysis.
+First, under the \SGER, facilities only faced compliance costs after their third full year of operations.
+Because facilities had to report federally during these periods, we can identify emissions intensities for some large facilities for 3 or 4 years before they face carbon pricing, and then for the remaining years in the sample when they are subject to carbon pricing.
+This is confounded slightly by the fact that the output-based allocations in future years are a function of the emissions intensity in the first years of operation.
+We account for this with an indicator variable in our analysis for the pre-compliance period.
+The data also provide the facility-specific rate at which output-based allocations were provided in each year, and the carbon price faced on emissions net of output-based allocations.
+The carbon price is constant for all Alberta-based facilities, but the combined-cycle plant in Fort Nelson, BC faces the BC carbon tax while being connected to the Alberta electricity grid, and so receives no output-based allocations and faced higher carbon prices in many years of the sample than comparable plants in Alberta.
 
@@
 We use NRGStream real-time data to compile real-time intertie flows between Alberta and British Columbia, Saskatchewan, and Montana and to source real-time generation (as opposed to net-to-grid metered volumes) from all generating units in the province where needed.
+These data are generally the same as metered volumes with the exception of cases where an industrial production facility is co-located with the power generation facility, in which case only the net-to-grid volumes from the facility are measured in the metered volumes data while the NRGStream scraped data captures actual production by hour for the generating facility.
 
-All data save the natural gas prices and NRGStream generation and trade date which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK.
+All data save the natural gas prices and NRGStream generation and trade data which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK.
 For the publicly available code, we substitute Henry Hub gas prices as a placeholder, to which we apply the average discount observed between Henry Hub and NIT to adjust the values so that regression results may be more closely replicated.
@@
 
-As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices, but are actually dominated by the changes in output-based allocations which occurred in 2018-2019 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe between 2016 and 2018.
-The largest cost increases affect coal-fired power plants and some simple-cycle generators, while have limited effects on combined heat and power plants.
+As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices, but are actually dominated by the changes in output-based allocations which occurred subsequent to the adoption of the CCIR in 2018 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe in 2016 and 2017.
+The largest cost increases affect coal-fired power plants and some simple-cycle generators, while limited effects were felt by combined heat and power plants.
 Compliance costs for natural gas combined-cycle plants dropped between 2017 and 2019 as a result of a change in the formula by which output-based allocations were calculated.
@@
 The next stage in our analysis is to transfer our information on facility compliance costs to the synthetic merit orders.
-Recall that, for the major emitters, we have specific facility-level information on their emissions-intensity their output-based allocations of emissions credits which determines their net emissions pricing liabilities.
+Recall that, for the major emitters, we have facility-year-specific information on their emissions-intensity and their output-based allocations of emissions credits which determines their net emissions pricing liabilities.
 Since we are interested in the separate effects of both components of emissions policies, and since both vary at the facility level through our sample, we can separately identify the impacts of each on price pass-through behaviour.
@@
 
-To these, synthetic merit orders, we add our hourly-level data on market status, weather, and fuel prices specified above.
-This then allows us to build regression equations for each percentile level of each synthetic merit order, asking the degree to which carbon price and output-based allocation changes impact the level of each step of the supply curves by facility, plant-type, by owner, or for the complete market.
+To these, synthetic merit orders, we add our hourly data on market status, weather, and fuel prices specified above.
+This then allows us to build regression equations for each percentile support point of each synthetic merit order, asking the degree to which carbon price and output-based allocation changes impact the level of each step of the supply curves by facility, plant-type, by owner, or for the complete market.
 

### 2021-07-22 aeso_scrapes.R

@@ -119,9 +119,6 @@ update_vols <- function(data_sent) {
   rbind(data_sent,process_data(data_store))
 }
 
-
-
-
 process_data <- function(data_sent) {
   #function to process all AESO data into useful load and trade volumes
   #testing

### 2021-07-22 merit_build.R

@@ -10,7 +10,7 @@ start_time<-Sys.time()
 options(scipen=999)
 
 update<-0 #add new data
-save<-0 #save files at the end
+save<-1 #save files at the end
 synth<-0 #synthetic plants?
   synth_type<-1 # 2 is by facility,1 is by plant_type, 0 is full merit as synthetic plant
 
@@ -74,9 +74,10 @@ load("data/forecast_data.RData")
   if(update==1)
    {
    load(file="data/metered_vols_data.Rdata" ) 
-    update_vols(all_vols)
+    #all_vols<-all_vols %>% filter(year<2019)
+    all_vols<-update_vols(all_vols)
     save(all_vols,file="data/metered_vols_data.Rdata" ) 
-    #isolate renewable volumes from metered volumes
+    #isolate renewable (non-hydro and biomass) volumes from metered volumes - the ones that default bid to zero
     renew_vols<- all_vols %>% 
     filter(Plant_Type=="WIND"| Plant_Type=="SOLAR") %>% select(date,he,asset_id,dispatched_mw=vol,pool_participant_id,effective_date_time=time)%>%
     #switch so that we have same data format as merit data
@@ -145,34 +146,15 @@ load("data/forecast_data.RData")
 #merit_aug<-merit_small
   
 #merit_aug<-merit_store  
+#merit_aug<-merit_small
 
-  
-#get sger info 
-load("data/emissions_data_merge.Rdata")
   
 merit_aug<-merit_aug %>% mutate(year=year(date))%>%
-  left_join(plant_data(),by=c("asset_id"="ID")) %>% #here, co2_est is in kg/MWh
-  #left_join(sger_emissions_data(),by=c("asset_id"))
-  left_join(sger_data%>%select(-co2_est),by=c("year","asset_id"))
-  
- 
-## merge_test<-merit_small %>% 
-#left_join(plant_data(),by=c("asset_id"="ID")) %>% #here, co2_est is in kg/MWh
-#  #left_join(sger_emissions_data(),by=c("asset_id"))
-#  left_join(sger_data,by=c("year","asset_id"))
-
- 
- 
- #sger data is t/mwh
-
-  #build sger allocations for 2010-2015,2016,2017
-  #allocations of zero and carbon tax of zero for pre-2017 firms outside of sger
-  #allocations of 0.37 for all facilities in 2018 and 2019
-  
-  merit_aug<-merit_aug %>% #filter(year(date)>=2012) %>%
[…]

### 2021-07-22 merit_scripts.R

@@ -74,6 +74,12 @@ ngx_data_read<-function(){
 }
 
 
+ghg_data<-function(){
+  load(file = "data/plant_data_merge.RData")
+  paper_data %>% select(ID,AESO_Name,Plant_Type,Plant_Fuel,Capacity,year,ei,oba_rate,ctax,sger_first_year,compliance_cost)
+}
+
+
 
 sger_emissions_data<-function(){
   keep_columns<-
@@ -180,30 +186,13 @@ plant_data<-function(){
   plant_info<-arrange(plant_info,NRG_Stream)
   
   #bring in ghg data
-  ghg_rates <- read.xlsx(xlsxFile = "data/AB_Plant_Info_New.xlsx", sheet = "GHG_Rates", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE)
-  #ghg_rates<-dcast(ghg_rates, formula = GHG_ID ~ ...,value.var = "Poln_rate")
-  ghg_rates<-ghg_rates %>% spread(Pollutant,Poln_rate) %>% select(GHG_ID,CO2)
+  #get sger info 
+  load("data/emissions_data_merge.Rdata")
+  plant_info<-plant_info%>% left_join(sger_data,by=c("ID"="asset_id"))
   
   
-  #bring in heat rates
-  heat_rates <- read.xlsx(xlsxFile = "data/AB_Plant_Info_New.xlsx", sheet = "Heat_Rates", startRow = 1,skipEmptyRows = TRUE,detectDates = TRUE) %>%
-    select(GHG_ID,Aurora_ID,Heat.Rate)
-  #combine all plant info, heat rates, and GHGs by plant ID
-  combined<-merge(ghg_rates,heat_rates,by="GHG_ID",all.y = T) # NA's match
-  coal_co2_btu<-100.4  #coal fuel factor GHGs/MMBTU
-  gas_co2_btu<-53.077752 #gas fuel factor GHGs/MMBTU
-  combined<-merge(plant_info,combined,suffixes = c(".info",".rates"),all.x=TRUE,by="Aurora_ID") # NA's match
-  combined$co2_est<-combined$CO2/2.20462*combined$Heat.Rate #convert to kg/mmbtu
-  combined$co2_est<-ifelse(combined$Plant_Fuel=="COAL",coal_co2_btu*combined$Heat.Rate,combined$co2_est)
[…]

### 2021-07-31 paper/paper_draft.tex
updates

@@
 There is a lot of day-to-day and hour-to-hour variability in how coal are offered into the market.
-Figure \ref{fig:synth_coal_2019}, which shows all of the 7pm merit orders (left panel), their synthetic equivalents, and the average synthetic merit order for the last year of our sample, provides a sense of the degree of variability in our sample.
+Figure \ref{fig:synth_coal_2019}, which shows all of the 7pm merit orders (left panel) and their synthetic equivalents, along with the average synthetic merit order for the last year of our sample (right panel), provides a sense of the variability in our sample.
 Similar variability exists in the offers of generation from other dispatchable sources.\footnote{See online graphical appendix, Figure A1, for a similar Figure for all dispatchable generating source types.} Figure \ref{fig:synth_coal_2019} shows the degree to which changes in total offered capacity would skew our results if we attempted to base our analysis on offered megawatts.
-Instead, ny using percentiles of offered capacity as our units of measure, we can more cleanly ask how different segments of the supply curve shift vertically in response to different events, while implicitly assuming that the total offered power available in the market in a given hour is independent of the carbon pricing policy on offer at the time.
+Instead, by using percentiles of offered capacity as our units of measure, we can more cleanly ask how different segments of the supply curve shift vertically in response to different events, while implicitly assuming that the total offered power available in the market in a given hour is independent of the carbon pricing policy on offer at the time.
 
@@
 
-\subsection{Facility-level impacts}
+\subsection{Market-wide impacts}
+The most general results derive from the analysis of the market as a whole.
+Here, we are calculating for each hour of our sample the supply curve of power offered into the market which we summarize using a step function with 15 points of support mapping offers at given percentiles of total power offers.
+At each of these points of support, we also track the marginal carbon pricing cost and marginal output-based allocation value along with other market and environmental variables.
+We find, as shown in Table X that...
 
 \subsection{Impacts By Generator Type}
+We know that different generator types tend to offer their power into the market in different ways, and so our most detailed results are based on a decomposition of the market-wide impacts into impacts by type of plant.
+Here, we find that...
 
-\subsection{Impacts By Generator Type}
+\subsection{Impacts By Offer Control}
+We can also use the data post-2013 to assess the degree to which different market participants might offer their power into the merit order and how they have passed through carbon policy values to power offers.
+We only use data after 2013 for this analysis because, prior to that time, the AESO did not provide offer control records along with their merit order data.
+The remains a sufficient sample, since we have three full years of data (2013-2015) in the initial \SGER pricing program, along with at least one full year under each of the altered policy regimes over which to assess changed in offers by different market participants.
 
-\subsection{Market-wide impacts}
+PPA explanation...
+
+Balancing Pool explanation...
+
+Return of PPAs to the Balancing Pool
+
+Declining ability to identify TransCanada impacts because of changes in resources.
 

### 2021-07-31 merit_build.R

@@ -11,8 +11,8 @@ options(scipen=999)
 
 update<-0 #add new data
 save<-1 #save files at the end
-synth<-0 #synthetic plants?
-  synth_type<-1 # 2 is by facility,1 is by plant_type, 0 is full merit as synthetic plant
+synth<-1 #synthetic plants?
+  synth_type<-4  #4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_type, 0 is full merit as synthetic plant
 
 
 load("data/all_merit.RData")  
@@ -110,15 +110,15 @@ load("data/forecast_data.RData")
     summarize(pool_participant_name=last(pool_participant_name))
   renew_vols<-renew_vols %>% left_join(aeso_assets,by="pool_participant_id")%>%
     rename(offer_control=pool_participant_name)%>% select(-pool_participant_id)
-  renew_vols<-renew_vols %>% mutate(key_firm=case_when(
-   grepl("TransAlta",offer_control)~TRUE,
-   grepl("TransCanada",offer_control)~TRUE,
-   grepl("ENMAX",offer_control)~TRUE,
-   grepl("Capital Power",offer_control)~TRUE,
-   grepl("ATCO",offer_control)~TRUE,
-   grepl("Balancing Pool",offer_control)~TRUE,
-   TRUE~FALSE #if it's not one of these, it's false
- ))
+  renew_vols<-renew_vols %>% mutate(offer_sum=case_when(
+   grepl("TransAlta",offer_control)~"TransAlta",
+   grepl("TransCanada",offer_control)~"TransCanada",
+   grepl("ENMAX",offer_control)~"ENMAX",
+   grepl("Capital Power",offer_control)~'Capital Power',
+   grepl("ATCO",offer_control)~"ATCO",
+   grepl("Balancing Pool",offer_control)~"Balancing Pool",
+   TRUE~"Other" #if it's not one of these, it's false
+ ),key_firm=(offer_sum!="Other"))
  
 
   #storage objects for testing purposes
@@ -149,25 +149,39 @@ load("data/forecast_data.RData")
 #merit_aug<-merit_small
 
   
-merit_aug<-merit_aug %>% mutate(year=year(date))%>%
-  left_join(ghg_data(),by=c("year","asset_id"="ID"))%>%
+merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
+  mutate(year=year(date))%>%
[…]

### 2021-07-31 merit_scripts.R

@@ -76,7 +76,7 @@ ngx_data_read<-function(){
 
 ghg_data<-function(){
   load(file = "data/plant_data_merge.RData")
-  paper_data %>% select(ID,AESO_Name,Plant_Type,Plant_Fuel,Capacity,year,ei,oba_rate,ctax,sger_first_year,compliance_cost)
+  paper_data %>% select(ID,year,ei,oba_rate,ctax,sger_first_year,compliance_cost)
 }
 
 
@@ -185,17 +185,12 @@ plant_data<-function(){
     mutate(NRG_Stream = ifelse(grepl("AB - NPC1 Denis St Pierre Hr Avg MW",NRG_Stream),"AB - NPC1 Denis St  Pierre Hr Avg MW",NRG_Stream))  
   plant_info<-arrange(plant_info,NRG_Stream)
   
-  #bring in ghg data
-  #get sger info 
-  load("data/emissions_data_merge.Rdata")
-  plant_info<-plant_info%>% left_join(sger_data,by=c("ID"="asset_id"))
-  
   
   plant_info %>%
     select(-Latitude,-Longitude,-Aurora_Name,-NRG_Stream,-Aurora_ID)
   
-  
 }
+
 
 
 

### 2021-07-31 paper_graphs.R

@@ -363,6 +363,7 @@ gen_plain <- df2 %>% mutate(Plant_Type=factor(Plant_Type,levels=AB_plant_order))
     )%>% filter(file_date==max(file_date))
   
   load(file = paste("data/",bids_files$value,sep=""))
+  paste("loaded file=data/",bids_files$value,sep="")
   merit_bids<-merit_aug
   
   
@@ -373,8 +374,8 @@ data_files<- list.files("data/") %>% as_tibble() %>% filter(grepl("merit_data",v
   mutate(file_date=gsub("merit_data_","",value),
          file_date=ymd_hm(gsub(".RData","",file_date))
          )%>% filter(file_date==max(file_date))
-
 load(file = paste("data/",data_files$value,sep=""))
+paste("loaded file=data/",data_files$value,sep="")
 
 #merit_aug<-merit_aug %>% filter(year==2019,Plant_Type=="COAL")%>% filter(actual_posted_pool_price>200)
 
@@ -498,10 +499,11 @@ merit_bids %>% filter(date==ymd("2019-02-04"),he=="19",Plant_Type=="COAL")%>% re
   scale_fill_grey("",end = .5,start=.5)+   
   scale_x_continuous(breaks=pretty_breaks(), expand=c(0,0))+
   scale_y_continuous(breaks=pretty_breaks(), expand=c(0,0))+
+  expand_limits(y=1000)+
   #scale_colour_manual(labe ls=c("Brent","WTI"),values=c("#41ae76","#238b45","#006d2c","#00441b","Black","Black","Black","Black"))
   paper_theme()+
   theme(
-    legend.position = c(.1,.8),
+    legend.position = c(.2,.9),
     legend.margin=margin(c(0,0,0,0),unit="cm"),
     legend.text = element_text(colour="black", size = 12),
     plot.caption = element_text(size = 16, face = "italic"),

### 2021-07-31 regression_analysis.R

@@ -1,0 +1,1173 @@
+library(skimr)
+library(broom)
+library(purrr)
+library(margins)
+source("power_paper_base.R")
+
+#merit all
+
+bids_files<- list.files("data/") %>% as_tibble() %>% filter(grepl("synth_all",value))%>% 
+  mutate(file_date=gsub("synth_all_","",value),
+         file_date=ymd_hm(gsub(".RData","",file_date))
+  )%>% filter(file_date==max(file_date))
+
+load(file = paste("data/",bids_files$value,sep=""))
+paste("loaded file=data/",bids_files$value,sep="")
+merit_bids_all<-merit_aug
+
+
+#pick 10% of the date/he combos at random  
+sample_share<-.5
+merit_bids_all<-merit_bids_all%>% sample_n(nrow(.)*sample_share, replace = F) #no replacement - don't duplicate
+
+
+# regression on all the plants
+all_plants <- merit_bids_all %>%  
+  select(bid,percentile,ctax,oba,supply_cushion,hourly_renewables,hour,on_peak,supply_cushion,day_ahead_forecasted_ail,forecast_pool_price,
+         day_ahead_forecasted_ail,total_export_capability,total_import_capability,nit_settle_cad_gj,hdd_YEG,hdd_YMM,hdd_YYC,
+         cdd_YEG,cdd_YMM,cdd_YYC,year,month_fac,he)%>%
+  na.omit()%>%
+    mutate(gas2=nit_settle_cad_gj^2,gas3=nit_settle_cad_gj^3,
+         year_fac=as.factor(year), yearmonth=interaction(month_fac,year_fac),
+         tight=(supply_cushion<=500),really_tight=(supply_cushion<200),he_fac=as.factor(he),pctl_fac=factor(percentile),
+         peak_fac=factor(on_peak),
+         out_mkt=(bid>forecast_pool_price)*(bid-forecast_pool_price) )
+  
+
+#all_plants%>%ggplot()+geom_point(aes(percentile,out_mkt),size=.025)
+
+all_plants %>% filter(percentile>79)%>%select(bid,percentile) %>% group_by(percentile) %>% 
+  arrange(bid) %>% mutate(share=row_number()/n())%>% ggplot()+geom_line(aes(share,bid,group=factor(percentile),color=factor(percentile)))
+
+all_plants %>% select(out_mkt,percentile) %>% group_by(percentile) %>% 
[…]

### 2021-07-31 tahoe_analysis.R

@@ -1,0 +1,715 @@
+#Mac
+if(R.version$platform ==  "x86_64-apple-darwin15.6.0")
+  setwd("/Users/aleach/Google Drive/alberta_power")
+#PC
+if(R.version$platform ==  "x86_64-w64-mingw32")
+  setwd("C:/Users/aleach/Google Drive/alberta_power")
+print(getwd())
+
+library(grid)
+library(gridExtra)
+library(ggpubr)
+library(gganimate)
+library(timeDate)
+
+
+#seasons
+if(!exists("ajl_hourly", mode="function")) source("../andrew_base.R")
+
+#load merit data for the paper
+
+load("merit_ghg_price.RData")  
+
+library(broom)
+library(purrr)
+library(skimr)
+
+
+merit_fossils<-bids_aug %>% filter(Plant_Fuel %in% c("COAL")) 
+bids_aug<-NULL
+merit_aug<-NULL
+by_type<-NULL
+
+
+save(merit_fossils,file="merit_fossils.RData")
+
+load("merit_fossils.RData")  
+#merit_fossils<-merit_fossils %>% select(-bid_10,-bid_30,-bid_40,-bid_60,-bid_70,-bid_90)
+# Melt the data frame to get bid percentile data
+id_names<-names(merit_fossils)[!grepl("bid_",names(merit_fossils))]
+
+#pick 10% of the date/he combos at random  
+sample_share<-.25
+merit_fossils<-merit_fossils%>% sample_n(nrow(.)*sample_share, replace = F) #no replacement - don't duplicate
+
+
+
+merit_fossils<-merit_fossils %>% 
+  melt(id=id_names,variable.name="percentile",value.name = "price")%>%
+  mutate(percentile=as.numeric(gsub("bid_","",as.character(percentile))))
+         
+         merit_fossils<-merit_fossils %>% mutate(he=ifelse(he=="02*","02",he))
+
+
+# Split into pieces, fit model to each piece
+by_type <- merit_fossils %>%  mutate(p2=percentile^2,p3=percentile^3,p4=percentile^4,gas2=nit_settle_cad_gj^2,gas3=nit_settle_cad_gj^3,
+         year_fac=as.factor(year), yearmonth=interaction(month_fac,year_fac),
+         tight=(supply_cushion<=500),really_tight=(supply_cushion<200),he_fac=as.factor(he))%>%
+#nest(-c(Plant_Type,percentile)) %>%
+nest(-c(Plant_Type,offer_control)) %>% 
+  mutate(
[…]

### 2021-08-12 paper/paper_draft.tex
updates

@@
 The most general results derive from the analysis of the market as a whole.
-Here, we are calculating for each hour of our sample the supply curve of power offered into the market which we summarize using a step function with 15 points of support mapping offers at given percentiles of total power offers.
+Here, we are calculating, for each hour of our sample, the supply curve of power offered into the market which we summarize using a step function with 15 points of support mapping offers at given percentiles of total power offers.
 At each of these points of support, we also track the marginal carbon pricing cost and marginal output-based allocation value along with other market and environmental variables.
-We find, as shown in Table X that...
+When we consider only the net marginal cost of carbon pricing (the hourly carbon charge net the hourly output-based allocation of credits), we find a pattern that will be familiar through the remainder of the results in this paper.
+As shown in Figure \ref{fig:net_all}, there is no impact of carbon pricing on offers at the low end of the merit, as this range is characterized by zero dollar offers in all hours of the year.
+The mid-merit range is where we see most of the impacts, and these are intuitive although smaller than might be expected.
+Our study design is such that, if the costs of carbon pricing were fully passed-through to offers, and if we measured we would expect to see an estimated coefficient exactly equal to 1 on the impact of carbon pricing.
+The maximum carbon price pass-through appears at the 60th through 70th percentile of offered power, with the maximum share of net carbon costs passed through to offered prices at just less than 40\%.
+In this formulation, there is no statistical support for the contention that carbon prices are, on average, fully passed-through at the margin in electricity markets.
+
+\begin{figure}[!htb]
+    \centering
+     \includegraphics[width=.9\textwidth]{../images/all_plants_net.png}
+    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types and all hours. Error bars denote 95\% confidence intervals}
+    \label{fig:net_all}
+\end{figure}
+
+Since the marginal cost of offered power becomes less elastic (the slope of the merit order increases) at higher levels of offered power, it's reasonable to expect that the pass-through would be altered at peak hours when demand for power is higher.
+In our base regression shown above, we include a peak-hours factor within the regression, but do not interact that term with carbon costs.
+In Figure \ref{fig:net_peak}, we re-calculate the regression results allowing the carbon pricing effect to vary across peak hours and non-peak hours, and we do find intuitive although only weakly statistically significant changes.
+The net effect of the carbon pricing cost is positive across all by the 80th percentile sampling point, with the largest effects occurring slightly lower in the merit order in off-peak hours, and we also see statistically significant pass-through of net carbon costs at the upper end of the merit order as well.
+
+\begin{figure}[!htb]
+    \centering
+     \includegraphics[width=.9\textwidth]{../images/all_plants_net_peak.png}
+    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types and all hours. Error bars denote 95\% confidence intervals}
+    \label{fig:net_peak}
+\end{figure}
+
+Our data allows a decomposition of the net effect of carbon pricing into a carbon tax and an OBA effect, although with some important caveats.
+Recall that what we are calling a carbon tax effect in any given hour is the average annual carbon tax per MWh of generation, and so we will miss impacts that occur as a result of hourly ramping or other intensive margin conditions.
+Also, for some facilities, we have emissions-intensity data at the aggregate level, and in some cases a facility may include as many as six generating units.
+Where the average emissions intensity varies across units within the facility, we will measure the individual unit intensities with error as well.
+As such, we knot that our marginal effective costs of carbon pricing are measured imprecisely.
+The output-based allocations are consistent through the year on a per-MWh basis, and are determined at the facility level, so we have more accurate measures of these values in our data.
+Again, our estimation strategy is such that a coefficient with an absolute value of 1 where perfect pass-through has occurred, with a value of negative 1 for the full pass through of an output-based allocation to lower offers and a value of 1 for the full pass-through of carbon taxes to higher power offers.
+
+Our results, shown in Figure \ref{fig:decomp_all_hours}, mostly reflect an intuitive decomposition of the results derived based on net values, with a few exceptions.
+Again, we see no impact below the 50th percentile of offered power, and in the mid-merit, an increased in offered prices as marginal carbon prices increase, but with less than full pass through.
+In the upper merit, it's clear that factors other than carbon prices are driving offers, and we see a positive impact of carbon prices with varied impacts of output-based allocations, each estimated much less precisely than the mid-merit impacts.
+
+\begin{figure}[!htb]
+    \centering
+     \includegraphics[width=.9\textwidth]{../images/all_plants_no_peaks.png}
+    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types and all hours. Error bars denote 95\% confidence intervals}
+    \label{fig:decomp_all_hours}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+     \includegraphics[width=.9\textwidth]{../images/all_plants.png}
+    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types and all hours. Error bars denote 95\% confidence intervals}
+    \label{fig:decomp_peaks}
+\end{figure}
 

### 2021-08-12 aeso_scrapes.R

@@ -641,7 +641,7 @@ get_AS_merit_report <- function(start_date, end_date,key_firms=NULL) {
 #clean_data<-get_AS_merit_report(start_date,end_date)
 
 all_as_merit<- function(){
-  years<-seq(2013,2018)
+  years<-seq(2018,2021)
   for(year_id in years){
     days<-seq.Date(as.Date(paste(year_id,"-01-01",sep="")),as.Date(paste(year_id,"-12-31",sep="")),by="1 day")
     if(year_id=="2012")
@@ -664,7 +664,7 @@ all_as_merit<- function(){
   }
 }
 
-
+#all_as_merit()
 
 
 year_as_merit<- function(year_id){

### 2021-08-12 paper_graphs.R

@@ -456,34 +456,8 @@ merit_aug %>% filter(year(date)==2019,he=="19",Plant_Type=="COAL")%>%
 ggsave("images/all_coal_merit.png",dpi=300,width = 14,height = 10)
 
 
-
-merit_aug %>% filter(date==ymd("2019-02-04"),hour==19,asset_id=="BR5" )%>% arrange(price,AESO_Name)%>%
-  mutate(merit=cumsum(size))%>%
-  ggplot(aes(merit,price,color=AESO_Name))+
-  geom_rect(mapping=aes(xmin=merit-size,xmax=merit,ymin=-5,ymax=price),fill=NA)+
-  scale_color_grey("Unit Name")+   
-  scale_x_continuous(breaks=pretty_breaks(), expand=c(0,0))+
-  scale_y_continuous(breaks=pretty_breaks(), expand=c(0,0))+
-  #scale_colour_manual(labe ls=c("Brent","WTI"),values=c("#41ae76","#238b45","#006d2c","#00441b","Black","Black","Black","Black"))
-  paper_theme()+
-  theme(
-    legend.position = "right",
-    legend.margin=margin(c(0,0,0,0),unit="cm"),
-    legend.text = element_text(colour="black", size = 12),
-    plot.caption = element_text(size = 16, face = "italic"),
-    plot.title = element_text(face = "bold"),
-    plot.subtitle = element_text(size = 16, face = "italic"),
-    panel.grid.minor = element_blank(),
-    text = element_text(size = 16,face = "bold"),
-    axis.text.y =element_text(size = 16,face = "bold", colour="black"),
-    axis.text.x=element_text(size = 16,face = "bold", colour="black",angle=0, hjust=1),
-  )+
-  labs(x=paste("Offered Generation (MW)"),y="Price ($/MWh)",
-       #title=paste("Alberta Energy Merit Order, Feb 4, 2019 at hour ending 7pm"),
-       #caption="Source: AESO Data, graph by Andrew Leach."
-  )
-
 #test_bids<-merit_bids %>% filter(date==ymd("2019-02-04"),he=="19",Plant_Type=="COAL")%>% rename(price=bid)
+
 
 merit_bids %>% filter(date==ymd("2019-02-04"),he=="19",Plant_Type=="COAL")%>% rename(price=bid)%>%
   mutate(merit=percentile/100*total_offers,size=merit-lag(merit,n=1),size=ifelse(is.na(size),merit[1],size))%>%
[…]

### 2021-08-12 power_paper_base.R

@@ -37,6 +37,7 @@ paper_theme<-function(caption_align=1){
     legend.position = "bottom",
     legend.text=element_text(size=12),
     legend.margin=margin(c(0,0,-0.25,0),unit="cm"),
+    plot.margin=margin(c(.5,.5,0.5,0.5),unit="cm"),
     plot.caption = element_text(size = 9, color = "gray40",hjust=1),
     plot.title = element_text(face = "bold"),
     plot.subtitle = element_text(size = 10, color = "gray40"),

### 2021-08-12 regression_analysis.R

@@ -17,21 +17,31 @@ merit_bids_all<-merit_aug
 
 
 #pick 10% of the date/he combos at random  
-sample_share<-.5
-merit_bids_all<-merit_bids_all%>% sample_n(nrow(.)*sample_share, replace = F) #no replacement - don't duplicate
+#sample_share<-.5
+#merit_bids_all<-merit_bids_all%>% sample_n(nrow(.)*sample_share, replace = F) #no replacement - don't duplicate
 
 
 # regression on all the plants
 all_plants <- merit_bids_all %>%  
-  select(bid,percentile,ctax,oba,supply_cushion,hourly_renewables,hour,on_peak,supply_cushion,day_ahead_forecasted_ail,forecast_pool_price,
+  select(bid,percentile,ctax,oba,net,supply_cushion,hourly_renewables,hour,on_peak,supply_cushion,day_ahead_forecasted_ail,forecast_pool_price,
          day_ahead_forecasted_ail,total_export_capability,total_import_capability,nit_settle_cad_gj,hdd_YEG,hdd_YMM,hdd_YYC,
-         cdd_YEG,cdd_YMM,cdd_YYC,year,month_fac,he)%>%
+         cdd_YEG,cdd_YMM,cdd_YYC,year,month_fac,date,he)%>%
   na.omit()%>%
-    mutate(gas2=nit_settle_cad_gj^2,gas3=nit_settle_cad_gj^3,
+    mutate(
+         gas2=nit_settle_cad_gj^2,gas3=nit_settle_cad_gj^3,
          year_fac=as.factor(year), yearmonth=interaction(month_fac,year_fac),
          tight=(supply_cushion<=500),really_tight=(supply_cushion<200),he_fac=as.factor(he),pctl_fac=factor(percentile),
          peak_fac=factor(on_peak),
-         out_mkt=(bid>forecast_pool_price)*(bid-forecast_pool_price) )
+         out_mkt=(bid>forecast_pool_price)*(bid-forecast_pool_price),
+         policy=case_when(year(date)<2016 ~ "SGER_15",
+                                  year(date)==2016 ~ "SGER_20",
+                                  year(date)==2017 ~ "SGER_30",
+                                  year(date)==2018 ~ "CCIR_30",
+                                  year(date)==2019 ~ "CCIR_30",
+                                  year(date)==2020 ~ "TIER_30",
+                                  year(date)==2021 ~ "TIER_40",
[…]

### 2022-01-18 electricity_deck.Rmd
updating

---
title: Alberta's Electricity Market
subtitle: Energy Industries and Markets Slide Pack
author: 
  - Andrew Leach 
institution: 
- Alberta School of Business and CABREE
fontsize: 10pt
output:
 beamer_presentation:
   #template: "C:/Users/aleach/Google Drive/default_beamer.tex"
   #Mac
   template: "ajl-latex-beamer.tex"
    #PC
   #template: "C:/Users/aleach/Google Drive/ajl-latex-beamer.tex"

   theme: "Darmstadt"  
   incremental: false
   slide_level: 2
    #keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)

library(ggplot2)
library(lubridate)
library(dplyr)
library(viridis)
library(zoo)
```

#Introduction

## Why should electricity interest you?

Price volatility:

- Electricity prices are an order of magnitude more volatile than oil or gas prices

- Both supply and demand in Alberta affect us more than global or regional supplies and demands in oil and gas markets

New technology:

- Electricity is, arguably, changing faster than any other energy market

- Alberta's electricity market is entering a period of market- and regulatory-driven transition

Economics 101 in action

- Nowhere else will you see supply and demand curves actually mapped out in real time determining prices as clearly as in Alberta's power market

```{r plotit, echo=FALSE,results=FALSE,message=FALSE}
options(warn=-1) #suppress warnings in Markdown

slide_theme<-function(){
  return( theme(panel.border = element_blank(),
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = "gray",linetype="dotted"),
        axis.line.x = element_line(color = "gray"),
        axis.line.y = element_line(color = "gray"),
        axis.text = element_text(size = 20),
        axis.text.x = element_text(margin = margin(t = 10)),
        axis.title = element_text(size = 20),
        #axis.label.x = element_text(size=20,vjust=+5),
        plot.subtitle = element_text(size = 20,hjust=0.5),
        plot.caption = element_text(face="italic",size = 16,hjust=0),
        legend.key.width=unit(2,"line"),
        legend.position = "bottom",
        #legend.direction = "horizontal",
        #legend.box = "horizontal",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust=0.5,size = 24)))
}

```

```{r gendata, echo=FALSE,results=FALSE,message=FALSE}

load("gen.RData") ## which is here *equivalent* to

gen_data<-gen_data[-is.na(gen_data$gen),]

gen_data$AIL<-as.numeric(gen_data$AIL)
gen_data$Price<-as.numeric(gen_data$Price)
gen_data$Demand<-as.numeric(gen_data$Demand)

```

```{r, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}
filename<-paste("forecast_data",".RData",sep = "")
#save(data_store, file= filename) 
load(file = filename)

load("all_merit.RData")  #load merit order data

#hourly prices  
forecast_data$day_ahead_forecasted_ail<-gsub(",","",forecast_data$day_ahead_forecasted_ail)
forecast_data$actual_ail<-gsub(",","",forecast_data$actual_ail)
forecast_data$forecasted_actual_ail_difference<-gsub(",","",forecast_data$forecasted_actual_ail_difference)
forecast_data$Year<-year(forecast_data$date)
forecast_data$actual_posted_pool_price<-as.numeric(forecast_data$actual_posted_pool_price)
forecast_data$forecast_pool_price<-as.numeric(forecast_data$forecast_pool_price)
forecast_data$actual_ail<-as.numeric(forecast_data$actual_ail)
forecast_data$day_ahead_forecasted_ail<-as.numeric(forecast_data$day_ahead_forecasted_ail)
sub_samp<-forecast_data
sub_samp$time<-NULL
sub_samp<-sub_samp[!is.na(sub_samp$actual_posted_pool_price),]
df1 <- sub_samp %>% group_by(he,Year) %>% summarise(AIL=mean(actual_ail),p_mean=mean(actual_posted_pool_price))

#png(file="hourly-prices.png", width = 1400, height = 750)
```

## Price volatility
```{r, fig.width=16,fig.height=11, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}
p<-ggplot(df1, aes(he,p_mean,group=as.factor(Year),colour=as.factor(Year))) +
  geom_line(data=df1,size=1.5) +
  geom_line(data=subset(df1,Year==2017),size=3) +
  #geom_point(size=1.5) +
  labs(colour="Year") +
  scale_color_viridis(discrete=TRUE)+   
  theme_classic() +
  theme(panel.border = element_blank(),
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = "gray",linetype="dotted"),
        axis.line.x = element_line(color = "gray"),
        axis.line.y = element_line(color = "gray"),
        axis.text = element_text(size = 20),
        axis.text.x = element_text(margin = margin(t = 10)),
        axis.title = element_text(size = 20),
        #axis.label.x = element_text(size=20,vjust=+5),
        plot.subtitle = element_text(size = 20,hjust=0.5),
        plot.caption = element_text(face="italic",size = 16,hjust=0),
        legend.key.width=unit(2,"line"),
        legend.position = "bottom",
        #legend.direction = "horizontal",
        #legend.box = "horizontal",
        legend.text = element_text(size = 20),
        plot.title = element_text(hjust=0.5,size = 24))+
  labs(y="Hourly Average Power Price ($/MWh)",x="Hour",
             title="Hourly Average Energy Prices ($/MWh, 2000-2018)",
             caption="Source: AESO Data, Accessed via NRGStream, Graph by Andrew Leach")

print(p)
```

## Price volatility
```{r, fig.width=16,fig.height=11, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}
sub_samp$month<-month(sub_samp$date)
sub_samp$Date_ID = as.yearmon(sub_samp$date)
df1 <- sub_samp %>% group_by(Date_ID) %>% summarise(cons =sum(actual_ail)/1000,AIL =mean(actual_ail),Peak_AIL =max(actual_ail), p_mean=sum(actual_posted_pool_price*actual_ail)/sum(actual_ail))
df1$Date<-as.Date(df1$Date_ID)
df1$m12_avg<-as.numeric(rollapply(df1$p_mean,12,mean,fill=NA,align = c("right")))
df1$m12_ail_avg<-as.numeric(rollapply(df1$AIL,12,mean,fill=NA,align = c("right")))

lims=c(min(df1$Date)+years(1),max(df1$Date)+months(3))
p<-ggplot(data=subset(df1,Date>="2001-01-01"), aes(Date,p_mean,colour="A"),size=2.5) +
  geom_line(size=1.5) +
  geom_line(data=subset(df1,Date>="2001-01-01"),aes(Date,m12_avg,colour="B"),size=2.5) +
  scale_color_viridis("",labels=c("Monthly Average Prices","12 Month Rolling Average Prices"),discrete=TRUE)+   
  scale_x_date(name=NULL,date_breaks = "1 year", date_labels =  "%b\n%Y",limits=lims,expand=c(0,0)) +
  theme_classic() +
  slide_theme()+
      labs(y="Monthly/Annual Average Power Price ($/MWh)",x="\nMonth",
             title="Load-Weighted Monthly Average Energy Prices ($/MWh, 2005-2017)",
             caption="Source: AESO Data, Accessed via NRGStream, Graph by Andrew Leach")

print(p)
```

## Supply and Demand

```{r, fig.width=16,fig.height=11, echo=FALSE,warnings=FALSE,results=FALSE,message=FALSE}

df1<-subset(merit_data, he==18 & date==as.Date("2017-10-05"))

p<-ggplot(df1,aes(merit,price,fill=offer_sum)) +
  geom_rect(mapping=aes(xmin=merit-size,xmax=merit,ymin=-20,ymax=price))+
  scale_fill_viridis("Offer\nControl",discrete = TRUE)+   
  theme_classic() +
  slide_theme()+
  labs(x=paste("Offered Generation (MW)"),y="Price ($/MWh)",
       title=paste("Alberta Energy Merit Order, ",max(df1$date)," ",max(df1$he),":00",sep = ""),
       caption="Source: AESO Data, graph by Andrew Leach.")
print(p)

```

## New Technology

```{r,echo=FALSE,message=FALSE, out.width = "320px"}
knitr::include_graphics("/Users/aleach/Google Drive/eclipse.png")

```

#The Market

## The Grid 2.0
```{r,echo=FALSE,message=FALSE, out.width = "320px"}
knitr::include_graphics("grid_ess.png")
```

## Market Participants
\begin{itemize}
\setlength\itemsep{2em}
\item Generation

\item Transmission

\item Distribution

\item Ancillary Services

\item Load

\item Storage
\end{itemize}

## Energy units - electricity 
Watts: measure of capacity (instantaneous production, installed capacity, or instantaneous demand)

- Alberta system demand: 7,200-10,700 MW (million watts)

- Capital Power's Genessee 3 power plant is: 450 MW

Watt hours: measure of energy (production or demand during a given period of time; i.e. flow through)
    -Production over a day, week, month, year

Volts are a measure of the electrical potential - the ability to convert charge to power: Watts=amps x volts
    -Transmission lines: 150-765 kV
    -Distribution lines: 13,800 Volts
    -Household wiring: 120-240 Volts

# Prices
## Energy Prices
Electricity prices are expressed in power delivered over time
Cents/kilowatt-hour (c/kWh)
Dollars per megawatt-hour ($/MWh)
Capacity costs are expressed in a cost per megawatt or cost of capacity
Genessee 3 cost approximately $1.5 million/MW or $1.50 per watt to build
Solar panel prices have declined to now lie under $1/W of capacity
Balance of system costs imply that a solar system costs $2-3/W of installed capacity
Levelized costs of electricity (supply costs) in $/MWh

## Other Prices
Other prices matter for electricity markets as well
Renewable energy credits (usually prices in $/MWh)
Emissions credits or permits ($/tonne)
Capacity payments ($/MW)
Air emissions permits or credits ($/tonne)

#Market Participants

## Test

#Generation Sources

## Test

#GHG Emissions

## Test

#Etc

## Test

### 2022-01-18 presentations/power_slides.tex
updating

%First Beamer Talk

\documentclass{beamer}

% Setup appearance:

\usetheme{Darmstadt}
%\usecolortheme{lily}
%\usefonttheme[onlylarge]{structurebold}
%\setbeamerfont*{frametitle}{size=\normalsize,series=\bfseries}
%\useinnertheme{circles}

% Standard packages

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
%\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{multirow}
\usepackage{capt-of}
\usepackage{graphicx}
\usepackage{array}
\usepackage{tikz}

\setbeamertemplate{blocks}[rounded] [shadow=false]

% Some optional colors. Change or add as you see fit.
%---------------------------------------------------
 \definecolor{ualbertagreen}{HTML}{007C41}
\definecolor{ualbertagold}{HTML}{FFDB05}

% Some optional color adjustments to Beamer. Change as you see fit.
%------------------------------------------------------------------
\setbeamercolor{frametitle}{fg=ualbertagreen,bg=white}
\setbeamercolor{title}{fg=ualbertagreen,bg=white}
\setbeamercolor{author}{fg=ualbertagreen,bg=white}
\setbeamercolor{date}{fg=ualbertagreen,bg=white}
\setbeamercolor{affiliation}{fg=ualbertagreen,bg=white}
\setbeamercolor{institute}{fg=ualbertagreen,bg=white}
\setbeamercolor{local structure}{fg=ualbertagreen}
\setbeamercolor{section in toc}{fg=ualbertagreen,bg=white}
% \setbeamercolor{subsection in toc}{fg=ualbertagreen,bg=white}
\setbeamercolor{footline}{fg=ualbertagreen!50, bg=white}
\setbeamercolor{block title}{fg=ualbertagreen,bg=white}
\setbeamercolor{upper separation line head}{bg=ualbertagreen}
\setbeamercolor{lower separation line head}{bg=ualbertagold}
\setbeamercolor{middle separation line head}{bg=ualbertagold}
\setbeamercolor{frametitle}{fg=ualbertagreen,bg=white}

\setbeamercolor{section in head/foot}{bg=white,fg=ualbertagreen}
\setbeamercolor{author in head/foot}{bg=white,fg=ualbertagreen}
\setbeamercolor{date in head/foot}{bg=white,,fg=ualbertagreen}
\setbeamercolor{title in head/foot}{bg=white,fg=ualbertagreen}

\setbeamercolor{headline}{bg=white,fg=ualbertagreen}

\setbeamercolor{middle separation line head}{bg=ualbertagreen}
\setbeamercolor{alerted text}{fg=red}
\setbeamercolor{example text}{fg=black}
\setbeamercolor{structure}{fg=black}

% Various cosmetic things, though I must confess I forget what exactly these do and why I included them.
%-------------------------------------------------------------------------------------------------------
\setbeamercolor{structure}{fg=ualbertagreen}

\setbeamercolor{local structure}{parent=structure}
\setbeamercolor{item projected}{parent=item,use=item,fg=ualbertagreen,bg=white}
\setbeamercolor{enumerate item}{parent=item}

\setbeamertemplate{title page}{%
  \vbox{}
%  \vfill
    \vspace{0cm}% NEW
  \begingroup
    \centering
    \begin{beamercolorbox}[sep=8pt,center]{title}
      \usebeamerfont{title}\inserttitle\par%
      \ifx\insertsubtitle\@empty%
      \else%
        \vskip0.05em%
        {\usebeamerfont{subtitle}\usebeamercolor[fg]{subtitle}\insertsubtitle\par}%
      \fi%
    \end{beamercolorbox}%
    \vskip1em\par
    \begin{beamercolorbox}[sep=8pt,center]{author}
      \usebeamerfont{author}\insertauthor
    \end{beamercolorbox}
    \begin{beamercolorbox}[sep=8pt,center]{institute}
      \usebeamerfont{institute}\insertinstitute
    \end{beamercolorbox}
    \vspace{0.5cm}% NEW
    \begin{beamercolorbox}[sep=8pt,center]{date}
      \usebeamerfont{date}\insertdate
    \end{beamercolorbox}\vskip0.05em
%    {\usebeamercolor[fg]{titlegraphic}\inserttitlegraphic\par}
  \endgroup
%  \vfill
}

\logo{
   \tikz [remember picture,overlay]
    \node[yshift=.3cm,xshift=1.5cm] at (current page.south west)
        %or: (current page.center)
        {\includegraphics[width=1in]{UA-ASB-COLOUR.png}};
%\includegraphics[height=0.8cm]{UA-ASB-COLOUR.png}\vspace{220pt}
}

\setbeamertemplate{headline}{%
\leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=\paperwidth,ht=5ex,dp=1.825ex]{white}%
    \usebeamerfont{headline}\hskip6pt\inserttitle\par%
    \insertsectionnavigationhorizontal{\paperwidth}{}{\hskip0pt plus1filll}
    \end{beamercolorbox}%
  }
}

\setbeamertemplate{sidebar right}{}

\logo{
   \tikz [remember picture,overlay]
    \node[yshift=.3cm,xshift=1.5cm] at (current page.south west)
        %or: (current page.center)
        {\includegraphics[width=1in]{UA-ASB-COLOUR.png}};
%\includegraphics[height=0.8cm]{UA-ASB-COLOUR.png}\vspace{220pt}
}

%\setbeamertemplate{footline}{%
%\hfill\usebeamertemplate***{navigation symbols}
%\hspace{1cm}\insertframenumber{}/\inserttotalframenumber}
\defbeamertemplate*{footline}{my footline}{%
    \ifnum\insertpagenumber=1
        \Tiny{%
            \hfill%
		\vspace*{1pt}%
            %\insertframenumber/\inserttotalframenumber \hspace*{0.1cm}%
            \newline%
            \color{ualbertagold}{\rule{\paperwidth}{0.4mm}}\newline%
            \color{ualbertagold}{\rule{\paperwidth}{.4mm}}%
        }
%    \hbox{%
%        \begin{beamercolorbox}[wd=\paperwidth,ht=.8ex,dp=1ex,center]{}%
%      % empty environment to raise height
%        \end{beamercolorbox}%
%    %}%
    %\vskip0pt%
    %no page number on the first page
    %    \Tiny{%
    %        \hfill%
   % 		\vspace*{1pt}%
    %        \color{ualbertagold}{\rule{\paperwidth}{0.4mm}}\newline%
    %        \color{ualbertagold}{\rule{\paperwidth}{.4mm}}%
%        }%
  \else%
        \Tiny{%
            \hfill%
		\vspace*{1pt}%
            \insertframenumber/\inserttotalframenumber \hspace*{0.1cm}%
            \newline%
            \color{ualbertagold}{\rule{\paperwidth}{0.4mm}}\newline%
            \color{ualbertagold}{\rule{\paperwidth}{.4mm}}%
        }%
    \fi%
}

\renewcommand{\(}{\begin{columns}}
\renewcommand{\)}{\end{columns}}
\newcommand{\<}[1]{\begin{column}{#1}}
\renewcommand{\>}{\end{column}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Author, Title, etc.

\title[Electricity Slide Pack]
{%
  Electricity Markets%
}

\author[Leach]
{
  Andrew~Leach
}

\institute[2017]
{
  Alberta School of Business, University of Alberta
 }

\date[5/28/2017]
{\today}

\newcommand{\degC}{$^o$C$\,$}
\newcommand{\co}{$\text{CO}_{2}\,$}
\newcommand{\tx}{$G_{2\times\text{CO}_2}$}
\newcommand{\Real}{\mathbb R}

% The main document

\begin{document}

\begin{frame}
   \tikz [remember picture,overlay]
    \node[yshift=-0.95cm,xshift=0cm] at (current page.north)
        %or: (current page.center)
        %\node[yshift=-0.75cm,xshift=4.5cm] at (current page.north west)
        %{\includegraphics[width=3in]{UA-ASB-COLOUR.png}};
        {\includegraphics[width=\paperwidth]{EE_logo.png}}; \vspace{1cm}
   \titlepage
   \vfill
\end{frame}

\section{Introduction}

\begin{frame}{Why should you care about electricity?}
\begin{itemize}
\setlength\itemsep{.5em}
\item Price volatility:
\begin{itemize}
\setlength\itemsep{.5em}
    \item Electricity prices are an order of magnitude more volatile than oil or gas prices
    \item
         Both supply and demand in Alberta affect us more than global or regional supplies and demands in oil and gas markets
\end{itemize}
\item New technology:
\begin{itemize}
\setlength\itemsep{.5em}
\item Electricity is, arguably, changing faster than any other energy market
\item Alberta's electricity market is entering a period of market- and regulatory-driven transition
\end{itemize}
\item Economics 101 in action
\begin{itemize}
\setlength\itemsep{.5em}
\item Nowhere else will you see supply and demand curves actually mapped out in real time determining prices as clearly as in Alberta's power market
\end{itemize}

\end{itemize}

\vfill
\end{frame}

\begin{frame}{Price Volatility}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/monthly_prices.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Price Volatility}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_2000_2019.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Price Volatility}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/hourly-prices.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Growth}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/load_time.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{New Technology}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/eclipse.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{New Technology}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/brooks_solar.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Economics 101}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/merit_type.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Economics 101}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/merit_offer.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}

\begin{tabular}{p{\linewidth}}
    \centering
    \includegraphics[width=.9\linewidth]{grid_ess.png} \\[\abovecaptionskip]
  Source: \url{http://www.mpoweruk.com/grid_storage.htm}
\end{tabular}

\end{frame}

\begin{frame}{Market Participants}
\begin{itemize}
\setlength\itemsep{1.7em}
\item Generation

\item Transmission

\item Distribution

\item Ancillary Services

\item Load

\item Storage

\item Microgeneration
\end{itemize}

\vfill \end{frame}

\section{Units}

\begin{frame}{Energy units - electricity}
\begin{itemize}
\setlength\itemsep{.75em}
\item Watts: measure of capacity (instantaneous production, installed capacity, or instantaneous demand)
\begin{itemize}
\setlength\itemsep{.5em}
\item Alberta system demand: 7,200-10,700 MW (million watts)
\item Capital Power's Genessee 3 power plant has a nameplate capacity of 450 MW
\end{itemize}
\item Watt hours: measure of energy (production or demand during a given period of time; i.e. flow through)
\begin{itemize}
\setlength\itemsep{.5em}
\item Production over a day, week, month, year
\end{itemize}
\item Volts: measure of the electrical potential or the ability to convert charge to power (Watts=amps x volts)
\begin{itemize}
\setlength\itemsep{.5em}
\item Transmission lines: 150-765 kV
\item Distribution lines: 13,800 Volts
\item Household wiring: 120-240 Volts
\end{itemize}
\end{itemize}
\vfill \end{frame}

\begin{frame}{Energy Prices}
\begin{itemize}
\setlength\itemsep{.25em}
\item Electricity prices: expressed in power delivered over time
\begin{itemize}
\setlength\itemsep{.15em}
\item Cents/kilowatt-hour (c/kWh)
\item Dollars per megawatt-hour (\$/MWh)
\item Levelized costs of electricity (supply costs) in \$/MWh
\end{itemize}

\item Capacity costs are expressed in a cost per megawatt or cost of capacity
\begin{itemize}
\setlength\itemsep{.15em}
\item Genessee 3 cost approximately \$1.5 million/MW or \$1.50 per watt to build
\item Solar panel prices have declined to now lie under \$1/W of capacity
\item Balance of system costs imply that a solar system costs \$2-3/W of installed capacity
\end{itemize}
\item Other prices matter for electricity markets as well
\begin{itemize}
\setlength\itemsep{.15em}
\item Renewable energy credits (usually prices in \$/MWh)
\item Emissions credits or permits (\$/tonne)
\item Capacity payments (\$/MW)
\item Air emissions permits or credits (\$/tonne)
\end{itemize}
\end{itemize}

\vfill \end{frame}

\section{Power Markets}

\begin{frame}{Regulatory characteristics}
\begin{itemize}
\setlength\itemsep{.25em}
\item Rate-regulated or or state-owned utilities
\item Competitive markets
\begin{itemize}
\setlength\itemsep{.15em}
\item Energy only markets: ERCOT and Alberta
\item Energy and capacity markets: MISO, PJM, soon-to-be Alberta
\item Real-time vs day-ahead prices: PJM and others have day-ahead market and then a real time differences market
\item Many other design characteristic differences between restructured or competitive markets
\end{itemize}
\end{itemize}

\vfill
\end{frame}

\begin{frame}{Alberta Market Design}
\begin{itemize}
\setlength\itemsep{.25em}
\item Energy-only market
\item Real time, spot pricing, no day-ahead market
\item Single node
\item Capacity market to be added in the near future
\item Transmission
\item Congestion free (no nodal pricing)
\item No transmission rights
\item Ancillary services: separate, competitive market for operating reserves, transmission-must-run, load-shed and black start
\end{itemize}

\vfill
\end{frame}

\begin{frame}{Nodal Pricing Example}

\begin{tabular}{p{\linewidth}}
    \centering
    \includegraphics[width=\linewidth]{../images/cali_nodes.png} \\[.1\abovecaptionskip]
  Source: CAISO
\end{tabular}
\vfill
\end{frame}

\section{Wholesale Market}

\begin{frame}{The Wholesale Market}
\begin{itemize}
\setlength\itemsep{2em}
\item Suppliers place offers of power at particular price
\item Demand-side bids placed for power with a maximum price
\item Supply offers are sorted from low to high
\item Demand offers are sorted from high to low
\item Marginal price is set at the price which equates supply and demand - economics 101 at work!
\item Import supply is bid-in at \$0, but receive the marginal price
\item Export demand is bid-in at \$999, so they do not set the price directly but pay the marginal price
\item Consumer default bid allows AESO to go up merit order to meet observed demand
\end{itemize}

\vfill \end{frame}

\begin{frame}{The Merit Order}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/merit_type.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{The Merit Order}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/merit_offer.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Offers}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/bids.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Offers}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/bids_alpac.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Hourly Loads}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/hourly-loads.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Monthly Peak Loads}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/monthly-peak-loads.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Monthly Average Loads}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/monthly-avg-loads.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Hourly Prices}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/hourly-prices.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Recent Hourly Prices}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/hourly-prices-recent.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Prices over time}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/monthly_prices.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Prices over time}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_2000_2019.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Prices over time}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_week.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Forecasting Prices and Loads}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/price_and_load.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Duration Curve of Prices}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/price_dist.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Duration Curve of Prices}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/price_dist_recent.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Forward Markets}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/forwards_all.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Forward Markets}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/forwards_monthly.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Forward Markets}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/forwards_kenney.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Forward Markets}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/forwards.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Forward Markets}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/forwards_2017.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Forward Markets}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/forwards_long.png}}; \vspace{1cm}
   \vfill
\end{frame}

\section{PPAs}

\begin{frame}{The Balancing Pool: What on earth does it do?}
\vspace{-.2cm}\centering
\hspace{-1.5cm}\begin{tabular}{p{1.1\linewidth}}
    \centering
    \includegraphics[width=.9\linewidth]{balancingpool.png} \\[\abovecaptionskip]
  Source: Capital Power
\end{tabular}

\vfill \end{frame}

\section{Alberta Market Evolution}

\begin{frame}{Alberta's Evolving Electricity Market}
\begin{itemize}
\setlength\itemsep{2em}
\item Capacity Market
\item Coal Phase Out
\item Renewables (the REP Program)
\item Carbon Pricing
\end{itemize}

\vfill \end{frame}

\begin{frame}{Costs of New Capacity Additions}
   \tikz [remember picture,overlay]
    \node[yshift=-.25cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/eia_lcoe_2019_avg.png}}; \vspace{1cm}
   \\[\abovecaptionskip]  \vspace{5.5cm}
   \tiny{See for more information: \url{https://www.lazard.com/media/451086/lazards-levelized-cost-of-energy-version-130-vf.pdf}}
   \vfill
\end{frame}

\begin{frame}{Price Capture by Technology}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/price_capture.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Price Capture by Technology}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/price_capture_avg.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Evolution of Technology}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/eia_aeo_solar.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Evolution of Technology}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/solar_eia.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Evolution of Technology}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.9\paperwidth]{../images/solar_costs.png}}; \vspace{1cm}
   \vfill
   \vspace{4cm}\tiny{Source: NREL}  \hfill
\end{frame}

\begin{frame}{GHG Policy and Electricity Supply}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.95\paperwidth]{../images/gen_ghg_price.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{GHG Policy and Electricity Supply}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.95\paperwidth]{../images/monthly_ghgs.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{GHG Policy and Electricity Supply}
   \tikz [remember picture,overlay]
    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.95\paperwidth]{../images/monthly_ghg_mwh.png}}; \vspace{1cm}
   \vfill
\end{frame}

%\begin{frame}{GHG Emissions from Electricity}
%   \tikz [remember picture,overlay]
%    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
%        {\includegraphics[width=.8\paperwidth]{../inventory_ghg_mwh.png}}; \vspace{1cm}
%   \vfill
%\end{frame}

%\begin{frame}{GHG Emissions from Electricity}
%   \tikz [remember picture,overlay]
%    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
%        {\includegraphics[width=.8\paperwidth]{../elec_share.png}}; \vspace{1cm}
%   \vfill
%\end{frame}

\end{document}

### 2022-01-18 presentations/power_slides_law.tex
updating

--- 2022-01-18 presentations/power_slides.tex
@@
 {
-  Alberta School of Business, University of Alberta
+  University of Alberta
  }
@@
 \begin{frame}
-   \tikz [remember picture,overlay]
-    \node[yshift=-0.95cm,xshift=0cm] at (current page.north)
+   %\tikz [remember picture,overlay]
+   % \node[yshift=-0.95cm,xshift=0cm] at (current page.north)
         %or: (current page.center)
@@
         %{\includegraphics[width=3in]{UA-ASB-COLOUR.png}};
-        {\includegraphics[width=\paperwidth]{EE_logo.png}}; \vspace{1cm}
+        %{\includegraphics[width=\paperwidth]{EE_logo.png}}; \vspace{1cm}
    \titlepage
@@
 
-\begin{frame}{Why should you care about electricity?}
+\begin{frame}{Why aspiring lawyers should care about electricity and what you need to know}
 \begin{itemize}
 \setlength\itemsep{.5em}
-\item Price volatility:
+\item Regulatory complexity:
 \begin{itemize}
 \setlength\itemsep{.5em}
-    \item Electricity prices are an order of magnitude more volatile than oil or gas prices
-    \item
-         Both supply and demand in Alberta affect us more than global or regional supplies and demands in oil and gas markets
+    \item Complex regulation, even in \textit{deregulated} jurisdictions
+    \item Each part of the market has its own set of regulatory constructs (distribution, transmission, generation, retail)
 \end{itemize}
-\item New technology:
+\item New technology means evolving market structures:
 \begin{itemize}
@@
 \item Electricity is, arguably, changing faster than any other energy market
-\item Alberta's electricity market is entering a period of market- and regulatory-driven transition
+\item Alberta's electricity market in the midst of a period of market- and regulatory-driven transition
 \end{itemize}
@@
 \setlength\itemsep{.5em}
-\item Nowhere else will you see supply and demand curves actually mapped out in real time determining prices as clearly as in Alberta's power market
+\item Supply and demand curves determine prices every hour in Alberta's power market
 \end{itemize}
-
+\item The next constitutional battleground?
+\begin{itemize}
+\setlength\itemsep{.5em}
+\item Fed/prov issues, environmental policy, section 125 jurisprudence, the oddity that is section 92(10) including the declaratory power and POGG all come into play!
 \end{itemize}
-
-\vfill
-\end{frame}
-
-\begin{frame}{Price Volatility}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/monthly_prices.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Price Volatility}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_2000_2019.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Price Volatility}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/hourly-prices.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Growth}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/load_time.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{New Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/eclipse.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{New Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/brooks_solar.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Economics 101}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/merit_type.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Economics 101}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/merit_offer.png}}; \vspace{1cm}
+\end{itemize}
 \vfill
@@
     \centering
-    \includegraphics[width=.9\linewidth]{grid_ess.png} \\[\abovecaptionskip]
+    \includegraphics[width=.9\linewidth]{../images/grid_storage.png} \\[\abovecaptionskip]
   Source: \url{http://www.mpoweruk.com/grid_storage.htm}
@@
 \item Microgeneration
+\end{itemize}
+
+\vfill \end{frame}
+
+\begin{frame}{Market Regulation in Alberta}
+\begin{itemize}
+\setlength\itemsep{1.7em}
+\item Generation is a competitive market competitive
+
+\item Transmission is regulated on a cost-of-service basis
+
+\item Distribution (wires) is regulated on a cost-of-service basis, but retail (billing) is competitive.
+
+\item Ancillary Services is competitive
+
+\item Load (i.e. customers) may contract for electricity supply
+
+\item Storage (Still a lot TBD)
+
+\item Microgeneration (free market for self-supply)
 \end{itemize}
@@
 \item Capacity payments (\$/MW)
-\item Air emissions permits or credits (\$/tonne)
+\item GHG or other emissions permits or credits (\$/tonne)
 \end{itemize}
@@
 \item Energy only markets: ERCOT and Alberta
-\item Energy and capacity markets: MISO, PJM, soon-to-be Alberta
-\item Real-time vs day-ahead prices: PJM and others have day-ahead market and then a real time differences market
+\item Energy and capacity markets: MISO, PJM
+\item Real-time vs day-ahead prices: PJM and others have day-ahead market and then a real-time differences market
 \item Many other design characteristic differences between restructured or competitive markets
@@
 \end{itemize}
-
 \vfill
@@
 
-\begin{frame}{The Wholesale Market}
+\begin{frame}{The Alberta Wholesale Market}
 \begin{itemize}
-\setlength\itemsep{2em}
-\item Suppliers place offers of power at particular price
+\item Suppliers place offers of power at particular price[-5em]
 \item Demand-side bids placed for power with a maximum price
@@
 \item Marginal price is set at the price which equates supply and demand - economics 101 at work!
-\item Import supply is bid-in at \$0, but receive the marginal price
+\item Import and renewable supply is bid-in at \$0, but everyone receives the market price
 \item Export demand is bid-in at \$999, so they do not set the price directly but pay the marginal price
@@
 
-\begin{frame}{Offers}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/bids_alpac.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
 \begin{frame}{Hourly Loads}
@@
 
-\begin{frame}{Monthly Peak Loads}
+\begin{frame}{Peak and Average Loads}
    \tikz [remember picture,overlay]
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/monthly-peak-loads.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Monthly Average Loads}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/monthly-avg-loads.png}}; \vspace{1cm}
+        {\includegraphics[width=.9\paperwidth]{../images/loads_clean.png}}; \vspace{1cm}
    \vfill
@@
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/hourly-prices.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Recent Hourly Prices}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/hourly-prices-recent.png}}; \vspace{1cm}
+        {\includegraphics[width=.9\paperwidth]{../images/hourly_prices_col.png}}; \vspace{1cm}
    \vfill
@@
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/monthly_prices.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Prices over time}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_2000_2019.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Prices over time}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_week.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Forecasting Prices and Loads}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/price_and_load.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Duration Curve of Prices}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/price_dist.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Duration Curve of Prices}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/price_dist_recent.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Forward Markets}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/forwards_all.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Forward Markets}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/forwards_monthly.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Forward Markets}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/forwards_kenney.png}}; \vspace{1cm}
+        {\includegraphics[width=.9\paperwidth]{../images/peak_prices.png}}; \vspace{1cm}
    \vfill
@@
 
-\begin{frame}{Forward Markets}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/forwards_2017.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
+\section{PPAs and the Balancing Pool}
 
-\begin{frame}{Forward Markets}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/forwards_long.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\section{PPAs}
-
-\begin{frame}{The Balancing Pool: What on earth does it do?}
+\begin{frame}{The Balancing Pool: What on earth does/did it do?}
 \vspace{-.2cm}\centering
@@
     \centering
-    \includegraphics[width=.9\linewidth]{balancingpool.png} \\[\abovecaptionskip]
+    \includegraphics[width=.9\linewidth]{../images/balancingpool.png} \\[\abovecaptionskip]
   Source: Capital Power
@@
 \setlength\itemsep{2em}
-\item Capacity Market
+\item Capacity Market False Start
 \item Coal Phase Out
-\item Renewables (the REP Program)
-\item Carbon Pricing
+\item Renewables (the REP Program and now commercial PPAs)
+\item Carbon Pricing (TIER and the federal GGPPA)
 \end{itemize}
@@
     \node[yshift=-.25cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/eia_lcoe_2019_avg.png}}; \vspace{1cm}
-   \\[\abovecaptionskip]  \vspace{5.5cm}
-   \tiny{See for more information: \url{https://www.lazard.com/media/451086/lazards-levelized-cost-of-energy-version-130-vf.pdf}}
+        {\includegraphics[width=.9\paperwidth]{../images/eia_lcoe.png}}; \vspace{1cm}
+   %\\[\abovecaptionskip]  \vspace{5.5cm}
+   %\tiny{See for more information: \url{https://www.lazard.com/media/451086/lazards-levelized-cost-of-energy-version-130-vf.pdf}}
+   \vfill
+\end{frame}
+
+\begin{frame}{Costs of New Capacity Additions}
+   \tikz [remember picture,overlay]
+    \node[yshift=-.25cm,xshift=0cm] at (current page.center)
+        {\includegraphics[width=.9\paperwidth]{../images/lazard_lcoe.png}}; \vspace{1cm}
+   %\\[\abovecaptionskip]  \vspace{5.5cm}
+   %\tiny{See for more information: \url{https://www.lazard.com/media/451086/lazards-levelized-cost-of-energy-version-130-vf.pdf}}
+   \vfill
+\end{frame}
+
+\begin{frame}{New Technology}
+   \tikz [remember picture,overlay]
+    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
+        {\includegraphics[width=.9\paperwidth]{../images/ab_solar.png}}; \vspace{1cm}
    \vfill
@@
         {\includegraphics[width=.9\paperwidth]{../images/price_capture.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Price Capture by Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/price_capture_avg.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{Evolution of Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/eia_aeo_solar.png}}; \vspace{1cm}
    \vfill
@@
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.95\paperwidth]{../images/gen_ghg_price.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\begin{frame}{GHG Policy and Electricity Supply}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
         {\includegraphics[width=.95\paperwidth]{../images/monthly_ghgs.png}}; \vspace{1cm}
@@
 
-\begin{frame}{GHG Policy and Electricity Supply}
+\begin{frame}{Deeper cuts: the roles of storage and transmission}
+\begin{itemize}
+\setlength\itemsep{2em}
+\item Renewables now offer some of the cheapest electricity we have ever seen
+\item Renewables are not (generally) dispatchable
+\item Renewables seasonal and/or daily generation patterns don't always match load
+\item The sun is always shining / the wind is always blowing somewhere
+\end{itemize}
+
+\vfill \end{frame}
+
+\begin{frame}{Deeper cuts: the roles of storage and transmission}
+\begin{itemize}
+\setlength\itemsep{2em}
+\item How do we overcome the need for more transmission or storage?
+\item Who will pay for the assets? How will the assets be paid for the services they provide?
+\item Renewables, storage, and even transmission can erode their own value proposition - with a lot of storage or transmission in place, it's harder to see the value for storage
+\item The value of transmission and storage assets may not be captured by the jurisdiction or the regulatory sector in which they are built
+
+\end{itemize}
+
+\vfill \end{frame}
+
+\begin{frame}{Readings and guest speaker}
+\begin{itemize}
+\setlength\itemsep{2em}
+\item Rivers and Dolter paper on decarbonizing Canada's supply
+\item van de Biezenbos paper on transmission.
+
+\end{itemize}
+
+\vfill \end{frame}
+
+\end{document}
+
+\begin{frame}{Price Volatility}
    \tikz [remember picture,overlay]
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.95\paperwidth]{../images/monthly_ghg_mwh.png}}; \vspace{1cm}
+        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_2000_2021.png}}; \vspace{1cm}
    \vfill
@@
 
-%\begin{frame}{GHG Emissions from Electricity}
-%   \tikz [remember picture,overlay]
-%    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-%        {\includegraphics[width=.8\paperwidth]{../inventory_ghg_mwh.png}}; \vspace{1cm}
-%   \vfill
-%\end{frame}
+\begin{frame}{Price Volatility}
+   \tikz [remember picture,overlay]
+    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
+        {\includegraphics[width=.9\paperwidth]{../images/hourly_prices_col.png}}; \vspace{1cm}
+   \vfill
+\end{frame}
 
-%\begin{frame}{GHG Emissions from Electricity}
-%   \tikz [remember picture,overlay]
-%    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-%        {\includegraphics[width=.8\paperwidth]{../elec_share.png}}; \vspace{1cm}
-%   \vfill
-%\end{frame}
+\begin{frame}{Growth}
+   \tikz [remember picture,overlay]
+    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
+        {\includegraphics[width=.9\paperwidth]{../images/load_time.png}}; \vspace{1cm}
+   \vfill
+\end{frame}
 
-\end{document}
+\begin{frame}{Economics 101}
+   \tikz [remember picture,overlay]
+    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
+        {\includegraphics[width=.9\paperwidth]{../images/coal_merit.png}}; \vspace{1cm}
+   \vfill
+\end{frame}
+
+\begin{frame}{Economics 101}
+   \tikz [remember picture,overlay]
+    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
+        {\includegraphics[width=.9\paperwidth]{../images/all_coal_merit.png}}; \vspace{1cm}
+   \vfill
+\end{frame}

### 2022-01-18 paper/paper_draft.tex

@@
 
+These results are consistent if a decomposition by peak- vs.
+off-peak hours is introduced, as shown in Figure \ref{fig:decomp_peaks}.
+The impact of carbon pricing is always statistically greater than or equal to zero, with a maximum pass-through of approximately 40\% of average carbon costs in the mid-merit, and less elsewhere.
+In peak hours, the pass-through is smaller in the 50th to 60th percentiles, and larger in the 60th to 75th percentiles, with a similar 40\% maximum pass-through.
+Carbon price pass-through remains low in the 80th percentile, and increases in the upper merit during peak hours, while being close to zero in off-peak hours except at the extreme top of the merit order.
+The upper-merit impacts of OBAs are counter-intuitive, especially at peak hours, with estimates suggesting that higher offers have occurred with more valuable OBAs in the 85th to 95th percentiles of power offers in peak hours.
+
 \begin{figure}[!htb]
@@
      \includegraphics[width=.9\textwidth]{../images/all_plants.png}
-    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types and all hours. Error bars denote 95\% confidence intervals}
+    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types. Error bars denote 95\% confidence intervals}
     \label{fig:decomp_peaks}
@@
 \subsection{Impacts By Generator Type}
-We know that different generator types tend to offer their power into the market in different ways, and so our most detailed results are based on a decomposition of the market-wide impacts into impacts by type of plant.
-Here, we find that...
+We know that different generator types tend to offer their power into the market in different ways, and we decompose our key results by plant type, concentrating on the fossil fuel generators, with the exception of the combined heat and power plants for whom we do not observe external drivers for power offers.
+The results by plant type are, in many cases, counter-intuitive.
+We find first that coal plants do not consistently pass-through carbon costs to their offers.
+In fact, in the mid-merit, they tend to shade their offers a bit, actually reducing their offers as carbon prices are increased.
+This is likely as a result of their direct competition with gas-fired power plants which have lower emissions-intensity.
+In order to remain in the market at their maximum efficiency, it appears that coal plants are not passing through any carbon costs and are, in fact, reducing their offers despite an increase in costs.
+In the upper ranges of the merit order, coal plants have higher than full pass through, indicating that their offers over the last megawatts of their portfolio increased by more than their net carbon costs during the sample period.
+This indicates a change in operating stategy, perhaps, from one of \textit{base load} to more peaking behaviour, only operating plants at full capacity when the market price is very high.
+This is also, likely, an artefact of our assumption of constant emissions intensity of output.
+If the marginal emissions-intensity of output at the 90th percentile of installed capacity is much higher than in the mid-range of capacity, we would expect our pass-through estimates to appear as though more than 100\% of these costs were being passed-through, although that alone is insufficient to explain offers increasing by 1-3 times the increase in net carbon costs.
+
+\begin{figure}[!htb]
+    \centering
+     \includegraphics[width=.9\textwidth]{../images/by_type.png}
+    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types. Error bars denote 95\% confidence intervals}
+    \label{fig:decomp_peaks}
+\end{figure}
+
+The natural gas, combined cycle power plants respond as we would have expected, at least qualitatively, with limited pass-through in the lower part of the merit order, where plants are generally offering in a minimum capacity at a lower strike price, an then with more pass-through in the mid- and upper-merit up to the 95th and 100th percentiles, where other strategic factors drive offers.
+Here, however, the quantitative effects are strikingly small.
+Natural gas power plants passed through less than 20\% of the net costs of carbon policies to their offers.
+
+For the simple-cycle plants, the results are the reverse of what we might expect.
+For these plants, some of which operate as peaking plants, we would expect to see a near-perfect pass through or, at the very least, something similar.
+Instead, we see almost all sampling points in the offers from simple cycle plants showing a negative relationship with net carbon costs: higher carbon prices were equated with not only lower offers but quantitatively significant reductions in the offered price level with increases in the carbon cost.
+
+There are a few factors at play that may explain this counter-intuitive result: first, some of the simple-cycle gas turbines are relatively new, so we do not observe all plants over the entire sample space.
+In and of itself, this would not be important given our sampling technique, but many of the newer simple-cycle plants have been added as a consequence of emissions policies, as alternatives to venting or flaring of methane from oil field operations.
+Additionally, many simple-cycle plants operate in the ancillary services market, capturing value as spinning reserves rather than when generating, and so this will skew our estimated emissions intensities since these plants would report emissions for periods in which they are burning fuel but not generating power.
+Finally, as coal plants have retired and/or shifted to higher offers, some simple-cycle plants would see an advantage to producing in more hours of the day, such that strategic considerations would lead them to lower their offers despite an increase in carbon costs in order to ensure a higher annual capacity factor.
+These plants will also be most sensitive to the hourly opportunity costs of natural gas and, while we observe hourly spot prices, individual entities may face different gas costs on an hourly basis than what we observe, in particular if facilities have long-term contracts for delivery in place.
 

### 2022-01-24 paper/paper_draft.tex
drafting

@@
     colorlinks,
-    citecolor= black,
-    filecolor= black,
-    linkcolor= black,
-    urlcolor= black
+    citecolor= blue,
+    filecolor= blue,
+    linkcolor= blue,
+    urlcolor= blue
 }
+
 \usepackage{setspace}
@@
 
-\usepackage[multiple]{footmisc}
+\usepackage[hang,flushmargin]{footmisc}
 
@@
 
+\urlstyle{same}
+
 \specialcomment{ALnote}{\begingroup\sffamily\colorbox{blue}{Andrew: } \color{blue}}{\endgroup}
@@
 %%%% TITLE %%%%
-\title{Carbon Price Pass-Through in Alberta's Electricity Market\thanks{We are grateful to Sam Harrison for excellent
+\title{Marginal and Average Carbon Price Pass-Through\thanks{We are grateful to Sam Harrison for excellent
 research assistance.
-We thank seminar and conference participants at the 2019 AERE Summer Meetings for comments and suggestions.
-Financial support for this work was provided by the Future Energy Systems Canada First Research Excellence Fund grant.
-}
+We thank seminar and conference participants at the AERE Summer Meetings for comments and suggestions.
+Financial support for this work was provided by the Future Energy Systems Canada First Research Excellence Fund grant held by the University of Alberta.}
 }
@@
 We evaluate the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.
-Between 2015 and 2019, legislative changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes, creating variation in average and marginal carbon prices both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
-We exploit this variance in treatment to provide unique evidence of offer response to carbon pricing in power markets.
+Between 2015 and 2019, legislative and regulatory changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes, creating variation in average and marginal carbon prices both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
+We exploit this variance in treatment to provide unique evidence of carbon price pass-through in power markets.
 We show that even after adjusting for costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.
+These changes have accelerated the phase-out of coal-fired power in the province.
 
-The Alberta wholesale power market provides an excellent laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis, so there are limited confounding factors beyond what is observable in our data.
-Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.\footnote{The 2021 maximum Alberta internal load per AESO 2021 data was 11,729 MW at time of writing, which is also the peak value ever recorded in Alberta.} Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
+The Alberta wholesale power market provides an excellent quasi-laboratory because it is small, isolated, and the market is settled on a real-time, energy-only basis.
+As such, there are limited confounding factors beyond what is observable in our data.
+Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.\footnote{The 2021 maximum Alberta internal load per AESO 2021 data was 11,729 MW at time of writing, which is also the peak value ever recorded in Alberta.} Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint operating capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
 The downside to this is that Alberta, a major oil-producing jurisdiction, was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies, and this economic downturn led to a period of low prices in the power market.
@@
 
-We compile data describing market offer behaviour and a wide range of relevant covariates.
-We measure and describe facility-level hourly offers of power into the market over slightly more than 10 years preceding the COVID-19 pandemic.
+We compile over 10 years of hourly data describing market offer behaviour and a wide range of relevant characteristics.
+We measure and describe facility-level hourly offers of power into the market for the time period preceding the COVID-19 pandemic.
 We combine these data with detailed plant characteristics compiled from regulatory data, emissions compliance data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data including hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.
@@
 
-In order to isolate the impact of the changes in policies on offer behaviour, we develop a unique empirical strategy based on synthetic power plants and synthetic power portfolios.
-For each hour of each day, we create synthetic power portfolios either by plant type .
-We are then able to examine, for example, how changes in policies affected the offer of power across all coal- or gas-fired plants in the province.
-
-or by which entity holds offer control on the units in question
-, or across all facilities owned by particular actors in the market.
-We normalize these portfolios by percentile, so we are able to compare behaviour across portfolios of different sizes without loss of generality.
+In order to isolate the impact of the changes in policies on offer behaviour, we develop a unique empirical strategy...
 
@@
 
-\section{Alberta's GHG Policy Changes}
+The paper proceeds as follows.
+In Section \ref{sec:policy}, we detail Alberta's greenhouse gas policy history.
+In Sections \ref{sec:data} and \ref{sec:market} we detail our data set and provide a statistical portrait of the province's power market.
+In Section \ref{sec:empirical} we detail our empirical strategy.
+Section \ref{sec:results} summarizes our results.
+Section \ref{sec:conclusion} concludes.
 
-Alberta has had carbon pricing in place since the \emph{Specified Gas Emitters Regulation} (\emph{SGER}) took effect on July 1, 2007.
-That regulation, the first industrial carbon price in North America, implemented a price of \$15/tonne, and allocated emissions credits to covered facilities at a rate equal to 88\% of an individual facility's historic (2003-2005) emissions intensity.
+\section{Alberta's GHG Policy Changes}\label{sec:policy}
+
+Alberta has had carbon pricing in place since the \emph{Specified Gas Emitters Regulation} (\emph{SGER}) took effect on July 1, 2007.\footnote{\textit{Specified Gas Emitters Regulation}, Alta Reg 139/2007, $<$\url{canlii.ca/t/52x2q}$>$.} That regulation, the first industrial carbon price in North America, implemented a price of \$15/tonne, and allocated emissions credits to covered facilities at a rate equal to 88\% of an individual facility's historic (2003-2005) emissions intensity.
 For new facilities built after 2005, they were allocated emissions credits per unit output at a rate equal to 88\% of their average year 3 emissions intensity.
@@
 
-This system remained in place until June, 2015 when the government of Premier Rachel Notley introduced a series of changes to the existing regulation.
-The first set of changes increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for the output-based allocation of emissions credits to 85\% and 80\% of historic facility level emissions intensity for 2016 and 2017 respectively.
+This system remained in place until June, 2015 when the government of Premier Rachel Notley introduced a series of changes to the existing regulation.\footnote{\textit{Specified Gas Emitters Amendment Regulation}, Alta Reg 105/2015, $<$\url{perma.cc/Z2ZM-WDWP}$>$.} The first set of changes increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for the output-based allocation of emissions credits to 85\% and 80\% of historic facility level emissions intensity for 2016 and 2017 respectively.
 Combined, these changes implied a material increase in the average cost of carbon in each of the years 2016 and 2017.
@@
 Two changes in this iteration of policies affected power markets.
-Most importantly, the \textit{Carbon Competitiveness Incentive Regulation (CCIR)} replaced the \emph{SGER} and these regulations levelized the output-based allocation of emissions credits across all power generators at 0.37t/MWh, the emissions intensity of the best-in-class combined cycle natural gas generation facility in the province.
-This implied that coal producers saw a steep increase in their average costs of carbon, while impacts on gas power plants varied depending on the heat rates of the facility.
+Most importantly, the \textit{Carbon Competitiveness Incentive Regulation (CCIR)} replaced the \emph{SGER} and these regulations levelized the output-based allocation of emissions credits across all power generators at 0.37t/MWh, the emissions intensity of the best-in-class combined cycle natural gas generation facility in the province.\footnote{\textit{Carbon Competitiveness Incentive Regulation}, Alta Reg 255/2017, $<$\url{canlii.ca/t/53h9v}$>$.} This implied that coal producers saw a steep increase in their average costs of carbon, while impacts on gas power plants varied depending on the heat rates of the facility.
 Both combined heat and power plants and existing renewable generation facilities saw the value of their emissions credits issued per MWh generated decrease under the \emph{CCIR}.
@@
 The \emph{CCIR} remained in place until it was replaced by the \emph{
-Technology Innovation and Emissions Reduction Regulation} (\emph{TIER}) on January 1, 2020, outside of our sample period.
-However, it's important to not that while \emph{TIER} altered the design of output-based allocations in most sectors, it did not impose material changes in the electricity industry and so we should not expect any changes late in our sample period which abuts the implementation of \emph{TIER}.
+Technology Innovation and Emissions Reduction Regulation} (\emph{TIER}) which took effect on January 1, 2020, outside of our sample period.\footnote{\textit{Technology Innovation and Emissions Reduction Regulation}, Alta Reg 133/2019, $<$\url{canlii.ca/t/54qgg}$>$.}  However, it's important to note that while \emph{TIER} altered the design of output-based allocations in most sectors, it did not impose material changes in the electricity industry and so we should not expect any changes late in our sample period which abuts the implementation of \emph{TIER}.
+\emph{TIER} remains in place as of this writing, with the price on carbon having increased to \$40 per tonne for 2021 and \$50 per tonne for 2022.\footnote{See Alberta Environment and Parks Ministerial Order 87-2021 under the \emph{Emissions Management and Climate Resilience Act}, SA 2003, c.
+E-7.8, $<$\url{perma.cc/3HN3-H9RC}$>$.}
 

### 2022-01-24 paper/paper_draft.tex
drafting

@@
 
-\section{Data}
+\section{Data}\label{sec:data}
 
 The majority of data for this paper comes from the Alberta Electricity System Operator (AESO).
-The AESO provides, with a 60-day lag, information on the offers made by power plant owners or controlling entities into the pool on an hourly basis.
+The AESO provides, with a 60-day lag, information on the offers made by power plant owners or controlling entities into the wholesale power pool on an hourly basis.
 Facilities offer their power in up to 7 increasing-price blocks, with a price cap of \$1000/MWh and a floor of \$0/MWh.
 Plants are dispatched according the merit order of these offer blocks, from lowest to highest offered prices.\footnote{Blocks may be either flexible or not, and non-flexible blocks will only be dispatched when demand allows the entire block to be used for the hour.
-If what would otherwise be the marginal block is not a flexible block, the next-highest-priced offer block will be dispatched.} We have hourly merit order and dispatch data from September 1, 2009 through May 31, 2021, but we truncate the data at December 31, 2019 to avoid impacts of the COVID-19 pandemic.\footnote{For a discussion of Alberta market responses to the COVID-19 pandemic, see Leach et al.
-(2020).} The data also allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.\footnote{The offer control identifiers are only published after 2013, so our analysis on these data takes place over a smaller sample.}
+If what would otherwise be the marginal block is not a flexible block, the next-highest-priced offer block will be dispatched.} We have hourly merit order and dispatch data from September 1, 2009 through December 31, 2021, but we truncate the data at December 31, 2019 to avoid confounding impacts of the COVID-19 pandemic.\footnote{For a discussion of Alberta market responses to the COVID-19 pandemic, see Leach et al.
+(2020).} The data also allows us to identify, by block, which entity had offer control in the market for that block of power in each hour.\footnote{The offer control identifiers are only published after 2013, so our analysis based on these data takes place over a smaller sample period.}
 
 We also use three other AESO data sets to build our analysis sample.
-First, the AESO issues hourly price and load data, including 3-hour ahead forecast and actual prices and loads, which we merge with merit order data.
-Next, we include intertie capability rating data which allow us to account for hours of limited import or export capacity.
+First, the AESO issues hourly price and load data, including 3-hour ahead forecast and actual prices and day-ahead forecast and actual loads, which we merge with merit order data.
+Next, we include intertie capability rating data which allows us to account for hours of limited import or export capacity.
 Finally, and most importantly, we use metered volumes data at the facility level to incorporate renewable generation into the merit order.
 The AESO treats non-dispatchable renewables (wind and solar power in our case) as negative load, but lists them in the merit order data as \$0 offers at full nameplate capacity.
-For each facility-hour pair in our data set, we merge actual metered generation into the merit order and treat renewable generators as having offered exactly their metered volumes each hour, still at a \$0MWh offer.
+For each facility-hour pair in our data set, we merge actual metered generation into the merit order and treat renewable generators as having offered exactly their metered volumes each hour, still at a \$0/MWh offer price.
 
@@
 Where we have no information at all, we assume an emissions-intensity of 0.55 tonnes per MWh for simple cycle natural gas plants, the only type of plant for which we lack information of this type, based on facilities for which we do have information in our sample.
-The second source of incompleteness comes from the fact that data are reported at the facility level, not by generating unit.
+
+The second source of incompleteness comes from the fact that emissions are reported annually at the facility level, not by generating unit.
 We apply the same emissions intensity to multiple generating units within a facility.\footnote{There are two exceptions to this.
-The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.} Since we have only annual data at the facility level, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units.
+The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.} Since we have only annual data at the facility level, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units nor with the capacity factor at any point in time.
+This will lead to under-reporting of emissions-intensity when units are ramping up or generating below their average efficiency, and over-reporting of emissions-intensity for units operating above their annual average operating efficiency.
+This is analogous to an assumption that operators do not account for variations in emissions intensity in making operating or offer decisions at the facility level, or that they account only for the average emissions intensity in calculating marginal costs.
+
 Finally, combined heat and power units present a challenge since many units report their emissions for the entire production facility, not simply for the power plant.
@@
 The emissions reporting data tell us the total electricity and heat produced in these facilities, and the emissions attributed to them, but do not tell us what the emissions intensity of an incremental unit of electricity would be.
-The output-based allocations for cogeneration facilities are determined by the carbon pricing policies, so we know the value of these for each regulated entity.
+The output-based allocations for these facilities are determined by the carbon pricing policies, so we know the value of these for each regulated entity.
 For the purposes of our analysis, we adopt an assumed emissions intensity of electricity of 0.418 tonnes per MWh for combined heat and power units.
@@
 The data also provide the facility-specific rate at which output-based allocations were provided in each year, and the carbon price faced on emissions net of output-based allocations.
-The carbon price is constant for all Alberta-based facilities, but the combined-cycle plant in Fort Nelson, BC faces the BC carbon tax while being connected to the Alberta electricity grid, and so receives no output-based allocations and faced higher carbon prices in many years of the sample than comparable plants in Alberta.
+The carbon price is constant for all Alberta-based facilities, but the combined-cycle plant in Fort Nelson, B.C.
+faces the BC carbon tax while being connected to the Alberta electricity grid, and so receives no output-based allocations and faced higher carbon prices in many years of the sample compared to plants in Alberta.
 
@@
 
-\section{Alberta's Power Market}
+\section{Alberta's Power Market}\label{sec:market}
 
@@
 Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_load}, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
-There was also a major, short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, which explains the markedly-lower spring trough in that year's load profile as shown in Figure \ref{fig:ab_load}.
+There was also a major, short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, leading to the markedly-lower spring trough in that year's load profile as shown in Figure \ref{fig:ab_load}.
 The 2020 COVID-19 demand shock is also evident in Figure \ref{fig:ab_load}, but lies outside the sample period for our analysis.
@@
 Our sample period covers ten years with no major regulatory changes in the power market design.
-Alberta briefly considered an addition of a capacity market, but this was never implemented.
+Alberta briefly considered an addition of a capacity market between 2016 and 2019, but this was never implemented.
 However, market conditions have varied over the sample period.
@@
 These conditions are reflected in peak and off-peak prices shown in Figure \ref{fig:ab_prices}.
-Some of these market changes coincide with policy changes of interest in our paper, in particular the changes to the \emph{SGER} introduced in June of 2015 which took effect in January of 2016 and 2017 respectively and are coincidental with a period of over-supply, and the introduction of the \emph{CCIR} in 2018 which occurred during a period of tightening reserve margins and lingering uncertainty over market structure.
+Some of these market changes coincide with policy changes of interest in our paper, in particular the changes to the \emph{SGER} introduced in June of 2015 which took effect in January of 2016 and 2017 respectively and are coincidental with a period of over-supply.
+Similarly, the introduction of the \emph{CCIR} in 2018 occurred during a period of tightening reserve margins and lingering uncertainty over market structure.
 From 2018 through pre-COVID 2020, both GHG policy and the wholesale electricity market were tightening, and so we need to be careful to disentangle these impacts.

### 2022-01-25 paper/paper_draft.tex
graph fixes

@@
 
-\begin{figure}[t]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_area_grey.png}
+\begin{figure}[h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel.png}
 \vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: AESO data provided by NRGStream, authors' graph.}
@@
 Within natural gas generation, the mix of plant types is also relevant to our study.
-By the end of our sample period, a significant increase in generating capacity from combined cycle (NGCC) plants had been installed, and this is reflected in higher generation from those sources.
-The installed capacity of combined heat and power (COGEN) facilities also grew during the sample period, and these plants tend to operate with high capacity factors but also tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.
-As will be discussed below, net-to-grid power from COGEN facilities is primarily offered into the market at a \$0 offer and accepts market price, and so we see limited evidence of carbon price pass-through for most of the power offered by these facilities.
-On the other hand, peaking capacity in the market during our sample period was largely met through ramping of coal and combined-cycle plants, as well as simple cycle (SCGT) natural gas turbines.
+During our sample period, a significant increase in generating capacity from combined cycle (NGCC) plants and combined heat and power (COGEN) facilities occurred.
+Combined heat and power plants tend to operate with high capacity factors but also tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.\footnote{As will be discussed below, net-to-grid power from COGEN facilities is primarily offered into the market at a \$0 offer and accepts market price, and so we see limited evidence of carbon price pass-through for most of the power offered by these facilities.} Combined-cycle plants, on the other hand, provided more flexible natural gas generation on the margin.
+Peaking capacity in the market during our sample period was largely met through ramping of coal and combined-cycle plants, as well as simple cycle (SCGT) natural gas turbines.
 Each of these would be exposed to carbon pricing, and so we expect to see pass through affect offers of power from these facilities into the market.
@@
 In the left-hand panel, we see that there is less variability through the day in Alberta as industrial load dominates cycles in residential and commercial load.
-Seasonal patterns are also both less pronounced and exhibit a winter peak, compare to summer-peaking ERCOT and MISO systems.
+Seasonal patterns are also both less pronounced and exhibit a winter peak, compared to the summer-peaking ERCOT and MISO systems.
 
@@
 
-The nature of Alberta's wholesale market is such that different generators will derive different levels of average revenue from the sale of their power.
+Alberta's wholesale market is such that different generators will derive different levels of average revenue from the sale of their power.
 Those that dispatch only in high-price times will have higher revenues on average than plants without dispatch control, or plants which run most hours of the day.
 Since prices have varied substantially over time, as shown above in Figure \ref{fig:ab_prices}, captured average prices at the plant level have varied as well, as shown in Figure \ref{fig:capture_prices}.
-We can see evidence here of how plants of different types offer their power into the market: hydro and combined- and simple-cycle gas turbines follow prices, and dispatch more capacity in higher-priced hours, just as imports tend to occur when prices are higher.
-Lower prices tend to prevail when wind generation is highest, and so captured prices for wind power facilities are lower than other forms of generation, and exports tend to be scheduled when prices are expected to be low.
+We can see evidence here of how plants of different types offer their power into the market: hydro and combined- and simple-cycle gas turbines follow prices, and dispatch more often in higher-priced hours, just as imports tend to occur when prices are higher.
+Prices tend to be negatively correlated with wind generation, and so captured prices for wind power facilities are lower than other forms of generation.
+Exports tend to be scheduled when prices are expected to be low, which is reflected in the graph.
 Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewable sources since its generation pattern coincides with daily and seasonal peak prices, although we would expect this to change as \emph{duck curve} effects take hold over time.
@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture.png}
-\vspace{-0.75cm}	\caption{Annual average revenue by plant type. source: AESO data provided by NRGStream, authors' graph.}
+\vspace{-0.75cm}	\caption{Annual average revenue by plant type. Source: AESO generation data provided by NRGStream, authors' graph.}
 \label{fig:capture_prices}
@@
 The net effect of these policies creates variation within and across facilities and facility types and over time.
-Consider, for example, Figure \ref{fig:sector_compliance_cost} which shows the mean value and range of annual compliance costs for larger generators in Alberta's fleet over the sample period.\footnote{Plants were only included in the sample for this figure if their annual emissions were sufficiently large ($\geq$100kt CO$_2$e per year) such that we have specific compliance cost data.
+Consider, for example, Figure \ref{fig:sector_compliance_cost} which shows the mean value and range of annual compliance costs for larger generators in Alberta's fleet over the sample period.\footnote{Plants were only included in the sample for this figure if we had facility-specific compliance cost data.
 For our estimation results, we impute compliance costs for a wider set of facilities using emissions intensities from the compliance cost data.}
@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../SGER/sector_compliance_cost.png}
-\vspace{-0.75cm}	\caption{Upper: Annual average compliance cost (line) and range (shaded area) by plant type.
+\vspace{-0.75cm}	\caption{Upper: Generation-weighted annual average compliance cost (black lines) and individual plant realizations (lighter grey lines) by plant type.
 Lower: Price charged on excess net annual emissions under Alberta's carbon pricing programs.
-Source: Alberta Environment compliance cost data}
+Source: Alberta Environment compliance cost data.}
 \label{fig:sector_compliance_cost}
@@
 
-As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices, but are actually dominated by the changes in output-based allocations which occurred subsequent to the adoption of the CCIR in 2018 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe in 2016 and 2017.
-The largest cost increases affect coal-fired power plants and some simple-cycle generators, while limited effects were felt by combined heat and power plants.
+As Figure \ref{fig:sector_compliance_cost} shows, changes in facility compliance costs were not all coincident with increased marginal carbon prices.
+In fact, the compliance cost changes are mostly determined  by the changes in output-based allocations which occurred subsequent to the adoption of the \emph{CCIR} in 2018 rather than the initial increase carbon prices from \$15 to \$30 per tonne \coe in 2016 and 2017.
+The largest cost increases were experienced by coal-fired power plants and some simple-cycle generators, while limited effects were felt by combined heat and power plants (not shown).
 Compliance costs for natural gas combined-cycle plants dropped between 2017 and 2019 as a result of a change in the formula by which output-based allocations were calculated.
+These changes in compliance costs, across and within generators, provide the variation through which we identify pass-through to power offers.
 
@@
 a 350MW plant vs a 200MW plant) while facing similar optimization constraints, and they may be jointly dispatched by the same operator.
-In particular for coal plants, plant-level minimum-must-run constraints imply that facilities will always offer some portion of their generation into the market at a \$0/MWh price, effectively acting as price takers for that hour for that block of power.
-They will offer marginal blocks of power, for capacity above their minimum-must-run level, at higher prices.
+In particular for coal plants, plant-level minimum-must-run constraints imply that facilities will always offer some portion of their generation into the market at a \$0/MWh price, effectively acting as price takers for an endogenously-determined share of their generation.
+They may offer marginal blocks of power, for capacity above their minimum-must-run level, at higher prices.
 However, given that plants are of different sizes, and blocks are endogenously determined, it's challenging to identify a change in offers looking only at block offers.
@@
 Finally, since multiple plants may be controlled by a single entity, the observed data are the product of a complex portfolio optimization problem and the offers of any individual plant may not reveal an overall portfolio-level attempt to pass through the costs of carbon prices to wholesale prices.
-We avoid these problems by constructing synthetic offer curves either by plant-type or by controlling entity, thus estimating the impact of policy changes on the supply behaviour of the aggregate fleet rather than on any individual plant.
-We can then use the synthetic merit order data we construct to examine how power has been offered into the system over time as climate policies have varied.
+We avoid these problems by constructing synthetic offer curves either by plant-type and by controlling entity, thus estimating the impact of policy changes on the supply behaviour of the aggregate fleet rather than on any individual plant.
+We then use the synthetic merit order data we construct to examine how power has been offered into the system over time as climate policies have varied.
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}[t]
     \centering
@@
 
-There is a lot of day-to-day and hour-to-hour variability in how coal are offered into the market.
+There is a lot of day-to-day and hour-to-hour variability in how coal units are offered into the market.
 Figure \ref{fig:synth_coal_2019}, which shows all of the 7pm merit orders (left panel) and their synthetic equivalents, along with the average synthetic merit order for the last year of our sample (right panel), provides a sense of the variability in our sample.
@@
 We use this information to construct synthetic supply curves for each of the major participants in the Alberta power market.
-This means that we have facility-level, generator-class-level and market-level synthetic merit orders to analyze.
+This means that we have facility-level, firm- or portfolio-level and market-level synthetic merit orders to analyze.
 

### 2022-01-25 aeso_scrapes.R

@@ -315,6 +315,8 @@ get_forecast_report <- function(start_date,end_date) {
   clean_data$date<-as_date(clean_data$time)
   clean_data$forecast_ail<- gsub("\\,", "", clean_data$forecast_ail)
   clean_data$actual_ail<- gsub("\\,", "", clean_data$actual_ail)
+  clean_data <-clean_data %>% rename(day_ahead_forecasted_ail=forecast_ail)
+  clean_data <-clean_data %>% rename(forecasted_actual_ail_difference=forecast_ail_actual_ail_difference)
   #clean_data<-clean_data %>% select(-he) %>% mutate_if(is.character,as.numeric)
   #clean_data<-clean_data %>% select(-he)%>% mutate_if(is.integer,as.numeric)
   #set numeric columns to numeric
@@ -350,7 +352,9 @@ all_forecasts<-function() {
 
 #all_forecasts()
 #load("forecast_data.Rdata")
-#save(forecast_data, file= filename) 
+#forecast_data <-forecast_data %>% rename(day_ahead_forecasted_ail=forecast_ail)
+#forecast_data <-forecast_data %>% rename(forecasted_actual_ail_difference=forecast_ail_actual_ail_difference)
+#save(forecast_data, file= "forecast_data.Rdata") 
 
 update_forecasts<-function() {
   #load the existing file

### 2022-01-25 nrgstream_graphs.R

@@ -98,7 +98,6 @@ sub_samp$Year<-year(sub_samp$time)
 df1 <- sub_samp %>% group_by(Plant_Type,time,Year) %>% summarise(sumcap = sum(Capacity),total_gen=sum(gen),p_mean=mean(Price))
 df1$Year_ID=as.character(df1$Year)
 # Histogram of Generation Densities
-set_png(file="images/wind_cdf.png")
 ggplot(df1,aes(total_gen))+
   #geom_density(aes(fill="Wind Power Generation",colour=year(time)),alpha=0.5)+
   #stat_density(geom="line",position="identity",aes(group=Year_ID,colour=Year_ID),size=1.5)+
@@ -110,7 +109,7 @@ ggplot(df1,aes(total_gen))+
   labs(x="Wind Generation (MW)",y="% of hours generation < X MW",
        title="Cumulative Density Function, Wind Energy (2010-2017 Avg)",
        caption="Source: AESO Data, Accessed via NRGStream, Graph by Andrew Leach")
-dev.off()
+ggsave(file="images/wind_cdf.png")
 
 
 
@@ -140,7 +139,7 @@ gen_set<-c("COAL","COGEN","HYDRO","NGCC", "OTHER", "SCGT","SOLAR","IMPORT","EXPO
 #test_samp2<-test_samp %>% filter(Plant_Type %in% gen_set,! NRG_Stream %in% trade_excl)
 
 
-df2 <- df1 %>% filter(Plant_Type %in% gen_set,year(time)<2021) %>%
+df2 <- df1 %>% filter(Plant_Type %in% gen_set,year(time)<2022) %>%
        group_by(Plant_Type,Year) %>% summarise(capture = sum(total_rev)/sum(total_gen),avg_rev = sum(total_rev)/sum(total_gen),p_mean=mean(p_mean))
 
 
@@ -203,7 +202,7 @@ ggplot(df2,aes(Year,capture/p_mean*100-100,colour=Plant_Type,fill=Plant_Type),al
 dev.off()
 
 
-df3 <- df1 %>% filter(Plant_Type %in% gen_set,year(time)<2021) %>%
+df3 <- df1 %>% filter(Plant_Type %in% gen_set,year(time)<2022) %>%
   group_by(Year) %>% summarise(capture = sum(total_rev)/sum(total_gen),avg_rev = sum(total_rev)/sum(total_gen),p_mean=mean(p_mean))%>%
[…]

### 2022-01-25 paper_graphs.R

@@ -134,34 +134,41 @@ gen_plain <- df2 %>% mutate(Plant_Type=factor(Plant_Type,levels=AB_plant_order))
   
   
   
+  AB_palette<- c("black","black","grey50","grey50")
+  
     
   gen_fuel <- df2 %>% mutate(Plant_Type=factor(Plant_Type,levels=AB_plant_order))%>%
     #filter(date<ymd("2020-09-01"))%>%
     mutate(Plant_Type=fct_collapse(Plant_Type,
-     #"OTHER"=c("WIND","OTHER","HYDRO"),
+     "RENEWABLES"=c("WIND","OTHER","HYDRO"),
      "NATURAL GAS"=c("SCGT","COGEN","NGCC"),
      "NET IMPORTS"="TRADE"
      ))%>% 
-    group_by(date,month,year,Plant_Type) %>% summarise(gen=sum(gen,na.rm = T))%>% 
+    group_by(date,month,year,Plant_Type) %>% summarise(gen=sum(gen,na.rm = T),
+                                                      )%>%
     ungroup() %>%
+    group_by(Plant_Type) %>%
+    mutate(gen12m=zoo::rollmean(gen,12,fill=NA))%>%
     #filter(date>ymd("2014-12-31"))%>%
-    ggplot(aes(date,gen, col = Plant_Type,fill = Plant_Type,shape=Plant_Type)) +
+    ggplot(aes(date,gen, col = Plant_Type,lty=Plant_Type)) +
     geom_line(size=1.25)+
-    geom_point(aes(date,gen*ifelse(month%%2==0,1,NA)),size=2.5)+
-    geom_dl(aes(label=Plant_Type),method=list("last.bumpup",dl.trans(x=x+0.3),cex = .85))+
+    #geom_line(aes(y=gen12m),size=1.25)+
+    #geom_point(aes(date,gen*ifelse(month%%2==0,1,NA)),size=2.5)+
+    #geom_dl(aes(label=Plant_Type),method=list("last.bumpup",dl.trans(x=x+0.3),cex = .85))+
     scale_color_manual("",values= AB_palette)+
-    scale_fill_manual("",values= AB_palette)+
-    
+    #scale_fill_manual("",values= AB_palette)+
+    scale_linetype_manual("",values= c("solid","1131","11","solid"))+
     #scale_color_manual("",values=grey.colors(9,start=0,end=.8))+
[…]

### 2022-01-27 paper/paper_draft.tex
fixing graphs

@@
 	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../SGER/sector_compliance_cost.png}
-\vspace{-0.75cm}	\caption{Upper: Generation-weighted annual average compliance cost (black lines) and individual plant realizations (lighter grey lines) by plant type.
-Lower: Price charged on excess net annual emissions under Alberta's carbon pricing programs.
-Source: Alberta Environment compliance cost data.}
+\vspace{-0.75cm}	\caption{Generation-weighted annual average compliance cost (dashed black lines) and individual plant realizations (lighter grey lines) by plant type (\$/MWh), combined with the price charged on excess net annual emissions (black lines) under Alberta's carbon pricing programs (\$/tonne).
+Coal plants and natural gas combined-cycle (NGCC) and simple-cycle (SCGT) plants with more than 20MW installed capacity are shown.
+Source: Alberta Environment compliance cost data combined with AESO generation data provided by NRGStream.}
 \label{fig:sector_compliance_cost}

### 2022-01-27 nrgstream_graphs.R

@@ -220,6 +220,8 @@ my_palette<-c("black",grey.colors(10,start=0.2,end = .95))
 
 plot_a<-ggplot(df2,aes(Year,capture,fill=Plant_Type))+
   geom_col(aes(Year,capture,fill=Plant_Type),position = position_dodge(width = .9),width = .6,color="black",size=.5)+
+  #geom_col(aes(Year,p_mean),position = "identity",fill=NA,color="black")+
+  #geom_line(dataaes(Year,capture,fill=Plant_Type),position = position_dodge(width = .9),width = .6,color="black",size=.5)+
   #geom_text(aes(y=-10,label=Plant_Type),angle=90,size=2)+
   #  scale_color_viridis("Plant Type",discrete=TRUE)+
 #  scale_fill_viridis("Plant Type",discrete=TRUE)+

### 2022-01-27 paper_graphs.R

@@ -163,7 +163,7 @@ gen_plain <- df2 %>% mutate(Plant_Type=factor(Plant_Type,levels=AB_plant_order))
     #scale_shape_manual("",values=c(15,16,17,18,0,1,2,3))+
     blake_theme()+theme(plot.margin =unit(c(1,1,1,1),"cm"))+
     scale_x_date(date_labels = "%b\n%Y",date_breaks = "24 months",expand=c(0,0))+
-    expand_limits(x = as.Date(c("2004-01-01", "2023-11-30")))+
+    expand_limits(x = as.Date(c("2004-01-01", "2022-1-30")))+
     expand_limits(y =-500)+
     scale_y_continuous(expand = c(0,0),breaks=pretty_breaks())+
     

### 2022-01-27 power_paper_base.R

@@ -30,17 +30,19 @@ graph_black<-"black"
 
 paper_theme<-function(caption_align=1){
   theme_minimal()+theme(
-    axis.title.x = element_text(size = 12,margin = margin(t = 2, b =2)),
-    axis.text.x = element_text(size = 12,margin = margin(t = 2, b = 2)),
-    axis.title.y = element_text(size = 12,margin = margin(r = 2, l = 2)),
-    axis.text.y = element_text(size = 12,margin = margin(r = 2, l = 2)),
+    text = element_text(size = rel(5),margin = margin(t = 2, b =2,l=2,r=2)),
+    axis.title = element_text(size = rel(.8)),
+    #axis.text.x = element_text(size = 12,margin = margin(t = 2, b = 2)),
+    #axis.title.y = element_text(size = 12,margin = margin(r = 2, l = 2)),
+    #axis.text.y = element_text(size = 12,margin = margin(r = 2, l = 2)),
     legend.position = "bottom",
-    legend.text=element_text(size=12),
+    legend.text=element_text(size=rel(3)),
     legend.margin=margin(c(0,0,-0.25,0),unit="cm"),
+    legend.key.width = unit(rel(2),"line"),
     plot.margin=margin(c(.5,.5,0.5,0.5),unit="cm"),
-    plot.caption = element_text(size = 9, color = "gray40",hjust=1),
-    plot.title = element_text(face = "bold"),
-    plot.subtitle = element_text(size = 10, color = "gray40"),
+    #plot.caption = element_text(size = 9, color = "gray40",hjust=1),
+    #plot.title = element_text(face = "bold"),
+    #plot.subtitle = element_text(size = 10, color = "gray40"),
     panel.grid.minor = element_blank(),
     panel.grid.major.x = element_blank()
   )

### 2022-01-27 regression_analysis.R

@@ -87,57 +87,62 @@ ctax_all<-all_plants_reg %>%
 
 #library(hrbrthemes)
 #hrbrthemes::import_roboto_condensed()
-ctax_all%>% filter(percentile<=90 & percentile>=40)%>%
-  filter(grepl("ctax",term)|grepl("oba",term))%>%ggplot(aes(x=percentile, y=estimate, ymin=conf.low, ymax=conf.high,group=term,color=term)) +
+ctax_all%>% filter(percentile<=100 & percentile>=0)%>%
+  filter(grepl("ctax",term)|grepl("oba",term))%>%
+  mutate(term=gsub("peak_facFALSE:","",term))%>%
+  mutate(term=gsub("peak_facTRUE:","",term))%>%
+  mutate(peak2=fct_relevel(peak,"Peak Hours"))%>%
+  ggplot(aes(x=percentile, y=estimate, ymin=conf.low, ymax=conf.high,group=term,color=term)) +
   #geom_pointrange() +
   #geom_line(size=1.25)+
-  geom_point()+
+  geom_point(size=1.25)+
   #geom_errorbar(width=2.85)+
-  geom_errorbar(width=rel(.5))+
-  scale_x_continuous(expand=c(0,0),breaks=pretty_breaks())+
-  #expand_limits(x=0)+
-  expand_limits(y=c(-.55,.55))+
+  geom_errorbar(width=rel(.75),size=.85)+
   scale_y_continuous(expand=c(0,0),breaks=pretty_breaks())+
-  facet_grid(cols=vars(Plant_Type),rows = vars(peak))+
-  geom_hline(yintercept = 0, col = "orange") +
+  expand_limits(y=c(-.5,.5))+
+  scale_x_continuous(expand=c(0,0),breaks=pretty_breaks())+
+  expand_limits(x=c(0,100))+
+  facet_grid(rows = vars(peak2))+
+  #facet_grid(rows=fct_relevel(peak,"Peak Hours"))+
+  geom_hline(yintercept = 0, col = "black") +
   labs(
     x = "Percentile of total offered power (%)", y = "Marginal effect (Δ in offer : Δ in $/MWh in cost or value)",
-    title = "Marginal effect of carbon tax cost and output-based allocation values on power offers by plant type" 
+    #title = "Marginal effect of carbon tax cost and output-based allocation values on power offers by plant type" 
     #subtitle = "Conditional on plant type"
   ) +
   paper_theme()+
   #theme_ipsum() + theme(legend.position = "bottom")+
[…]

### 2022-02-17 paper/paper_draft.tex
fixing pass-through

@@
 
+Finally, Alberta's market is highly concentrated, with a market structure that has evolved in ways relevant to our analysis during the relevant time period.
+When the market was restructured in 2000, a system of power purchase arrangements (PPAs) were used as an alternative to divestment.
+Large coal-fired plants owned by legacy generators were covered by these PPAs which approximated the conditions for plant owners which had been present under regulation, while auctioning off the rights to offer control over these facilities to other entities.
+PPAs not sold in the auction were held by an entity known as the Balancing Pool.
+The PPAs included a cumbersomely-worded change-of-law clause which was interpreted to imply a option for PPA owners to exit these arrangements should they become unprofitable or ``more unprofitable" due to a change in law, and several PPA owners exercised this right during our period of study, specifically in response to the changes in carbon pricing policy announced in June of 2015.
+As detailed in Leach and Tombe (2016)
+
+As shown in Table \ref{tab:PPA}, a substantial amount of Alberta's generating capacity (5066 MW) was offered into the market under these PPAs during our study period, and the majority (3866 MW) of that was transferred back to the Balancing Pool in early 2016.
+The reason for this change was the erosion in the profitability of the PPAs contributed to by changes in carbon pricing.
+But, the coincidence of changes in offer control with changes in carbon pricing in early 2016 have the potential to confound our inference.
+
+\begin{table}
+\begin{tabular}{lccccc}
+  \hline\\[-1.75ex]
+Facility&Plant&Unit& Capacity& End of& Returned to\\
+&Owner&         ID   &  (MW) &  PPA Term &Balancing\\
+&&                 &    &      & Pool\\[+0.5ex]
+\hline \\[-1.5ex]
+Battle River& ATCO& BR3 &147& Dec-31-2013&NA\\
+&&BR4 &147& Dec-31-2013&NA\\
+&&BR5 &368& Dec-31-2020&Jan-1-2016\\[+0.75ex]
+Genesee &Capital Power &GN1 &381 &Dec-31-2020&NA\\
+&&GN2 &381 &Dec-31-2020&NA\\[+0.75ex]
+HR Milner &ATCO& HRM& 144& Jan-01-2013&NA\\[+0.75ex]
+Keephills& TransAlta & KH1 &383& Dec-31-2020&May-5-2016\\
+&&KH2& 383& Dec-31-2020&May-5-2016\\[+0.75ex]
+Sheerness &ATCO (50\%) & SH1& 378& Dec-31-2020&Mar-7-2016\\
+&TransAlta (50\%)&SH2& 378& Dec-31-20200&Mar-7-2016\\[+0.75ex]
+Sundance A &TransAlta&  SD1 &280& Dec-31-2020&Mar-7-2016\\
+&&SD2 &280& Dec-31-20200&Mar-7-2016\\[+0.75ex]
+Sundance B &TransAlta & SD3& 353& Dec-31-2020&Mar-7-2016\\
+&&SD4 &353 &Dec-31-20200&Mar-7-2016\\[+0.75ex]
+Sundance C &TransAlta& SD5& 353& Dec-31-2020&Mar-24-2016\\
+&&SD6& 357& Dec-31-20200&Mar-7-2016\\[+0.75ex]
+  \hline
+\end{tabular}
+\caption{Alberta coal-fired plant Power Purchase Arrangement (PPA) expiry dates and dates of return to the Balancing Pool}\label{tab:PPA}
+\end{table}
+
+\begin{figure}[t]
+    \centering
+    \begin{minipage}{\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{../images/mkt_power.png}
+        \caption{Share of offer control and share of non-zero offers from key firms including and not including the Balancing Pool government-owned corporation.
+Key firms include TransAlta, Capital Power, ENMAX, ATCO, Heartland Generation, and TransCanada, with the addition of the Balancing Pool in the left-hand panel.}
+    \label{fig:synth_illust}
+        %\caption{$dt=0.1$}
+        %\label{fig:prob1_6_2}
+    \end{minipage}%
+\end{figure}
+
 \section{The Impact of Carbon Pricing Policy Changes}
@@
 \section{Estimation Strategy}
-There are two factors which confound our ability to draw inference from the merit order data.
+The focus of our estimation strategy is to identify the magnitude of the vertical shift in the electricity supply curves for the market as a whole or aggregated by generator fuel or controlling entity in response to changes in carbon pricing policies.
+
+There are multiple factors which confound our ability to draw inference from the merit order data.
 The first is that plants of the same type (e.g.
@@
 In particular for coal plants, plant-level minimum-must-run constraints imply that facilities will always offer some portion of their generation into the market at a \$0/MWh price, effectively acting as price takers for an endogenously-determined share of their generation.
+Given that shutting down in any given hour would imply multiple hours out of the market, the true marginal opportunity cost of this power is negative.
 They may offer marginal blocks of power, for capacity above their minimum-must-run level, at higher prices.
@@
 Finally, since multiple plants may be controlled by a single entity, the observed data are the product of a complex portfolio optimization problem and the offers of any individual plant may not reveal an overall portfolio-level attempt to pass through the costs of carbon prices to wholesale prices.
-We avoid these problems by constructing synthetic offer curves either by plant-type and by controlling entity, thus estimating the impact of policy changes on the supply behaviour of the aggregate fleet rather than on any individual plant.
+We avoid these problems by constructing synthetic offer curves either by plant-type and by controlling entity, or for the market as a whole, thus estimating the impact of policy changes on the supply behaviour of the aggregate fleet rather than on any individual plant.
 We then use the synthetic merit order data we construct to examine how power has been offered into the system over time as climate policies have varied.
 
-Our empirical strategy is best understood visually through a series of figures.
-Consider Figure \ref{fig:synth_merit}.
-In the left-hand panel, we show the merit order for a particular hour for all of the coal facilities in the province, with offers from multi-unit facilities combined for ease of visualization.
+To understand our approach, consider the traditional textbook treatment of carbon pricing or output-based allocations, which would imply a vertical shift in the supply curve corresponding to the increased marginal cost or the carbon tax, with a negative shift in the case of output-based allocations which reduce net costs of generation.
+The horizontal summation of all of the individual firm supply curves should imply that, all else equal, the vertical shift in the market supply curve at any point is equal to the marginal cost impact of the carbon policy on the firm which would supply that particular marginal unit of power.
+We have data on these horizontally-summed supply curves and, by sampling specific points on the supply curve over time, we identify the magnitude of pass-through.
+
+\begin{figure}[t]
+    \centering
+    \begin{minipage}{\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{../images/sampling.png}
+        \caption{Illustration of two firms' (thicker and thinner lines) offers of power in no-, high-, and low-carbon-tax scenarios, with lighter grey lines indicating a higher carbon price.}
+    \label{fig:synth_illust}
+        %\caption{$dt=0.1$}
+        %\label{fig:prob1_6_2}
+    \end{minipage}%
+\end{figure}
+
+To measure this effect, we look at hourly merit orders and sample at multiple percentile points, using more sampling points at the upper end of the merit order.
+The variation in carbon pricing over time, as well as variation in which firms are participating in the market at different times allows us to develop a sample of offer values at each of our chosen percentile nodes over our entire sample.
+For each sampled point, we identify facility and market conditions, as well as carbon policy compliance costs, which we use to explain changes in the level offer levels over time.
+Consider the illustrative example in Figure \ref{fig:synth_illust} which shows shifts in a hypothetical two-firm supply curve in response to increasing carbon prices.
+At the 55th percentile, the first sampling point shown, the vertical shift is driven entirely by pass-through from Firm 2 of their marginal carbon costs over-and-above the base case.
+At the 80th percentile, the second sampling point shown, the increase under the low-tax scenario is caused by the pass-through from Firm 2, but in the higher carbon tax scenario, the vertical shift is a smaller than would be the case for a pure pass-through by Firm 2 since there is also a shift in the ordering of offered blocks such that a different firm is providing the 80th percentile supply into the market with the high carbon tax.
+
+To show what this implies for actual data, consider Figure \ref{fig:synth_merit}.
+In the left-hand panel, we show the merit order for one hour for all of the coal-powered facilities in the province, with offers from multi-unit facilities combined for ease of visualization.
 We see facilities bidding some of their power at \$0, ensuring they are in the market, with marginal generation blocks offered at increasing marginal prices.
 In this particular hour, the price was such that most of the offered coal generation was dispatched, with only blocks offered above the \$802/MWh market clearing price not being dispatched.
-In the right hand panel, we show our synthetic merit order which summarizes all of the coal power offered into the market in a given hour using the 15 points shown in the Figure, which correspond to the value of the merit order at prescribed percentiles of the total offered power in that hour.\footnote{We store the 10th through 40th percentiles in 10\% intervals and the 50th percentile and above in 5\% intervals to better capture the curvature of the merit order.} Summarizing the data this way reduces our stored information by a factor of three, but more importantly it allows us to answer the relevant question in our study - by how much did the supply curve shift in response to changes in the value of carbon pricing costs and output-based allocation revenue - by looking at changes in the value of our sampling points across the portfolio of offered power.
+In the right hand panel, we show our sampling of the merit order which summarizes all of the coal power offered into the market in a given hour using the 15 points shown in the Figure, which correspond to the value of the merit order at prescribed percentiles of the total offered power in that hour.\footnote{We store the 10th through 40th percentiles in 10\% intervals and the 50th percentile and above in 5\% intervals to better capture the curvature of the merit order.} Summarizing the data this way reduces our stored information by a factor of three, but more importantly it allows us to answer the relevant question in our study - by how much did the supply curve shift in response to changes in the value of carbon pricing costs and output-based allocation revenue - by looking at changes in the value of our sampling points across the portfolio of offered power.
 Or, visually, we want to ask how, on average, each of the points in the right-hand panel of Figure \ref{fig:synth_merit} change when carbon prices are increased or when output-based allocations of emissions credits change.
+For each of these sampled points, we include relevant facility- and market-level data as well as carbon tax compliance costs for that facility on that date.
 
-\begin{figure}[t]
+\begin{figure}[tb]
     \centering
@@
 There is a lot of day-to-day and hour-to-hour variability in how coal units are offered into the market.
-Figure \ref{fig:synth_coal_2019}, which shows all of the 7pm merit orders (left panel) and their synthetic equivalents, along with the average synthetic merit order for the last year of our sample (right panel), provides a sense of the variability in our sample.
+Figure \ref{fig:synth_coal_2019}, which shows all of the 7pm merit orders (left panel) and their synthetic equivalents based only on the sample points, along with the average synthetic merit order for the last year of our sample (right panel), provides a sense of the variability in our sample.
 Similar variability exists in the offers of generation from other dispatchable sources.\footnote{See online graphical appendix, Figure A1, for a similar Figure for all dispatchable generating source types.} Figure \ref{fig:synth_coal_2019} shows the degree to which changes in total offered capacity would skew our results if we attempted to base our analysis on offered megawatts.
@@
 
-The next stage in our analysis is to transfer our information on facility compliance costs to the synthetic merit orders.
-Recall that, for the major emitters, we have facility-year-specific information on their emissions-intensity and their output-based allocations of emissions credits which determines their net emissions pricing liabilities.
-Since we are interested in the separate effects of both components of emissions policies, and since both vary at the facility level through our sample, we can separately identify the impacts of each on price pass-through behaviour.
-As such, for each of the synthetic merit orders we create for analysis purposes, we also capture and store both the marginal emissions charge and the marginal output-based allocation at each percentile support point.
-
-For our analysis sample, we create three other types of synthetic merit orders.
+For our analysis sample, we create four types of synthetic merit orders.
 First, for very general results, we characterize the entire merit order using the methodology outlined above.
+Next, we decompose the results for dispatchable generation types (coal and natural gas plants).
 We also use the fact that, from 2013 through 2019, the AESO provides offer control by unit as part of their merit order releases.
 We use this information to construct synthetic supply curves for each of the major participants in the Alberta power market.
+We also look at the behaviour of some of the largest facilities in the province.
 This means that we have facility-level, firm- or portfolio-level and market-level synthetic merit orders to analyze.
 
-To these, synthetic merit orders, we add our hourly data on market status, weather, and fuel prices specified above.
-This then allows us to build regression equations for each percentile support point of each synthetic merit order, asking the degree to which carbon price and output-based allocation changes impact the level of each step of the supply curves by facility, plant-type, by owner, or for the complete market.
+For parts of our sample, censoring is also an issue.
+In the Alberta market, negative price offers are not allowed, and offers are capped at \$1000/MWh.
+This implies that we may under-estimate pass-through where either power offers have a less negative implied value below zero or where the carbon tax leads to a higher true costs of offer above \$1000, as we would not observe either of these effects.
+To correct for this, we augment our ordinary least squares estimates of pass-through with Tobit estimates with two-sided censoring, estimated via maximum likelihood.
+We do not find the censoring to have a marked effect on our estimates.
 
@@
     \centering
-     \includegraphics[width=.9\textwidth]{../images/by_type.png}
+     \includegraphics[width=.9\textwidth]{../images/by_type_net.png}
     \caption{Estimated marginal effect of net carbon charges on power offers, all plant types. Error bars denote 95\% confidence intervals}
@@
 Declining ability to identify TransCanada impacts because of changes in resources.
+
+\begin{figure}[!htb]
+    \centering
+     \includegraphics[width=.9\textwidth]{../images/offer.png}
+    \caption{Estimated marginal effect of net carbon charges on power offers, all plant types, by controlling entity. Error bars denote 95\% confidence intervals}
+    \label{fig:offer}
+\end{figure}
 

### 2022-02-17 aeso_scrapes.R

@@ -326,8 +326,8 @@ get_forecast_report <- function(start_date,end_date) {
 
 
 #sample code to use this scraper
-day<-as.Date(Sys.Date()-years(1))
-xdf<-get_forecast_report(as.Date(day), as.Date(day)+months(1))
+#day<-as.Date(Sys.Date()-years(1))
+#xdf<-get_forecast_report(as.Date(day), as.Date(day)+months(1))
 
 all_forecasts<-function() {
   years<-seq(2000,year(Sys.Date()))
@@ -382,12 +382,12 @@ update_forecasts<-function() {
 
 #code to use this scraper
 
-forecast_data<-forecast_data%>%rename(#forecast_ail=day_ahead_forecasted_ail,
-  forecast_ail_actual_ail_difference= forecast_actual_ail_difference)
+#forecast_data<-forecast_data%>%rename(#forecast_ail=day_ahead_forecasted_ail,
+#  forecast_ail_actual_ail_difference= forecast_actual_ail_difference)
 #filename<-paste("data/forecast_data",".RData",sep = "")
 #save(forecast_data, file= filename) 
 #update_forecasts()
-load("data/forecast_data.Rdata")
+#load("data/forecast_data.Rdata")
 
 
 

### 2022-02-17 firm_theory_graphs_new.R

@@ -1,0 +1,586 @@
+library(fOptions)
+#using plot3d and plot3Drgl packages, color palette previously created using viridis
+library(plot3D)
+library(plot3Drgl)
+library(tidyverse)
+library(ggthemes)
+library(ggrepel)
+library(patchwork)
+library(scales)
+library(directlabels)
+
+class_theme<-function(){
+  theme_classic()+
+    theme(
+      plot.subtitle = element_text(color="grey10",size=rel(.7)),
+      plot.title = element_text(face="bold",size=rel(.8)),
+      plot.caption = element_text(color="grey50",size=rel(.5),hjust=0),
+      legend.title = element_text(color="grey10",size=rel(.5)),
+      legend.text = element_text(color="grey10",size=rel(.5)),
+      axis.text = element_text(size=rel(1.5)),
+      axis.title = element_text(size=rel(1.5)),
+      panel.grid.major = element_line(size = .2,colour = "grey60"), 
+      #plot.margin = margin(t = .1, r = .1, b = .1, l = .1,unit= "cm"),
+      plot.margin = margin(t = .5, r = 1, b = .25, l = 1,unit= "cm"),
+      #axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
+      NULL)
+}
+
+
+
+class_clean<-function(){
+  theme_classic()+
+    theme(
+      plot.subtitle = element_text(color="grey10",size=rel(.7)),
+      plot.title = element_text(face="bold",size=rel(.8)),
+      plot.caption = element_text(color="grey50",size=rel(.5),hjust=0),
+      legend.title = element_text(color="grey10",size=rel(.5)),
+      legend.text = element_text(color="grey10",size=rel(.5)),
+      axis.text = element_text(size=rel(1.5)),
+      axis.title = element_text(size=rel(1.5)),
+      #panel.grid.major = element_line(size=0,colour = "black"), 
+      #plot.margin = margin(t = .1, r = .1, b = .1, l = .1,unit= "cm"),
[…]

### 2022-02-17 merit_build.R

@@ -12,10 +12,10 @@ options(scipen=999)
 update<-0 #add new data
 save<-1 #save files at the end
 synth<-1 #synthetic plants?
-  synth_type<-3  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_type, 0 is full merit as synthetic plant
+  synth_type<-0  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_fuel, 0 is full merit as synthetic plant
 
   if(synth_type==5)
-    focus_id<-"BR3"
+    focus_id<-c("SH1","SH2")
 
 load("data/all_merit.RData")  
 if(update==1){
@@ -28,6 +28,9 @@ if(update==1){
 #small_testing_sample 
 #merit_small<-merit_data%>%filter(date==ymd("2019-10-05"))
 #merit_data<-merit_small
+
+
+
 
 
 #bring in market data
@@ -271,6 +274,38 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
   
   
   
+  #Repair offer control
+  
+  key_firms<-c("ATCO","TransAlta","TransCanada","ENMAX","Capital Power","Heartland","Balancing Pool")
+  
+  merit_aug<-merit_aug%>% #filter(date<ymd("2013-06-05"))%>%
+    mutate(offer_sum=case_when(
+      grepl("TransAlta",offer_control)~"TransAlta",
+      grepl("TransCanada",offer_control)~"TransCanada",
+      grepl("ENMAX",offer_control)~"ENMAX",
+      grepl("URICA",offer_control)~"URICA",
+      grepl("Shepard",offer_control)~"ENMAX",
+      grepl("Calgary Energy",offer_control)~"ENMAX", #calgary energy centre
+      grepl("Capital Power",offer_control)~'Capital Power',
+      grepl("ATCO",offer_control)~"ATCO",
+      grepl("Heartland",offer_control)~"Heartland",
+      grepl("Balancing Pool",offer_control)~"Balancing Pool",
+      grepl("Canadian Natural",offer_control)~"CNRL",
+      grepl("Genalta",offer_control)~"Genalta",
+      grepl("Imperial Oil",offer_control)~"Imperial Oil",
+      grepl("Suncor",offer_control)~"Suncor",
+      asset_id=="WB4"~"University of Alberta", #never appears so assigned
[…]

### 2022-02-17 regression_analysis.R

@@ -4,6 +4,114 @@ library(purrr)
 library(margins)
 source("power_paper_base.R")
 
+#make a market power indicator
+
+load("data/all_merit.RData")  
+if(update==1){
+  merit_data<-rbind(merit_data,update_merit(merit_data))
+  #remove the 02* hours
+  #  merit_data<-merit_data%>%filter(he!="02*")
+  save(merit_data, file="data/all_merit.RData")  
+}
+
+mkt_power<-merit_data %>%filter(year(date)==2018) %>% select(date,he,asset_id,size,available_mw,dispatched_mw,offer_control,dispatched,)%>%
+  group_by(date,he,offer_control)%>%
+  summarize(avail=sum(available_mw),dispatched=sum(dispatched_mw),id=list(asset_id))
+
+offer_test<-merit_data %>% 
+  #filter(year(date)>=2013)%>%
+  #mutate(year=year(date))%>%
+  #select(asset_id,block=block_number,offer_control,year)%>%
+  #filter(!is.na(offer_control),offer_control!="")%>% #take out NA and blank
+  mutate(offer_sum=case_when(
+    grepl("TransAlta",offer_control)~"TransAlta",
+    grepl("TransCanada",offer_control)~"TransCanada",
+    grepl("ENMAX",offer_control)~"ENMAX",
+    grepl("URICA",offer_control)~"URICA",
+    grepl("Shepard",offer_control)~"ENMAX",
+    grepl("Calgary Energy",offer_control)~"ENMAX", #calgary energy centre
+    grepl("Capital Power",offer_control)~'Capital Power',
+    grepl("ATCO",offer_control)~"ATCO",
+    grepl("Heartland",offer_control)~"Heartland",
+    grepl("Balancing Pool",offer_control)~"Balancing Pool",
+    grepl("Canadian Natural",offer_control)~"CNRL",
+    grepl("Genalta",offer_control)~"Genalta",
+    grepl("Imperial Oil",offer_control)~"Imperial Oil",
+    grepl("Suncor",offer_control)~"Suncor",
+    asset_id=="WB4"~"University of Alberta", #never appears so assigned
+    asset_id=="TMR"~"Transmission Must Run", #never appears so assigned
+    asset_id=="RB2"~"ATCO", #never appears so assigned
+    TRUE~offer_control #if it's not one of these, leave it the same
+  ))%>%
+  group_by(asset_id)%>%
[…]

### 2022-02-17 sampling_illust.R

@@ -1,0 +1,168 @@
+
+
+t<-20
+supply_1<-function(x,tax=0){x^2+tax}
+
+#y=x^2+t -> sqrt(y-t)=x
+
+supply_2<-function(x,tax=0){.5*x^2+tax}
+#y=.5x^2+t -> sqrt((y-t)/.5)=x
+
+
+
+supply_mkt<-function(x,tax=0){x^2/(1+sqrt(2))^2+tax}
+#x_mkt= sqrt(y-t)+sqrt((y-t)/.5)
+#x_mkt=(1+sqrt(2))*sqrt(y-t)
+#x/(1+sqrt(2))
+
+
+
+#set up mock two-firm merit with and without ctax
+#ifelse(row_number()==1, 0, time2)
+
+d_1=tibble(size=c(200,100,100,0),
+           price=c(0,200,500,500),
+           ghgs=.3,
+           firm="Firm 1"
+           )
+
+
+d_2=tibble(size=c(100,60,60,0),
+           price=c(0,50,400,400),
+           ghgs=1,
+           firm="Firm 2")
+
+
+d<-bind_rows(d_1,d_2)%>% mutate(scenario="Base",ctax=0)%>%
+  bind_rows(bind_rows(d_1,d_2) %>% mutate(scenario="Low tax",ctax=50))%>%
+  bind_rows(bind_rows(d_1,d_2) %>% mutate(scenario="High tax",ctax=180))%>%
+  group_by(firm,scenario)%>%
+  mutate(price=ifelse(row_number()==1, 0,price+ghgs*ctax))%>%
+  mutate(
+  start=ifelse(row_number()==1, 0,lag(cumsum(size))),                    
+  end=cumsum(size))%>%ungroup()%>%
+  group_by(scenario)%>%
+  arrange(price)%>%
+  mutate(merit_start=ifelse(row_number()==1, 0,lag(cumsum(size))),                    
+  merit_end=cumsum(size))%>%
+  arrange(scenario,merit_start)%>%
+  mutate(scenario=as_factor(scenario),
+         scenario=fct_relevel(scenario,"Base","Low tax","High tax"),
+         )
+  
+  
+ggplot(d) +
+  geom_step(aes(y=price,x=merit_start/max(merit_start)*100,group=scenario,color=scenario,size=firm))+
+  geom_point(data=tibble(x=c(55,55,55),y=c(50,100,230)),aes(x=x,y=y),size=5)+
[…]

### 2022-02-18 merit_build.R
fixing some bits

@@ -12,7 +12,7 @@ options(scipen=999)
 update<-0 #add new data
 save<-1 #save files at the end
 synth<-1 #synthetic plants?
-  synth_type<-0  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_fuel, 0 is full merit as synthetic plant
+  synth_type<-1  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_fuel, 0 is full merit as synthetic plant
 
   if(synth_type==5)
     focus_id<-c("SH1","SH2")
@@ -118,6 +118,7 @@ load("data/forecast_data.RData")
   renew_vols<-renew_vols %>% mutate(offer_sum=case_when(
    grepl("TransAlta",offer_control)~"TransAlta",
    grepl("TransCanada",offer_control)~"TransCanada",
+   grepl("Heartland",offer_control)~"Heartland",
    grepl("ENMAX",offer_control)~"ENMAX",
    grepl("Capital Power",offer_control)~'Capital Power',
    grepl("ATCO",offer_control)~"ATCO",
@@ -354,10 +355,12 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
     
      #feb 16, 2022 - adding carry-through of plant offer control measure
       merit_aug <- merit_aug%>% 
-        mutate(offer_store=factor(offer_sum))
+        mutate(offer_store=factor(offer_sum),
+               plant_store=factor(Plant_Type))
         
       
       offer_levels<-levels(merit_aug$offer_store)
+      plant_levels<-levels(merit_aug$plant_store)
                #use offer-store as the carry-through for offer control at a given point
       if(synth_type==5){ #if we're doing a specific unit or PPA
         merit_aug <- merit_aug%>% filter(asset_id %in% focus_id)%>%
@@ -400,7 +403,7 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
                )%>%
         summarize(
           #place offer percentiles and prices in lists of vectors
-          offers=list(offer_store),
+          offers=list(offer_store),plants=list(plant_store),
           total_offers=sum(size),available_mw=sum(available_mw),dispatched_mw=sum(dispatched_mw),renew_gen=sum(renew_gen,na.rm = T),
[…]

### 2022-02-18 regression_analysis.R

@@ -6,19 +6,19 @@ source("power_paper_base.R")
 
 #make a market power indicator
 
-load("data/all_merit.RData")  
+load("data/all_merit.RData")
 if(update==1){
   merit_data<-rbind(merit_data,update_merit(merit_data))
   #remove the 02* hours
   #  merit_data<-merit_data%>%filter(he!="02*")
-  save(merit_data, file="data/all_merit.RData")  
+  save(merit_data, file="data/all_merit.RData")
 }
 
 mkt_power<-merit_data %>%filter(year(date)==2018) %>% select(date,he,asset_id,size,available_mw,dispatched_mw,offer_control,dispatched,)%>%
   group_by(date,he,offer_control)%>%
   summarize(avail=sum(available_mw),dispatched=sum(dispatched_mw),id=list(asset_id))
 
-offer_test<-merit_data %>% 
+offer_test<-merit_data %>%
   #filter(year(date)>=2013)%>%
   #mutate(year=year(date))%>%
   #select(asset_id,block=block_number,offer_control,year)%>%
@@ -64,19 +64,19 @@ offer_test<-merit_data %>%
             cr_nz=sum(gz*(key_firm==TRUE))/sum(gz),
             cr_nz_nbp=sum(gz*(key_firm_no_bp==TRUE))/sum(gz))%>%
   mutate(time=ymd_h(paste(date,he)))
-  
+
 
 offer_graph<-offer_test %>% pivot_longer(-c(date,he,time,total),names_to = "measure")%>%
-  mutate(no_bp=grepl("nbp",measure)) %>% 
+  mutate(no_bp=grepl("nbp",measure)) %>%
   mutate(no_bp=as_factor(no_bp),
-                                  no_bp=fct_recode(no_bp,"Balancing Pool Omitted"="TRUE",
-                                                   "Balancing Pool Included"="FALSE"))%>%
+         no_bp=fct_recode(no_bp,"Balancing Pool Omitted"="TRUE",
+                          "Balancing Pool Included"="FALSE"))%>%
   mutate(non_zero=grepl("nz",measure),
          non_zero=as_factor(non_zero),
-                non_zero=fct_recode(non_zero,"Share of Non-Zero Offers"="TRUE",
-                                 "Share of Offers"="FALSE"),
+         non_zero=fct_recode(non_zero,"Share of Non-Zero Offers"="TRUE",
+                             "Share of Offers"="FALSE"),
          non_zero=fct_relevel(non_zero,"Share of Offers")
-         )
+  )
 
 
 
[…]

### 2022-04-28 paper/paper_draft.tex
fixing some bits
update

@@
 %%%% TITLE %%%%
-\title{Marginal and Average Carbon Price Pass-Through\thanks{We are grateful to Sam Harrison for excellent
+\title{Carbon Price Pass-Through in Alberta's Power Market\thanks{We are grateful to Sam Harrison for excellent
 research assistance.
@@
 \begin{abstract}
-\noindent We evaluate the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.
-Between 2015 and 2019, legislative changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes, creating variation in average and marginal carbon prices both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
-We exploit this variance in treatment to provide unique evidence of offer response to carbon pricing in power markets.
-We show that even after adjusting for costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.\\ \\
+\noindent\\ \\
 Keywords: climate change, carbon pricing, electricity, coal\\ \\

### 2022-04-28 merit_build.R

@@ -12,7 +12,7 @@ options(scipen=999)
 update<-0 #add new data
 save<-1 #save files at the end
 synth<-1 #synthetic plants?
-  synth_type<-1  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_fuel, 0 is full merit as synthetic plant
+  synth_type<-2  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_fuel, 0 is full merit as synthetic plant
 
   if(synth_type==5)
     focus_id<-c("SH1","SH2")
@@ -513,10 +513,17 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
         
         
         
-        ungroup() %>% select(-merit,-price,-co2_est,-merit_func,-ghg_func,-ctax_func,-oba_func,-offer_func,-offers,
-                             -plant_func,-plants,
-                             -oba_val,-ctax_cost)
-      print(paste("Build bids, cleaned data frame, elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
+        ungroup() 
+      
+      print(paste("Built bids, elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
+      
+        
+      merit_aug<-merit_aug%>%
+        select(-merit,-price,-co2_est,-merit_func,-ghg_func,
+               -ctax_func,-oba_func,-offer_func,-offers,
+               -plant_func,-plants,-oba_val,-ctax_cost)
+      
+      print(paste("Cleaned data frame, elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
       
       #turn these into the appropriate format for later analysis
       
@@ -557,6 +564,7 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
       
       
       print(paste("Market Data Merged. Elapsed time is",time_length(interval(start_time, Sys.time()), "seconds"),"seconds"))
+
       
       
       if(save==1)

### 2022-06-11 paper/paper_draft.tex
cea deck

@@
 
-\specialcomment{ALnote}{\begingroup\sffamily\colorbox{blue}{Andrew: } \color{blue}}{\endgroup}
-\specialcomment{BBnote}{\begingroup\sffamily\colorbox{red}{Branko: }\color{red}}{\endgroup}
-
 \setlength{\abovecaptionskip}{10pt}
@@
 \addtolength{\parskip}{5pt}
+\setlength{\belowcaptionskip}{-10pt}
 
@@
 %%%% TITLE %%%%
-\title{Carbon Price Pass-Through in Alberta's Power Market\thanks{We are grateful to Sam Harrison for excellent
+\title{Carbon Price Pass-Through in Electricity Markets\thanks{We are grateful to Sam Harrison for excellent
 research assistance.
@@
 These changes have accelerated the phase-out of coal-fired power in the province.
+We also show, in contrast to much of the recent literature, that net marginal carbon price costs were not fully pass-through to offers.
+In fact, our results suggest that less than half of the changes in carbon policy costs was reflected in shifts in the merit order of electricity price offers in Alberta.
 
-The Alberta wholesale power market provides an excellent quasi-laboratory because it is small, isolated, and the market is settled on a real-time, energy-only basis.
-As such, there are limited confounding factors beyond what is observable in our data.
-Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.\footnote{The 2021 maximum Alberta internal load per AESO 2021 data was 11,729 MW at time of writing, which is also the peak value ever recorded in Alberta.} Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint operating capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share, and implies that extra-jurisdictional confounding factors are less likely to influence our results.
-The downside to this is that Alberta, a major oil-producing jurisdiction, was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies, and this economic downturn led to a period of low prices in the power market.
+The traditional view of competitive firm response to carbon pricing would be an upward shift in each firm's supply function of an amount equal to the net incremental costs imposed by the policy.
+For some facilities, there may be dynamic gains to participating in the market in a given hour (for example, if shutdown implies multi-hour restarts, as can be the case for coal generators), in which case the supply curve may lie below zero up to some positive quantity supplied in each hour.
+For these facilities, especially if they are inframarginal, we might not expect to see a carbon price change their bidding behaviour up to some threshold quantity.
+However, above such a threshold, we should expect to see carbon price pass-through into power offers.
+However, our results in the Alberta context suggest less than perfect pass-through, even for units likely to be on the margin in the market in a given hour.
+
+The degree to which carbon prices are passed-through to wholesale electricity markets has been extensively investigated.
+The studies of \cite{fabra_reguant} and \cite{hintermann16_germany} are most-closely related to our analysis, and both find that carbon costs imposed in the European Union emissions trading system are almost completely passed-through to market prices.
+In the case of \citeauthor{fabra_reguant}, they find more-than-complete (110\%) pass-through of carbon costs in peak hours, and less-than-complete (60\%) pass-through in off-peak hours.
+\citeauthor{hintermann16_germany} finds pass-through rates of 81-111\% depending on the hour of the day, and also finds that costs are fully passed-through on average.
+Our paper is unique in that we can measure facility-specific carbon policy costs (prices and the value of output-based allocations) precisely for long periods of time, which allows us to separately identify changes due to carbon pricing from changes due to other coincident market or climate factors.
+Furthermore, the changes in Alberta's climate change policies were such that the average and net marginal costs of carbon pricing policies changed across facilities for the same time periods.
+For example, in 2018, the marginal carbon price did not change from 2017, but there were substantial changes in output-based allocations of emissions credits such that marginal net carbon costs went up for some facilities and down for others.
+We use this variation to identify pass-through more precisely than is possible with daily market fluctuations in a cap-and-trade market.
+Our results estimate much lower pass-through of carbon prices to electricity offers and thus to market prices than either \cite{fabra_reguant} or \cite{hintermann16_germany}.
+
+Average carbon policy costs varied over time and between fuel types in Alberta through the various changes to emissions policies since 2015.
+This links our work to \cite{cullen_mansur} which uses variations in fuel costs (coal vs.
+natural gas) to identify impacts on emissions.
+They find substantial impacts from small changes in fuel costs which make natural-gas-fired plants more competitive relative to coal plants.
+The policy changes we study in Alberta, both the changes in output-based allocations and the carbon price increases, have had a similar effect on the coal-to-gas cost ratio as a reduction in natural gas prices.
+Their results also inform our results in the sense that the same gas price reductions seen in their study occurred in Alberta during our sample period, and we adjust for these impacts in our study design.
+We show that the degree to which carbon prices were passed through was most consistent with the theory for more flexible, combined-cycle natural gas facilities, while coal plants passed through a much lower share of increased emissions costs.
+
+The Alberta electricity market is highly concentrated, which links our paper to the literature following from the seminal paper of \cite{buchanan1969external} on externalities and related policies in the presence of market power, as well as to myriad studies of market power in electricity, e.g.
+\cite{borenstein_bushnell_wolak_2002}.
+For example, \cite{fowlie_reguant_ryan_2016} study the interaction of market structure in the cement industry and greenhouse gas mitigation policies.
+\cite{miller_cement}, also looking at the cement industry, shows that fuel cost changes were more-than-fully reflected in downstream pricing.
+Similarly, as in \cite{holland_emissions_2012}, incomplete regulation or regulations which provide competing incentives will distort price-pass through.
+The latter also applies in our work since there are changes in market structure coincident with the policy changes we study for which we must account.
+
+Another unique attribute of our study is that the policy variation over time and across facilities in Alberta also allows us to directly identify the pass-through of changes in output-based allocations of emissions credits (see \cite{fischerfox07} and \cite{fischerfox12}) to wholesale prices, as allocation rates did not vary systematically with changes in carbon prices.
+We cannot directly speak to the optimal allocations of emissions credits as in \cite{burtraw_palmer_2008}, but our results suggest that there is symmetric and incomplete pass-through of the marginal value of output-based allocations and carbon prices.
+To our knowledge, this has not been confirmed in other studies of carbon price pass-through.
+There are limits to this result as well, since the changes in output-based allocations systematically disadvantaged coal plants relative to other fuel sources, and we coincident with substantial declines in natural gas prices.
+This implies that Alberta's coal plants were already at a substantial competitive disadvantage, and thus unable to pass through higher net carbon policy costs associated with reduced output-based allocations of emissions credits.
+
+The Alberta wholesale power market provides an excellent quasi-laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis.
+As such, there are few confounding factors beyond what is observable in our data.
+Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.
+Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint operating capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share.
+The downside to this isolation is that Alberta, a major oil-producing jurisdiction, was hit with a major, oil-price-induced recession contemporaneously with some of the changes in carbon policies, and this economic downturn led to a period of low prices in the power market.
 Alberta's generation mix is also somewhat unique, with a significant portion of internal load served by combined heat and power (or co-generation) plants associated with industrial facilities, primarily located in the oil sands region in Northern Alberta.
@@
 
-We compile over 10 years of hourly data describing market offer behaviour and a wide range of relevant characteristics.
-We measure and describe facility-level hourly offers of power into the market for the time period preceding the COVID-19 pandemic.
-We combine these data with detailed plant characteristics compiled from regulatory data, emissions compliance data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data including hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.
+We compile over 13 years of hourly data describing market offer behaviour and a wide range of relevant characteristics.
+We measure and describe facility-level hourly offers of power into the market for the time period preceding the COVID-19 pandemic, although we also test this truncation of our data in later results.
+We combine power market data with detailed plant characteristics compiled from regulatory filings, emissions compliance data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data including hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.
 Finally, we include both observed carbon policy costs and the value of output-based allocations at the facility level.
 
-In order to isolate the impact of the changes in policies on offer behaviour, we develop a unique empirical strategy...
+In order to isolate the impact of the changes in policies on offer behaviour, we develop a unique empirical strategy based on merit order sampling.
+For each of three different aggregation levels (market, plant type, and controlling entity), we build hourly merit orders of offered generation.
+We sample these merit orders at various support points which allows us to build a synthetic merit order with substantially-reduced dimensionality, and also to make the date comparable across time period with different total market sizes.
+We also include measures of marginal carbon emissions policy costs along the merit order which we sample in parallel with the offered prices.
+As such, we derive a sample of merit order levels and the impacts of the marginal costs of emissions policies embedded in those offers.
+If firms act as they are predicted to by standard economic models, passing marginal costs through into their power offers, we will see a 1:1 relationship between these costs and upward shifts in our synthetic merit orders.
 
-Results preview...
+We find that firms pass through carbon policy costs as standard economic models would predict.
+Carbon prices changes are passed-through in the mid-merit at a maximum of 0.4:1 at the 60th percentile of offered power with smaller pass-through values at higher and lower points in the merit order.
+The impact of carbon pricing in the upper merit is smaller, and measured with less precision in part due to the censoring of the offers by a \$1000/MWh offer limit in the market.
+We also show that this effect is roughly replicated when carbon pricing net revenues are decomposed into carbon price and output-based allocation values, and across peak and off-peak hours.
+Our results by plant type are somewhat counter-intuitive.
+We see small but positive pass-through for natural gas combined cycle plants and for coal plants, but we estimate that simple cycle natural gas plants reduced their offer levels with higher carbon prices, all else equal.
+We expect that this is a combined effect of a strategic interaction with coal plants whereby the higher carbon prices increased costs more for coal plants than for gas plants, giving the gas plants the potential to shift coal higher up the merit order with shaded offers.
+Our results by controlling entity show two interesting results.
+First, we show that the Crown (government-owned) entity acted in the market more aggressively, reducing its offer prices even as carbon pricing increased in stringency for power offers most likely to be infra-marginal.
+
+The results show that traditional models of marginal cost pass-through are subject to substantial revision with respect to strategic incentives, portfolio optimization, and other impacts.
+Our results suggest that the market price impacts of Alberta's carbon pricing changes was substantially less than would be forecast by an assumed marginal net-cost pass-through, even though such an assumption would have been reasonable based on the evidence in the literature to date.
+We do that for firms with flexible fleets of generating units, the pass-through is close to what an assumption of marginal cost pass-through would predict.
+This suggests that technological, contractual, or market power implications play a substantial role in the degree to which carbon prices are reflected in offers of power.
 
@@
 In Sections \ref{sec:data} and \ref{sec:market} we detail our data set and provide a statistical portrait of the province's power market.
-In Section \ref{sec:empirical} we detail our empirical strategy.
+In Section \ref{sec:est} we detail our empirical strategy.
 Section \ref{sec:results} summarizes our results.
@@
 
-This system remained in place until June, 2015 when the government of Premier Rachel Notley introduced a series of changes to the existing regulation.\footnote{\textit{Specified Gas Emitters Amendment Regulation}, Alta Reg 105/2015, $<$\url{perma.cc/Z2ZM-WDWP}$>$.} The first set of changes increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for the output-based allocation of emissions credits to 85\% and 80\% of historic facility level emissions intensity for 2016 and 2017 respectively.
+This system remained in place until June, 2015 when the government of Premier Rachel Notley introduced a series of changes to the existing regulation.\footnote{\textit{Specified Gas Emitters Amendment Regulation}, Alta O.C.
+105/2015, $<$\url{perma.cc/Z2ZM-WDWP}$>$.} The first set of changes increased the carbon price to \$20/tonne for 2016 and to \$30/tonne for 2017, while also reducing the benchmarks for the output-based allocation of emissions credits to 85\% and 80\% of historic facility level emissions intensity for 2016 and 2017 respectively.
 Combined, these changes implied a material increase in the average cost of carbon in each of the years 2016 and 2017.
@@
 
-The \emph{CCIR} remained in place until it was replaced by the \emph{
-Technology Innovation and Emissions Reduction Regulation} (\emph{TIER}) which took effect on January 1, 2020, outside of our sample period.\footnote{\textit{Technology Innovation and Emissions Reduction Regulation}, Alta Reg 133/2019, $<$\url{canlii.ca/t/54qgg}$>$.}  However, it's important to note that while \emph{TIER} altered the design of output-based allocations in most sectors, it did not impose material changes in the electricity industry and so we should not expect any changes late in our sample period which abuts the implementation of \emph{TIER}.
+The \emph{CCIR} remained in place until it was replaced by the \emph{Technology Innovation and Emissions Reduction Regulation} (\emph{TIER}) which took effect on January 1, 2020, outside of our sample period.\footnote{\textit{Technology Innovation and Emissions Reduction Regulation}, Alta Reg 133/2019, $<$\url{canlii.ca/t/54qgg}$>$.}  However, it's important to note that while \emph{TIER} altered the design of output-based allocations in most sectors, it did not impose material changes in the electricity industry and so we should not expect any changes late in our sample period which abuts the implementation of \emph{TIER}.
 \emph{TIER} remains in place as of this writing, with the price on carbon having increased to \$40 per tonne for 2021 and \$50 per tonne for 2022.\footnote{See Alberta Environment and Parks Ministerial Order 87-2021 under the \emph{Emissions Management and Climate Resilience Act}, SA 2003, c.
@@
 We apply the same emissions intensity to multiple generating units within a facility.\footnote{There are two exceptions to this.
-The Keephills 3 and Genessee 3 coal-fired generating units report separately under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.} Since we have only annual data at the facility level, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units nor with the capacity factor at any point in time.
+The Keephills 3 and Genessee 3 coal-fired generating units report separately from other units at the same site under Alberta's carbon pricing rules, and so we have separate emissions-intensity data for them.} Since we have only annual data at the facility level, we do not allow emissions intensity to vary with the intensity of use of a particular unit within a facility, nor do we allow emissions intensities to vary with ramping of units nor with the capacity factor at any point in time.
 This will lead to under-reporting of emissions-intensity when units are ramping up or generating below their average efficiency, and over-reporting of emissions-intensity for units operating above their annual average operating efficiency.
@@
 
-Finally, combined heat and power units present a challenge since many units report their emissions for the entire production facility, not simply for the power plant.
+Finally, combined heat and power units present a challenge since emissions are reported for the entire production facility, not simply for the power plant, in many cases.
 There are some stand-alone units, but these also vary in terms of the emissions-intensity of net-to-grid electricity depending on the design of the specific unit and what share of the produced heat is used to generate electricity vs process steam.
@@
 
-We add weather data from Environment Canada weather stations in Edmonton, Fort McMurray, Cold Lake, and Calgary, covering the major demand centers in Alberta.
+We collect weather data from Environment Canada weather stations in Edmonton, Fort McMurray, Cold Lake, and Calgary, covering the major demand centers in Alberta.
 We take the average of available measurements in each area in order to maximize the number of hours we can cover with weather data.
 With this approach, we lose fewer hours from our data set than if we allow unavailable weather data from one station to remove an hour from our sample.
-There is also no reason to believe that weather data availability is correlated with our variables of interest, and so the lost observations should have no leverage on the results.
+There is also no reason to believe that weather data availability is correlated with our variables of interest, and so any lost observations should have no leverage on the results.
 We convert temperature data to heating and cooling degree days for each region, which we then use in our model.
@@
 Finally, we supplement our publicly-available data with several series from NRGStream, a commercial data aggregation service.
-Most importantly, this service provides daily natural gas prices for the Nova Inventory Transfer hub which are a better proxy for Alberta natural gas prices than publicly-available series for Henry Hub.
-NRGStream also scrapes the real-time data from the AESO, and thus provides slightly different data from the aggregate, publicly-available historic metered volumes provided by the grid operator.
-We use NRGStream real-time data to compile real-time intertie flows between Alberta and British Columbia, Saskatchewan, and Montana and to source real-time generation (as opposed to net-to-grid metered volumes) from all generating units in the province where needed.
-These data are generally the same as metered volumes with the exception of cases where an industrial production facility is co-located with the power generation facility, in which case only the net-to-grid volumes from the facility are measured in the metered volumes data while the NRGStream scraped data captures actual production by hour for the generating facility.
+Most importantly, this service provides daily natural gas prices for the Nova Inventory Transfer (NIT) system which are a better proxy for Alberta natural gas prices than publicly-available series for Henry Hub.
+NRGStream also scrapes the real-time data from the AESO, and thus provides a second source for the aggregate, publicly-available historic metered volumes provided by the grid operator.
+We use NRGStream data to compile real-time intertie flows between Alberta and British Columbia, Saskatchewan, and Montana and to source real-time generation (as opposed to net-to-grid metered volumes) from all generating units in the province where needed.\footnote{The generation data are generally the same as metered volumes with the exception of cases where an industrial production facility is co-located with the power generation facility, in which case only the net-to-grid volumes from the facility are measured in the metered volumes data while the NRGStream scraped data captures actual production by hour for the generating facility.}
 
-All data save the natural gas prices and NRGStream generation and trade data which we are not authorized to redistribute are available HERE and can be replicated using code for the paper which is available at GITHUB LINK.
+All data save the natural gas prices and NRGStream generation and trade data which we are not authorized to redistribute are available DATA LINK TO BE ADDED and the results in the paper can be replicated using code available at GITHUB LINK TO BE ADDED.
 For the publicly available code, we substitute Henry Hub gas prices as a placeholder, to which we apply the average discount observed between Henry Hub and NIT to adjust the values so that regression results may be more closely replicated.
+\begin{figure}[h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/loads_clean.png}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices.png}
+\vspace{-0.75cm}	\caption{Alberta internal load (top) and wholesale power prices, peak and off-peak hours (bottom).
+Peak hours are 8am to 11pm other than on statutory holidays or Sundays.
+Source: AESO data, authors' graph.}
+\label{fig:ab_prices_and_loads}
+\end{figure}
 
@@
 
-A statistical portrait of Alberta's power market provides context for our study.
 Alberta's wholesale market is a single price, energy-only market.
 There is no day-ahead energy market, but there is a separate ancillary services market which we do not address beyond a couple of tangential references in this paper.
-Alberta's record internal load, reached on February 9, 2021, is 11,729MW, so Alberta's is a relatively small power market compared to the similarly-structured Energy Reliability Council of Texas (ERCOT) market, for example, which had a 58,597 MW peak 2021 load.
-Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_load}, although growth has been substantially slower since 2014-2015 due to the effects of depressed oil prices on the Alberta economy discussed further below.
-There was also a major, short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, leading to the markedly-lower spring trough in that year's load profile as shown in Figure \ref{fig:ab_load}.
-The 2020 COVID-19 demand shock is also evident in Figure \ref{fig:ab_load}, but lies outside the sample period for our analysis.
+Alberta's power market is structured similarly to the Energy Reliability Council of Texas (ERCOT) market, but is much smaller.
+Alberta's record peak load is 11,939MW, reached in 2021, or about one sixth of ERCOT's 74,820 MW record load.
+Alberta has seen significant growth in average and peak loads, as shown in Figure \ref{fig:ab_prices_and_loads}, although growth has been substantially slower since 2014 due to the effects of depressed oil prices on the Alberta economy discussed further below.
+There was also a major, short-term market event with a wildfire in Fort McMurray in 2016 which took a lot of oil sands generation and load off-line, leading to the markedly-lower spring trough in that year's load profile.
+Similarly, the 2020-22 COVID-19 demand shock is evident in Figure \ref{fig:ab_prices_and_loads}, but lies outside the sample period for most of our analysis.
 
-\begin{figure}[t]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/loads_clean.png}
-\vspace{-0.75cm}	\caption{Alberta Internal Load. Source: Alberta Electric System Operator data.}
-\label{fig:ab_load}
+\begin{figure}[th]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel_policies.png}
+\vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: AESO data provided by NRGStream, authors' graph.}
+\label{fig:gen_mix}
 \end{figure}
@@
 Our sample period covers ten years with no major regulatory changes in the power market design.
-Alberta briefly considered an addition of a capacity market between 2016 and 2019, but this was never implemented.
+Alberta briefly considered an addition of a capacity market between 2016 and 2019, but this was never implemented \cite{alberta_electricity_system_operator_capacity-market-brochurepdf_2017,alberta_electricity_system_operator_overview_2017}.
 However, market conditions have varied over the sample period.
 There are three distinct periods during our sample - relatively high load growth and tight market conditions from 2009 through 2013, followed by sharply constrained growth and high reserve margins after 2014 through most of 2018, followed by a return to tighter market conditions after the spring of 2018.
-These conditions are reflected in peak and off-peak prices shown in Figure \ref{fig:ab_prices}.
+These conditions are reflected in peak and off-peak prices shown in Figure \ref{fig:ab_prices_and_loads}.
 Some of these market changes coincide with policy changes of interest in our paper, in particular the changes to the \emph{SGER} introduced in June of 2015 which took effect in January of 2016 and 2017 respectively and are coincidental with a period of over-supply.
@@
 
-\begin{figure}[t]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/peak_prices.png}
-\vspace{-0.75cm}	\caption{Wholesale power prices, peak and off-peak hours. Peak hours are 8am to 11pm other than on statutory holidays or Sundays. Source: AESO data, authors' graph.}
-\label{fig:ab_prices}
-\end{figure}
-
-\begin{figure}[h]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/gen_fuel.png}
-\vspace{-0.75cm}	\caption{Generation mix in Alberta.  Source: AESO data provided by NRGStream, authors' graph.}
-\label{fig:gen_mix}
-\end{figure}
-
-The generation mix in Alberta's power market is dominated by fossil fuels although, as shown in Figure \ref{fig:gen_mix}, the dominant fossil fuel has changed from coal to natural gas.
+Alberta's generation mix is dominated by fossil fuels although, as shown in Figure \ref{fig:gen_mix}, the dominant fossil fuel has changed from coal to natural gas.
 Within natural gas generation, the mix of plant types is also relevant to our study.
@@
 Combined heat and power plants tend to operate with high capacity factors but also tend to be price-takers in the market, with very little flexibility at the margin since the industrial processes with which they are associated rely on them for process heat.\footnote{As will be discussed below, net-to-grid power from COGEN facilities is primarily offered into the market at a \$0 offer and accepts market price, and so we see limited evidence of carbon price pass-through for most of the power offered by these facilities.} Combined-cycle plants, on the other hand, provided more flexible natural gas generation on the margin.
-Peaking capacity in the market during our sample period was largely met through ramping of coal and combined-cycle plants, as well as simple cycle (SCGT) natural gas turbines.
+Peaking capacity in the market during our sample period was largely met through ramping of coal and simple- and combined-cycle natural gas plants.
 Each of these would be exposed to carbon pricing, and so we expect to see pass through affect offers of power from these facilities into the market.
@@
 Those that dispatch only in high-price times will have higher revenues on average than plants without dispatch control, or plants which run most hours of the day.
-Since prices have varied substantially over time, as shown above in Figure \ref{fig:ab_prices}, captured average prices at the plant level have varied as well, as shown in Figure \ref{fig:capture_prices}.
-We can see evidence here of how plants of different types offer their power into the market: hydro and combined- and simple-cycle gas turbines follow prices, and dispatch more often in higher-priced hours, just as imports tend to occur when prices are higher.
+Since prices have varied substantially over time, captured average prices at the plant level have varied as well, as shown in Figure \ref{fig:capture_prices}.
 Prices tend to be negatively correlated with wind generation, and so captured prices for wind power facilities are lower than other forms of generation.
 Exports tend to be scheduled when prices are expected to be low, which is reflected in the graph.
-Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewable sources since its generation pattern coincides with daily and seasonal peak prices, although we would expect this to change as \emph{duck curve} effects take hold over time.
+Solar, a relatively new entrant in the Alberta market, has captured higher prices than other variable renewable sources since its generation pattern coincides with daily and seasonal peak prices, although we would expect this to change as \emph{duck curve} effects take hold over time \cite{rivers_stretching_2020}.
 
-\begin{figure}[t]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture.png}
+\begin{figure}[!h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/price_capture_all.png}
 \vspace{-0.75cm}	\caption{Annual average revenue by plant type. Source: AESO generation data provided by NRGStream, authors' graph.}
@@
 
-Finally, Alberta's market is highly concentrated, with a market structure that has evolved in ways relevant to our analysis during the relevant time period.
-When the market was restructured in 2000, a system of power purchase arrangements (PPAs) were used as an alternative to divestment.
-Large coal-fired plants owned by legacy generators were covered by these PPAs which approximated the conditions for plant owners which had been present under regulation, while auctioning off the rights to offer control over these facilities to other entities.
-PPAs not sold in the auction were held by an entity known as the Balancing Pool.
-The PPAs included a cumbersomely-worded change-of-law clause which was interpreted to imply a option for PPA owners to exit these arrangements should they become unprofitable or ``more unprofitable" due to a change in law, and several PPA owners exercised this right during our period of study, specifically in response to the changes in carbon pricing policy announced in June of 2015.
-As detailed in Leach and Tombe (2016)
-
-As shown in Table \ref{tab:PPA}, a substantial amount of Alberta's generating capacity (5066 MW) was offered into the market under these PPAs during our study period, and the majority (3866 MW) of that was transferred back to the Balancing Pool in early 2016.
-The reason for this change was the erosion in the profitability of the PPAs contributed to by changes in carbon pricing.
-But, the coincidence of changes in offer control with changes in carbon pricing in early 2016 have the potential to confound our inference.
-
-\begin{table}
+\renewcommand{\arraystretch}{.85}
+\begin{table}[!h]
+\vspace{.5cm}
+\caption{Alberta coal-fired plant Power Purchase Arrangement (PPA) expiry dates and dates of return to the Balancing Pool}\label{tab:PPA}
 \begin{tabular}{lccccc}
@@
 Sheerness &ATCO (50\%) & SH1& 378& Dec-31-2020&Mar-7-2016\\
-&TransAlta (50\%)&SH2& 378& Dec-31-20200&Mar-7-2016\\[+0.75ex]
+&TransAlta (50\%)&SH2& 378& Dec-31-2020&Mar-7-2016\\[+0.75ex]
 Sundance A &TransAlta&  SD1 &280& Dec-31-2020&Mar-7-2016\\
-&&SD2 &280& Dec-31-20200&Mar-7-2016\\[+0.75ex]
+&&SD2 &280& Dec-31-2020&Mar-7-2016\\[+0.75ex]
 Sundance B &TransAlta & SD3& 353& Dec-31-2020&Mar-7-2016\\
-&&SD4 &353 &Dec-31-20200&Mar-7-2016\\[+0.75ex]
+&&SD4 &353 &Dec-31-2020&Mar-7-2016\\[+0.75ex]
 Sundance C &TransAlta& SD5& 353& Dec-31-2020&Mar-24-2016\\
-&&SD6& 357& Dec-31-20200&Mar-7-2016\\[+0.75ex]
+&&SD6& 357& Dec-31-2020&Mar-7-2016\\[+0.75ex]
   \hline
 \end{tabular}
-\caption{Alberta coal-fired plant Power Purchase Arrangement (PPA) expiry dates and dates of return to the Balancing Pool}\label{tab:PPA}
 \end{table}
 
-\begin{figure}[t]
+\begin{figure}[!h]
     \centering
@@
 Key firms include TransAlta, Capital Power, ENMAX, ATCO, Heartland Generation, and TransCanada, with the addition of the Balancing Pool in the left-hand panel.}
-    \label{fig:synth_illust}
+    \label{fig:mkt_power}
         %\caption{$dt=0.1$}
@@
 \end{figure}
+
+Finally, Alberta's market is highly concentrated, with a market structure that has evolved in ways relevant to our analysis during the relevant time period.
+When the market was restructured in 2000, a system of power purchase arrangements (PPAs) were used as an alternative to divestment \cite{kleit_electricity_2007}.
+Large coal-fired plants owned by legacy generators were covered by these PPAs which approximated the conditions for plant owners which had been present under regulation, while auctioning off the rights to offer control over these facilities to other entities.
+PPAs not sold in the auction were held by an entity known as the Balancing Pool.
+A substantial amount of Alberta's generating capacity (5066 MW) was offered into the market under these PPAs during our study period.
+The PPAs included a cumbersomely-worded change-of-law clause which was interpreted to imply an option for PPA owners to exit these arrangements should they become unprofitable or ``more unprofitable" due to a change in law.
+In response to the changes in carbon pricing policy announced in June of 2015, all PPA owners exercised rights of return \cite{bankes_termination_2016}, and offer control over all of the legacy plants (3866 MW) reverted to the Balancing Pool between January 1, 2016 and May 5, 2016 (see Table \ref{tab:PPA}).
+As detailed in \cite{leach_tombe_ppa}, this was caused nominally by the increase in stringency of the \textit{SGER}, but the adoption of the Climate Leadership Plan and the \textit{CCIR} meant that the expected value of the PPAs was negative for their owners \cite{leach_tombe_ppa}.
+The fact that these changes in offer control, and implied changes in market power shown in Figure \ref{fig:mkt_power}, happened coincidentally with changes in carbon pricing in 2016 and 2017 have the potential to confound our inference.
+We discuss this in our results section.
 
@@
 \begin{figure}[t]%%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../SGER/sector_compliance_cost.png}
-\vspace{-0.75cm}	\caption{Generation-weighted annual average compliance cost (dashed black lines) and individual plant realizations (lighter grey lines) by plant type (\$/MWh), combined with the price charged on excess net annual emissions (black lines) under Alberta's carbon pricing programs (\$/tonne).
-Coal plants and natural gas combined-cycle (NGCC) and simple-cycle (SCGT) plants with more than 20MW installed capacity are shown.
-Source: Alberta Environment compliance cost data combined with AESO generation data provided by NRGStream.}
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../SGER/compliance_cost_policies.png}
+\vspace{-0.75cm}	\caption{Generation-weighted annual average compliance cost (solid black lines) and individual plant realizations (lighter dashed lines) by plant type (\$/MWh).
+Coal plants and natural gas combined-cycle (CC) and simple-cycle (SC) plants with more than 20MW installed capacity are included in this figure.
+The outlier NGCC plant is the Fort Nelson facility which is subject to BC's carbon tax but is connected to the Alberta electricity grid.
+Source: Government of Canada and Alberta Environment compliance cost data combined with AESO generation data provided by NRGStream.}
 \label{fig:sector_compliance_cost}
@@
 
-\section{Estimation Strategy}
+\section{Estimation Strategy}\label{sec:est}
 The focus of our estimation strategy is to identify the magnitude of the vertical shift in the electricity supply curves for the market as a whole or aggregated by generator fuel or controlling entity in response to changes in carbon pricing policies.
@@
 
-To measure this effect, we look at hourly merit orders and sample at multiple percentile points, using more sampling points at the upper end of the merit order.
-The variation in carbon pricing over time, as well as variation in which firms are participating in the market at different times allows us to develop a sample of offer values at each of our chosen percentile nodes over our entire sample.
-For each sampled point, we identify facility and market conditions, as well as carbon policy compliance costs, which we use to explain changes in the level offer levels over time.
 Consider the illustrative example in Figure \ref{fig:synth_illust} which shows shifts in a hypothetical two-firm supply curve in response to increasing carbon prices.
@@
 
-\begin{figure}[tb]
+\begin{figure}[t]
     \centering
@@
 
-For parts of our sample, censoring is also an issue.
+\begin{figure}[h]%%
+	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../images/expectation.png}
+\vspace{-0.75cm}	\caption{\textit{Ex ante} expected outcome for estimates of $\zeta$ or (-) $\kappa$ for offer percentiles $j\in (0,100)$ in (\ref{eq:regression}).}
+\label{fig:expectation}
+\end{figure}
+
+The variation in carbon pricing over time, as well as variation in which firms are participating in  the market, and which facilities' offers are present at a sampling point in the merit order at different times all allow us to identify the impact of carbon costs on offer values.
+For each sampled point in our synthetic merit order, we identify facility ($F$) and market conditions ($M$), as well as carbon policy compliance costs, which we use to explain changes in the level offer levels over time.
+Our regression equation is given by:
+\begin{align}\label{eq:regression}
+ \underbrace{P_{j,t}}_{\substack{\text{Offer value} \\ \text{at percentile }j\\ \text{at time }t}}&=
+ \underbrace{\alpha_{j}}_{\substack{\text{Constant for} \\ \text{percentile }j}}+
+ \underbrace{\beta_j M_t}_{\substack{\text{Market, policy, and} \\ \text{climate variables} \\\text{at time }t}}+
+ \underbrace{\delta_j F_{j,t}}_{\substack{\text{Facility} \\ \text{characteristics} \\ \text{at percentile }j\\ \text{at time }t}}+
+ \underbrace{\zeta_j t_{j,t}}_{\substack{\text{Carbon cost} \\ \text{at percentile j}\\ \text{at time }t}}+
+ \underbrace{\kappa_j O_{j,t}}_{\substack{\text{OBA value} \\ \text{at percentile j}\\ \text{at time }t}}+
+ \underbrace{\epsilon_{j,t}.}_{\substack{\text{Error term}}}
+\end{align}
+\noindent We estimate this equation for each level of aggregation that we consider: for the full market, for portfolios merit orders by fuel, and for portfolios by offer control.
+By construction, if there is full net carbon price pass-through, the coefficients $\zeta$ and $\kappa$ should be equal to 1 and our estimates of these coefficients are the focus of the balance of the paper.
+In setting up the analysis in this way, we take advantage of the fact that the specific facility' offer which occupies a particular sampling point is random, and firms would have no information (until hours after the fact) with respect to their exact position in the merit order, nor do they know the degree to which a particular percentile offer is assured to be infra- or extra-marginal in a given hour.
+Of course, firms do have information and expectations about the expected demand and will estimate the likely merit order in calculating their offers.
+
+If there is complete pass-through, at least at the margin and during peak hours as seen in \cite{fabra_reguant} and \cite{hintermann16_germany}, we would expect our estimates to look like those shown in Figure \ref{fig:expectation} in which some offers are constrained by considerations like minimum stable generation or contractual heat supply agreements, but that offers close to the margin reflect complete, near-complete, or more-than-complete pass-through of carbon policy costs to offers.
+While we do find this in some cases, this is not the case systematically through our analysis.
+To correct in part for facilities which face different incentives, we build an indicator of overall market-responsive behaviour which requires two elements to be true of a firm's offer strategy: a facility must have variability in the average value of power offered to the market (inter-quartile range of offer values greater than zero) and it much have variability in its upper quartile (80th percentile) offers as well.
+If both of these conditions are met, we deem a facility to be market-responsive.
+Figure \ref{fig:price_sense} shows how this indicator allows up to separate facilities in our data, with the implication that we should only expect to see pass-through from market-responsive facilities.
+We find that there are 11 facilities which do not meet either of these criteria and so we designated them non-market-responsive.
+
+For parts of our sample, censoring is an issue.
 In the Alberta market, negative price offers are not allowed, and offers are capped at \$1000/MWh.
+As such, for some percentiles $j$, the values of $P_{j,t}$ may be censored from below, as negative opportunity costs of power do not show up in our data, or they may be censored from above where a firm would offer higher than \$1000/MWh if that were allowed.
 This implies that we may under-estimate pass-through where either power offers have a less negative implied value below zero or where the carbon tax leads to a higher true costs of offer above \$1000, as we would not observe either of these effects.
 To correct for this, we augment our ordinary least squares estimates of pass-through with Tobit estimates with two-sided censoring, estimated via maximum likelihood.
-We do not find the censoring to have a marked effect on our estimates.
+As discussed further below, censoring does not appear to have a marked effect on our estimates.\footnote{Still working on error terms here for the Tobits.}
 
-\section{Results and Discussion}
+\section{Results and Discussion}\label{sec:results}
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!ht]
     \centering
@@
 
-These results are consistent if a decomposition by peak- vs.
-off-peak hours is introduced, as shown in Figure \ref{fig:decomp_peaks}.
-The impact of carbon pricing is always statistically greater than or equal to zero, with a maximum pass-through of approximately 40\% of average carbon costs in the mid-merit, and less elsewhere.
-In peak hours, the pass-through is smaller in the 50th to 60th percentiles, and larger in the 60th to 75th percentiles, with a similar 40\% maximum pass-through.
-Carbon price pass-through remains low in the 80th percentile, and increases in the upper merit during peak hours, while being close to zero in off-peak hours except at the extreme top of the merit order.
-The upper-merit impacts of OBAs are counter-intuitive, especially at peak hours, with estimates suggesting that higher offers have occurred with more valuable OBAs in the 85th to 95th percentiles of power offers in peak hours.
-
-\begin{figure}[!htb]
+\begin{figure}[!ht]
     \centering
@@
 \end{figure}
+
+These results are consistent across peak- and off-peak hours, as shown in Figure \ref{fig:decomp_peaks}.
+The impact of carbon pricing is always statistically greater than or equal to zero, with a maximum pass-through of approximately 40\% of average carbon costs in the mid-merit, and less elsewhere.
+In peak hours, the pass-through is smaller in the 50th to 60th percentiles, and larger in the 60th to 75th percentiles, with a similar 40\% maximum pass-through.
+Carbon price pass-through remains low in the 80th percentile, and increases in the upper merit during peak hours, while being close to zero in off-peak hours except at the extreme top of the merit order.
+The upper-merit impacts of OBAs are counter-intuitive, especially at peak hours, with estimates suggesting that higher offers have occurred when more valuable OBAs were available in the 85th to 95th percentiles of power offers in peak hours.
+This may be a consequence of the fact that OBA values decreased substantially for coal plants after 2018, at a time when fuel costs for natural gas plants also declined.
+As a result, the marginal (90-100\%) of generation from a coal unit would no longer be competitive with simple- or combined-cycle gas generators if declining OBA value were compensated-for with increased offer prices (see \cite{cullen_mansur} for a similar dynamic in US electricity markets subsequent to the decline in natural gas costs for generators).
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}[h!]
     \centering
@@
 
-\subsection{Impacts By Offer Control}
-We can also use the data post-2013 to assess the degree to which different market participants might offer their power into the merit order and how they have passed through carbon policy values to power offers.
-We only use data after 2013 for this analysis because, prior to that time, the AESO did not provide offer control records along with their merit order data.
-The remains a sufficient sample, since we have three full years of data (2013-2015) in the initial \SGER pricing program, along with at least one full year under each of the altered policy regimes over which to assess changed in offers by different market participants.
-
-PPA explanation...
-
-Balancing Pool explanation...
-
-Return of PPAs to the Balancing Pool
-
-Declining ability to identify TransCanada impacts because of changes in resources.
-
-\begin{figure}[!htb]
+\begin{figure}[ht!]
     \centering
@@
 
-\section{Conclusion}
+\subsection{Impacts By Offer Control}
+We can also use the data post-2013 to assess the degree to which different market participants might offer their power into the merit order and how they have passed through carbon policy values to power offers.
+Prior to 2013, the AESO did not provide offer control records along with their merit order data.
+We use two approaches to test offer control impacts: first, we use only the post-2013 data, which remains a sufficient sample, since we have three full years of data (2013-2015) in the initial \SGER pricing program, along with at least one full year under each of the altered policy regimes over which to assess changed in offers by different market participants.
+Second, we impute offer control backwards to use our entire sample, although we find that this does not significantly alter our results.
+
+Before we turn to the results, there is a potential endogeneity issue with respect to offer control that affects our results..
+
+The dependent variable in our regression analysis is the offered price in a given hour, and so we don't have a direct, endogenous relationship between offer control and our dependent variable either before or after the 2016 changes in offer control.
+We do, however, have a sample selection issue in our analysis by offer control, since the selection of units into the Balancing Pool portfolio, or our of the portfolios of TransCanada, Enmax, and Capital Power.
+Ownership is always a choice variable, and so portfolios are determined as a function of expectations of future market conditions in all cases, so the impact of the PPA returns are not in a category of their own.
+Other ownership changes, including ATCO's decision to divest its fossil-fueled assets to Heartland Generation September 30, 2019 also occurred during our analysis sample.
+
+The results by offer control show that:
+\begin{itemize}
+  \item Balancing pool is very strange - appears to have operated counter to incentives. Talk about lawsuit, etc. including political constraints
+  \item ENMAX - prototypical results partly because its fleet is largely highly-flexible combined-cycle gas
+  \item Other firms - TA, TC, Capital saw big changes in offer controlled fleet during the sample period, so likely not as informative
+  \item Capital upper merit is interesting and might be a result of simple-cycle plants in the Ancillary services market so that denominator in percentile calculations is over-estimated.
+  \item Might focus this section just on BP and ENMAX, and then add a graph with just ENMAX Shepard which offers into the market EXACTLY as we would expect.
+\end{itemize}
+
+\section{Conclusion}\label{sec:conclusion}
+
+To be written.
 
 \bibliographystyle{chicago}
-\bibliography{../env}
+\bibliography{../../env}
 
 \end{document}
-
-\section{Introduction}
-We examine the offer behaviour of competitive electricity suppliers in response to carbon prices.
-We exploit a unique set of significant policy variations over time and across facilities to measure carbon pricing pass-through to electricity prices.
-Beginning in 2007, the province of Alberta, Canada, instituted carbon pricing which affected generators in its energy-only wholesale electricity market.
-Changes in provincial governments in 2015 and 2019 led to the emissions pricing scheme being changed 3 times in 4 years, with the first 3 changes all announced during the 2015 calendar year, and a subsequent change announced in 2018.
-These changes were each implemented in such a way as to have firm- and facility-type-specific impacts which generates variation that we can exploit to separately characterize the role of carbon pricing and output-based allocations (OBAs) on pricing behaviour.
-
-Compare to Mar and other literature.
-
-First, in June 2015, it was announced that the carbon price would increase to \$20 per tonne and that output-based allocations would decrease beginning in 2016.
-The price was also set to increase further, to \$30 per tonne, in 2017.
-Later in 2015, following the recommendations of an expert panel, the government announced that it would alter the formula by which output-based allocations were provided to one which was consistent across all electricity generators beginning in 2018.
-This change was material, with coal plant allocations reduced by more than 0.5t/MWh in many cases, leading to changes in average costs of output of more than \$15/MWh for some facilities.
-This combination of policy changes implies that the effective marginal cost of emissions and the average emissions cost per unit output both changed significantly, with changes having different impacts on generators within and across fuel types.
-We show the effect of the policy changes on average revenue by fuel type in Figure 1.
-
-\begin{figure}[!h]%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../alberta_power/price_capture_ctax.png}
-\label{fig:masnadi_1}
-\vspace{-0.75cm}	\caption{Impact of changes in carbon pricing on plant revenues by type}
-\end{figure}
-
- Fossil-fuel-fired generators were charged a carbon price of \$15/tonne on emissions net of allocations of emissions credits.
-Emissions credits were allocated per unit output at rates based on facility-specific emissions-intensity histories.
-Renewable generators were also able to qualify for offset credits which provided an alternative revenue stream.
-
-\noindent
-
-\newpage \noindent We use data including hourly electricity prices and loads, plant availability, as well as offer functions at the plant level, historic emissions intensities, and emissions policy compliance records to assess the degree to which carbon prices have been passed through to wholesale electricity prices.
-We also assess the degree to which changes in the composition of the market including changes to offer control and firm concentration, as well as changes due to facility exit, have altered the pass-through of carbon prices.
-Our work builds largely on the analysis of electricity price pass-through of \citet{fabra_reguant}, but also draws on work by \citet{hintermann17} on market power, \citet{hintermann15} and \citet{hintermann16} on impacts of carbon pricing in the EU-ETS, \citet{nazifi16} on pass-through in Australia's electricity market, and \citet{woo_et_al} on pass-through in California's market.
-
-\noindent Our preliminary analysis suggests that there has been both a significant effect on emissions (see Figure \ref{fig:monthly_ghg}) and on the composition of the market (see Figure \ref{fig:monthly_gen}) with significant reduction in the generation from coal plants and significant increase in the hourly daily cycling of coal plants as the change in carbon prices push coal plants higher in the merit order relative to gas generation, a dynamic examined in \citet{cullen_mansur}.
-
-\begin{figure}[p]%
-	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../../alberta_power//monthly_ghg_mwh.png}
-\vspace{-0.75cm}	\caption{Impact of changes in carbon pricing on GHG emissions}\label{fig:monthly_ghg}
-\end{figure}
-
-\begin{figure}[p]%
-	\centering \vspace{-.25cm} \includegraphics[width=5.5in]{../../alberta_power/gen_ghg_price.png}
-\vspace{-.5cm}	\caption{Impact of changes in carbon pricing on market share by fuel.
-Combined heat and power (COGEN), simple-cycle gas turbines (SCGT) and natural gas combined cycle (NGCC) facilities are each natural-gas-fired.}\label{fig:monthly_gen}
-\end{figure}
-
-%\begin{figure}[!h]%
-	%\centering \vspace{-.5cm} \includegraphics[width=6.5in]{CJE/masnadi_fig_1.png} %\label{fig:masnadi_1}
-%	\caption{Figure 1 from Masnadi et al. (2018)}
-%	\vspace{-.5cm}
-%\end{figure}
-\newpage
-\vspace{10cm}
-
-\begin{figure}[!h]%
-	\centering \vspace{-.25cm} \includegraphics[width=6.5in]{../../alberta_power/peak_prices_2009_2019.png}
-\label{fig:masnadi_1}
-\vspace{-0.75cm}	\caption{Peak and off-peak power prices, 2007-2019}
-\end{figure}
-
-\newpage

### 2022-06-11 ep_paper_graphs.R

@@ -7,9 +7,9 @@ library(cowplot)
 
 
 load(file="data/metered_vols_data.Rdata")
-#load(file="nrgstream/nrgstream_gen.Rdata") 
+load(file="nrgstream/nrgstream_gen.Rdata") 
 
-#update_forecasts()
+update_forecasts()
 load(file="data/forecast_data.Rdata") 
 forecast_data <- forecast_data %>% filter (he!="02*")
 
@@ -18,78 +18,109 @@ forecast_data <- forecast_data %>% filter (he!="02*")
 
 #sample period prices and loads
 
-peak_data<-forecast_data %>%filter(!is.na(actual_posted_pool_price),!is.na(actual_ail))%>% group_by(year,month) %>% summarize(ail=mean(actual_ail,na.rm = T),peak_ail=max(actual_ail),trough_ail=min(actual_ail),
-                                                                                                                              mean_peak_price=sum(actual_posted_pool_price*actual_ail*(on_peak==TRUE),na.rm = T)/sum(actual_ail*(on_peak==TRUE),na.rm = T),
-                                                                                                                              mean_off_peak_price=sum(actual_posted_pool_price*actual_ail*(on_peak==FALSE),na.rm = T)/sum(actual_ail*(on_peak==FALSE),na.rm = T),
-                                                                                                                              mean_peak_ail=sum(actual_ail*(on_peak==TRUE),na.rm = T)/sum((on_peak==TRUE),na.rm = T),
-                                                                                                                              mean_off_peak_ail=sum(actual_ail*(on_peak==FALSE),na.rm = T)/sum((on_peak==FALSE),na.rm = T),
-                                                                                                                              mean_price=sum(actual_posted_pool_price*actual_ail,na.rm = T)/sum(actual_ail,na.rm = T),peak_price=max(actual_posted_pool_price),trough_price=min(actual_posted_pool_price)
-)%>%
[…]

### 2022-06-11 merit_build.R

@@ -9,14 +9,15 @@ start_time<-Sys.time()
 
 options(scipen=999)
 
-update<-1 #add new data
+update<-0 #add new data
 save<-1 #save files at the end
 synth<-0 #synthetic plants?
-  synth_type<-2  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_fuel, 0 is full merit as synthetic plant
+  synth_type<-0  #5 is a target facility, focus_id,4 is facility,3 is offer control by type, 2 is by offer_control,1 is by plant_fuel, 0 is full merit as synthetic plant
 
   if(synth_type==5)
     focus_id<-c("SH1","SH2")
 
+  
 load("data/all_merit.RData")  
 if(update==1){
   merit_data<-rbind(merit_data,update_merit(merit_data))
@@ -79,7 +80,7 @@ load("data/forecast_data.RData")
   if(update==1)
    {
    load(file="data/metered_vols_data.Rdata" ) 
-    #all_vols<-all_vols %>% filter(year<2019)
+    #all_vols<-all_vols %>% filter(year<2022)
     all_vols<-update_vols(all_vols)
     save(all_vols,file="data/metered_vols_data.Rdata" ) 
     #isolate renewable (non-hydro and biomass) volumes from metered volumes - the ones that default bid to zero
@@ -148,12 +149,14 @@ load("data/forecast_data.RData")
  
 #testing storage objects
 #merit_store<-merit_aug  
-#merit_small<-merit_aug%>%filter(year(date)==2019)
+#merit_small<-merit_aug%>%filter(year(date)==2022)
 #merit_aug<-merit_small
   
 #merit_aug<-merit_store  
 #merit_aug<-merit_small
 
+  
+plant<-plant_data()
   
 merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
   mutate(year=year(date))%>%
@@ -166,11 +169,61 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
                            year(date)==2019 ~ "CCIR_30",
                            year(date)==2020 ~ "TIER_30",
                            year(date)==2021 ~ "TIER_40",
-                           TRUE ~ "TIER_40")
[…]

### 2022-06-11 nrgstream_gen.R

@@ -253,9 +253,40 @@ get_plant_info()
 #nrgstream_gen<-all_gen_data()
 #rm(gen_data)
 #if we're changing over to a new year, run both of these two lines of code:
-nrgstream_gen<-filter(nrgstream_gen,year(Time)<=2020)
-nrgstream_gen<-data_update(nrgstream_gen,2021)
+nrgstream_gen<-filter(nrgstream_gen,year(Time)<=2021)
+#nrgstream_gen<-data_update(nrgstream_gen,2021)
 nrgstream_gen<-data_update(nrgstream_gen,2022)
+
+
+#Repair coal-to-gas-conversions
+nrgstream_gen<-nrgstream_gen %>% 
+  mutate(Plant_Type=case_when( 
+    (ID=="HRM") & (date>=ymd("2020-05-08")) ~ "SCGT",  #Milner change to gas effective 
+    #SH1 Sheerness #1 and SH2 Sheerness #2 -July 30, 2021.https://www.aeso.ca/market/market-updates/2021/sh1-sheerness-1-and-sh2-sheerness-2-change-in-fuel-type-notice/
+    (ID %in% c("SH1","SH2")) & (date>=ymd("2021-07-30")) ~ "NGCONV",
+    #KH3 January 11, 2022
+    (ID =="KH1") & (date>=ymd("2022-01-11")) ~ "NGCONV",
+    #KH2 July 27, 2021.
+    (ID =="KH2") & (date>=ymd("2021-07-21")) ~ "NGCONV",
+    #Battle River #4 (BR4)	March 8, 2022
+    (ID =="BR4") & (date>=ymd("2022-03-08")) ~ "NGCONV",
+    #Battle River #5 (BR5)	 November 19, 2021
+    (ID =="BR5") & (date>=ymd("2021-11-19")) ~ "NGCONV",
+    #Sundance #6 (SD6)	401	0	0 February 19, 2021
+    (ID =="SD6") & (date>=ymd("2021-02-19")) ~ "NGCONV",
+    TRUE ~ Plant_Type),
+    Capacity=case_when(
+      (ID=="HRM") & (date>=ymd("2020-04-23"))&(date<ymd("2020-05-08")) ~ 185,  #Milner change to gas effective 
+      (ID=="HRM") & (date>=ymd("2020-05-08"))&(date<ymd("2021-12-09")) ~ 208,  #Milner change to gas effective 
[…]

### 2022-06-11 paper_graphs.R

@@ -388,23 +388,24 @@ paste("loaded file=data/",data_files$value,sep="")
 
 #test<-merit_aug %>% filter(date==ymd("2019-02-04"),he=="19",Plant_Type=="COAL")%>% select(actual_posted_pool_price)
 
-merit_aug %>% filter(date==ymd("2019-02-04"),he=="19",Plant_Type=="COAL")%>% 
+coal_merit<-merit_aug %>% filter(date==ymd("2019-02-04"),he=="19",Plant_Type=="COAL")%>% 
   mutate(facility=gsub(" #","_",AESO_Name))%>% separate(facility,into = c("facility","number"), sep="_(?=[^_]+$)")%>%
   group_by(facility,price,dispatched) %>% summarize(size=sum(size),dispatched_mw=sum(dispatched_mw))%>% ungroup()%>%
   arrange(price,facility)%>%
   mutate(merit=cumsum(size),total_offers=sum(size),total_dispatch=sum(dispatched_mw),dispatch_limit=max(merit*(dispatched=="Y")))%>%
   ggplot()+
   geom_rect(aes(xmin=merit-size,xmax=merit,ymin=-20,ymax=price,fill=facility),color="black")+
-  geom_vline(aes(xintercept=max(dispatch_limit)))+
-  annotate("text",3950,y=815,label="All offers to the\nleft of this line\nwere dispatched",size=rel(6),hjust=1,vjust=0)+
+  #geom_vline(aes(xintercept=max(dispatch_limit)))+
   #geom_label_repel(aes(merit-size/2,y=price,label=facility))+
-  scale_fill_grey("",end = 1,start=0)+   
+  #scale_fill_grey("",end = 1,start=0)+   
+  scale_fill_manual("",values=c(grey.colors(6, end = 1,start=0)))+
   scale_x_continuous(breaks=pretty_breaks(), expand=c(0,0))+
   scale_y_continuous(breaks=pretty_breaks(), expand=c(0,0))+
   #scale_colour_manual(labe ls=c("Brent","WTI"),values=c("#41ae76","#238b45","#006d2c","#00441b","Black","Black","Black","Black"))
   paper_theme()+
   theme(
-    legend.position = c(.1,.8),
+    legend.justification = c("left", "top"),
+    legend.position = c(.05,.99),
     legend.margin=margin(c(0,0,0,0),unit="cm"),
     legend.text = element_text(colour="black", size = 18),
[…]

### 2022-06-11 regression_analysis.R

@@ -67,7 +67,7 @@ offer_test<-merit_data %>%
 
 
 offer_graph<-offer_test %>% pivot_longer(-c(date,he,time,total),names_to = "measure")%>%
-  mutate(no_bp=grepl("nbp",measure)) %>%
+  mutate(no_bp=grepl("nbp",measure)) %>% 
   mutate(no_bp=as_factor(no_bp),
          no_bp=fct_recode(no_bp,"Balancing Pool Omitted"="TRUE",
                           "Balancing Pool Included"="FALSE"))%>%
@@ -80,7 +80,7 @@ offer_graph<-offer_test %>% pivot_longer(-c(date,he,time,total),names_to = "meas
 
 
 
-ggplot(offer_graph%>%filter(year(date)<2020)%>%
+ggplot(offer_graph%>%#filter(year(date)<2022)%>%
          mutate(year=year(date),month=month(date))%>%
          group_by(year,month,non_zero,no_bp)%>%
          #group_by(date,non_zero,no_bp)%>%
@@ -90,11 +90,11 @@ ggplot(offer_graph%>%filter(year(date)<2020)%>%
   facet_grid(cols = vars(no_bp))+
   scale_color_manual("",values=c("black","grey50","grey80"))+
   scale_linetype_manual("",values=c("solid","11"))+
-  expand_limits(y=c(0,100))+
-  expand_limits(x=ymd("2020-3-01"))+
+  expand_limits(y=100)+
+  expand_limits(x=ymd("2022-05-01"))+
   #guides(color="none")+
   scale_y_continuous(expand = c(0,0),breaks=pretty_breaks())+
-  scale_x_date(expand = c(0,0),breaks=pretty_breaks())+
+  scale_x_date(expand = c(0,0),breaks=pretty_breaks(n=8))+
   paper_theme()+
   theme(panel.spacing = unit(2,"lines"),
         legend.key.width = unit(3,"lines"),
@@ -106,11 +106,17 @@ ggplot(offer_graph%>%filter(year(date)<2020)%>%
   labs(x="",y=expression("Key Firms' Average Share of Hourly Offers"*phantom(1)*'(%)'),
        NULL
   )+
-  NULL
+  annotate("rect",xmin=ymd("2016-01-01"),xmax=ymd("2016-5-5"),ymin=-Inf,ymax=93.5, alpha=0.1, fill="black")+
[…]

### 2022-06-11 regression_analysis_bak.R

@@ -80,7 +80,7 @@ offer_graph<-offer_test %>% pivot_longer(-c(date,he,time,total),names_to = "meas
 
 
 
-ggplot(offer_graph%>%filter(year(date)<2020)%>%
+ggplot(offer_graph%>%#filter(year(date)<2022)%>%
          mutate(year=year(date),month=month(date))%>%
          group_by(year,month,non_zero,no_bp)%>%
          #group_by(date,non_zero,no_bp)%>%
@@ -90,11 +90,11 @@ ggplot(offer_graph%>%filter(year(date)<2020)%>%
   facet_grid(cols = vars(no_bp))+
   scale_color_manual("",values=c("black","grey50","grey80"))+
   scale_linetype_manual("",values=c("solid","11"))+
-  expand_limits(y=c(0,100))+
-  expand_limits(x=ymd("2020-3-01"))+
+  expand_limits(y=100)+
+  expand_limits(x=ymd("2022-05-01"))+
   #guides(color="none")+
   scale_y_continuous(expand = c(0,0),breaks=pretty_breaks())+
-  scale_x_date(expand = c(0,0),breaks=pretty_breaks())+
+  scale_x_date(expand = c(0,0),breaks=pretty_breaks(n=8))+
   paper_theme()+
   theme(panel.spacing = unit(2,"lines"),
         legend.key.width = unit(3,"lines"),
@@ -106,12 +106,18 @@ ggplot(offer_graph%>%filter(year(date)<2020)%>%
   labs(x="",y=expression("Key Firms' Average Share of Hourly Offers"*phantom(1)*'(%)'),
        NULL
   )+
+  annotate("rect",xmin=ymd("2016-01-01"),xmax=ymd("2016-5-5"),ymin=-Inf,ymax=93.5, alpha=0.1, fill="black")+
+  annotate("text", x = ymd("2016-1-1")+days(round(125/2)), y = 94, label = "PPAs\nreturned\nto BP",size=3.5,hjust = 0.5,vjust = 0)+  
+  annotate("rect",xmin=ymd("2020-12-15"),xmax=ymd("2021-1-15"),ymin=-Inf,ymax=93.5, alpha=0.1, fill="black")+
+  annotate("text", x = ymd("2020-12-31"), y = 94, label = "PPA\nexpiration\nDec. 2020",size=3.5,hjust = 0.5,vjust = 0)+  
[…]

### 2022-06-16 presentations/cea_creea_deck.tex

\documentclass[aspectratio=169]{beamer}
\usepackage{fontawesome} % Required for FontAwesome icons
\usepackage{graphicx,ualberta,url}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
%\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{multirow}
\usepackage{capt-of}
\usepackage{graphicx}
\usepackage{array}
\usepackage{tikz}
\usepackage{ifpdf}

\urlstyle{same}

\hypersetup{
    colorlinks,
    citecolor= blue,
    filecolor= blue,
    linkcolor= ualbertagold,
    urlcolor= ualbertagreen
}

\newcommand{\TextUnderscore}{\rule{.4em}{.4pt}}

\newcommand*{\vcenteredhbox}[1]{\begingroup
\setbox0=\hbox{#1}\parbox{\wd0}{\box0}\endgroup}

\newcommand{\icon}[4][black]{% The first parameter is the FontAwesome icon name, the second is the box size and the third is the text
    \vcenteredhbox{{
    \makebox(#3, #3){\textcolor{#1}{\large\csname fa#2\endcsname}}
    }}% Icon and box
	\hspace{0.2cm}% Whitespace
    \vcenteredhbox{
    \textcolor{#1}{#4}
    }% Text
}

%----------------------------------------------------------------------------------------
%	GRAPHICS DEFINITIONS
%----------------------------------------------------------------------------------------

\definecolor{bg-color}{HTML}{4b6a78}
\definecolor{grayish}{HTML}{DDDDDD}
\definecolor{code}{HTML}{e83e8c}

\begin{document}

\title[Carbon Price Pass-Through]{Carbon Price Pass-Through in Alberta's Electricity Market}
%\author{Branko Bo\v skovi\'{c} \and Andrew Leach \and Charles F.\ Mason}
%\institute{University of Alberta  \and University of Alberta  \and University of Wyoming}

%\author[shortname]{author1 \inst{1} \and author2 \inst{2}}
%\institute[shortinst]{\inst{1} affiliation for author1 \and %
 %                     \inst{2} affiliation for author2}

\author[Leach and Shaffer]{%
  \texorpdfstring{%
    \begin{columns}
      \column{.1\linewidth}
      \column{.4\linewidth}
      \centering
      Andrew Leach \\University of Alberta
      \column{.4\linewidth}
      \centering
      Blake Shaffer \\University of Calgary
      \column{.1\linewidth}
      \end{columns}
 }
 {Leach and Shaffer (2022)}
}
\date{CEA Conference 2022}

%\frame[plain]{
%  \titlepage
%}

\frame{\vspace{-2cm}
  \titlepage
}

\section{Introduction}

\begin{frame}{Carbon Pricing's Effect on Electricity Markets}
    \begin{itemize}
    \item We want to understand how firms react to carbon pricing: do they pass through marginal prices as our models assume?
    \item Electricity markets are a real-life supply and demand model
    \item Alberta GHG policies changed carbon prices and output-based credit allocations across time and facilities
    \item We build a rich data set of relevant firm and market covariates
    \item Our results show weaker pass through than seen in Fabra and Reguant (2014) or Hintermann (2016) which each find near-complete pass-through of emissions costs
    \item We are left with the question of why costs might not be fully passed-through to market offer curves, at least for marginal blocks of power
    \end{itemize}
\vfill
\end{frame}

\section{Motivation and Context}

%\begin{frame}{Oil sands reserves\hspace{3cm} \small Source: AER (2013)}
%\pgfputat{\pgfxy(0,.25)}{\pgfimage[width=4.25in]{AB_reserves.pdf}}
%\end{frame}

\begin{frame}{Alberta's Power Market is an Ideal Laboratory}
    \begin{itemize}
    \item Alberta's power market is relatively small, isolated, with minimal intertie capacity:
    \small\begin{itemize}
    \item WECC (1325 MW export and 1500 MW import) and Saskatchewan (150 MW)
        \end{itemize}
    \end{itemize}
\vfill
\end{frame}

\begin{frame}{Alberta Power Market}
   \tikz [remember picture,overlay]
    \node[yshift=-.85cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.5\paperwidth]{../images/alberta_grid.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Alberta's Power Market is an Ideal Laboratory}
    \begin{itemize}
    \item Alberta's power market is relatively small, isolated, with minimal intertie capacity:
    \small\begin{itemize}
    \item WECC (1325 MW export and 1500 MW import) and Saskatchewan (150 MW)
        \end{itemize}
    \item Alberta's market is winter-peaking (load) but often has high summer prices
  \end{itemize}
\vfill
\end{frame}

\begin{frame}{Alberta Power Market Prices and Loads}
   \tikz [remember picture,overlay]
    \node[yshift=-0.6cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.75\paperwidth]{../images/prices_and_loads.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Alberta's Power Market is an Ideal Laboratory}
    \begin{itemize}
    \item Alberta's power market is relatively small, isolated, with minimal intertie capacity:
    \small\begin{itemize}
    \item WECC (1325 MW export and 1500 MW import) and Saskatchewan (150 MW)
        \end{itemize}
    \item Alberta's market is winter-peaking (load) but often has high summer prices
    \item Alberta's market has a wide variety of sources of generation
  \end{itemize}
\vfill
\end{frame}

\begin{frame}{Alberta Power Generation Sources}
   \tikz [remember picture,overlay]
    \node[yshift=-0.6cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.85\paperwidth]{../images/gen_fuel_policies.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Alberta's Power Market is an Ideal Laboratory}
    \begin{itemize}
    \item Alberta's power market is relatively small, isolated, with minimal intertie capacity:
    \small\begin{itemize}
    \item WECC (1325 MW export and 1500 MW import) and Saskatchewan (150 MW)
        \end{itemize}
    \item Alberta's market is winter-peaking (load) but often has high summer prices
    \item Alberta's market has a wide variety of sources of generation
    \item Carbon emissions reporting covers most facilities in the market
    \small\begin{itemize}
    \item Carbon pricing has been in place since 2007
    \item Carbon pricing design changed repeatedly from 2015-2022
        \end{itemize}
  \end{itemize}
\vfill
\end{frame}

\begin{frame}{Alberta's Policy Changes as Treatments}
    \begin{itemize}
    \item 2007-2015: \textit{Specified Gas Emitters Regulation (SGER)}
    \small\begin{itemize}
    \item Price of \$15/t, OBAs at 88\% of historic GHG/MWh
    \item Covered only large (>100kt/yr) facilities
    \end{itemize}
    \item 2016-2017: \textit{Specified Gas Emitters Regulation (SGER)}
    \small\begin{itemize}
    \item 2016: price \$20/t, OBAs at 85\% of historic GHG/MWh
    \item 2017: price \$30/t, OBAs at 80\% of historic GHG/MWh
    \item Covered only large (>100kt/yr) facilities
    \end{itemize}
    \item 2018-2019: \textit{Carbon Competitiveness Incentive Regulation (CCIR)}
    \small\begin{itemize}
    \item 2018-19: price \$30/t, OBAs at 0.37t/MWh for all facilities
    \item Covered large (>100kt/yr) facilities, with opt-in provision
    \end{itemize}
    \item 2019-present: \textit{Technology Innovation and Emissions Reduction Regulation (TIER)}
    \small\begin{itemize}
    \item 2020: price \$30/t, OBAs at 0.37t/MWh for all facilities
    \item 2021: price \$40/t, OBAs at 0.37t/MWh for all facilities
    \item 2022: price \$50/t, OBAs at 0.37t/MWh for all facilities
    \item Covers large (>100kt/yr) facilities, with opt-in provision
    \end{itemize}
    \end{itemize}
\vfill
\end{frame}

\begin{frame}{Carbon pricing impacts across facilities over time}
    \tikz [remember picture,overlay]
    \node[yshift=-.1cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.85\paperwidth]{../../sger/compliance_cost_policies.png}}; \vspace{4.85cm}
   \begin{itemize}
    \item different impacts of carbon pricing on plants of the same type at the same time, and changes in pricing for the same plants over time combine to allow us to identify expected pass-through of carbon tax costs and OBA value
    \end{itemize}
   \vfill
\end{frame}

\section{Empirical Framework}
\begin{frame}{Empirical Framework}
    \begin{itemize}
    \item We estimate GHG cost pass through using samples drawn from merit orders at each hour $t$ at each of $J$ percentile support points to estimate:
    \small\begin{align*}\label{eq:regression}
 \underbrace{P_{j,t}}_{\substack{\text{Offer value} \\ \text{at percentile }j\\ \text{at time }t}}&=
 \underbrace{\alpha_{j}}_{\substack{\text{Constant for} \\ \text{percentile }j}}+
 \underbrace{\beta_j M_t}_{\substack{\text{Market, policy, and} \\ \text{climate variables} \\\text{at time }t}}+
 \underbrace{\delta_j F_{j,t}}_{\substack{\text{Facility} \\ \text{characteristics} \\ \text{at percentile }j\\ \text{at time }t}}+\\&
 \underbrace{\zeta_j t_{j,t}}_{\substack{\text{Carbon cost} \\ \text{at percentile j}\\ \text{at time }t}}+
 \underbrace{\kappa_j O_{j,t}}_{\substack{\text{OBA value} \\ \text{at percentile j}\\ \text{at time }t}}+
 \underbrace{\epsilon_{j,t}}_{\substack{\text{Error term}}}
\end{align*}
     \end{itemize}
   \vfill
\end{frame}

\begin{frame}{Empirical Framework}
       \tikz [remember picture,overlay]
    \node[yshift=-.7cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.8\paperwidth]{../images/sampling.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Empirical Framework}
       \tikz [remember picture,overlay]
    \node[yshift=-.7cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.99\paperwidth]{../images/coal_synth_cea.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Empirical Framework}
    \begin{itemize}
    \item We estimate GHG cost pass through using samples drawn from merit orders at each hour $t$ at each of $J$ percentile support points to estimate:
    \small\begin{align*}
 \underbrace{P_{j,t}}_{\substack{\text{Offer value} \\ \text{at percentile }j\\ \text{at time }t}}&=
 \underbrace{\alpha_{j}}_{\substack{\text{Constant for} \\ \text{percentile }j}}+
 \underbrace{\beta_j M_t}_{\substack{\text{Market, policy, and} \\ \text{climate variables} \\\text{at time }t}}+
 \underbrace{\delta_j F_{j,t}}_{\substack{\text{Facility} \\ \text{characteristics} \\ \text{at percentile }j\\ \text{at time }t}}+\\&
 \underbrace{\zeta_j t_{j,t}}_{\substack{\text{Carbon cost} \\ \text{at percentile j}\\ \text{at time }t}}+
 \underbrace{\kappa_j O_{j,t}}_{\substack{\text{OBA value} \\ \text{at percentile j}\\ \text{at time }t}}+
 \underbrace{\epsilon_{j,t}}_{\substack{\text{Error term}}}
\end{align*}
\item The estimated values of parameters $\zeta_j$ and $\kappa_j$ will tell us the degree of pass-through of carbon tax costs and OBA values at each percentile $j$ of each merit order we examine.
     \end{itemize}
   \vfill
\end{frame}

\begin{frame}{Expected Estimates of  $\zeta_j$}
       \tikz [remember picture,overlay]
    \node[yshift=-.6cm,xshift=0cm] at (current page.center)
       {\includegraphics[width=.7\paperwidth]{../images/expectation.png}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Results}

\begin{frame}{Carbon pricing pass through, full market}
   \tikz [remember picture,overlay]
    \node[yshift=-.6cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.7\paperwidth]{../images/all_plants_net.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Carbon pricing pass through, full market}
   \tikz [remember picture,overlay]
    \node[yshift=-.6cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.7\paperwidth]{../images/all_plants_no_peaks.png}}; \vspace{1cm}
   \vfill
\end{frame}

\begin{frame}{Carbon pricing pass through, full market}
   \tikz [remember picture,overlay]
    \node[yshift=-.6cm,xshift=0cm] at (current page.center)
        {\includegraphics[width=.7\paperwidth]{../images/all_plants_net_peak.png}}; \vspace{1cm}
   \vfill
\end{frame}

\section{Discussion}

\begin{frame}{Discussion}
   \begin{itemize}
   \item Alberta's energy-only power market provides an ideal environment to study carbon price pass-through
    \item Alberta policy changes alter both carbon prices and output-based credit allocations provide facility-specific treatment
    \item Results imply much lower pass-through rates than previous work in this area
    \item The question is why this would be the case?
    \small\begin{itemize}
    \item Coal-to-gas price ratios (Cullen and Mansur (2017))
    \item Strategic behaviour
    \item Omitted variables (facilities with invariant bids)
    \item Changes in ownership and the role of the Balancing Pool during our sample period
    \end{itemize}
     \end{itemize}
   \vfill
\end{frame}

\section{Contact}

%\begin{frame}{Oil sands reserves\hspace{3cm} \small Source: AER (2013)}
%\pgfputat{\pgfxy(0,.25)}{\pgfimage[width=4.25in]{AB_reserves.pdf}}
%\end{frame}

\begin{frame}{Contact information}
\vspace{-.5 cm}
\begin{center}
Andrew Leach, Alberta School of Business, University of Alberta\\[2ex]

\begin{minipage}[t]{0.15\textwidth} % 27.5% of the page width for the first row of icons
	\vspace{-\baselineskip} % Required for vertically aligning minipages
	\end{minipage}
      \hfill
\begin{minipage}[t]{0.5\textwidth} % 27.5% of the page width for the first row of icons
	\vspace{-\baselineskip} % Required for vertically aligning minipages
	%\icon{Phone}{12}{\textcolor[rgb]{0.00,0.00,0.00}{+1 780-492-8489}}\\[-0.5ex]
	\icon{EnvelopeO}{12}{\href{mailto:aleach@ualberta.ca}{aleach@ualberta.ca}}\\[-0.5ex]
    \icon{Globe}{12}{\href{https://aleach.ca}{aleach.ca}}
    \end{minipage}
      \hfill
\begin{minipage}[t]{0.4\textwidth} % 27.5% of the page width for the second row of icons
	\vspace{-\baselineskip} % Required for vertically aligning minipages

	% The first parameter is the FontAwesome icon name, the second is the box size and the third is the text
	% Other icons can be found by referring to fontawesome.pdf (supplied with the template) and using the word after \fa in the command for the icon you want
    %\icon{Linkedin}{12}{\href{https://linkedin.com/in/leach-econ}{leach-econ}}\\	[-0.5ex]	
    \icon{Github}{12}{\href{https://github.com/leachandrew}{leachandrew}}\\[-0.5ex]
    \icon{Twitter}{12}{\href{https://twitter.com/andrew_leach}{@andrew\TextUnderscore leach}}
\end{minipage}
\\[+6ex]
Blake Shaffer, Dept of Economics and School of Public Policy, University of Calgary\\[2ex]

\begin{minipage}[t]{0.15\textwidth} % 27.5% of the page width for the first row of icons
	\vspace{-\baselineskip} % Required for vertically aligning minipages
	\end{minipage}
      \hfill
\begin{minipage}[t]{0.5\textwidth} % 27.5% of the page width for the first row of icons
	\vspace{-\baselineskip} % Required for vertically aligning minipages
	\icon{EnvelopeO}{12}{\href{mailto:blake.shaffer@ucalgary.ca}{blake.shaffer@ucalgary.ca}}\\[-0.5ex]
    \icon{Globe}{12}{\href{https://blakeshaffer.ca}{blakeshaffer.ca}}
    \end{minipage}
      \hfill
\begin{minipage}[t]{0.4\textwidth} % 27.5% of the page width for the second row of icons
	\vspace{-\baselineskip} % Required for vertically aligning minipages

	% The first parameter is the FontAwesome icon name, the second is the box size and the third is the text
	% Other icons can be found by referring to fontawesome.pdf (supplied with the template) and using the word after \fa in the command for the icon you want
    \icon{Github}{12}{\href{https://github.com/blakeshaffer}{blakeshaffer}}\\[-0.5ex]
    \icon{Twitter}{12}{\href{https://twitter.com/bcshaffer}{@bcshaffer}}
\end{minipage}
\end{center}

\vfill
\end{frame}

\end{document}

### 2022-10-05 presentations/power_slides_law.tex
fixing charts and deck

@@
 
-\documentclass{beamer}
+\documentclass[aspectratio=169]{beamer}
 
@@
 
+% Bayliss Fancy fit image command with optional caption
+\usepackage{adjustbox} % Shrink stuff
+
+\makeatletter
+
+\newcommand{\fitimage}[2][\@nil]{
+  \begin{figure}
+    \begin{adjustbox}{width=0.95\textwidth, totalheight=\textheight-2\baselineskip-2\baselineskip,keepaspectratio}
+      \includegraphics{#2}
+    \end{adjustbox}
+    \def\tmp{#1}%
+   \ifx\tmp\@nnil
+      \else
+      \newline{\tiny{#1}}
+    \fi
+  \end{figure}
+}
+
+\makeatother
+
 % The main document
@@
 \setlength\itemsep{.5em}
-\item Fed/prov issues, environmental policy, section 125 jurisprudence, the oddity that is section 92(10) including the declaratory power and POGG all come into play!
+\item Fed/prov issues, environmental policy, section 125 jurisprudence, the oddity that is section 92(10) including the declaratory power, and POGG all come into play!
 \end{itemize}
@@
     \centering
-    \includegraphics[width=.9\linewidth]{../images/grid_storage.png} \\[\abovecaptionskip]
+    \includegraphics[width=.75\linewidth]{../images/grid_storage.png} \\[\abovecaptionskip]
   Source: \url{http://www.mpoweruk.com/grid_storage.htm}
@@
 \begin{itemize}
-\setlength\itemsep{1.7em}
+\setlength\itemsep{1.2em}
 \item Generation
@@
 \setlength\itemsep{1.7em}
-\item Generation is a competitive market competitive
+\item Generation is a competitive market
 
@@
 
-\item Distribution (wires) is regulated on a cost-of-service basis, but retail (billing) is competitive.
+\item Distribution (local wires) is regulated on a cost-of-service basis, but retail (billing) is competitive.
 
@@
 \setlength\itemsep{.5em}
-\item Alberta system demand: 7,200-10,700 MW (million watts)
+\item Alberta system demand: 7,200-10,700 MW (megawatts or million watts)
 \item Capital Power's Genessee 3 power plant has a nameplate capacity of 450 MW
@@
 \item Production over a day, week, month, year
+\item A 7.5W LED bulb on all day (24hr) uses 180Wh of electricity (.18kWh)
 \end{itemize}
@@
 \begin{itemize}
-\setlength\itemsep{.25em}
+\setlength\itemsep{1.25em}
 \item Rate-regulated or or state-owned utilities
+\begin{itemize}
+\setlength\itemsep{.15em}
+\item EPCOR (legacy) or BC Hydro
+\item PG \& E in California
+\end{itemize}
+
 \item Competitive markets
@@
 
-\begin{frame}{Alberta Market Design}
+\begin{frame}{Alberta Wholesale Energy Market Design}
 \begin{itemize}
@@
 \item Single node
-\item Capacity market to be added in the near future
+\item Capacity market was contemplated, but not pursued
 \item Transmission
+\begin{itemize}
+\setlength\itemsep{.25em}
+\item Regulated monopoly
 \item Congestion free (no nodal pricing)
 \item No transmission rights
+\item Paid for (mostly) by load (consumers)
+\end{itemize}
 \item Ancillary services: separate, competitive market for operating reserves, transmission-must-run, load-shed and black start
@@
     \centering
-    \includegraphics[width=\linewidth]{../images/cali_nodes.png} \\[.1\abovecaptionskip]
+    \includegraphics[width=.7\linewidth]{../images/cali_nodes.png} \\[.1\abovecaptionskip]
   Source: CAISO
@@
 \begin{frame}{The Merit Order}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/merit_type.png}}; \vspace{1cm}
+   \fitimage{../images/merit_type.png}
    \vfill
@@
 \begin{frame}{The Merit Order}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/merit_offer.png}}; \vspace{1cm}
+   \fitimage{../images/merit_offer.png}
    \vfill
@@
 \begin{frame}{Offers}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/bids.png}}; \vspace{1cm}
+   \fitimage{../images/bids.png}
    \vfill
@@
 \begin{frame}{Hourly Loads}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/hourly-loads.png}}; \vspace{1cm}
+   \fitimage{../images/hourly_loads.png}
    \vfill
@@
 \begin{frame}{Peak and Average Loads}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/loads_clean.png}}; \vspace{1cm}
+   \fitimage{../images/loads_clean.png}
    \vfill
@@
 \begin{frame}{Hourly Prices}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/hourly_prices_col.png}}; \vspace{1cm}
+   \fitimage{../images/hourly_prices.png}
    \vfill
@@
 \begin{frame}{Prices over time}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/peak_prices.png}}; \vspace{1cm}
+  \fitimage{../images/peak_prices.png}
    \vfill
@@
 \begin{frame}{Forward Markets}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/forwards.png}}; \vspace{1cm}
+   \fitimage{../images/forwards.png}
    \vfill
@@
 \begin{frame}{The Balancing Pool: What on earth does/did it do?}
-\vspace{-.2cm}\centering
-\hspace{-1.5cm}\begin{tabular}{p{1.1\linewidth}}
-    \centering
-    \includegraphics[width=.9\linewidth]{../images/balancingpool.png} \\[\abovecaptionskip]
-  Source: Capital Power
-\end{tabular}
-
+\fitimage[Source: Capital Power]{../images/balancingpool.png}
 \vfill \end{frame}
@@
 \begin{frame}{Costs of New Capacity Additions}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.25cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/eia_lcoe.png}}; \vspace{1cm}
+   \fitimage{../images/lcoe_22.png}
    %\\[\abovecaptionskip]  \vspace{5.5cm}
@@
 \begin{frame}{Costs of New Capacity Additions}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.25cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/lazard_lcoe.png}}; \vspace{1cm}
-   %\\[\abovecaptionskip]  \vspace{5.5cm}
-   %\tiny{See for more information: \url{https://www.lazard.com/media/451086/lazards-levelized-cost-of-energy-version-130-vf.pdf}}
+   \fitimage{../images/lcoe_lazard_22.png}
    \vfill
@@
 \begin{frame}{New Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/ab_solar.png}}; \vspace{1cm}
+   \fitimage{../images/ab_solar.png}
+   \vfill
+\end{frame}
+
+\begin{frame}{New Technology}
+   \fitimage{../images/solar.png}
+   \vfill
+\end{frame}
+
+\begin{frame}{New Technology}
+   \fitimage{../images/solar_year.png}
+   \vfill
+\end{frame}
+
+\begin{frame}{New Technology}
+   \fitimage{../images/small_solar.png}
    \vfill
@@
 \begin{frame}{Price Capture by Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/price_capture.png}}; \vspace{1cm}
+\fitimage{../images/price_capture_renew.png}
    \vfill
@@
 \begin{frame}{Evolution of Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/solar_eia.png}}; \vspace{1cm}
+   \fitimage{../images/eia_solar.png}
    \vfill
@@
 \begin{frame}{Evolution of Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/solar_costs.png}}; \vspace{1cm}
+   \fitimage{../images/eia_coal.png}
    \vfill
-   \vspace{4cm}\tiny{Source: NREL}  \hfill
+\end{frame}
+
+\begin{frame}{Evolution of Technology}
+   \fitimage[Source: Lazard]{../images/lazard_decline.png}
+   \vfill
 \end{frame}
@@
 \begin{frame}{GHG Policy and Electricity Supply}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.95\paperwidth]{../images/monthly_ghgs.png}}; \vspace{1cm}
+   \fitimage{../images/monthly_ghgs.png}
    \vfill
@@
 \item Rivers and Dolter paper on decarbonizing Canada's supply
-\item van de Biezenbos paper on transmission.
+\item van de Biezenbos paper on transmission
+
+\item Clean Electricity Standard discussion paper
 

### 2022-10-05 rep_surplus.Rmd
fixing charts and deck

---
title: "AB REP Surplus from Hastings-Simon et al (2022)"
output:
  html_document:
    code_folding: hide
date: '2022-08-12'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,warning = FALSE,message=FALSE)
library(tidyverse)
library(lubridate)
library(viridis)
library(scales)

paper_theme<-function(){
  theme_minimal()+theme(
    text = element_text(size = rel(5),margin = margin(t = 2, b =2,l=2,r=2)),
    axis.title = element_text(size = rel(.8)),
    legend.position = "bottom",
    legend.text=element_text(size=rel(3)),
    strip.text=element_text(size=rel(3)),
    legend.margin=margin(c(0,0,-0.25,0),unit="cm"),
    legend.key.width = unit(rel(2),"line"),
    plot.margin=margin(c(.5,.5,0.5,0.5),unit="cm"),
    #plot.caption = element_text(size = 9, color = "gray40",hjust=1),
    #plot.title = element_text(face = "bold"),
    #plot.subtitle = element_text(size = 10, color = "gray40"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )
}

```

## REP CFD Valuation

The contracts-for-differences procured through the REP have been net positive for the government to date, driven by an increase in wholesale power prices in Alberta in 2021 and 2022 year-to-date. We estimate the value of the REP contract portfolio to date starting with the hourly surplus value generated on power sales relative to project-specific, inflation-adjusted contract strike prices, to which we add an estimate of the value of greenhouse gas emissions credits (EC) transferred to the government under the REP contracts at market carbon prices (τ_t). These are each multiplied by hourly production for each project from the first hour in which a facility was in commercial operation, through to the end-of-day July 31, 2022.

We make three important assumptions to value the portfolio of contracts.

First, we use the capacity-weighted average strike prices for each REP round as a proxy for project-specific strike prices, as project-specific information was not disclosed.

Second, ww assume that the lowest strike price in REP round 1 (CA\$30.90/MWh) was associated with the Sharp Hills project that terminated its support agreement and, based on this assumption, we adjust the capacity-weighted average strike price for first round contracts to CA\$41.36/MWh, which we apply to the remaining three round 1 projects.

Finally, we estimate the value of surrendered environmental attributes based on the annual value of emissions performance credits that would have otherwise been allocated to the projects at a rate of 0.37t/MWh under Alberta’s carbon pricing policy, the Technology and Innovation Emissions Reduction Regulation (2019), multiplied by annual carbon prices established in the regulation.

This document details how we set the preliminaries for our analysis.

```{r rep_prelims}
#CPI Adjustments for REP1 projects. REP3 still at 1, since the only project in-service came on-line this year.
cpi<-function(year,REP="REP1"){
      ret<-case_when(
        (as.character(year)==2019) ~ 0.97801,
        (as.character(year)==2020) ~ 1.00000, 
        (as.character(year)==2021) ~  1.00733,
        (as.character(year)==2022) &  (REP=="REP1") ~  1.05572,
        (as.character(year)=="2022 (YTD)") &  (REP=="REP1")~ 1.05572,
        TRUE ~ 1)
      ret
}

#define the REP projects
    REP_projects<-c("RIV1","CRR2","WHT1","WRW1")

```

This document uses publicly-avaialble AESO data. The paper uses the same data for this analysis, but for longer-term history we use AESO data as scraped and provided by NRGStream.

```{r aeso_data}
#use updated AESO MV data to calculate REP contract values

aeso_file<-read.csv("https://www.aeso.ca/assets/Uploads/data-requests/Hourly_Metered_Volumes_and_Pool_Price_and_AIL.csv")         
#make a local version if you want to run it faster or make changes
#save(aeso_file,file="aeso_file_local.RData")
#load("aeso_file_local.RData")

  aeso_rep<-aeso_file %>% 
      select(-Date_Begin_GMT)%>%
      rename("Market"=ACTUAL_AIL)%>%
      pivot_longer(-c(Date_Begin_Local,"ACTUAL_POOL_PRICE","DAY_AHEAD_POOL_PRICE",
                   "EXPORT_BC","EXPORT_MT","EXPORT_SK","IMPORT_BC",
                   "IMPORT_MT","IMPORT_SK"),
                   names_to = "ID", values_to = "gen")%>%
      filter(ID %in% c(REP_projects,"Market"))%>%
      mutate(time=mdy_hm(Date_Begin_Local))%>%
      select(-Date_Begin_Local)%>%
      filter(!is.na(gen))%>%
      select(time,price=ACTUAL_POOL_PRICE,ID,gen)%>%
      mutate(year=year(time),
             power_value=gen*price)%>%
      filter(as.numeric(as.character(year))>=2019)%>%
      #value carbon credits
      mutate(env_attr=case_when(
        (ID %in% REP_projects)& year==2019 ~30*.37,
        (ID %in% REP_projects)& year==2020 ~30*.37,
        (ID %in% REP_projects)& year==2021 ~40*.37,
        (ID %in% REP_projects)& year==2022 ~50*.37,
        TRUE~0
      ))%>%
      #assign rep series
      mutate(rep=case_when(
        (ID=="WHT1") ~ "REP1",
        (ID=="WRW1") ~ "REP3",
        (ID=="RIV1") ~ "REP1",
        (ID=="CRR2") ~ "REP1",
        TRUE ~ "Not in REP"))%>%
      #assign REP strike prices
      mutate(raw_strike=case_when(
        (rep=="REP1") ~ 41.36,
        (rep=="REP2") ~ 38.69,
        (rep=="REP3") ~ 40.14,
        TRUE ~ 0))%>%
      #adjust strikes for inflation, then calculate net positions
      group_by(year)%>%
      #calculate generation by year for the graph labels
      mutate(
       rep_gen_year=sum(gen*(ID %in% REP_projects))
       )%>%
      ungroup()%>%
      mutate(
        strike=raw_strike*(.8+.2*cpi(year,rep)),
        #value the cfd
        cfd_val=strike*gen,
        #hourly value on power
        net_govt_power=(price-strike)*gen,
        #hourly environmental attr value
        env_ttl=env_attr*gen,
        #hourly net
        net_govt_ttl=net_govt_power+env_ttl,
        #set up the years so we can label the YTD
        year=as.factor(year),
        year_mod=fct_recode(year,"2022 (YTD)"="2022"),
        label=paste(as.character(year_mod),"\n\n","(",round(rep_gen_year/10^3,0)," GWh of REP-\nsupported generation)",sep="")
     )%>%
      #create yearly summaries by ID. Preserve the labels
      group_by(year,ID,label)%>%
      summarize(cfd_surplus=sum(net_govt_power),
                env_surplus=sum(env_ttl),
                #net_govt_ttl=net_govt_power+env_ttl,        
                env_ttl=sum(env_ttl,na.rm = T)/sum(gen,na.rm = T),
                power_val=sum(price*gen)/sum(gen),
                total_val=power_val+env_ttl
                )%>%
      ungroup()%>%
      #get the order by COD, with market on the left
      mutate(
        ID=fct_relevel(ID,"WHT1","CRR2","RIV1","WRW1"),
        ID=fct_relevel(ID,"Market"),
        ID=fct_recode(ID,"Market Generation-Weighted Average Price"="Market"),
        ID=fct_recode(ID,"Whitla 1 (REP1)"="WHT1","Castle Rock Ridge 2 (REP1)"="CRR2","Riverview Wind (REP1)"="RIV1","Windrise Wind (REP3)"="WRW1"),
        #Plant_Type=fct_reorder(Plant_Type,startup,min),
      )
    ```

And, so using these data, we can summarize the annual surplus (net of and including the value of environmental attributes surrendered by the project proponents):

```{r aeso_graph,fig.width=16,fig.height=8}
    #create the figure
    aeso_rep_plant_cpi_graph<-ggplot(aeso_rep)+
      geom_col(aes(ID,power_val,fill=ID),size=.25,position=position_dodge(width=0.9),color="black",alpha=0.5)+
      geom_col(aes(ID,y=total_val,color="Deemed value of foregone emissions credits (0.37t/MWh, valued at Alberta TIER compliance prices)"),fill=NA,alpha=0.5,size=.5,position=position_dodge(width=0.9))+
      facet_wrap(~label,nrow=1)+
      #scale_fill_grey("")+
      scale_fill_viridis("",option = "B",discrete = T,direction = 1)+
      scale_color_manual("",values=c("black"))+
      scale_linetype_manual("",values=c("solid","21","22"))+
      scale_y_continuous(breaks = pretty_breaks(),expand=c(0,0))+
      expand_limits(y=0)+
      paper_theme()+
      theme(legend.position = "bottom",legend.box="vertical", legend.margin=margin(),
            legend.text=element_text(size=rel(2)),
            axis.text.x = element_blank(),
            NULL)+
      expand_limits(y=120)+
      guides(#fill="none",
        fill=guide_legend(order = 1,nrow = 1,label.position = "right",keywidth = 1,keyheight = 1),
        linetype=guide_legend(order = 3,nrow = 1,keywidth = 3),
        color=guide_legend(order = 2,ncol = 1,label.position = "right",keywidth = 1,keyheight = 1))+
      labs(x="",y="Value of generated electricity and environmental attributes ($/MWh)",
           #title="Energy Price Capture ($/MWh, 2010-2021)",
           #caption="Source: AESO Data, accessed via NRGStream\nGraph by @andrew_leach"
      )

    aeso_rep_plant_cpi_graph+
      geom_col(data=aeso_rep%>% filter(year!="2022"),aes(ID,power_val,fill=ID),size=.25,position=position_dodge(width=0.9),
               color="black",alpha=.95)+
      geom_col(data=aeso_rep%>% filter(year!="2022"),aes(ID,total_val,
               color="Deemed value of foregone emissions credits (0.37t/MWh, valued at Alberta TIER compliance prices)"),
               fill=NA,alpha=0.25,size=.5,position=position_dodge(width=0.9))+
      geom_hline(aes(yintercept=37*(.8+.2*cpi(year)),lty="REP1 Average CfD Strike Price"),color="black",size=1.5)+
      #omit REP 2 since no projects included
      #geom_hline(aes(yintercept=38.69*(.8+.2*cpi(year)),lty="REP2 Average CfD Strike Price"),color="black",size=1.5)+
      geom_hline(aes(yintercept=40.14*(.8+.2*cpi(year)),lty="REP3 Average CfD Strike Price"),color="black",size=1.5)+
      NULL
```

And, these data also yield the cumulative totals cited in the paper:

* Total CfD portfolio surplus: `r as.numeric(aeso_rep%>%filter(ID != "Market Generation-Weighted Average Price")%>% summarize(power_surplus=sum(cfd_surplus,na.rm=T))/10^6)` million

* Total value of GHG emissions credits: `r as.numeric(aeso_rep%>%filter(ID != "Market Generation-Weighted Average Price")%>% summarize(power_surplus=sum(env_surplus,na.rm=T))/10^6)` million

* Total combined surplus: `r as.numeric(aeso_rep%>%filter(ID != "Market Generation-Weighted Average Price")%>% summarize(power_surplus=sum(cfd_surplus+env_surplus,na.rm=T))/10^6)` million

### 2022-10-05 aeso_data_proc.R

@@ -16,7 +16,7 @@ if(!exists("getSeason", mode="function")) source("get_season.R")
 
 #update prices and loads
 
-update<-0
+update<-1
 if(update==1)
   update_forecasts()
 #re-load
@@ -29,6 +29,8 @@ forecast_data<-assign_peaks(forecast_data)
 
 load(file="data/metered_vols_data.Rdata" ) 
 if(update==1){
+  all_vols<-all_vols %>% filter(date<=ymd("2021-12-31"))
+  #test<-update_vols(all_vols)
   all_vols<-update_vols(all_vols)
   save(all_vols, file="data/metered_vols_data.Rdata" ) 
   }

### 2022-10-05 aeso_scrapes.R

@@ -117,6 +117,7 @@ update_vols <- function(data_sent) {
   }
   
   #test<-process_data(data_store)
+  #test2<-data_sent %>% bind_rows(test)
   #return the data
   rbind(data_sent,process_data(data_store))
 }
@@ -181,7 +182,10 @@ process_data <- function(data_sent) {
     mutate(asset_id = ifelse(grepl("DOW1",asset_id),"DOWG",asset_id))%>% #dow
     mutate(asset_id = ifelse(grepl("GEN1",asset_id),"GEN6",asset_id))%>% #gen1 in asset id is judy creek (gen6)
     mutate(asset_id = ifelse(grepl("GEN3",asset_id),"WCD1",asset_id))%>% #gen3 is cadotte
-    mutate(asset_id = ifelse(grepl("SCTG",asset_id),"APS1",asset_id)) #SCTG is sctoford, but it's shell-offered power
+    mutate(asset_id = ifelse(grepl("SCTG",asset_id),"APS1",asset_id))%>% #SCTG is sctoford, but it's shell-offered power
+    {.}
+    
+    
   #merge in asset names and pool participant ID combos
   clean2<-clean2 %>% left_join(select(aeso_assets,asset_name,asset_id,pool_participant_id),by=c("asset_id","pool_participant_id"))%>%
     #now we need to fix some of the joint marketing of power from assets
@@ -214,6 +218,7 @@ process_data <- function(data_sent) {
   combined<-merge(ghg_rates,heat_rates,by="GHG_ID",all.y = T) # NA's match
   coal_co2_btu<-100.4  #coal fuel factor GHGs/MMBTU
   gas_co2_btu<-53.077752 #gas fuel factor GHGs/MMBTU
+  
   combined<-merge(plant_info,combined,suffixes = c(".info",".rates"),all.x=TRUE,by="Aurora_ID") # NA's match
   combined$co2_est<-combined$CO2/2.20462*combined$Heat.Rate #convert to kg/mmbtu
   combined$co2_est<-ifelse(combined$Plant_Fuel=="COAL",coal_co2_btu*combined$Heat.Rate,combined$co2_est)

### 2022-10-05 ep_paper_graphs.R

@@ -9,13 +9,69 @@ library(cowplot)
 load(file="data/metered_vols_data.Rdata")
 load(file="nrgstream/nrgstream_gen.Rdata") 
 
+
+
+
+
 update_forecasts()
 load(file="data/forecast_data.Rdata") 
 forecast_data <- forecast_data %>% filter (he!="02*")
 
+hourly_patterns<-forecast_data %>% 
+  filter(!is.na(actual_posted_pool_price),!is.na(actual_ail))%>% 
+  assign_date_time_days()%>%
+  group_by(hour,year)%>%
+  summarize(ail=mean(actual_ail),price=mean(actual_posted_pool_price))%>%
+  mutate(hour=paste(hour,":00",sep=""),hour=factor(hour))
+
+library(gghighlight)
+
+hourly_graph<-  ggplot(hourly_patterns%>%filter(year>=2004)) +
+  geom_line(aes(factor(hour),price,group=factor(year),color=factor(year)),size=.85)+
+  geom_line(data=hourly_patterns %>% filter(year==2022),aes(factor(hour),price,group=factor(year),color=factor(year)),size=1.5)+
+  scale_color_viridis("",option = "C",discrete = T,direction=-1)+
+  scale_y_continuous(expand=c(0,0))+
+  paper_theme()+
+  scale_x_discrete(breaks=c("1:00","4:00","7:00","10:00",
+                            "13:00","16:00","19:00","22:00"))+
+  expand_limits(y=0)+ #make sure you get the zero line
+  #guides(color = guide_legend())+
+  theme(#legend.position="bottom",
+        legend.margin=margin(c(0,0,0,0),unit="cm")
+        #legend.text = element_text(colour="black", size = 12, face = "bold")
+        #axis.title.y = element_text(margin = margin(t = 0, r = 5, b = 0, l = 0, unit = "pt")
+  )+
+  labs(y="Pool Prices ($/MWh)",x="Hour ending",
+       #title=paste("Alberta Hourly Wholesale Power Prices and Alberta Internal Load",sep=""),
+       #subtitle=paste(month.name[month(period_start)],", ",year(period_start)," to ",month.name[month(end_date)],", ",year(end_date),sep="")
+       NULL)
+hourly_graph 
[…]

### 2022-10-05 nrgstream_gen.R

@@ -253,46 +253,3 @@ get_plant_info()
 #nrgstream_gen<-all_gen_data()
 #rm(gen_data)
 #if we're changing over to a new year, run both of these two lines of code:
-nrgstream_gen<-filter(nrgstream_gen,year(Time)<=2021)
-#nrgstream_gen<-data_update(nrgstream_gen,2021)
-nrgstream_gen<-data_update(nrgstream_gen,2022)
-
-
-#Repair coal-to-gas-conversions
-nrgstream_gen<-nrgstream_gen %>% 
-  mutate(Plant_Type=case_when( 
-    (ID=="HRM") & (date>=ymd("2020-05-08")) ~ "SCGT",  #Milner change to gas effective 
-    #SH1 Sheerness #1 and SH2 Sheerness #2 -July 30, 2021.https://www.aeso.ca/market/market-updates/2021/sh1-sheerness-1-and-sh2-sheerness-2-change-in-fuel-type-notice/
-    (ID %in% c("SH1","SH2")) & (date>=ymd("2021-07-30")) ~ "NGCONV",
-    #KH3 January 11, 2022
-    (ID =="KH1") & (date>=ymd("2022-01-11")) ~ "NGCONV",
-    #KH2 July 27, 2021.
-    (ID =="KH2") & (date>=ymd("2021-07-21")) ~ "NGCONV",
-    #Battle River #4 (BR4)	March 8, 2022
-    (ID =="BR4") & (date>=ymd("2022-03-08")) ~ "NGCONV",
-    #Battle River #5 (BR5)	 November 19, 2021
-    (ID =="BR5") & (date>=ymd("2021-11-19")) ~ "NGCONV",
-    #Sundance #6 (SD6)	401	0	0 February 19, 2021
-    (ID =="SD6") & (date>=ymd("2021-02-19")) ~ "NGCONV",
-    TRUE ~ Plant_Type),
-    Capacity=case_when(
-      (ID=="HRM") & (date>=ymd("2020-04-23"))&(date<ymd("2020-05-08")) ~ 185,  #Milner change to gas effective 
-      (ID=="HRM") & (date>=ymd("2020-05-08"))&(date<ymd("2021-12-09")) ~ 208,  #Milner change to gas effective 
-      (ID=="HRM") & (date>=ymd("2021-12-09")) ~ 300,  #Milner change to gas effective 
-      TRUE~Capacity
-    )
-    
-    
-  )
-
-
[…]

### 2022-10-05 nrgstream_graphs.R

@@ -1018,7 +1018,8 @@ df1 <- daily_forwards %>% group_by(Trade_Date,Inst_Date) %>% summarise(Settle = 
 #df1=melt(df1,id=c("Trade_Date","Inst_Year"),variable.name = "term",value.name = "settle")
 today_date<-max(daily_forwards$Trade_Date)
 dates<-c(today_date,today_date-weeks(1),today_date-weeks(2),today_date-weeks(3),today_date-weeks(4))
-df1 <- df1 %>% filter(Inst_Date<=min(dates)+)
+
+df1 <- df1 %>% filter(Inst_Date<=min(dates))
 
 png<-1
 if(png==1)
@@ -1825,3 +1826,139 @@ for(i in c(3:170))
         caption="Source: AESO 2017 Long Term Outlook\nGraph by @andrew_leach")
  dev.off()
  
+
+ 
+ monthly_solar<-nrgstream_gen %>% 
+   filter(Plant_Type=="SOLAR",!is.na(gen),he=="13")%>%
+   mutate(month=month(Time),year=year(Time))%>%
+   group_by(month,year,ID,AESO_Name)%>%summarize(gen=mean(gen,na.rm = T))%>%
+   mutate(date=ymd(paste(year,"-",month,"-1",sep="")))%>%
+   ungroup()
+ 
+ 
+monthly_solar %>%
+  ggplot()+
+  geom_col(aes(date,gen,group=ID,fill=ID),position = "stack")+
+  scale_x_date(expand=c(0,0),breaks="3 months",labels = date_format("%b\n%Y",tz="America/Denver"))+
+  scale_y_continuous(expand=c(0,0))+
+  expand_limits(x=c(ymd("2017-12-31"),Sys.Date()+months(2)),y=800)+
+  scale_fill_viridis("",option="C",discrete = T)+
+  guides(fill=guide_legend(ncol = 1))+
+   theme(panel.border = element_blank(),
+        panel.grid = element_blank(),
+        panel.grid.major.y = element_line(color = "gray",linetype="dotted"),
+        axis.line.x = element_line(color = "gray"),
+        axis.line.y = element_line(color = "gray"),
+        axis.text = element_text(size = 16),
+        axis.text.x = element_text(margin = margin(t=0),hjust = .50,vjust = .50),
+        axis.title = element_text(size = 16),
+        #axis.label.x = element_text(size=20,vjust=+5),
[…]

### 2022-10-05 power_paper_base.R

@@ -266,7 +266,7 @@ assign_peaks<-function(data_orig,time_var=time){
     #Victoria day - monday before May 25, so earliest it can be is day 18, latest is day 24
     Victoria_Day=ifelse(month(!!temp_time)==5 & day(!!temp_time)<=24 & day(!!temp_time)>=18 & wday(!!temp_time,label = T)=="Mon",T,F) #Monday before May 25
   ) %>% mutate(
-    stat = select(., holiday_list) %>% rowSums()>0
+    stat = select(., all_of(holiday_list)) %>% rowSums()>0
     #stat = select(all_of(holiday_list)) %>% rowSums()>0
   )
   

### 2023-05-11 presentations/power_slides.tex
updates

@@
         %or: (current page.center)
-        {\includegraphics[width=1in]{UA-ASB-COLOUR.png}};
+        {\includegraphics[width=1in]{UA-COLOUR.png}};
 %\includegraphics[height=0.8cm]{UA-ASB-COLOUR.png}\vspace{220pt}
@@
 {
-  Alberta School of Business, University of Alberta
+  Department of Economics and Faculty of Law, University of Alberta
  }
@@
    \tikz [remember picture,overlay]
-    \node[yshift=-0.95cm,xshift=0cm] at (current page.north)
+    \node[yshift=-0.5cm,xshift=-1.7cm] at (current page.north east)
         %or: (current page.center)
@@
         %{\includegraphics[width=3in]{UA-ASB-COLOUR.png}};
-        {\includegraphics[width=\paperwidth]{EE_logo.png}}; \vspace{1cm}
+        {\includegraphics[width=1in]{UA-COLOUR.png}}; \vspace{1cm}
    \titlepage
@@
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/peak_prices_2000_2019.png}}; \vspace{1cm}
+        {\includegraphics[width=.9\paperwidth]{../images/prices_and_loads.png}}; \vspace{1cm}
    \vfill
@@
 
-\begin{frame}{New Technology}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-        {\includegraphics[width=.9\paperwidth]{../images/brooks_solar.png}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
 \begin{frame}{Economics 101}
@@
     \centering
-    \includegraphics[width=.9\linewidth]{grid_ess.png} \\[\abovecaptionskip]
+    \includegraphics[width=.9\linewidth]{../images/grid_ess.png} \\[\abovecaptionskip]
   Source: \url{http://www.mpoweruk.com/grid_storage.htm}

### 2023-05-11 aeso_lto.R

@@ -22,18 +22,18 @@ df1<-lto_data %>% pivot_longer(-date, names_to = "forecast",values_to = "peak") 
 #set_png(file="images/AESO_LTO.png", width = 1400, height = 750)
 #jpeg(file="AESO_LTO.jpg", width = 1400, height = 750)
 ggplot(df1) +
-  geom_line(data=filter(df1,grepl("Reference",forecast)|grepl("Forecast",forecast))%>% filter(year!=2021),aes(date,peak,group = forecast,color="Previous AESO Reference Case Forecasts"),size=.7) +
-  geom_line(data=filter(df1,forecast=="Actual"),aes(date,peak,group = forecast,colour="Historic Peak Loads"),size=3) +
-  geom_line(data=df1%>%filter(year==2021)%>%mutate(forecast=fct_relevel(forecast,"2021 Reference Case",after = Inf)),aes(date,peak,group = forecast,colour=forecast),size=1.7) +
-  #geom_point(size=1) +
-  #scale_color_viridis(labels=c("2004 Forecast","2006 Forecast", "2007 Forecast" ,"2008 Forecast","2009 Forecast","2012 Forecast" , "2014 Reference Case", "2016 Reference Case","Actual AIL"),discrete=TRUE)+   
-  #scale_colour_brewer(labels=c("2004 Forecast","2006 Forecast", "2007 Forecast" ,"2008 Forecast","2009 Forecast","2012 Forecast" , "2014 Reference Case", "2016 Reference Case","Actual AIL"),type = "seq", palette = "Greens", direction = 1)+
-  #scale_colour_brewer(type = "seq", palette = "Greens", direction = 1)+
-  scale_colour_manual(NULL,#labels=c("2004 Forecast","2005 Forecast","2006 Forecast", "2007 Forecast" ,"2008 Forecast","2009 Forecast","2012 Forecast" , "2014 Reference Case", "2016 Reference Case","2017 Reference Case","2019 Reference Case","Actual AIL"),
-                      values=c(brewer.pal(5,"Blues")[2:5],"Black","grey60"))+
-  #scale_linetype_manual(NULL,values=c("11","solid","22","33"))+
-  
+  geom_line(data=filter(df1,grepl("Reference",forecast)|grepl("Forecast",forecast))%>% filter(year!=2021),
+            aes(date,peak,group = forecast,color="Previous AESO Reference Case Forecasts",linetype="Previous AESO Reference Case Forecasts"),size=.7) +
[…]

### 2023-05-11 ep_paper_graphs.R

@@ -8,7 +8,9 @@ library(cowplot)
 
 load(file="data/metered_vols_data.Rdata")
 load(file="nrgstream/nrgstream_gen.Rdata") 
-
+nrgstream_gen <- nrgstream_gen %>% 
+  #ng_conv(id_sent=ID)%>%
+  rename(time=Time)
 
 
 
@@ -26,7 +28,7 @@ hourly_patterns<-forecast_data %>%
 
 library(gghighlight)
 
-hourly_graph<-  ggplot(hourly_patterns%>%filter(year>=2004)) +
+hourly_graph<-  ggplot(hourly_patterns%>%filter(year>=2004,year<2023)) +
   geom_line(aes(factor(hour),price,group=factor(year),color=factor(year)),size=.85)+
   geom_line(data=hourly_patterns %>% filter(year==2022),aes(factor(hour),price,group=factor(year),color=factor(year)),size=1.5)+
   scale_color_viridis("",option = "C",discrete = T,direction=-1)+
@@ -50,13 +52,13 @@ ggsave(file=paste("images/hourly_prices.png",sep=""),width = 14,,height=7,dpi=30
 
 hourly_graph<-  ggplot(hourly_patterns%>%filter(year>=2004)) +
   geom_line(aes(factor(hour),ail,group=factor(year),color=factor(year)),size=.85)+
-  geom_line(data=hourly_patterns %>% filter(year==2022),aes(factor(hour),ail,group=factor(year),color=factor(year)),size=1.5)+
+  geom_line(data=hourly_patterns %>% filter(year==2023),aes(factor(hour),ail,group=factor(year),color=factor(year)),size=2.5)+
   scale_color_viridis("",option = "C",discrete = T,direction=-1)+
   scale_y_continuous(expand=c(0,0))+
   paper_theme()+
   scale_x_discrete(breaks=c("1:00","4:00","7:00","10:00",
                             "13:00","16:00","19:00","22:00"))+
-  expand_limits(y=c(6000,10500))+ #make sure you get the zero line
+  expand_limits(y=c(6000,11200))+ #make sure you get the zero line
   #guides(color = guide_legend())+
   theme(#legend.position="bottom",
     legend.margin=margin(c(0,0,0,0),unit="cm")
@@ -228,7 +230,7 @@ ggsave("images/prices_and_loads_ep.png",width = 14,height=8,bg="white",dpi = 300
 
 #align variable names
[…]

### 2023-05-11 merit_build.R

@@ -30,7 +30,8 @@ if(update==1){
 #merit_small<-merit_data%>%filter(date==ymd("2019-10-05"))
 #merit_data<-merit_small
 
-
+#fix the ANC1 issue if it exists
+merit_data<-merit_data %>% mutate(asset_id=gsub("ANCI","ANC1",asset_id))
 
 
 
@@ -156,8 +157,6 @@ load("data/forecast_data.RData")
 #merit_aug<-merit_small
 
   
-plant<-plant_data()
-  
 merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
   mutate(year=year(date))%>%
   left_join(ghg_data(),by=c("asset_id"="ID","year"))%>%
@@ -170,6 +169,7 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
                            year(date)==2020 ~ "TIER_30",
                            year(date)==2021 ~ "TIER_40",
                            year(date)==2022 ~ "TIER_50",
+                           year(date)==2023 ~ "TIER_65",
                            TRUE ~ "TIER_50")
     )%>%
   select(-year)# take out the year value since it will create duplication with older code snippets later for full data set
@@ -177,7 +177,7 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
 
 #testing storage objects
 #merit_store<-merit_aug  
-#merit_small<-merit_aug%>%filter(year(date)>=2020)
+#merit_small<-merit_aug%>%filter(year(date)==2020)
 #merit_aug<-merit_small
 
 #merit_aug<-merit_store  
@@ -186,43 +186,44 @@ merit_aug<-merit_aug %>% left_join(plant_data(),by=c("asset_id"="ID"))%>%
 
 
 
-#Repair coal-to-gas-conversions
+#Repair coal-to-gas-conversions with script in power_paper_base
 merit_aug<-merit_aug %>% 
-  mutate(Plant_Type=case_when( 
-    (asset_id=="HRM") & (date>=ymd("2020-05-08")) ~ "SCGT",  #Milner change to gas effective 
[…]

### 2023-05-11 merit_scripts.R

@@ -197,7 +197,7 @@ plant_data<-function(){
 
 clean_merit_trade<-function(data_sent,id_tag){
   #for testing
-  #data_sent<-head(merit_data,1000) #%>% assign_time %>% assign_date_time_days %>% assign_peaks
+  #data_sent<-merit_data) #%>% assign_time %>% assign_date_time_days %>% assign_peaks
   #data_sent<-merit_day #%>% assign_time %>% assign_date_time_days %>% assign_peaks
   
   #read in all AESO asset names so we're up-to-date
@@ -233,8 +233,8 @@ clean_merit_trade<-function(data_sent,id_tag){
     mutate(key_firm=TRUE,offer_control="TRADE",offer_sum="TRADE",merit=0,
            from=0,to=available_mw,block_number=0,) %>% 
     ungroup()%>%
-    mutate(dest=ifelse(import_export=="I",paste(dest,"_IMP",sep = ""),paste(dest,"_EXP",sep = "")))
-  names(trade)[names(trade) == "dest"] <- "asset_id"
+    mutate(dest=ifelse(import_export=="I",paste(dest,"_IMP",sep = ""),paste(dest,"_EXP",sep = "")))%>%
+    rename(asset_id=dest)
   clean<-data_sent%>%filter(!import_export %in% c("E","I")) %>% #select the non-imports and non-exports from the merit data
     rbind(trade)
   clean

### 2023-05-11 nrgstream_gen.R

@@ -148,7 +148,7 @@ data_update <- function(data_sent,fix_year=0) {
   #build current year data and append to gen.Rdata
   #testing
   #data_sent<-filter(gen_data,year(Time)==2018)
-  #fix_year<-2019
+  #fix_year<-2023
   
   get_plant_info()
   stack<-NULL
@@ -285,11 +285,16 @@ load(file="nrgstream/nrgstream_gen.RData")
 get_plant_info()
 #use this to update
 
-nrgstream_gen<-filter(nrgstream_gen,year(Time)<=2020)
-nrgstream_gen<-data_update(nrgstream_gen,2021)
+nrgstream_gen<-filter(nrgstream_gen,year(Time)<=2021)
+#nrgstream_gen<-data_update(nrgstream_gen,2021)
 nrgstream_gen<-data_update(nrgstream_gen,2022)
+nrgstream_gen<-data_update(nrgstream_gen,2023)
 
 nrgstream_gen<-ng_conv(nrgstream_gen)
+
+
+#test_data<-nrgstream_gen %>% filter(is.na(Plant_Type))
+
 
 save(nrgstream_gen, file= "nrgstream/nrgstream_gen.RData")
 

### 2023-05-11 nrgstream_graphs.R

@@ -50,6 +50,13 @@ ctax_year<-function(year_sent){
     year_sent <= 2015 ~ 15,
     year_sent == 2016 ~ 20,
     year_sent == 2017 ~ 30,
+    year_sent == 2018 ~ 30,
+    year_sent == 2019 ~ 30,
+    year_sent == 2020 ~ 30,
+    year_sent == 2021 ~ 40,
+    year_sent == 2022 ~ 50,
+    year_sent == 2023 ~ 65,
+    
     TRUE                      ~  30 
   )
 }
@@ -87,7 +94,19 @@ nrgstream_gen<-nrgstream_gen[!is.na(nrgstream_gen$time),]
 #take out AB-BC and AB-MON
 #nrgstream_gen<-filter(nrgstream_gen, AESO_Name !="AB - BC Hydro  Imp/Exp Hr Avg MW",AESO_Name !="AB - Montana Imp/Exp Hr Avg MW")
 
-
+# capacities
+# nrgstream_gen %>% filter(year(time)==2022,Plant_Fuel%in%c("SOLAR","WIND"))%>% 
+#   filter(time>=ymd("2022-12-31"))%>%
+#   group_by(time,Plant_Fuel)%>%
+#   summarize(Capacity=sum(Capacity))%>%
+#   group_by(Plant_Fuel)%>%
+#   summarize(Capacity=max(Capacity))
+# 
+# 
+# caps<-nrgstream_gen %>% filter(year(time)==2022,Plant_Fuel%in%c("WIND"))%>% 
+#   filter(time>=ymd("2022-12-31"))%>%
+#   group_by(ID)%>%
+#   summarize(Capacity=max(Capacity))
 
 sub_samp<-subset(nrgstream_gen, time > as.Date("2010-01-1"))
 #sub_samp<-subset(sub_samp, time < as.Date("2017-12-31"))
@@ -126,7 +145,8 @@ sub_samp<-filter(nrgstream_gen, time >= as.Date("2010-01-1"))
 
 df_test<-filter(sub_samp,is.na(gen))
 
-df1 <- sub_samp %>% filter(! NRG_Stream %in% trade_excl)%>% group_by(Plant_Type,time) %>% summarise(meancap = mean(Cap_Fac),total_gen=sum(gen,na.rm = T),total_rev=sum(Revenue,na.rm = T),p_mean=mean(Price)) %>% ungroup()
[…]

### 2023-05-11 paper_graphs.R

@@ -9,12 +9,60 @@ forecast_data <- forecast_data %>% filter (he!="02*")
 
 
 
+# load profile
+
+peak_data<-forecast_data %>%filter(!is.na(actual_posted_pool_price),!is.na(actual_ail),year(time)==2022)%>% 
+  assign_date_time_days()%>%
+  assign_peaks()%>%
+  group_by(he) %>% 
+  summarize(mean_ail=mean(actual_ail,na.rm = T),med_ail=median(actual_ail,na.rm=T),trough_ail=min(actual_ail),max_ail=max(actual_ail),
+            q95_ail=quantile(actual_ail, probs=c(.995)),
+            q05_ail=quantile(actual_ail, probs=c(.005))
+  )%>%ungroup()%>%
+  mutate(he=as.numeric(he)) 
+
+
+bottom_panel<-
+  ggplot(peak_data) +
+  #geom_line(aes(date,ail,colour="basic"),size=1.5)+
+  geom_rect(aes(xmin=1, xmax=24, ymin=0, ymax=min(peak_data$trough_ail)), fill="grey80",color="black", alpha=0.1) +
+  geom_rect(aes(xmin=1, xmax=24, ymax=max(peak_data$max_ail), ymin=min(peak_data$trough_ail)), fill="grey90",color="black", alpha=.1) +
+    geom_ribbon(aes(he,ymax=q95_ail,ymin=q05_ail,fill="Two-tailed 99th percentile range"),alpha=1)+
+  geom_line(aes(he,mean_ail,linetype="A"),size=.85,color="black")+
+  geom_line(aes(he,med_ail,linetype="B"),size=.85,color="blue")+
+  #geom_ribbon(aes(date,ymax=peak_ail,ymin=trough_ail,fill="Range of Monthly Values"),alpha=.5)+
+  
+  #scale_color_manual("",values = c("black"),labels=c("Average Monthly Price"))+
+  scale_fill_manual("",values = c("grey50"))+
+  scale_linetype_manual("",values = c("solid","11"),labels=c("Average load","Median Load"))+
+  scale_x_continuous(expand=c(0,0),breaks = pretty_breaks(6))+
+  scale_y_continuous(expand=c(0,0),breaks = pretty_breaks())+
+  expand_limits(y=c(0,12000))+
+  guides(linetype = guide_legend(override.aes = list(color = c("black","blue"))),color="none")+
+  paper_theme()+
+  annotate("text",x=11,y=4000,label="'Base Load'",
[…]

### 2023-05-11 power_paper_base.R

@@ -371,3 +371,48 @@ colors_tableau10_medium <- function()
            "#ED97CA", "#A2A2A2", "#CDCC5D", "#6DCCDA"))
 }
 
+
+ng_conv<-function(data_sent,id_sent="ID",time_var="Time"){
+  #index_sym<- rlang::ensym(index_id)
+  #data_sent %>% 
+  #  #group_by(across(all_of(index_id)))%>%
+  #  mutate(
+  #    raw=ifelse(!!index_sym=="a",0,raw),
+  #    raw_2=case_when(
+  #      (!!index_sym=="a")~0,
+  #      TRUE~raw)
+  #  )
+#}
+  
+  ID<-rlang::ensym(id_sent)
+  Time<-rlang::ensym(time_var)
+  #Repair coal-to-gas-conversions
+  #data_sent<-merit_small
+  data_sent %>% 
+    mutate(Plant_Type=case_when( 
+      (!!ID=="HRM") & (!!Time>=ymd("2020-05-08")) ~ "SCGT",  #Milner change to gas effective 
+      #SH1 Sheerness #1 and SH2 Sheerness #2 -July 30, 2021.https://www.aeso.ca/market/market-upTimes/2021/sh1-sheerness-1-and-sh2-sheerness-2-change-in-fuel-type-notice/
+      (!!ID %in% c("SH1","SH2")) & (Time>=ymd("2021-07-30")) ~ "NGCONV",
+      #KH3 January 11, 2022
+      (!!ID =="KH1") & (!!Time>=ymd("2022-01-11")) ~ "NGCONV",
+      #KH2 July 27, 2021.
+      (!!ID =="KH2") & (!!Time>=ymd("2021-07-21")) ~ "NGCONV",
+      (!!ID =="KH3") & (!!Time>=ymd("2022-01-11")) ~ "NGCONV",
+      #Battle River #4 (BR4)	March 8, 2022
+      (!!ID =="BR4") & (!!Time>=ymd("2022-03-08")) ~ "NGCONV",
+      #Battle River #5 (BR5)	 November 19, 2021
+      (!!ID =="BR5") & (!!Time>=ymd("2021-11-19")) ~ "NGCONV",
+      #Sundance #6 (SD6)	401	0	0 February 19, 2021
+      (!!ID =="SD6") & (!!Time>=ymd("2021-02-19")) ~ "NGCONV",
[…]

### 2024-03-13 presentations/cea_2022_deck.tex
cea
update

--- 2021-06-05 presentations/aere_deck.tex
@@
 
-\usepackage{graphicx,hyperref,fes,url}
+\usepackage{graphicx,hyperref,url,fes}
 \usepackage[english]{babel}
@@
 
-% Add a date and possibly the name of the event to the slides
-%  - again first a short version to be shown at the bottom of each slide
-%  - second the full date and event name for the title slide
-
-\renewcommand{\(}{\begin{columns}}
-\renewcommand{\)}{\end{columns}}
-\newcommand{\<}[1]{\begin{column}{#1}}
-\renewcommand{\>}{\end{column}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% The main document
 
@@
 \newcommand{\Real}{\mathbb R}
-
-% The main document
 
@@
  }
- {Leach and Shaffer (2019)}
+ {Leach and Shaffer (2022)}
 }
@@
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-       {\includegraphics[width=.9\paperwidth]{../../alberta_power/gen_ghg_price.png}}; \vspace{1cm}
+       {\includegraphics[width=.9\paperwidth]{../../alberta_power/images/gen_ghg_price.png}}; \vspace{1cm}
    \vfill
@@
     \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-       {\includegraphics[width=.9\paperwidth]{../../alberta_power/monthly_ghg_mwh.png}}; \vspace{1cm}
+       {\includegraphics[width=.9\paperwidth]{../../alberta_power/images/monthly_ghg_mwh.png}}; \vspace{1cm}
    \vfill
@@
 
-\begin{frame}{Secanell Group}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-      {\includegraphics[height=3in]{../../secanell_1}}; \vspace{1cm}
+\section{Other FES Theme 06 Research}
+\begin{frame} Secanell group
+\begin{block}{}
+\footnotesize{\begin{itemize}
+  \item Design, procurement and installation of low and high temperature electrolysis fabrication and testing facilities
+  \item Fabrication of single-cell low temperature electrolyzer prototypes using iridium and iridium oxide catalysts
+  \item Demonstrated, for the first time, the use of inkjet printing as a feasible technology to fabricate low temperature electrolyzer cells  (https://doi.org/10.1149/2.1101807jes)
+   \item Development of a novel nano-pompon-like iridium catalyst superstructure for low temperature fuel cell electrolyzers (https://doi.org/10.1016/j.jcat.2019.01.018)
+   \item Development of a novel technique for characterization of iridium catalyst degradation (https://doi.org/10.1016/j.jcat.2019.01.018)
+ \item Development of novel micro-structures for high temperature solid oxide electrolyzers and fuel cells (https://doi.org/10.1016/j.electacta.2018.02.055)
+  \end{itemize}}
+The project currently involves 5 PhD students and 1 MSc, 5 faculty (1 in MECE, 3 in CHEME and 1 in Chem). We are also collaborating with Ionomr, a firm in BC (https://ionomr.com/).
+  \end{block}
    \vfill
@@
   \end{block}
-   \vfill
-\end{frame}
-
-\begin{frame}{Secanell Group}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-      {\includegraphics[height=3in]{../../secanell_2}}; \vspace{1cm}
-   \vfill
-\end{frame}
-
-\section{Other FES Theme 06 Research}
-\begin{frame} Secanell group
-\begin{block}{}
-\footnotesize{\begin{itemize}
-  \item Design, procurement and installation of low and high temperature electrolysis fabrication and testing facilities
-  \item Fabrication of single-cell low temperature electrolyzer prototypes using iridium and iridium oxide catalysts
-  \item Demonstrated, for the first time, the use of inkjet printing as a feasible technology to fabricate low temperature electrolyzer cells  (https://doi.org/10.1149/2.1101807jes)
-   \item Development of a novel nano-pompon-like iridium catalyst superstructure for low temperature fuel cell electrolyzers (https://doi.org/10.1016/j.jcat.2019.01.018)
-   \item Development of a novel technique for characterization of iridium catalyst degradation (https://doi.org/10.1016/j.jcat.2019.01.018)
- \item Development of novel micro-structures for high temperature solid oxide electrolyzers and fuel cells (https://doi.org/10.1016/j.electacta.2018.02.055)
-  \end{itemize}}
-The project currently involves 5 PhD students and 1 MSc, 5 faculty (1 in MECE, 3 in CHEME and 1 in Chem). We are also collaborating with Ionomr, a firm in BC (https://ionomr.com/).
-  \end{block}
-   \vfill
-\end{frame}
-
-\begin{frame}{Secanell Group}
-   \tikz [remember picture,overlay]
-    \node[yshift=-.5cm,xshift=0cm] at (current page.center)
-      {\includegraphics[width=4.5in]{../../secanell_3}}; \vspace{1cm}
    \vfill

### 2024-03-13 class_deck_slides.R

@@ -1,0 +1,2647 @@
+source("power_paper_base.R")
+source("aeso_scrapes.R")
+
+library(directlabels)
+library(cowplot)
+
+
+econ_course<-"../../Courses/econ_366_w24/slides/elec/"
+
+
+#load(file="data/metered_vols_data.Rdata")
+load(file="nrgstream/nrgstream_gen.Rdata") 
+nrgstream_gen <- nrgstream_gen %>% clean_names()
+  #ng_conv(id_sent=ID)%>%
+  #rename(time=time)
+
+renew_gen<-nrgstream_gen%>%filter(plant_type%in%c("WIND","SOLAR"))%>%
+  select(time,plant_type,gen)%>%
+  group_by(time,plant_type)%>%
+  summarize(gen=sum(gen,na.rm = T))%>%
+  group_by(time)%>%
+  mutate(renew_gen=sum(gen,na.rm = T))%>%
+  ungroup()%>%
+  pivot_wider(names_from = plant_type,values_from=gen)
+
+
+update_forecasts()
+load(file="data/forecast_data.Rdata") 
+forecast_data <- forecast_data %>% filter (he!="02*")
+
+
+hourly_patterns<-forecast_data %>% 
+  filter(!is.na(actual_posted_pool_price),!is.na(actual_ail))%>% 
+  assign_date_time_days()%>%
+  group_by(hour,year)%>%
+  summarize(ail=mean(actual_ail),price=mean(actual_posted_pool_price))%>%
+  mutate(hour=paste(hour,":00",sep=""),hour=factor(hour))
+
+library(gghighlight)
+
+hourly_graph<-  ggplot(hourly_patterns%>%filter(year>=year(Sys.Date())-11)) +
+  geom_line(aes(factor(hour),price,group=factor(year),color=factor(year)),size=.85)+
+  geom_line(data=hourly_patterns %>% filter(year==2023),aes(factor(hour),price,group=factor(year),color=factor(year)),size=1.5)+
+  geom_line(data=hourly_patterns %>% filter(year==2024),aes(factor(hour),price,group=factor(year),color=factor(year)),size=2.5)+
+  scale_color_viridis("",option = "C",discrete = T,direction=-1)+
+  scale_y_continuous(expand=c(0,0))+
+  paper_theme()+
+  scale_x_discrete(breaks=c("1:00","4:00","7:00","10:00",
+                            "13:00","16:00","19:00","22:00"))+
[…]

### 2024-03-13 ep_paper_graphs.R

@@ -8,18 +8,18 @@ library(cowplot)
 
 #load(file="data/metered_vols_data.Rdata")
 load(file="nrgstream/nrgstream_gen.Rdata") 
-nrgstream_gen <- nrgstream_gen %>% 
+nrgstream_gen <- nrgstream_gen %>% clean_names()
   #ng_conv(id_sent=ID)%>%
-  rename(time=Time)
+  #rename(time=time)
 
-renew_gen<-nrgstream_gen%>%filter(Plant_Type%in%c("WIND","SOLAR"))%>%
-  select(time,Plant_Type,gen)%>%
-  group_by(time,Plant_Type)%>%
+renew_gen<-nrgstream_gen%>%filter(plant_type%in%c("WIND","SOLAR"))%>%
+  select(time,plant_type,gen)%>%
+  group_by(time,plant_type)%>%
   summarize(gen=sum(gen,na.rm = T))%>%
   group_by(time)%>%
   mutate(renew_gen=sum(gen,na.rm = T))%>%
   ungroup()%>%
-  pivot_wider(names_from = Plant_Type,values_from=gen)
+  pivot_wider(names_from = plant_type,values_from=gen)
 
 
 
@@ -338,32 +338,32 @@ ggsave("images/prices_and_loads_ep.png",width = 14,height=8,bg="white",dpi = 300
 
 #align variable names
 df2 <- nrgstream_gen %>% filter(date<floor_date(max(date),"month"))%>% #trim to last full month of data
-  select(time,vol=gen,ID,AESO_Name,Plant_Type,Plant_Fuel,Capacity) %>%
-  filter(Plant_Type %in% c("COAL","COGEN","NGCC","WIND","SCGT","NGCONV","TRADE","HYDRO","SOLAR","OTHER"))%>%
+  select(time,vol=gen,id,aeso_name,plant_type,plant_fuel,capacity) %>%
+  filter(plant_type %in% c("COAL","COGEN","NGCC","WIND","SCGT","NGCONV","TRADE","HYDRO","SOLAR","OTHER"))%>%
   #filter(year(time)>=2010) %>% 
   left_join(forecast_data) %>% filter(!is.na(date))%>%
   #strip the AB-WECC tie since it's duplicate of AB-MT and AB-BC
-  filter(!ID %in% c("AB_WECC","AB_WECC_Exp","AB_WECC_Imp"))%>%
[…]

### 2024-03-13 nrgstream_gen.R

@@ -287,16 +287,18 @@ load(file="nrgstream/nrgstream_gen.RData")
 get_plant_info()
 #use this to update
 
-nrgstream_gen<-filter(nrgstream_gen,year(Time)<2023)
+nrgstream_gen<-filter(nrgstream_gen,year(Time)<2024)
 ##nrgstream_gen<-data_update(nrgstream_gen,2021)
 #nrgstream_gen<-data_update(nrgstream_gen,2022)
-nrgstream_gen<-data_update(nrgstream_gen,2023)
+#nrgstream_gen<-data_update(nrgstream_gen,2023)
 nrgstream_gen<-data_update(nrgstream_gen,2024)
 
 nrgstream_gen<-ng_conv(nrgstream_gen)
 
 
-#test_data<-nrgstream_gen %>% filter(is.na(Plant_Type))
+#nrgstream_gen<-nrgstream_gen%>%clean_names()
+
+#test_data<-nrgstream_gen %>% filter(is.na(plant_type))
 
 
 save(nrgstream_gen, file= "nrgstream/nrgstream_gen.RData")

### 2025-02-24 paper/paper_draft.tex
paper edits

@@
 \author{Andrew Leach\\
- University of Alberta
- \and
-Blake Shaffer\\University of Calgary}
+ University of Alberta}
  \begin{document}
@@
 
-We evaluate the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.
-Between 2015 and 2019, legislative and regulatory changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes, creating variation in average and marginal carbon prices both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
-We exploit this variance in treatment to provide unique evidence of carbon price pass-through in power markets.
-We show that even after adjusting for costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.
+This paper evaluates the response of wholesale power market participants in Alberta to a series of changes in greenhouse gas (GHG) emissions policies.
+Between 2015 and 2024, legislative and regulatory changes have affected carbon price levels, the rates of output-based allocations of emissions credits, and the coverage of carbon pricing across facilities of different sizes, creating variation in average and marginal carbon prices both across and within facility types and across and within portfolios held by major players in Alberta's wholesale power market.
+I exploit this variance in treatment to provide unique evidence of carbon price pass-through in power markets, and show that even after adjusting for costs imposed by carbon policies at the facility level, statistically significant changes in portfolio-level offer curves were induced leading to a shift in the merit order toward gas and away from coal which exceeded that which would have been predicted by carbon prices alone.
 These changes have accelerated the phase-out of coal-fired power in the province.
-We also show, in contrast to much of the recent literature, that net marginal carbon price costs were not fully pass-through to offers.
-In fact, our results suggest that less than half of the changes in carbon policy costs was reflected in shifts in the merit order of electricity price offers in Alberta.
+My results also demonstrate that, in contrast to much of the recent literature, net marginal carbon pricing costs were not fully passed-through to offers in the wholesale market.
+In fact, the results of my analysis suggest that less than half of the changes in carbon policy costs was reflected in shifts in the merit order of electricity price offers in Alberta.
 
-The traditional view of competitive firm response to carbon pricing would be an upward shift in each firm's supply function of an amount equal to the net incremental costs imposed by the policy.
+The traditional view of competitive firm response to carbon pricing would be an upward shift in each firm's supply function of an amount equal to the net, incremental costs imposed by the policy.
 For some facilities, there may be dynamic gains to participating in the market in a given hour (for example, if shutdown implies multi-hour restarts, as can be the case for coal generators), in which case the supply curve may lie below zero up to some positive quantity supplied in each hour.
@@
 However, above such a threshold, we should expect to see carbon price pass-through into power offers.
-However, our results in the Alberta context suggest less than perfect pass-through, even for units likely to be on the margin in the market in a given hour.
+However, my results in the Alberta context suggest less than perfect pass-through, even for units likely to be on the margin in the market in a given hour.
 
@@
 In the case of \citeauthor{fabra_reguant}, they find more-than-complete (110\%) pass-through of carbon costs in peak hours, and less-than-complete (60\%) pass-through in off-peak hours.
-\citeauthor{hintermann16_germany} finds pass-through rates of 81-111\% depending on the hour of the day, and also finds that costs are fully passed-through on average.
+Similarly, \citeauthor{hintermann16_germany} finds pass-through rates of 81-111\% depending on the hour of the day, and also finds that costs are fully passed-through on average.
 Our paper is unique in that we can measure facility-specific carbon policy costs (prices and the value of output-based allocations) precisely for long periods of time, which allows us to separately identify changes due to carbon pricing from changes due to other coincident market or climate factors.
@@
 For example, in 2018, the marginal carbon price did not change from 2017, but there were substantial changes in output-based allocations of emissions credits such that marginal net carbon costs went up for some facilities and down for others.
-We use this variation to identify pass-through more precisely than is possible with daily market fluctuations in a cap-and-trade market.
-Our results estimate much lower pass-through of carbon prices to electricity offers and thus to market prices than either \cite{fabra_reguant} or \cite{hintermann16_germany}.
+I use this variation to identify pass-through more precisely than is possible with daily market fluctuations in a cap-and-trade market.
+My results estimate much lower pass-through of carbon prices to electricity offers and thus to market prices than either \cite{fabra_reguant} or \cite{hintermann16_germany}.
 
 Average carbon policy costs varied over time and between fuel types in Alberta through the various changes to emissions policies since 2015.
-This links our work to \cite{cullen_mansur} which uses variations in fuel costs (coal vs.
+This links this paper to \cite{cullen_mansur} which uses variations in fuel costs (coal vs.
 natural gas) to identify impacts on emissions.
 They find substantial impacts from small changes in fuel costs which make natural-gas-fired plants more competitive relative to coal plants.
-The policy changes we study in Alberta, both the changes in output-based allocations and the carbon price increases, have had a similar effect on the coal-to-gas cost ratio as a reduction in natural gas prices.
-Their results also inform our results in the sense that the same gas price reductions seen in their study occurred in Alberta during our sample period, and we adjust for these impacts in our study design.
-We show that the degree to which carbon prices were passed through was most consistent with the theory for more flexible, combined-cycle natural gas facilities, while coal plants passed through a much lower share of increased emissions costs.
+The policy changes I study in Alberta, both the changes in output-based allocations and the carbon price increases, have had a similar effect on the coal-to-gas cost ratio as a reduction in natural gas prices.
+Their results also inform my results in the sense that the same gas price reductions seen in their study occurred in Alberta during our sample period, and we adjust for these impacts in our study design.
+I show that the degree to which carbon prices were passed through was most consistent with the theory for more flexible, combined-cycle natural gas facilities, while coal plants passed through a much lower share of increased emissions costs.
 
-The Alberta electricity market is highly concentrated, which links our paper to the literature following from the seminal paper of \cite{buchanan1969external} on externalities and related policies in the presence of market power, as well as to myriad studies of market power in electricity, e.g.
+The Alberta electricity market is highly-concentrated, which links our paper to the literature following from the seminal paper of \cite{buchanan1969external} on externalities and related policies in the presence of market power, as well as to myriad studies of market power in electricity, e.g.
 \cite{borenstein_bushnell_wolak_2002}.
@@
 \cite{miller_cement}, also looking at the cement industry, shows that fuel cost changes were more-than-fully reflected in downstream pricing.
-Similarly, as in \cite{holland_emissions_2012}, incomplete regulation or regulations which provide competing incentives will distort price-pass through.
-The latter also applies in our work since there are changes in market structure coincident with the policy changes we study for which we must account.
+And, as in \cite{holland_emissions_2012}, incomplete regulation or regulations which provide competing incentives will distort price-pass through.
+The latter also applies in the present study since there are changes in market structure coincident with the policy changes we study for which we must account.
 
-Another unique attribute of our study is that the policy variation over time and across facilities in Alberta also allows us to directly identify the pass-through of changes in output-based allocations of emissions credits (see \cite{fischerfox07} and \cite{fischerfox12}) to wholesale prices, as allocation rates did not vary systematically with changes in carbon prices.
-We cannot directly speak to the optimal allocations of emissions credits as in \cite{burtraw_palmer_2008}, but our results suggest that there is symmetric and incomplete pass-through of the marginal value of output-based allocations and carbon prices.
-To our knowledge, this has not been confirmed in other studies of carbon price pass-through.
+Another unique attribute of this study is that the policy variation over time and across facilities in Alberta also allows us to directly identify the pass-through of changes in output-based allocations of emissions credits (see \cite{fischerfox07} and \cite{fischerfox12}) to wholesale prices, as allocation rates did not vary systematically with changes in carbon prices.
+The analysis cannot directly speak to the optimal allocations of emissions credits as in \cite{burtraw_palmer_2008}, but my results suggest that there is symmetric and incomplete pass-through of the marginal value of output-based allocations and carbon prices.
+To my knowledge, this has not been confirmed in other studies of carbon price pass-through.
 There are limits to this result as well, since the changes in output-based allocations systematically disadvantaged coal plants relative to other fuel sources, and we coincident with substantial declines in natural gas prices.
@@
 
-The Alberta wholesale power market provides an excellent quasi-laboratory for our study because it is small, isolated, and the market is settled on a real-time, energy-only basis.
-As such, there are few confounding factors beyond what is observable in our data.
-Alberta's market is winter-peaking, with peak 2021 load of just under 12,000 MW.
+The Alberta wholesale power market provides an excellent quasi-laboratory for this study because it is small, isolated, and the market is settled on a real-time, energy-only basis.
+As such, there are few confounding factors beyond what is observable in publicly-available data.
+Alberta's market is winter-peaking, with peak load of just over 12,000 MW.
 Alberta has only three small interconnections to British Columbia, Montana, and Saskatchewan, with total capacity of less than 15\% of peak internal load.\footnote{The British Columbia and Montana tie lines have a joint operating capacity of 1500 MW for imports and 1325 MW for export, while the Saskatchewan tie line has a path rating of 150 MW.} The relative isolation of Alberta's market means that local prices and generation will be more affected by Alberta's policies than would be the case if traded power had a more important market share.
@@
 
-We compile over 13 years of hourly data describing market offer behaviour and a wide range of relevant characteristics.
-We measure and describe facility-level hourly offers of power into the market for the time period preceding the COVID-19 pandemic, although we also test this truncation of our data in later results.
-We combine power market data with detailed plant characteristics compiled from regulatory filings, emissions compliance data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data including hourly renewable generation, imports and exports, import and export capabilities and total and forecast internal loads.
+The analysis in this paper is based on over 15 years of hourly data describing market offer behaviour and a wide range of relevant characteristics.
+We combine power market data with detailed plant characteristics compiled from regulatory filings, emissions compliance data from Alberta's and Canada's air emissions reporting, weather data, commodity price information, and power market data including hourly renewable generation, imports and exports, import and export capabilities, and total and forecast internal loads.
 Finally, we include both observed carbon policy costs and the value of output-based allocations at the facility level.
@@
 For each of three different aggregation levels (market, plant type, and controlling entity), we build hourly merit orders of offered generation.
-We sample these merit orders at various support points which allows us to build a synthetic merit order with substantially-reduced dimensionality, and also to make the date comparable across time period with different total market sizes.
+We sample these merit orders at various support points which allows us to build a synthetic merit order with substantially-reduced dimensionality, and also to make the data comparable across time periods with different total market sizes.
 We also include measures of marginal carbon emissions policy costs along the merit order which we sample in parallel with the offered prices.
@@
 
-Finally, Alberta's market is highly concentrated, with a market structure that has evolved in ways relevant to our analysis during the relevant time period.
+Finally, Alberta's market is highly-concentrated, with a market structure that has evolved in ways relevant to our analysis during our sample period.
 When the market was restructured in 2000, a system of power purchase arrangements (PPAs) were used as an alternative to divestment \cite{kleit_electricity_2007}.
@@
 The PPAs included a cumbersomely-worded change-of-law clause which was interpreted to imply an option for PPA owners to exit these arrangements should they become unprofitable or ``more unprofitable" due to a change in law.
-In response to the changes in carbon pricing policy announced in June of 2015, all PPA owners exercised rights of return \cite{bankes_termination_2016}, and offer control over all of the legacy plants (3866 MW) reverted to the Balancing Pool between January 1, 2016 and May 5, 2016 (see Table \ref{tab:PPA}).
+In response to the changes in carbon pricing policy announced in June of 2015, all private PPA owners exercised rights of return \cite{bankes_termination_2016}, and offer control over all of the legacy plants (3866 MW) reverted to the Balancing Pool between January 1, 2016 and May 5, 2016 (see Table \ref{tab:PPA}).
 As detailed in \cite{leach_tombe_ppa}, this was caused nominally by the increase in stringency of the \textit{SGER}, but the adoption of the Climate Leadership Plan and the \textit{CCIR} meant that the expected value of the PPAs was negative for their owners \cite{leach_tombe_ppa}.
@@
 
-Alberta's three carbon pricing regimes each affected generating facilities in two ways.
-First, each of three (\emph{SGER}, \emph{CCIR}, and \emph{TIER}) regimes imposes a price on emissions.
-Second, each regime allocates emissions credits based on output or deemed avoided emissions.
-The net effect of these policies creates variation within and across facilities and facility types and over time.
+Alberta's three carbon pricing regimes each affected generating facilities through two channels.
+First, each of the three (\emph{SGER}, \emph{CCIR}, and \emph{TIER}) regimes imposes a price on emissions.
+Second, each regime allocated emissions credits based on output or deemed avoided emissions.
+The net effects of these policies create variation within and across facilities and facility types and over time.
 Consider, for example, Figure \ref{fig:sector_compliance_cost} which shows the mean value and range of annual compliance costs for larger generators in Alberta's fleet over the sample period.\footnote{Plants were only included in the sample for this figure if we had facility-specific compliance cost data.

### 2025-02-24 cdn_weather.R

@@ -1,5 +1,5 @@
 #get weather info for power papers
-
+source("power_paper_base.R")
 hourly_data_pull<-function(stn_id,min_yr,max_yr,stn_string="NA"){
 #hourly data
   #year_index<-2022
@@ -21,14 +21,16 @@ for(year_index in seq(min_yr,max_yr)) {
   for(month_index in seq(1,12)) {
   
   url = paste("http://climate.weather.gc.ca/climate_data/bulk_data_e.html?format=csv&stationID=",stn_id,"&Year=",year_index,"&Month=",sprintf("%02d",month_index),"&timeframe=1",sep="")
-  test<-read.csv(url, skip = 0)%>% clean_names() %>% select(long=longitude_x,lat=latitude_y,
-                                          stn="station_name",climate_id,
-                                          date_time="date_time_lst",
-                                          year,                
-                                          month,
-                                          day,
-                                          time="time_lst",
-                                          temp="temp_c")
+  test<-read.csv(url, skip = 0)%>% clean_names() %>% 
+                                          # select(long=longitude_x,lat=latitude_y,
+                                          # stn="station_name",climate_id,
+                                          # date_time="date_time_lst",
+                                          # year,                
+                                          # month,
+                                          # day,
+                                          # time="time_lst",
+                                          # temp="temp_c")%>%
+    I()
   data_store[[paste(month.abb[month_index],"-",year_index,sep = "")]]<-test
   #index_id<-index_id+1
 }}
@@ -40,6 +42,11 @@ temp_data$stn_id<-stn_id
 temp_data$stn<-stn_string
 temp_data
 }
+
+blatch<-hourly_data_pull(27214,min_yr = 2019,max_yr = 2023)
+save(blatch,file="blatch.Rdata")
+#blatch<-hourly_data_pull(27214,min_yr = 2024,max_yr = 2024)
+
 
 all_weather_data<-function(min_yr=2009,max_yr=2021){
   #min_yr<-2009
@@ -167,6 +174,8 @@ update_weather_data<-function(data_sent){
 }
 
 #save(data_sent,file="ab_power_temps.RData")  
+
+
 
 
 
